### 2018-07-30 documentation/paper.tex
Added TeX to .gitignore; created stub for LaTeX paper

\documentclass[12pt]{article}
\begin{document}
\title{Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector}
\author{Brendon Matusch}
\date{\today}
\maketitle
\end{document}

### 2018-07-30 documentation/paper.tex
Added brief description of methods

@@
 \documentclass[12pt]{article}
+
+% Use 1-inch margins even though the font is 12-point
+\usepackage[margin=1in]{geometry}
+\usepackage{textgreek}
+
 \begin{document}
-\title{Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector}
-\author{Brendon Matusch}
-\date{\today}
-\maketitle
+
+% A description of the various techniques used to discriminate between neutrons and alphas
+\section{Methods}
+
+Two key techniques have been used for the task of discrimination between neutrons/WIMP candidates and alpha particles in the PICO-60 bubble chamber.
+The first, Acoustic Parameter (AP), is a function of the 8-band Fourier transform \textbeta \textsubscript{8}.
+It makes use of the increased resonant energy in certain frequency bands, relative to neutrons, present in the majority of alpha particles.
+The second, Deep Learning (DL) consists of a wide variety of neural networks, which are trained on either the raw waveform \textomega{}, the arbitrary-resolution Fourier transform \textbeta \textsubscript{N}, or the image \textiota.
+
+% A section about AP, how it works, and its limitations
+\subsection{Acoustic Parameter}
+Placeholder
+
 \end{document}

### 2018-07-30 documentation/paper.tex
Added information on data formats

@@
 \usepackage{textgreek}
+\usepackage{textcomp}
 
 \begin{document}
+
+% Details on the data formats used as inputs for alpha/neutron discriminators
+\section{Data Formats}
+
+Alpha/neutron discriminators used a variety of different data formats to make predictions.
+These are collected from the two essential sensors present in the PICO-60 apparatus: piezoelectric microphones (piezos) and cameras.
+
+\subsection{Piezo-Derived}
+
+\subsubsection{Raw Waveform}
+
+The lowest-level data derived from the piezos is the raw waveform \textomega.
+\textomega{} consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
+They are each represented as an integer between \textminus65536 and 65535.
+Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because before this there is background noise, and afterwards there is a clipped signal produced by hydraulics repressurizing the vessel.
+(These sections contain no information and can be undesirably fit on by neural networks.)
+
+\subsubsection{Fourier Transform}
+
+The raw waveform \textomega{} is converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input, which is implemented in \texttt{numpy} as \texttt{rfft}.
+Since DFTs produce sequences of complex numbers, the magnitude of each element is computed.
+The full-resolution Fourier transform \textbeta \textsubscript{50,000} consists of a sequence of complex numbers, half the length of the raw waveform \textomega.
+
+Beyond this, arbitrary-resolution banded Fourier transforms \textbeta \textsubscript{N\textless50,000} can be computed by integrating the resonant energy over all
 

### 2018-07-30 training/gravitational_grid_search.py
Use only GTX 1060 for gravitational grid search

@@ -3,8 +3,13 @@
 # Created by Brendon Matusch, July 2018
 
 import copy
+import os
 
 import numpy as np
+
+# Use only the second GPU for training (GTX 1060)
+os.environ['CUDA_VISIBLE_DEVICES'] = '1'
+
 from keras.optimizers import SGD
 
 from data_processing.bubble_data_point import RunType

### 2018-07-31 visualization/visualize_harmonics.py
Added warning to protect against hearing damage

@@ -26,4 +26,5 @@ frequency_domain[:3000] = 0
 # Convert the modified frequency domain back into the time domain
 time_domain = frequency_to_time_domain(frequency_domain)
 # Visualize and save the time and frequency domains
+# WARNING: AUDIO IS EXTREMELY LOUD
 visualize_and_save_audio(time_domain, frequency_domain)

### 2018-07-31 documentation/paper.tex
Added information on 3D position

@@
 \textomega{} consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
-They are each represented as an integer between \textminus65536 and 65535.
+They are each represented as a 16-bit integer.
 Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because before this there is background noise, and afterwards there is a clipped signal produced by hydraulics repressurizing the vessel.
@@
 
-Beyond this, arbitrary-resolution banded Fourier transforms \textbeta \textsubscript{N\textless50,000} can be computed by integrating the resonant energy over all
+Beyond this, arbitrary-resolution banded Fourier transforms \textbeta \textsubscript{N\textless50,000} can be computed by integrating the resonant energy over all frequencies within each of a set of bands.
+They can be thought of as a method of signal downscaling, compressing the signal into a smaller number of data points.
+Particular band resolutions used include \textbeta \textsubscript{8} (which is the input to Acoustic Parameter and initial neural network experiments), \textbeta \textsubscript{5}, \textbeta \textsubscript{10}, \textbeta \textsubscript{20}, and \textbeta \textsubscript{40}.
 
-% A description of the various techniques used to discriminate between neutrons and alphas
-\section{Methods}
+\subsection{Camera-Derived}
 
-Two key techniques have been used for the task of discrimination between neutrons/WIMP candidates and alpha particles in the PICO-60 bubble chamber.
-The first, Acoustic Parameter (AP), is a function of the 8-band Fourier transform \textbeta \textsubscript{8}.
-It makes use of the increased resonant energy in certain frequency bands, relative to neutrons, present in the majority of alpha particles.
-The second, Deep Learning (DL) consists of a wide variety of neural networks, which are trained on either the raw waveform \textomega{}, the arbitrary-resolution Fourier transform \textbeta \textsubscript{N}, or the image \textiota.
+\subsubsection{Image Window Sequence}
 
-% A section about AP, how it works, and its limitations
-\subsection{Acoustic Parameter}
-Placeholder
+For each event that is captured, the 4 cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected.
+These raw images contain a large amount of extraneous information; they encompass the entire vessel.
+To reduce the input information, the image window sequence \textiota includes 50\texttimes50 cropped windows around the position of the bubble.
+The 71 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
+\textiota includes 5 frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
+
+\subsubsection{3D Position}
+
+The 3-dimensional position \textchi{} of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera.
+This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects \textbeta \textsubscript{8} for variations in amplitude which depend on the position of the bubble.
 

### 2018-07-31 documentation/paper.tex
Switched in-text Greek letters to math mode

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{textgreek}
 \usepackage{textcomp}
@@
 
-The lowest-level data derived from the piezos is the raw waveform \textomega.
-\textomega{} consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
+The lowest-level data derived from the piezos is the raw waveform $\omega$.
+$\omega$ consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
 They are each represented as a 16-bit integer.
@@
 
-The raw waveform \textomega{} is converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input, which is implemented in \texttt{numpy} as \texttt{rfft}.
+The raw waveform $\omega$ is converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input, which is implemented in \texttt{numpy} as \texttt{rfft}.
 Since DFTs produce sequences of complex numbers, the magnitude of each element is computed.
-The full-resolution Fourier transform \textbeta \textsubscript{50,000} consists of a sequence of complex numbers, half the length of the raw waveform \textomega.
+The full-resolution Fourier transform $\beta _{50,000}$ consists of a sequence of complex numbers, half the length of the raw waveform $\omega$.
 
-Beyond this, arbitrary-resolution banded Fourier transforms \textbeta \textsubscript{N\textless50,000} can be computed by integrating the resonant energy over all frequencies within each of a set of bands.
+Beyond this, arbitrary-resolution banded Fourier transforms $\beta _{N<50,000}$ can be computed by integrating the resonant energy over all frequencies within each of a set of bands.
 They can be thought of as a method of signal downscaling, compressing the signal into a smaller number of data points.
-Particular band resolutions used include \textbeta \textsubscript{8} (which is the input to Acoustic Parameter and initial neural network experiments), \textbeta \textsubscript{5}, \textbeta \textsubscript{10}, \textbeta \textsubscript{20}, and \textbeta \textsubscript{40}.
+Particular band resolutions used include $\beta _{8}$ (which is the input to Acoustic Parameter and initial neural network experiments), $\beta _{5}$, $\beta _{10}$, $\beta _{20}$, and $\beta _{40}$.
 
@@
 These raw images contain a large amount of extraneous information; they encompass the entire vessel.
-To reduce the input information, the image window sequence \textiota includes 50\texttimes50 cropped windows around the position of the bubble.
+To reduce the input information, the image window sequence $\iota$ includes 50\texttimes50 cropped windows around the position of the bubble.
 The 71 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
-\textiota includes 5 frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
+$\iota$ includes 5 frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
 
@@
 
-The 3-dimensional position \textchi{} of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera.
-This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects \textbeta \textsubscript{8} for variations in amplitude which depend on the position of the bubble.
+The 3-dimensional position $\chi$ of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera.
+This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
 

### 2018-07-31 documentation/paper.tex
Added cuts for basic quality and bubble multiplicity

@@
 
+\section{Data Cuts}
+
+Techniques in this study are trained and validated on a number of different data sets.
+The selection of these sets focused on the fundamental trade-off between quality and quantity of data; by setting a higher standard for the validity of training data, one has less data to train on.
+
+The initial data set $D$ to which these cuts are applied consists of all events recorded during PICO-60 run 2.
+
+\subsection{Basic Quality Cut}
+
+A certain number of cuts are necessarily applied to all data to ensure meaningful results.
+Otherwise, significant overfitting on biases in the data is likely.
+This basic quality cut $QualCut(D)$ consists of the following restrictions:
+\begin{itemize}
+    \item The run was not collected during engineering or testing ($\texttt{run\_type}\neq99$)
+    \item Recording was triggered by the camera ($\texttt{trigger\_main}=0$)
+    \item Acoustic Parameter is not erroneously large and negative ($log_{10}(\texttt{acoustic\_bubnum})>-100$)
+    \item Recorded more than 25 seconds after reaching target pressure ($\texttt{te}>25$)
+    \item The bubble position $\chi$ was successfully calculated ($[X, Y, Z]\neq[-100, -100, -100]$)
+\end{itemize}
+
+\subsection{Bubble Multiplicity Cut}
+
+PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles and WIMP candidates never scatter.
+Thus, no discriminator is required to handle these events, so they are removed.
+The bubble multiplicity cut $MultiCut(D)$ consists of the following restrictions:
+\begin{itemize}
+    \item Either 0 or 1 bubbles are detected based on images from the camera ($\texttt{nbub}<2$)
+    \item The number of bubbles approximated using the pressure transducer is close to 1 ($0.7<\texttt{dytranCZ}<1.3$)
+\end{itemize}
+
 \end{document}

### 2018-07-31 documentation/paper.tex
Partially documented wall cuts

@@
 This basic quality cut $QualCut(D)$ consists of the following restrictions:
+
 \begin{itemize}
@@
 The bubble multiplicity cut $MultiCut(D)$ consists of the following restrictions:
+
 \begin{itemize}
@@
 
+\subsection{Wall Cut}
+
+Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel.
+It can be desirable for a discriminator to handle these events correctly; however, Acoustic Parameter does not, and neither does the neural network used in the previous PICO-60 paper.
+Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.
+
+The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the pressure transducer), and the acoustic wall cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
+
 \end{document}

### 2018-07-31 documentation/paper.tex
Defined fiducial cut

@@
 
+\subsubsection{Fiducial Cut}
+
+The fiducial cut $FidWallCut(D)$ defines an spatial area along the walls of the vessel within which no events are accepted.
+It makes use of the bubble position $\chi$, the distance from the bubble to the center of the vessel $R$, and the distance from the bubble to the nearest wall, which is defined as $min(\texttt{Dwall}, \texttt{Dwall\_horiz})$.
+It is defined as follows:
+
+\begin{itemize}
+    \item $Z \leq 523$
+    \item Any of the following 4 restrictions are true:
+    \begin{itemize}
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (0 < Z \leq 400)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (Z \leq 0) \land (R \leq 100)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (Z \leq 0) \land (R > 100)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (Z > 400)$
+    \end{itemize}
+\end{itemize}
+
 \end{document}

### 2018-07-31 documentation/paper.tex
Use subscripts for position
Defined pressure and acoustic wall cuts

@@
     \item Recorded more than 25 seconds after reaching target pressure ($\texttt{te}>25$)
-    \item The bubble position $\chi$ was successfully calculated ($[X, Y, Z]\neq[-100, -100, -100]$)
+    \item The bubble position $\chi$ was successfully calculated
+
+    ($[\chi_{X}, \chi_{Y}, \chi_{Z}]\neq[-100, -100, -100]$)
 \end{itemize}
@@
 
-The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the pressure transducer), and the acoustic wall cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
+The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the pressure transducer), and the acoustic cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
 
@@
     \begin{itemize}
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (0 < Z \leq 400)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (Z \leq 0) \land (R \leq 100)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (Z \leq 0) \land (R > 100)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (Z > 400)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (0 < \chi_{Z} \leq 400)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (\chi_{Z} \leq 0) \land (R \leq 100)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (\chi_{Z} \leq 0) \land (R > 100)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (\chi_{Z} > 400)$
     \end{itemize}
@@
 
+\subsubsection{Pressure Cut}
+
+The pressure cut $PresWallCut(D)$ restricts the pressure detected by the pressure transducer, without any position corrections, to be close to 1 ($0.7<\texttt{dytranC}<1.3$).
+In combination with the acoustic cut $AcWallCut(D)$, this acts as a backup to the fiducial cut.
+
+\subsubsection{Acoustic Cut}
+
+The acoustic cut $AcWallCut(D)$ is defined using the banded Fourier transform $\beta_{8}$ as follows:
+
+$45 < mean(\beta_{8}[(0, 2), (0, 1)]) < 300$
+
+This takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
+
 \end{document}

### 2018-07-31 documentation/paper.tex
Reduced font size to 10pt; removed comments
Created stub for supervised learning section
Added section from interim report on performance analysis

@@
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 
-% Use 1-inch margins even though the font is 12-point
 \usepackage[margin=1in]{geometry}
@@
 
-% Details on the data formats used as inputs for alpha/neutron discriminators
 \section{Data Formats}
@@
 
+\section{Performance Analysis}
+
+The performance of neural networks relative to the Acoustic Parameter and the network used in the PICO-60 paper was done using the class-wise standard deviation.
+This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the binary discriminator in question (Acoustic Parameter or a neural network), corresponding to the sets of neutrons and alpha particles respectively.
+
+$S=std(N \cup A)$
+
+$C=(std(N \div S) + std(A \div S)) \div 2$
+
+The first step is to calculate the standard deviation S of the union of the two sets.
+This gives an indication of the scale of the overall distribution.
+When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
+While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, Acoustic Parameter has a significantly wider range.
+Normalization of the union prevents this from creating a bias where Acoustic Parameter would produce a higher standard deviation with a similarly proportioned error.
+
+The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually.
+This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are.
+Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off of those specific values, will produce a low class-wise standard deviation.
+
+This performance metric is used rather than a simple accuracy score because it takes in to account not only whether the discriminator is correct or incorrect, but how confident it is.
+The intent is for a decisive discriminator, which produces a wide separation between the two classes, to be preferred over one that produces a nebulous cloud of outputs with a seemingly arbitrary decision boundary.
+
+\section{Supervised Learning}
+
+Supervised learning was used to directly replace Acoustic Parameter.
+It takes advantage of labeled sets of neutron and alpha events, which are used as training.
+Several different network configurations were applied to this task.
+
+\subsection{Convolutional Neural Network for Raw Waveform Analysis}
+
 \end{document}

### 2018-07-31 documentation/paper.tex
Use spaces between paragraphs rather than indentation
Created bibliography; added information on fully convolutional networks; updated .gitignore with LaTeX temp files

@@
 \usepackage{textcomp}
+\usepackage{parskip}
+\usepackage[backend=biber, style=alphabetic, sorting=ynt]{biblatex}
+
+\addbibresource{paper.bib}
 
@@
 
+A convolutional neural network was used for analyzing the raw waveform directly, without any preprocessing whatsoever.
+This avoids any destruction of information, and should be possible, in theory, for a neural network to handle.
+
+For this task, a very deep 1-dimensional fully convolutional neural network was applied. The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.
+
+\printbibliography
+
 \end{document}

### 2018-07-31 documentation/paper.tex
More information on fully convolutional network

@@
 It takes advantage of labeled sets of neutron and alpha events, which are used as training.
-Several different network configurations were applied to this task.
+Several different neural network configurations were applied to this task.
 
@@
 This avoids any destruction of information, and should be possible, in theory, for a neural network to handle.
+For this task, a very deep 1-dimensional fully convolutional neural network was applied.
+The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.
 
-For this task, a very deep 1-dimensional fully convolutional neural network was applied. The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.
+Two different configurations were tested: one that operates strictly on the waveform with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it can learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
 

### 2018-08-02 documentation/paper.tex
Created objective section describing problems and essence of solution

@@
 \begin{document}
+
+\section{Background}
+
+\subsection{Objective}
+
+In experiments for dark matter detection, one of the most important and challenging hurdles to overcome is that of background radiation.
+Unwanted particles from a variety of sources can produce event signatures that appear very similar to those of dark matter candidates.
+In the PICO-60 bubble chamber experiment for detection of weakly interacting massive particles (WIMPs), these came in the form of neutrons (from unknown sources) and alpha particles (from decay of radon and associated daughter isotopes).
+
+To resolve this, physicists must determine a discriminator function that can, based on some sensory data, separate dark matter candidates from background radiation.
+Two general problems stand in the way:
+
+\begin{enumerate}
+    \item A detailed simulation of the physical environment is often used to produce an accurate discriminator.
+This takes a long time even for a team of experienced physicists, and will have to be updated whenever the physical variables of the experimental apparatus change.
+    \item The primary intention of many dark matter experiments, which is to reduce background radiation to the lowest level possible, means that there will be a very small number of background events detected.
+This places a severe limit on the amount of data that is available to optimize such a discriminator.
+    \item Furthermore, what little data is available to optimize a discriminator very often contains impurities, because whatever background radiation is present during WIMP detection runs is also present during calibration runs.
+These are difficult to separate without already having access to a functional discriminator, creating a "chicken or the egg" problem (from which the most common way to escape is to create a time- and labor-intensive simulation).
+\end{enumerate}
+
+The objective of this study is to solve all 3 of these problems using machine learning.
+This is done with a semi-supervised algorithm that simultaneously learns an effective discriminator and improves the purity of its own training data set.
+
+% \subsection{Existing Data}
+
+% During PICO-60 analysis, a function known as Acoustic Parameter was derived from simulations.
+It discriminates between alpha particles and nuclear recoils based on audio data (specifically the banded Fourier transform $\beta_{8}$, which is detailed in section \textbf{2.1.2}).
+
+% WIMP candidates in the cross-section and mass ranges that the PICO-60 detector is sensitive to are predicted to produce nuclear recoils with the same properties as those created by neutrons.
+The key difference is that neutrons frequently (but not always) scatter and produce multiple bubbles, where the extremely small predicted cross sections of WIMP candidates mean that they should almost invariably create just one bubble.
+These two properties together mean that neutron events producing one bubble can be used to calibrate a discriminator.
 

### 2018-08-02 documentation/paper.tex
Completed section on existing work; cleanup

@@
     \item Furthermore, what little data is available to optimize a discriminator very often contains impurities, because whatever background radiation is present during WIMP detection runs is also present during calibration runs.
-These are difficult to separate without already having access to a functional discriminator, creating a "chicken or the egg" problem (from which the most common way to escape is to create a time- and labor-intensive simulation).
+These are difficult to separate without already having access to a functional discriminator, creating a ``chicken or the egg'' problem (from which the most common way to escape is to create a time- and labor-intensive simulation).
 \end{enumerate}
@@
 
-% \subsection{Existing Data}
+\subsection{Existing Work}
 
-% During PICO-60 analysis, a function known as Acoustic Parameter was derived from simulations.
+During PICO-60 analysis, a function known as Acoustic Parameter was derived from simulations.
 It discriminates between alpha particles and nuclear recoils based on audio data (specifically the banded Fourier transform $\beta_{8}$, which is detailed in section \textbf{2.1.2}).
 
-% WIMP candidates in the cross-section and mass ranges that the PICO-60 detector is sensitive to are predicted to produce nuclear recoils with the same properties as those created by neutrons.
+WIMP candidates in the cross-section and mass ranges that the PICO-60 detector is sensitive to are predicted to produce nuclear recoils with the same properties as those created by neutrons.
 The key difference is that neutrons frequently (but not always) scatter and produce multiple bubbles, where the extremely small predicted cross sections of WIMP candidates mean that they should almost invariably create just one bubble.
-These two properties together mean that neutron events producing one bubble can be used to calibrate a discriminator.
+This means single-bubble neutron events can be used to optimize a discriminator to detect WIMP events.
+
+There was no explicit alpha particle calibration data collected as part of the PICO-60 experiment.
+However, the previous analysis revealed that there were no WIMP candidates detected.
+Also, based on the known ratio of multiple- to single-bubble neutron events, it is unlikely that there is more than 1 single-bubble neutron event throughout the background runs.
+This means they can be used as a high-purity alpha calibration set.
+
+While these two known event sets were only used for minor calibrations of Acoustic Parameter, there is no reason they cannot be used to optimize a machine learning discriminator from scratch.
 
@@
 
-Alpha/neutron discriminators used a variety of different data formats to make predictions.
+Discriminators can use a variety of different data formats to make predictions.
 These are collected from the two essential sensors present in the PICO-60 apparatus: piezoelectric microphones (piezos) and cameras.

### 2018-08-02 documentation/paper.tex
Added details on optimization practices; renamed introduction

@@
 
-\section{Background}
+\section{Introduction}
 
@@
 The objective of this study is to solve all 3 of these problems using machine learning.
-This is done with a semi-supervised algorithm that simultaneously learns an effective discriminator and improves the purity of its own training data set.
+This is done with a semi-supervised algorithm called \textit{iterative cluster nucleation}, which simultaneously learns an effective discriminator and improves the purity of its own training data set.
+
+\subsection{Hyperparameter Optimization}
+
+Learning algorithms in this study are optimized using grid searches, in which many different combinations of hyperparameters are tested.
+The computational work required for a grid search increases exponentially with the number of hyperparameters $n$ and the number of options per hyperparameter $m$; to be exact, its complexity is $O(m^{n})$.
+Thus, it is important to minimize the number of hyperparameters that are optimized at once.
+
+Since iterative cluster nucleation incorporates a neural network model, optimization is a much simpler task if the network's hyperparameters are already optimized.
+Thus, the optimization task is split into two distinct phases:
+
+\begin{enumerate}
+    \item Optimize the hyperparameters of a neural network on validation accuracy, to accurately reproduce the imperfect ground truths it is trained on.
+While this does not produce an optimal discriminator, it does reveal architectures that are capable, in essence, of learning to accomplish the discrimination task.
+    \item Using neural network hyperparameters that were determined to be effective, optimize the hyperparameters of the iterative cluster nucleation algorithm to reproduce Acoustic Parameter as closely as possible.
+Since Acoustic Parameter is not used to optimize the neural network directly at any point, models that excel at this task demonstrate an ability to learn the desired discriminator while ignoring the impurities in the training set.
+\end{enumerate}
 

### 2018-08-02 documentation/paper.tex
Removed specific mention of iterative cluster nucleation

@@
 The objective of this study is to solve all 3 of these problems using machine learning.
-This is done with a semi-supervised algorithm called \textit{iterative cluster nucleation}, which simultaneously learns an effective discriminator and improves the purity of its own training data set.
+This is done using semi-supervised algorithms that simultaneously learn an effective discriminator and improves the purity of its own training data set.
 
@@
 
-Since iterative cluster nucleation incorporates a neural network model, optimization is a much simpler task if the network's hyperparameters are already optimized.
+Since the semi-supervised learning algorithms used incorporate  neural network models, optimization is a much simpler task if the network's hyperparameters are already optimized.
 Thus, the optimization task is split into two distinct phases:
@@
 While this does not produce an optimal discriminator, it does reveal architectures that are capable, in essence, of learning to accomplish the discrimination task.
-    \item Using neural network hyperparameters that were determined to be effective, optimize the hyperparameters of the iterative cluster nucleation algorithm to reproduce Acoustic Parameter as closely as possible.
+    \item Using neural network hyperparameters that were determined to be effective, optimize the hyperparameters of the semi-supervised learning algorithm to reproduce Acoustic Parameter as closely as possible.
 Since Acoustic Parameter is not used to optimize the neural network directly at any point, models that excel at this task demonstrate an ability to learn the desired discriminator while ignoring the impurities in the training set.
@@
 
+\subsection{For Supervised Learning}
+
 The performance of neural networks relative to the Acoustic Parameter and the network used in the PICO-60 paper was done using the class-wise standard deviation.
@@
 
+\subsection{For Semi-Supervised Learning}
+
+Semi-supervised learning algorithms were scored based on how well they were able to learn to reproduce Acoustic Parameter. ADD MORE INFORMATION HERE BASED ON HOW ANALYSIS IS DONE.
+
 \section{Supervised Learning}

### 2018-08-02 documentation/paper.tex
Wording fix in raw waveform section

@@
 
-Two different configurations were tested: one that operates strictly on the waveform with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it can learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
+Two different configurations were tested: one that operates strictly on the waveform with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
 

### 2018-08-03 documentation/paper.tex
Started documenting iterative cluster nucleation
Added way more details on ICN

@@
 These are difficult to separate without already having access to a functional discriminator, creating a ``chicken or the egg'' problem (from which the most common way to escape is to create a time- and labor-intensive simulation).
+    \item When a discriminator is optimized on the impure training data available, for which ground truths are defined based on the calibration run of origin, it is really being optimized to discriminate between calibration runs, rather than between classes of particles.
+Without careful consideration, these two goals may not match up.
+For instance, a discriminator can undesirably fit on noise that varies between calibration runs.
 \end{enumerate}
@@
 The objective of this study is to solve all 3 of these problems using machine learning.
-This is done using semi-supervised algorithms that simultaneously learn an effective discriminator and improves the purity of its own training data set.
+This is done using semi-supervised algorithms that simultaneously learn an effective discriminator and improves the purity of their own training data sets.
 
@@
 
+\section{Semi-Supervised Learning}
+
+\subsection{Iterative Cluster Nucleation}
+
+Iterative cluster nucleation is a semi-supervised learning algorithm that takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
+The algorithm is initialized as follows:
+
+\begin{enumerate}
+    \item Take a subset of the training data, referred to as the seed set $\varsigma$. Use this as the beginning of the training set.
+    \item Remove classifications from any other available training examples. These create the unlabeled set $\upsilon$.
+    \item Compile and randomly initialize the weights of a neural network discriminator $NN(x)$ (where $x$ is the input data format).
+\end{enumerate}
+
+Parameterized by the initial seed threshold $j$ (on the order of 0.01) and the seed threshold multiplier $k$ (on the order of 1.05), the algorithm follows these iterative steps:
+
+\begin{enumerate}
+    \item Train $NN(x)$ for 30 epochs on $\varsigma$, using the imperfect ground truth values available.
+    \item Using the partially trained weights of $NN(x)$, run inference on the entirety of $\upsilon$, producing a set of predictions $p$.
+    \item Find predictions within $p$ that are within a distance of $j$ of either 0 or 1; as this is a binary classifier, such a prediction represents high confidence.
+Remove any such examples from $\upsilon$ and add them to $\varsigma$.
+The principle is that, when $NN(x)$ is trained on a relatively small set for short period of time, the few predictions within $p$ that are very confident are highly likely to be correct.
+    \item If no examples have been removed from $\upsilon$ and added to $\varsigma$, multiply $j$ by $k$ in place.
+This is done to increase the acceptance rate later in the training process, when most easily classifiable examples have been added to $\varsigma$.
+Otherwise, gridlock would occur, where certain examples could not be confidently classified given the training set.
+\end{enumerate}
+
+To arrive at an effective discriminator, the network first trains on a smaller amount of imperfect data.
+Because the ground truth data is imperfect, training on a relatively small set for a short period of time (potentially using regularization, in addition) ensures that the network will be less likely to overfit and its learned function will be simpler.
+This is in line with the objective, since, assuming the training inputs are normalized correctly, it should be simpler to discriminate between classes of particles than calibration runs.
+
+The essence of this algorithm, then, is a positive feedback loop in which, by including the network's most confident predictions in the training set, the simpler function that is desired is reinforced, even when the original ground truths teach the more complex function that would likely have been arrived at by a conventional supervised learning system.
+
 \printbibliography

### 2018-08-03 training/gravitational_grid_search.py
Renamed incorrect distortion root to distortion power

@@ -18,24 +18,24 @@ from data_processing.experiment_serialization import save_test
 from models.banded_frequency_network import create_model
 
 
-def gravitational_ground_truth_offsets(predictions: np.ndarray, distortion_root: float, gravity_multiplier: float) -> np.ndarray:
+def gravitational_ground_truth_offsets(predictions: np.ndarray, distortion_power: float, gravity_multiplier: float) -> np.ndarray:
     """Get an array of ground truth offsets based on a gravitational model where examples classified very close to one edge will be pulled toward that edge, and examples near the middle will make little difference"""
     # The function should pass through (0.5, 0), should change very little near that point, and should rapidly asymptote in the negative or positive directions as the prediction comes close to 0 or 1
     # First, scale the predictions to the range of -1 to 1
     predictions_scaled = (predictions - 0.5) * 2
     # Take the hyperbolic tangent so examples in the middle are affected minimally
     hyperbolic_tangent = np.tanh(predictions_scaled)
-    # Take the Nth root (removing the sign and multiplying it back in after so the negative side is the same as the positive side) of the hyperbolic tangent so that the area around 0 is squashed
-    root_distorted = np.sign(hyperbolic_tangent) * np.power(np.abs(hyperbolic_tangent), distortion_root)
+    # Take the Nth power (removing the sign and multiplying it back in after so the negative side is the same as the positive side) of the hyperbolic tangent so that the area around 0 is squashed
+    power_distorted = np.sign(hyperbolic_tangent) * np.power(np.abs(hyperbolic_tangent), distortion_power)
     # Multiply it by a constant so the gravitational offset does not dominate the training process
-    return root_distorted * gravity_multiplier
+    return power_distorted * gravity_multiplier
 
 
-# Iterate over different values for the number of definitive training examples, the gravity multiplier increment, the distortion root, and the stochastic gradient descent learning rate
+# Iterate over different values for the number of definitive training examples, the gravity multiplier increment, the distortion power, and the stochastic gradient descent learning rate
 for definitive_training_examples in [128, 256]:
     for gravity_multiplier_increment in [0.0005, 0.001, 0.003, 0.005, 0.008]:
         for learning_rate in [0.001, 0.003, 0.01, 0.03]:
-            for distortion_root in [3, 5, 7, 9, 11]:
+            for distortion_power in [3, 5, 7, 9, 11]:
                 # Print a few blank lines for separation
                 for _ in range(3):
                     print()
@@ -44,9 +44,9 @@ for definitive_training_examples in [128, 256]:
[…]

### 2018-08-03 training/gravitational_train.py

@@ -18,8 +18,8 @@ DEFINITIVE_TRAINING_EXAMPLES = 128
 # The value (greater than 1) to add to the gravity multiplier every epoch
 GRAVITY_MULTIPLIER_INCREMENT = 0.003
 
-# The root to use to flatten out the middle of the gravity function of the prediction spectrum
-DISTORTION_ROOT = 9
+# The power to use to flatten out the middle of the gravity function of the prediction spectrum
+DISTORTION_POWER = 9
 
 # Create an instance of the fully connected neural network
 model = create_model()
@@ -54,17 +54,17 @@ training_ground_truths = training_ground_truths.astype(float)
 training_ground_truths[DEFINITIVE_TRAINING_EXAMPLES:] = 0.5
 
 
-def gravitational_ground_truth_offsets(predictions: np.ndarray, distortion_root: float, gravity_multiplier: float) -> np.ndarray:
+def gravitational_ground_truth_offsets(predictions: np.ndarray, distortion_power: float, gravity_multiplier: float) -> np.ndarray:
     """Get an array of ground truth offsets based on a gravitational model where examples classified very close to one edge will be pulled toward that edge, and examples near the middle will make little difference"""
     # The function should pass through (0.5, 0), should change very little near that point, and should rapidly asymptote in the negative or positive directions as the prediction comes close to 0 or 1
     # First, scale the predictions to the range of -1 to 1
     predictions_scaled = (predictions - 0.5) * 2
     # Take the hyperbolic tangent so examples in the middle are affected minimally
     hyperbolic_tangent = np.tanh(predictions_scaled)
-    # Take the Nth root (removing the sign and multiplying it back in after so the negative side is the same as the positive side) of the hyperbolic tangent so that the area around 0 is squashed
-    root_distorted = np.sign(hyperbolic_tangent) * np.power(np.abs(hyperbolic_tangent), distortion_root)
+    # Take the Nth power (removing the sign and multiplying it back in after so the negative side is the same as the positive side) of the hyperbolic tangent so that the area around 0 is squashed
+    power_distorted = np.sign(hyperbolic_tangent) * np.power(np.abs(hyperbolic_tangent), distortion_power)
     # Multiply it by a constant so the gravitational offset does not dominate the training process
-    return root_distorted * gravity_multiplier
+    return power_distorted * gravity_multiplier
 
 
 # The gravity multiplier should start at 0 and is added to every epoch
[…]

### 2018-08-03 documentation/paper.tex
Defined gravitational differentiation function

@@
 
+\subsection{Gravitational Differentiation}
+
+Gravitational differentiation is a novel technique for gradient computation in the final layer of a neural network.
+It is intended for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
+
+It makes use of the gravitational differentiation function $GravDiff(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$.
+The function is defined as follows:
+
+$GravDiff(p, \psi, g) = g \cdot sgn(p) \cdot abs(tanh(2(p - 0.5))) ^ \psi$
+
 \subsection{Iterative Cluster Nucleation}

### 2018-08-03 documentation/paper.tex
Explained the absolute heck out of gravitational differentiation

@@
 
+The essential purpose of $GravDiff(p, \psi, g)$ is to allow confidently classified unlabeled examples to influence the training process, pushing them further towards the classification they are already near (like gravity), while ensuring examples with low-confidence classifications do not have a significant adverse effect.
+
+It can be considered a distortion of the hyperbolic tangent that flattens the central range and exaggerates the asymptote on either side.
+The network's prediction $p$ is first transformed so its range is 1 to -1 rather than 0 to 1 ($2(p - 0.5)$).
+Next, the sigmoidal hyperbolic tangent is applied, producing a value of 0 in the center and asymptotic slopes to -1 and 1 at the edges.
+
+However, what is really desired is a shallow slope in the center (so low-confidence network outputs in the range of, for instance, 0.2 to 0.8 will produce very shallow output gradients).
+Applying a relatively large exponent (on the order of 9) accomplishes this, because values less than 1 will rapidly approach 0.
+However, this only works with odd integer exponents (because they produce negative outputs for negative inputs), limiting the ability to fine-tune this function by changing $\psi$.
+To resolve this, a piecewise exponential function is used, where the absolute value is taken prior to applying the power and the sign is multiplied in afterwards.
+This permits use of the full range of curves produced by even and non-integer values of $\psi$.
+
 \subsection{Iterative Cluster Nucleation}

### 2018-08-03 documentation/paper.tex
Explained iterative algorithm

@@
 
+This function is applied in a training algorithm that is parameterized by the set of imperfectly labeled data $\varsigma$, the set of unlabeled training data $\upsilon$, a binary classification neural network $NN(x)$ (where $x$ is the input data format), and the gravitational multiplier increment $\delta_g$.
+The distortion power $\psi$ is assumed to be constant throughout a training run, and the gravitational multiplier $g$ is initialized at 0, gradually increasing during training.
+It iterates as follows:
+
+\begin{enumerate}
+    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for $\upsilon$.
+    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $GravDiff(p, \psi, g)$ and record the resulting gradients for the next iteration.
+    \item Increment $g$ by $\delta_g$.
+At the beginning, when $g = 0$, training will be entirely based on the labeled set $\varsigma$, and progressively, over the course of a run, the gravitational effect increases with $g$.
+\end{enumerate}
+
 \subsection{Iterative Cluster Nucleation}
@@
     \item Remove classifications from any other available training examples. These create the unlabeled set $\upsilon$.
-    \item Compile and randomly initialize the weights of a neural network discriminator $NN(x)$ (where $x$ is the input data format).
+    \item Compile and randomly initialize the weights of a binary classification neural network $NN(x)$.
 \end{enumerate}

### 2018-08-03 documentation/paper.tex
Rearranged semi-supervised learning section

@@
 
+Semi-supervised learning is the idea of training a machine learning model on a set of labeled data in addition to a set of unlabeled data.
+This has the advantage of requiring fewer labels to be collected than conventional supervised learning.
+However, it also has a very powerful advantage when training data is impure and the function to be learned is relatively simple (not simple in that it can be easily calculated by a person, but simple in that it does not require a huge set of training data).
+It allows a neural network to reinforce its own decisions.
+
+In both of the semi-supervised learning algorithms applied, to arrive at an effective discriminator, the network first trains on a smaller amount of imperfect data.
+Because the ground truth data is imperfect, training on a relatively small set for a short period of time (potentially using regularization, in addition) ensures that the network will be less likely to overfit and its learned function will be simpler.
+This is in line with the objective, since, assuming the training inputs are normalized correctly, it should be simpler to discriminate between classes of particles than calibration runs.
+
+The essence of this algorithm, then, is a positive feedback loop in which, by using the network's most confident predictions in the training process, the desired (simpler) function is reinforced, when a conventional supervised learning system would have learned the more complex function that is taught by the ground truths.
+
 \subsection{Gravitational Differentiation}
@@
     \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $GravDiff(p, \psi, g)$ and record the resulting gradients for the next iteration.
-    \item Increment $g$ by $\delta_g$.
+    \item Increment $g$ by $\delta_{g}$.
 At the beginning, when $g = 0$, training will be entirely based on the labeled set $\varsigma$, and progressively, over the course of a run, the gravitational effect increases with $g$.
@@
 
-To arrive at an effective discriminator, the network first trains on a smaller amount of imperfect data.
-Because the ground truth data is imperfect, training on a relatively small set for a short period of time (potentially using regularization, in addition) ensures that the network will be less likely to overfit and its learned function will be simpler.
-This is in line with the objective, since, assuming the training inputs are normalized correctly, it should be simpler to discriminate between classes of particles than calibration runs.
-
-The essence of this algorithm, then, is a positive feedback loop in which, by including the network's most confident predictions in the training set, the simpler function that is desired is reinforced, even when the original ground truths teach the more complex function that would likely have been arrived at by a conventional supervised learning system.
-
 \printbibliography

### 2018-08-03 documentation/paper.tex
Scrapped section on performance analysis

@@
 
-\section{Performance Analysis}
-
-\subsection{For Supervised Learning}
-
-The performance of neural networks relative to the Acoustic Parameter and the network used in the PICO-60 paper was done using the class-wise standard deviation.
-This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the binary discriminator in question (Acoustic Parameter or a neural network), corresponding to the sets of neutrons and alpha particles respectively.
-
-$S=std(N \cup A)$
-
-$C=(std(N \div S) + std(A \div S)) \div 2$
-
-The first step is to calculate the standard deviation S of the union of the two sets.
-This gives an indication of the scale of the overall distribution.
-When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
-While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, Acoustic Parameter has a significantly wider range.
-Normalization of the union prevents this from creating a bias where Acoustic Parameter would produce a higher standard deviation with a similarly proportioned error.
-
-The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually.
-This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are.
-Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off of those specific values, will produce a low class-wise standard deviation.
-
-This performance metric is used rather than a simple accuracy score because it takes in to account not only whether the discriminator is correct or incorrect, but how confident it is.
-The intent is for a decisive discriminator, which produces a wide separation between the two classes, to be preferred over one that produces a nebulous cloud of outputs with a seemingly arbitrary decision boundary.
-
-\subsection{For Semi-Supervised Learning}
-
-Semi-supervised learning algorithms were scored based on how well they were able to learn to reproduce Acoustic Parameter. ADD MORE INFORMATION HERE BASED ON HOW ANALYSIS IS DONE.
-
 \section{Supervised Learning}

### 2018-08-03 documentation/paper.tex
Added results for raw waveform CNN

@@
 
-Supervised learning was used to directly replace Acoustic Parameter.
-It takes advantage of labeled sets of neutron and alpha events, which are used as training.
-Several different neural network configurations were applied to this task.
+Supervised learning was the first step in the optimization procedure.
+It makes use of fully labeled calibration sets, which are used for training.
+Performance of these supervised neural networks is judged based on its accuracy at replication of the provided labels on a validation set (which is set aside and not trained on).
 
@@
 The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.
+(Its hyperparameters were modified significantly through manual empirical testing as well as grid searches.)
 
 Two different configurations were tested: one that operates strictly on the waveform with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
+The second configuration was introduced because the position corrections applied to $\beta_{8}$ in previous PICO-60 research were applied only to broad frequency bands and are thus not practical to apply to a full resolution waveform.
+
+During early tests with variations on $DeepConv(\omega)$, it was revealed to have an incredible capability to separate run types based on unwanted biases.
+It was able to separate the AmBe calibration run from combined background runs with 96\% accuracy, and to separate all calibration runs from background runs with 91\% accuracy.
+However, during these tests, the white noise at the beginning and hydraulic recompression sounds at the end were not cropped off, and the network was shown to be overfitting on this information.
+
+Once the audio was cropped properly, $DeepConv(\omega)$ achieved a maximum of 77\% accuracy on validation data.
+Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ managed 85\%.
+The maximum obtained on a grid search with position correction was 90\%.
+While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
+That said, neither of these accuracy statistics are particularly good, which does not bode well for this input format as a whole.
 

### 2018-08-03 documentation/paper.tex
Introduced MLP for FFT analysis

@@
 
-A convolutional neural network was used for analyzing the raw waveform directly, without any preprocessing whatsoever.
+A convolutional neural network was used for analyzing the raw waveform $\omega$ directly, without any preprocessing whatsoever.
 This avoids any destruction of information, and should be possible, in theory, for a neural network to handle.
@@
 
-Two different configurations were tested: one that operates strictly on the waveform with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
+Two different configurations were tested: one that operates strictly on $\omega$ with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
 The second configuration was introduced because the position corrections applied to $\beta_{8}$ in previous PICO-60 research were applied only to broad frequency bands and are thus not practical to apply to a full resolution waveform.
@@
 That said, neither of these accuracy statistics are particularly good, which does not bode well for this input format as a whole.
+
+\subsection{Multi-Layer Perceptron for Fourier Transform Analysis}
+
+A multi-layer perceptron (a neural network composed of fully connected layers) was applied to analysis of audio data in the frequency domain.
+Preprocessing by applying a Fourier transform makes it an easier task for the neural network to accomplish in some ways.
+If the frequency distribution and overall volume are indeed the most important factors for discrimination, the network can gather them straight from the Fourier transform $\beta_{N}$ rather than having to analyze $\omega$ to extract this information.
 

### 2018-08-03 documentation/paper.tex
Added details and statistics for FFT MLP

@@
 
-The raw waveform $\omega$ is converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input, which is implemented in \texttt{numpy} as \texttt{rfft}.
+The raw waveform $\omega$ is converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input.
 Since DFTs produce sequences of complex numbers, the magnitude of each element is computed.
@@
 
-Beyond this, arbitrary-resolution banded Fourier transforms $\beta _{N<50,000}$ can be computed by integrating the resonant energy over all frequencies within each of a set of bands.
+Beyond this, arbitrary-resolution banded Fourier transforms $\beta _{N<50,001}$ can be computed by integrating the resonant energy over all frequencies within each of a set of bands.
 They can be thought of as a method of signal downscaling, compressing the signal into a smaller number of data points.
@@
 The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.
+L2 regularization was used to alleviate overfitting, and batch normalization was applied to the input.
 (Its hyperparameters were modified significantly through manual empirical testing as well as grid searches.)
@@
 If the frequency distribution and overall volume are indeed the most important factors for discrimination, the network can gather them straight from the Fourier transform $\beta_{N}$ rather than having to analyze $\omega$ to extract this information.
+
+Once again, configurations with and without position input were tested: $FourierMLP(\beta_{N})$ and \\ $FourierMLPWithPos(\beta_{N}, \chi)$.
+Several different resolutions $N$ of $\beta_{N}$ were used.
+While there were variety of network architectures tested, most had a small number of layers (on the order of 3), applied dropout and L2 regularization, and used batch normalization once again.
+
+It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
+A perceptron trained on $\beta_{8}$ managed an excellent 97\% validation accuracy by this metric.
+Increasing the number of bands seemed initially not to provide any improvement; the best accuracy reached with $\beta_{20}$ was 93\%, and with $\beta_{40}$ it was 94\%.
+
+That experimentation applied $WallCut(D)$ to the training and validation sets.
+Where the higher resolution demonstrated significant power was when $WallCut(D)$ was not applied, and only $QualCut(MultiCut(D))$ was used.
+The network trained on $\beta_{40}$ still managed 92\% validation accuracy.
+However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%.
 

### 2018-08-03 documentation/paper.tex
Mentioned performance with position corrections; fixed some statistics

@@
 Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ managed 85\%.
-The maximum obtained on a grid search with position correction was 90\%.
+The maximum obtained on a grid search with position correction was 91\%.
 While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
@@
 It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
-A perceptron trained on $\beta_{8}$ managed an excellent 97\% validation accuracy by this metric.
+A perceptron trained on $\beta_{8}$ managed an excellent 98\% validation accuracy by this metric.
 Increasing the number of bands seemed initially not to provide any improvement; the best accuracy reached with $\beta_{20}$ was 93\%, and with $\beta_{40}$ it was 94\%.
+
+Unlike for the convolutional neural network trained on $\omega$, wall cuts did not improve performance at all.
+While $FourierMLP(\beta_{N})$ peaked at 98\% validation accuracy, $FourierMLPWithPos(\beta_{N})$ reached only 95\%.
+Nevertheless, they both performed better than any of the network architectures trained on $\omega$.
 

### 2018-08-03 documentation/paper.tex
Fixed statistics; added section on image CNN

@@
 While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
-That said, neither of these accuracy statistics are particularly good, which does not bode well for this input format as a whole.
+That said, neither of these accuracy statistics are particularly good (graphs of 85\% accuracy often appear effectively random), which does not bode well for this input format as a whole.
 
@@
 
-A multi-layer perceptron (a neural network composed of fully connected layers) was applied to analysis of audio data in the frequency domain.
+A multi-layer perceptron (a neural network composed of dense layers) was applied to analysis of audio data in the frequency domain.
 Preprocessing by applying a Fourier transform makes it an easier task for the neural network to accomplish in some ways.
@@
 However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%.
+
+\subsection{Convolutional Neural Network for Image Window Analysis}
+
+It is an open question whether or not there is any information in the image data $\iota$ that could be used to distinguish between particle classes.
+Relative to the extremely short period of time in which the bubble forms (on the scale of nanoseconds), the framerate of the camera (340Hz) is extremely slow.
+While the very early stages of bubble formation (when the sound is produced) are known to differ dependong on whether the bubble was created by a nuclear recoil or an alpha particle, it is unknown whether any visually apparent differences persist when the bubble is visible.
+
+In an effort to resolve this, a 2-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$.
+The network architecture consists of a moderate number of convolutional layers (on the scale of 9) and 3 dense layers at the end.
+L2 regularization was used on all layers, and dropout was additionally used on the dense layers.
+No position input is used (since there are no microphones that would require position correction), so the basic configuration is $ImageConv(\iota)$.
+
+Throughout a variety of different network architectures, the best validation accuracy obtained was 85\%.
+(The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.) It is likely fitting on some form of noise within $\iota$, since the network's predictions seem almost random when represented graphically.
+The fact that throughout many trials (including a grid search), no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.
 

### 2018-08-03 documentation/paper.tex
Added notices for pending results
Fixed equation formatting
Added citations; detailed software used

@@
 \usepackage{parskip}
+\usepackage{amsmath}
 \usepackage[backend=biber, style=alphabetic, sorting=ynt]{biblatex}
@@
     \item Recorded more than 25 seconds after reaching target pressure ($\texttt{te}>25$)
-    \item The bubble position $\chi$ was successfully calculated
-
-    ($[\chi_{X}, \chi_{Y}, \chi_{Z}]\neq[-100, -100, -100]$)
+    \item The bubble position $\chi$ was successfully calculated ($[\chi_{X}, \chi_{Y}, \chi_{Z}]\neq[-100, -100, -100]$)
 \end{itemize}
@@
 
+RESULTS PENDING
+
 \subsection{Iterative Cluster Nucleation}
@@
 
+RESULTS PENDING
+
+\section{Technologies}
+
+\subsection{Software}
+
+All programming for this study was done in Python 3.
+Keras \cite{keras} was used for all machine learning tasks.
+NumPy and SciPy were used for linear algebra and signal processing.
+ROOT, scikit-image and scikit-learn were used for data loading and storage.
+Matplotlib was used for data visualization.
+
 \printbibliography

### 2018-08-05 visualization/visualize_audio.py
Removed unnecessary imports

@@ -2,16 +2,13 @@
 """A tool for saving the audio from a bubble event as a graph in the time and frequency domains, as well as an audio clip that can be listened to"""
 # Created by Brendon Matusch, June 2018
 
-import itertools
 import os
 import sys
-import time
 
 import matplotlib.pyplot as plt
 import numpy as np
 from scipy.io import wavfile
 
-from data_processing.audio_domain_processing import time_to_frequency_domain
 from data_processing.bubble_data_point import load_bubble_audio, load_bubble_frequency_domain
 from data_processing.event_data_set import EventDataSet
 from utilities.verify_arguments import verify_arguments

### 2018-08-05 visualization/save_event_images.py
Created script for saving images in a grid

@@ -1,0 +1,40 @@
+#!/usr/bin/env python3
+"""A script for saving the first list images corresponding to a certain event, tiled together"""
+# Created by Brendon Matusch, August 2018
+
+import os
+import sys
+
+import numpy as np
+from skimage.io import imsave
+
+from data_processing.bubble_data_point import load_bubble_images
+from data_processing.event_data_set import EventDataSet
+from utilities.verify_arguments import verify_arguments
+
+# A unique bubble index is required
+verify_arguments('unique bubble index')
+# Load that one bubble from the data file
+identifier = int(sys.argv[1])
+bubble = EventDataSet.load_specific_indices([identifier])[0]
+# Load the images associated with this bubble, taking the first array
+images = load_bubble_images(bubble)[0]
+# These should be shaped into a 5 by 2 grid of images
+# Create such an empty array, using the side length of the square images
+window_side_length = images.shape[0]
+# Use the same data type as the source images
+grid = np.zeros((window_side_length * 5, window_side_length * 2), dtype=images.dtype)
+# Iterate over each axis of the grid
+for grid_x in range(5):
+    for grid_y in range(2):
+        # Calculate the corresponding top left corner of the image, using the window side length
+        left_x = grid_x * window_side_length
+        top_y = grid_y * window_side_length
+        # Get the corresponding index in the array of images
+        image_index = grid_x + (grid_y * 5)
+        # Get the corresponding window and place it in the grid image
+        grid[left_x:left_x + window_side_length, top_y:top_y + window_side_length] = images[:, :, image_index]
+# Transpose the grid because it is expected in (X, Y) dimension order
+grid = np.transpose(grid)
+# Save the grid image to disk
+imsave(os.path.expanduser('~/image_grid.png'), grid)

### 2018-08-05 documentation/paper.tex
Added section on hardware
Fixed slight error
Added images to show audio and image data

@@
 \usepackage[backend=biber, style=alphabetic, sorting=ynt]{biblatex}
+\usepackage{graphicx}
+\graphicspath{{./images/}}
+\usepackage{wrapfig}
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{audio}
+    \caption{\label{} An example of the cropped audio waveform $\omega$ and full resolution Fourier transform $\beta_{50,001}$.}
+\end{figure}
+
 \subsection{Camera-Derived}
@@
 $\iota$ includes 5 frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.7\textwidth]{image_grid}
+    \caption{\label{} An example of the image sequence $\iota$.}
+\end{figure}
 
@@
 
+\subsection{Hardware}
+
+Training was done on the Graham supercomputer at the University of Waterloo, accessed through Compute Canada and SHARCNET.
+I would like to thank Ken Clark and SNOLAB very much for generously providing access to this hardware.
+
 \printbibliography

### 2018-08-05 visualization/graph_wall_event_positions.py
Added legend for visual position graph

@@ -3,6 +3,8 @@
 # Created by Brendon Matusch, July 2018
 
 import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+from matplotlib.patches import Patch
 
 from data_processing.event_data_set import EventDataSet
 
@@ -36,6 +38,14 @@ for x, y, color, shape in zip(x_positions, y_positions, fiducial_cut_colors, aud
         c=color,
         marker=shape
     )
+# Create patches for the 2 colors used to describe the status of the fiducial cut
+passes_fiducial_patch = Patch(color='b', label='Passes fiducial cut')
+fails_fiducial_patch = Patch(color='r', label='Fails fiducial cut')
+# Create line objects to represent the 2 markers used for audio and pressure cuts
+passes_audio_line = Line2D([], [], marker='o', label='Passes audio and pressure wall cuts')
+fails_audio_line = Line2D([], [], marker='x', label='Fails audio and/or pressure wall cuts')
+# Display them in a legend
+plt.legend(handles=[passes_fiducial_patch, fails_fiducial_patch, passes_audio_line, fails_audio_line])
 # Label the X and Y axes
 plt.xlabel('X Position')
 plt.ylabel('Y Position')

### 2018-08-05 visualization/graph_wall_event_positions.py
Filter out extreme outliers where the position is clearly wrong

@@ -15,6 +15,11 @@ events = [
     event for event in events
     if EventDataSet.passes_standard_cuts(event)
 ]
+# Filter out extreme outlier events where the Y position's magnitude is greater than or equal to 150
+events = [
+    event for event in events
+    if abs(event.y_position) < 150
+]
 # Get the X and Y positions of the events
 x_positions = [event.x_position for event in events]
 y_positions = [event.y_position for event in events]

### 2018-08-05 visualization/graph_binary_predictions.py
Added legend to binary prediction graphing script

@@ -7,6 +7,7 @@ import random
 import sys
 
 import matplotlib.pyplot as plt
+from matplotlib.patches import Patch
 import numpy as np
 
 from data_processing.experiment_serialization import load_test
@@ -64,6 +65,11 @@ plt.scatter(
     y=network_outputs,
     c=point_colors
 )
+# Create patches that describe the 2 differently colored classes
+background_patch = Patch(color='r', label='Background radiation runs')
+calibration_patch = Patch(color='b', label='Neutron calibration runs')
+# Display them in a legend
+plt.legend(handles=[background_patch, calibration_patch])
 # Label the X and Y axes
 plt.xlabel('Logarithmic Acoustic Parameter')
 plt.ylabel('Neural Network Prediction')

### 2018-08-05 visualization/graph_binary_predictions.py
Specified figure size

@@ -59,6 +59,8 @@ point_colors = [
     'r' if ground_truth else 'b'
     for ground_truth in ground_truths
 ]
+# Set the size of the resulting graph (it should be standard across all such graphs)
+plt.figure(figsize=(8, 6))
 # Scatter plot the acoustic parameter on the X axis and the neural network's predictions on the Y axis
 plt.scatter(
     x=acoustic_parameters,

### 2018-08-05 documentation/paper.tex
Added wall event positions graph
Added graph for waveform grid search
Clarified lack of decisiveness

@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{wall_event_positions}
+    \caption{\label{} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
+\end{figure}
+
 \section{Supervised Learning}
@@
 
-A convolutional neural network was used for analyzing the raw waveform $\omega$ directly, without any preprocessing whatsoever.
+A convolutional neural network was used for analyzing the raw waveform $\omega$ directly, without any preprocessing whatsoever, and without application of $WallCut(D)$.
 This avoids any destruction of information, and should be possible, in theory, for a neural network to handle.
@@
 While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
-That said, neither of these accuracy statistics are particularly good (graphs of 85\% accuracy often appear effectively random), which does not bode well for this input format as a whole.
+
+It is difficult to say with a high degree of confidence whether or not this is an effective discriminator, since Acoustic Parameter is not decisive on data without $WallCut(D)$.
+However, the lack of decisiveness observed even when accuracy was relatively high, as seen in Figure \ref{waveform_best_validation}, implies this is a suboptimal technique.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{waveform_best_validation}
+    \caption{\label{waveform_best_validation} Accuracy of $DeepConvWithPos(\omega, \chi)$ compared to Acoustic Parameter at discriminating between run types without $WallCut(D)$ (91\% accuracy).}
+\end{figure}
 

### 2018-08-05 documentation/paper.tex
Added information and graph on Fourier transform MLP

@@
 
-Once again, configurations with and without position input were tested: $FourierMLP(\beta_{N})$ and \\ $FourierMLPWithPos(\beta_{N}, \chi)$.
+Once again, configurations with and without position input were tested.
+This time, position corrections were optionally included when position input was not used.
+The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$, $FourierMLP(PosCor(\beta_{N}))$ and $FourierMLPWithPos(\beta_{N}, \chi)$.
 Several different resolutions $N$ of $\beta_{N}$ were used.
@@
 It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
-A perceptron trained on $\beta_{8}$ managed an excellent 98\% validation accuracy by this metric.
+$FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy by this metric.
 Increasing the number of bands seemed initially not to provide any improvement; the best accuracy reached with $\beta_{20}$ was 93\%, and with $\beta_{40}$ it was 94\%.
 
-Unlike for the convolutional neural network trained on $\omega$, wall cuts did not improve performance at all.
-While $FourierMLP(\beta_{N})$ peaked at 98\% validation accuracy, $FourierMLPWithPos(\beta_{N})$ reached only 95\%.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{fourier_mlp_no_correction}
+    \caption{\label{fourier_mlp_no_correction} Accuracy of $FourierMLP(\beta_{8})$ compared to Acoustic Parameter at discriminating between run types (98\% accuracy).}
+\end{figure}
+
+Unlike for the convolutional neural network trained on $\omega$, position input did not improve performance at all.
+In fact, $FourierMLP(\beta_{8})$, performed better than $FourierMLPWithPos(\beta_{8}, \chi)$, which reached only 95\% validation accuracy.
 Nevertheless, they both performed better than any of the network architectures trained on $\omega$.
 
-That experimentation applied $WallCut(D)$ to the training and validation sets.
-Where the higher resolution demonstrated significant power was when $WallCut(D)$ was not applied, and only $QualCut(MultiCut(D))$ was used.
-The network trained on $\beta_{40}$ still managed 92\% validation accuracy.
+In all previous experimentation, $WallCut(D)$ was applied to the training and validation sets.
+The higher resolution demonstrated significant power when $WallCut(D)$ was not applied, and only $QualCut(MultiCut(D))$ was used.
+$FourierMLPWithPos(\beta_{40}, \chi)$ still managed 92\% validation accuracy.
 However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%.

### 2018-08-05 documentation/paper.tex
Added graph for image window classification

@@
 Relative to the extremely short period of time in which the bubble forms (on the scale of nanoseconds), the framerate of the camera (340Hz) is extremely slow.
-While the very early stages of bubble formation (when the sound is produced) are known to differ dependong on whether the bubble was created by a nuclear recoil or an alpha particle, it is unknown whether any visually apparent differences persist when the bubble is visible.
+While the very early stages of bubble formation (when the sound is produced) are known to differ depending on whether the bubble was created by a nuclear recoil or an alpha particle, it is unknown whether any visually apparent differences persist when the bubble is visible.
 
@@
 Throughout a variety of different network architectures, the best validation accuracy obtained was 85\%.
-(The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.) It is likely fitting on some form of noise within $\iota$, since the network's predictions seem almost random when represented graphically.
+(The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.) It is likely fitting on some form of noise within $\iota$, since the network's predictions seem almost random when represented graphically (as seen in Figure \ref{image_window}).
 The fact that throughout many trials (including a grid search), no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{image_window}
+    \caption{\label{image_window} Accuracy of $ImageConv(\iota)$ compared to Acoustic Parameter at discriminating between run types (85\% accuracy).}
+\end{figure}
 

### 2018-08-05 documentation/paper.tex
Removed unnecessary hardware section

@@
 
-\section{Technologies}
-
-\subsection{Software}
+\section{Software}
 
 All programming for this study was done in Python 3.
-Keras \cite{keras} was used for all machine learning tasks.
+Keras \cite{keras}, running on a TensorFlow backend, was used for all machine learning tasks.
 NumPy and SciPy were used for linear algebra and signal processing.
@@
 
-\subsection{Hardware}
-
-Training was done on the Graham supercomputer at the University of Waterloo, accessed through Compute Canada and SHARCNET.
-I would like to thank Ken Clark and SNOLAB very much for generously providing access to this hardware.
-
 \printbibliography

### 2018-08-05 documentation/paper.tex
Added some ICN information to paper

@@
 
-The essence of this algorithm, then, is a positive feedback loop in which, by using the network's most confident predictions in the training process, the desired (simpler) function is reinforced, when a conventional supervised learning system would have learned the more complex function that is taught by the ground truths.
+The essence of these 2 algorithms, then, is a positive feedback loop in which, by using the network's most confident predictions in the training process, the desired (simpler) function is reinforced, when a conventional supervised learning system would have learned the more complex function that is taught by the ground truths.
+
+Performance of the supervised learning systems was quantified by how accurately they reproduced the labels provided to them (on a validation set).
+They were observed to learn to separate the calibration runs from background runs, including the percentage of alpha particles present in the calibration runs.
+One of the main intents of semi-supervised learning is to address the impure training data.
+Therefore, their performance is quantified by how well it is able to replicate Acoustic Parameter, rather than simply reproduce the labels provided.
 
@@
 
-RESULTS PENDING
-
 \subsection{Iterative Cluster Nucleation}
@@
 
-RESULTS PENDING
+During a grid search of these parameters (in addition to regularization parameters of the network), the best accuracy achieved for replication of Acoustic Parameter was exactly 100\%.
+The mean of all epochs within that training run was 98\%.
+Further demonstrating this is a powerful algorithm and that high performance was not a one-off, 35\% of the training runs replicated Acoustic Parameter with a mean accuracy greater than 92\%.
 

### 2018-08-05 statistics/grid_search_maximum_ap_similarity.py
Clarified names of grid search statistics scripts; created stub for maximum AP similarity

@@ -1,0 +1,16 @@
+#!/usr/bin/env python3
+"""Given a grid search AP similarity statistics file, calculate the maximum accuracy (as opposed to mean) for every run"""
+# Created by Brendon Matusch, August 2018
+
+import os
+import sys
+
+from utilities.verify_arguments import verify_arguments
+
+# A path to a grid search AP similarity statistics file is expected
+verify_arguments('path to grid search AP similarity statistics file')
+# Load the full contents of the file
+with open(os.path.expanduser(sys.argv[1])) as statistics_file:
+    statistics_file_contents = statistics_file.read()
+# Split it on the separator lines that contain the mean disagreements, removing the last element because it only contains mean data
+text_by_run = statistics_file_contents.split('Mean')[:-1]

### 2018-08-05 statistics/grid_search_maximum_ap_similarity.py
Removed pointless statistics script

@@ -1,16 +1,0 @@
-#!/usr/bin/env python3
-"""Given a grid search AP similarity statistics file, calculate the maximum accuracy (as opposed to mean) for every run"""
-# Created by Brendon Matusch, August 2018
-
-import os
-import sys
-
-from utilities.verify_arguments import verify_arguments
-
-# A path to a grid search AP similarity statistics file is expected
-verify_arguments('path to grid search AP similarity statistics file')
-# Load the full contents of the file
-with open(os.path.expanduser(sys.argv[1])) as statistics_file:
-    statistics_file_contents = statistics_file.read()
-# Split it on the separator lines that contain the mean disagreements, removing the last element because it only contains mean data
-text_by_run = statistics_file_contents.split('Mean')[:-1]

### 2018-08-05 statistics/grid_search_ap_similarity_statistics.py
Got maximum as well as mean disagreements

@@ -46,7 +46,10 @@ for file_path in file_paths:
 # Iterate over the run identifiers in the dictionary
 for run_identifier in disagreements_by_hyperparameters:
     # Print the mean number of disagreements throughout the entire run
-    print(f'Mean {np.mean(list(disagreements_by_hyperparameters[run_identifier].values()))} disagreements in run {run_identifier}')
+    disagreement_values = list(disagreements_by_hyperparameters[run_identifier].values())
+    print(f'Mean {np.mean(disagreement_values)} disagreements in run {run_identifier}')
+    # Print the maximum number of disagreements throughout the run
+    print(f'Maximum {np.amax(disagreement_values)} disagreements in run {run_identifier}')
     # Iterate over the specific file paths in this run
     for file_path in disagreements_by_hyperparameters[run_identifier]:
         # If there is a small number of disagreements, print out the path and number

### 2018-08-05 documentation/paper.tex
Added additional information about ICN graphs and updated image

@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{icn_grid_search}
+    \caption{\label{icn_grid_search} Accuracy of best iterative cluster nucleation training run compared to Acoustic Parameter and original ground truths (100\% accuracy).}
+\end{figure}
+
+As seen in Figure \ref{icn_grid_search}, the network learned to replicate Acoustic Parameter very accurately.
+Interestingly, the examples with Acoustic Parameter values between the neutron and alpha particle ranges were also classified as such by the network.
+
 \section{Software}

### 2018-08-06 documentation/paper.tex
Reorganized introduction

@@
 \usepackage{graphicx}
+\usepackage{wrapfig}
+
 \graphicspath{{./images/}}
-\usepackage{wrapfig}
 
@@
 
-\section{Introduction}
-
-\subsection{Objective}
+\section{Objective}
 
@@
 Unwanted particles from a variety of sources can produce event signatures that appear very similar to those of dark matter candidates.
-In the PICO-60 bubble chamber experiment for detection of weakly interacting massive particles (WIMPs), these came in the form of neutrons (from unknown sources) and alpha particles (from decay of radon and associated daughter isotopes).
 
@@
 \begin{enumerate}
-    \item A detailed simulation of the physical environment is often used to produce an accurate discriminator.
+    \item A detailed model of the physical environment is often used to produce an accurate discriminator.
 This takes a long time even for a team of experienced physicists, and will have to be updated whenever the physical variables of the experimental apparatus change.
+
     \item The primary intention of many dark matter experiments, which is to reduce background radiation to the lowest level possible, means that there will be a very small number of background events detected.
 This places a severe limit on the amount of data that is available to optimize such a discriminator.
-    \item Furthermore, what little data is available to optimize a discriminator very often contains impurities, because whatever background radiation is present during WIMP detection runs is also present during calibration runs.
-These are difficult to separate without already having access to a functional discriminator, creating a ``chicken or the egg'' problem (from which the most common way to escape is to create a time- and labor-intensive simulation).
-    \item When a discriminator is optimized on the impure training data available, for which ground truths are defined based on the calibration run of origin, it is really being optimized to discriminate between calibration runs, rather than between classes of particles.
-Without careful consideration, these two goals may not match up.
-For instance, a discriminator can undesirably fit on noise that varies between calibration runs.
+
+    Furthermore, what little data is available to optimize a discriminator very often contains impurities, because whatever background radiation is present during WIMP detection runs is also present during calibration runs.
+These are difficult to separate without already having access to a functional discriminator, creating a ``chicken or the egg'' problem (from which one of the most common ways to escape is to create a time- and labor-intensive simulation).
 \end{enumerate}
 
-The objective of this study is to solve all 3 of these problems using machine learning.
-This is done using semi-supervised algorithms that simultaneously learn an effective discriminator and improves the purity of their own training data sets.
+The objective of this study is to explore a potential methodology that resolves both of these problems using techniques based on machine learning.
+These include supervised learning algorithms that learn to separate calibration data sets, and semi-supervised learning algorithms that simultaneously learn an effective discriminator and improve the purity of their own training data sets.
+
+\section{Previous PICO-60 Analysis}
+
+\subsection{Background}
+
+The PICO-60 bubble chamber experiment for detection of weakly interacting massive particles (WIMPs) was completed in 2017.
+It provides an excellent basis on which to develop and test the efficacy of such techniques as a potential tool for event discrimination, relative to the conventional techniques used during initial analysis of the experiment.
+
+WIMP candidates in the cross section and mass ranges that the PICO-60 detector is sensitive to are predicted to produce nuclear recoils with the same properties as those created by neutrons.
+The key difference is that neutrons frequently (but not always) scatter and produce multiple bubbles, where the extremely small predicted cross sections of WIMP candidates mean that they should almost invariably create just one bubble.
+This means single-bubble neutron events can be used to optimize a discriminator to detect WIMP events.
+
+The approximate expected ratio between single-bubble and multiple-bubble events generated by neutrons is known.
+A measured quantity of single-bubble events generated by nuclear recoils, in excess of this prediction, would be indicative of WIMP interactions.
+Consequently, it is imperative to isolate only those events associated with nuclear recoils.
+
+\subsection{Discrimination Techniques}
+
+In this study, newly developed methodologies were compared to the following discrimination techniques developed previously for the PICO-60 experiment.
+
+\subsubsection{Acoustic Parameter}
+
+Acoustic Parameter is the technique used for event discrimination in the original PICO-60 analysis.
+It is a function that discriminates between alpha particles and nuclear recoils based on audio data.
+It was derived based on a physical model of the bubble chamber, and was tuned using calibration runs using neutron sources to produce events in the PICO-60 apparatus.
+
+For Acoustic Parameter to be sufficiently accurate, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$), and preprocessed with the position correction function $PosCor(\chi)$.
+In addition, a set of data cuts must be applied to remove certain classes of events.
+These conversion, preprocessing, and data cutting steps were re-applied during this study, and are described in section \ref{data_formats}.
+
+\subsubsection{Machine Learning}
+
+As part of the PICO-60 experiment, some testing was conducted to determine whether machine learning could be a viable technique to consider for a discrimination function.
+The results, while not definitive, indicated that the general technique warranted further investigation.
+
+\subsection{Data Sources and Formats} \label{data_formats}
+
+Discriminators can use a variety of different data formats to make predictions.
+These are collected from the two essential sensors present in the PICO-60 apparatus: piezoelectric microphones (piezos) and cameras.
+
+\subsubsection{Piezo-Derived}
+
+\textit{Raw Waveform}
+
+The lowest-level data derived from the piezos is the raw waveform $\omega$.
+$\omega$ consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
+They are each represented as a 16-bit integer.
+Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because before this there is background noise, and afterwards there is a clipped signal produced by hydraulics repressurizing the vessel.
+(These sections contain no information and can be undesirably fit on by neural networks.)
+
+\textit{Fourier Transform}
+
+The raw waveform $\omega$ is converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input.
+Since DFTs produce sequences of complex numbers, the magnitude of each element is computed.
+The full-resolution Fourier transform $\beta _{50,000}$ consists of a sequence of complex numbers, half the length of the raw waveform $\omega$.
+
+Beyond this, arbitrary-resolution banded Fourier transforms $\beta _{N<50,001}$ can be computed by integrating the resonant energy over all frequencies within each of a set of bands.
+They can be thought of as a method of signal downscaling, compressing the signal into a smaller number of data points.
+Particular band resolutions used include $\beta _{8}$ (which is the input to Acoustic Parameter and initial neural network experiments), $\beta _{5}$, $\beta _{10}$, $\beta _{20}$, and $\beta _{40}$.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{audio}
+    \caption{\label{} An example of the cropped audio waveform $\omega$ and full resolution Fourier transform $\beta_{50,001}$.}
+\end{figure}
+
+\subsubsection{Camera-Derived}
+
+\textit{Image Window Sequence}
+
+For each event that is captured, the 4 cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected.
+These raw images contain a large amount of extraneous information; they encompass the entire vessel.
+To reduce the input information, the image window sequence $\iota$ includes 50\texttimes50 cropped windows around the position of the bubble.
+The 71 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
+$\iota$ includes 5 frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.7\textwidth]{image_grid}
+    \caption{\label{} An example of the image sequence $\iota$.}
+\end{figure}
+
+\textit{3D Position}
+
+The 3-dimensional position $\chi$ of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera.
+This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
+
+\subsection{Data Cuts}
+
+Techniques in this study are trained and validated on a number of different data sets.
+The selection of these sets focused on the fundamental trade-off between quality and quantity of data; by setting a higher standard for the validity of training data, one has less data to train on.
+
+The initial data set $D$ to which these cuts are applied consists of all events recorded during PICO-60 run 2.
+
+\subsubsection{Basic Quality Cut}
+
+A certain number of cuts are necessarily applied to all data to ensure meaningful results.
+Otherwise, significant overfitting on biases in the data is likely.
+This basic quality cut $QualCut(D)$ consists of the following restrictions:
+
+\begin{itemize}
+    \item The run was not collected during engineering or testing ($\texttt{run\_type}\neq99$)
+    \item Recording was triggered by the camera ($\texttt{trigger\_main}=0$)
+    \item Acoustic Parameter is not erroneously large and negative ($log_{10}(\texttt{acoustic\_bubnum})>-100$)
+    \item Recorded more than 25 seconds after reaching target pressure ($\texttt{te}>25$)
+    \item The bubble position $\chi$ was successfully calculated ($[\chi_{X}, \chi_{Y}, \chi_{Z}]\neq[-100, -100, -100]$)
+\end{itemize}
+
+\subsubsection{Bubble Multiplicity Cut}
+
+PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles and WIMP candidates never scatter.
+Thus, no discriminator is required to handle these events, so they are removed.
+The bubble multiplicity cut $MultiCut(D)$ consists of the following restrictions:
+
+\begin{itemize}
+    \item Either 0 or 1 bubbles are detected based on images from the camera ($\texttt{nbub}<2$)
+    \item The number of bubbles approximated using the pressure transducer is close to 1 ($0.7<\texttt{dytranCZ}<1.3$)
+\end{itemize}
+
+\subsubsection{Wall Cut}
+
+Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel.
+It can be desirable for a discriminator to handle these events correctly; however, Acoustic Parameter does not, and neither does the neural network used in the previous PICO-60 paper.
+Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.
+
+The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the pressure transducer), and the acoustic cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
+
+\textit{Fiducial Cut}
+
+The fiducial cut $FidWallCut(D)$ defines an spatial area along the walls of the vessel within which no events are accepted.
+It makes use of the bubble position $\chi$, the distance from the bubble to the center of the vessel $R$, and the distance from the bubble to the nearest wall, which is defined as $min(\texttt{Dwall}, \texttt{Dwall\_horiz})$.
+It is defined as follows:
+
+\begin{itemize}
+    \item $Z \leq 523$
+    \item Any of the following 4 restrictions are true:
+    \begin{itemize}
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (0 < \chi_{Z} \leq 400)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (\chi_{Z} \leq 0) \land (R \leq 100)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (\chi_{Z} \leq 0) \land (R > 100)$
+        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (\chi_{Z} > 400)$
+    \end{itemize}
+\end{itemize}
+
+\textit{Pressure Cut}
+
+The pressure cut $PresWallCut(D)$ restricts the pressure detected by the pressure transducer, without any position corrections, to be close to 1 ($0.7<\texttt{dytranC}<1.3$).
+In combination with the acoustic cut $AcWallCut(D)$, this acts as a backup to the fiducial cut.
+
+\textit{Acoustic Cut}
+
+The acoustic cut $AcWallCut(D)$ is defined using the banded Fourier transform $\beta_{8}$ as follows:
+
+$45 < mean(\beta_{8}[(0, 2), (0, 1)]) < 300$
+
+This takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{wall_event_positions}
+    \caption{\label{} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
+\end{figure}
+
+\section{Machine Learning Techniques}
 
@@
 \end{enumerate}
-
-\subsection{Existing Work}
-
-During PICO-60 analysis, a function known as Acoustic Parameter was derived from simulations.
-It discriminates between alpha particles and nuclear recoils based on audio data (specifically the banded Fourier transform $\beta_{8}$, which is detailed in section \textbf{2.1.2}).
-
-WIMP candidates in the cross-section and mass ranges that the PICO-60 detector is sensitive to are predicted to produce nuclear recoils with the same properties as those created by neutrons.
-The key difference is that neutrons frequently (but not always) scatter and produce multiple bubbles, where the extremely small predicted cross sections of WIMP candidates mean that they should almost invariably create just one bubble.
-This means single-bubble neutron events can be used to optimize a discriminator to detect WIMP events.
-
-There was no explicit alpha particle calibration data collected as part of the PICO-60 experiment.
-However, the previous analysis revealed that there were no WIMP candidates detected.
-Also, based on the known ratio of multiple- to single-bubble neutron events, it is unlikely that there is more than 1 single-bubble neutron event throughout the background runs.
-This means they can be used as a high-purity alpha calibration set.
-
-While these two known event sets were only used for minor calibrations of Acoustic Parameter, there is no reason they cannot be used to optimize a machine learning discriminator from scratch.
-
-\section{Data Formats}
-
-Discriminators can use a variety of different data formats to make predictions.
-These are collected from the two essential sensors present in the PICO-60 apparatus: piezoelectric microphones (piezos) and cameras.
-
-\subsection{Piezo-Derived}
-
-\subsubsection{Raw Waveform}
-
-The lowest-level data derived from the piezos is the raw waveform $\omega$.
-$\omega$ consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
-They are each represented as a 16-bit integer.
-Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because before this there is background noise, and afterwards there is a clipped signal produced by hydraulics repressurizing the vessel.
-(These sections contain no information and can be undesirably fit on by neural networks.)
-
-\subsubsection{Fourier Transform}
-
-The raw waveform $\omega$ is converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input.
-Since DFTs produce sequences of complex numbers, the magnitude of each element is computed.
-The full-resolution Fourier transform $\beta _{50,000}$ consists of a sequence of complex numbers, half the length of the raw waveform $\omega$.
-
-Beyond this, arbitrary-resolution banded Fourier transforms $\beta _{N<50,001}$ can be computed by integrating the resonant energy over all frequencies within each of a set of bands.
-They can be thought of as a method of signal downscaling, compressing the signal into a smaller number of data points.
-Particular band resolutions used include $\beta _{8}$ (which is the input to Acoustic Parameter and initial neural network experiments), $\beta _{5}$, $\beta _{10}$, $\beta _{20}$, and $\beta _{40}$.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{audio}
-    \caption{\label{} An example of the cropped audio waveform $\omega$ and full resolution Fourier transform $\beta_{50,001}$.}
-\end{figure}
-
-\subsection{Camera-Derived}
-
-\subsubsection{Image Window Sequence}
-
-For each event that is captured, the 4 cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected.
-These raw images contain a large amount of extraneous information; they encompass the entire vessel.
-To reduce the input information, the image window sequence $\iota$ includes 50\texttimes50 cropped windows around the position of the bubble.
-The 71 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
-$\iota$ includes 5 frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.7\textwidth]{image_grid}
-    \caption{\label{} An example of the image sequence $\iota$.}
-\end{figure}
-
-\subsubsection{3D Position}
-
-The 3-dimensional position $\chi$ of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera.
-This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
-
-\section{Data Cuts}
-
-Techniques in this study are trained and validated on a number of different data sets.
-The selection of these sets focused on the fundamental trade-off between quality and quantity of data; by setting a higher standard for the validity of training data, one has less data to train on.
-
-The initial data set $D$ to which these cuts are applied consists of all events recorded during PICO-60 run 2.
-
-\subsection{Basic Quality Cut}
-
-A certain number of cuts are necessarily applied to all data to ensure meaningful results.
-Otherwise, significant overfitting on biases in the data is likely.
-This basic quality cut $QualCut(D)$ consists of the following restrictions:
-
-\begin{itemize}
-    \item The run was not collected during engineering or testing ($\texttt{run\_type}\neq99$)
-    \item Recording was triggered by the camera ($\texttt{trigger\_main}=0$)
-    \item Acoustic Parameter is not erroneously large and negative ($log_{10}(\texttt{acoustic\_bubnum})>-100$)
-    \item Recorded more than 25 seconds after reaching target pressure ($\texttt{te}>25$)
-    \item The bubble position $\chi$ was successfully calculated ($[\chi_{X}, \chi_{Y}, \chi_{Z}]\neq[-100, -100, -100]$)
-\end{itemize}
-
-\subsection{Bubble Multiplicity Cut}
-
-PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles and WIMP candidates never scatter.
-Thus, no discriminator is required to handle these events, so they are removed.
-The bubble multiplicity cut $MultiCut(D)$ consists of the following restrictions:
-
-\begin{itemize}
-    \item Either 0 or 1 bubbles are detected based on images from the camera ($\texttt{nbub}<2$)
-    \item The number of bubbles approximated using the pressure transducer is close to 1 ($0.7<\texttt{dytranCZ}<1.3$)
-\end{itemize}
-
-\subsection{Wall Cut}
-
-Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel.
-It can be desirable for a discriminator to handle these events correctly; however, Acoustic Parameter does not, and neither does the neural network used in the previous PICO-60 paper.
-Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.
-
-The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the pressure transducer), and the acoustic cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
-
-\subsubsection{Fiducial Cut}
-
-The fiducial cut $FidWallCut(D)$ defines an spatial area along the walls of the vessel within which no events are accepted.
-It makes use of the bubble position $\chi$, the distance from the bubble to the center of the vessel $R$, and the distance from the bubble to the nearest wall, which is defined as $min(\texttt{Dwall}, \texttt{Dwall\_horiz})$.
-It is defined as follows:
-
-\begin{itemize}
-    \item $Z \leq 523$
-    \item Any of the following 4 restrictions are true:
-    \begin{itemize}
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (0 < \chi_{Z} \leq 400)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (\chi_{Z} \leq 0) \land (R \leq 100)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (\chi_{Z} \leq 0) \land (R > 100)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (\chi_{Z} > 400)$
-    \end{itemize}
-\end{itemize}
-
-\subsubsection{Pressure Cut}
-
-The pressure cut $PresWallCut(D)$ restricts the pressure detected by the pressure transducer, without any position corrections, to be close to 1 ($0.7<\texttt{dytranC}<1.3$).
-In combination with the acoustic cut $AcWallCut(D)$, this acts as a backup to the fiducial cut.
-
-\subsubsection{Acoustic Cut}
-
-The acoustic cut $AcWallCut(D)$ is defined using the banded Fourier transform $\beta_{8}$ as follows:
-
-$45 < mean(\beta_{8}[(0, 2), (0, 1)]) < 300$
-
-This takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{wall_event_positions}
-    \caption{\label{} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
-\end{figure}
 

### 2018-08-06 documentation/paper.tex
Clarified ML in intro

@@
 
-As part of the PICO-60 experiment, some testing was conducted to determine whether machine learning could be a viable technique to consider for a discrimination function.
+As part of the PICO-60 experiment, some testing was conducted to determine whether machine learning (in the form of a multi-layer perceptron) could be a viable technique to consider for a discrimination function.
 The results, while not definitive, indicated that the general technique warranted further investigation.

### 2018-08-06 documentation/paper.tex
Added background on machine learning

@@
 In addition, a set of data cuts must be applied to remove certain classes of events.
-These conversion, preprocessing, and data cutting steps were re-applied during this study, and are described in section \ref{data_formats}.
+These conversion, preprocessing, and data cutting steps were reapplied during this study, and are described in section \ref{data_formats}.
 
@@
 \section{Machine Learning Techniques}
+
+\subsection{Background}
+
+While machine learning has existed for many years, recent developments in artificial neural networks have opened up a wide variety of applications and fields in which it can now be used.
+
+One of the key benefits of machine learning systems is that they have the capacity to approximate arbitrary functions (such as Acoustic Parameter) without human intervention.
+This greatly alleviates the need for human programmers to define and optimize the exact algorithm used.
+
+This characteristic has already shown wide-ranging implications in many fields.
+For one, it has the potential to drastically increase the speed of iteration for the people working on a project, since retraining a machine learning system when a hardware or software variable changes is much faster than calibrating a human-designed model.
 

### 2018-08-06 documentation/paper.tex
Clarified PosCor and multiple resolutions

@@
 
-For Acoustic Parameter to be sufficiently accurate, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$), and preprocessed with the position correction function $PosCor(\chi)$.
-In addition, a set of data cuts must be applied to remove certain classes of events.
+Before Acoustic Parameter can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$).
+It must also be preprocessed with the position correction function $PosCor(\beta_{8}, \chi)$, which adjusts the overall volume of an audio recording according to its position in the vessel.
+This is done because the resonance characteristics differ significantly depending on the position of an event.
+
+Finally, a set of data cuts must be applied to remove certain classes of events which are not accurately classified by Acoustic Parameter. These classes include wall event and multiple-bubble events.
+
 These conversion, preprocessing, and data cutting steps were reapplied during this study, and are described in section \ref{data_formats}.
@@
 
-The raw waveform $\omega$ is converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input.
-Since DFTs produce sequences of complex numbers, the magnitude of each element is computed.
+The raw waveform $\omega$ was converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input.
+Since DFTs produce sequences of complex numbers, the magnitude of each element was computed.
 The full-resolution Fourier transform $\beta _{50,000}$ consists of a sequence of complex numbers, half the length of the raw waveform $\omega$.
 
-Beyond this, arbitrary-resolution banded Fourier transforms $\beta _{N<50,001}$ can be computed by integrating the resonant energy over all frequencies within each of a set of bands.
-They can be thought of as a method of signal downscaling, compressing the signal into a smaller number of data points.
-Particular band resolutions used include $\beta _{8}$ (which is the input to Acoustic Parameter and initial neural network experiments), $\beta _{5}$, $\beta _{10}$, $\beta _{20}$, and $\beta _{40}$.
+Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to Acoustic Parameter and the original neural analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands.
+This can be thought of as a method of signal downscaling, compressing the signal into a smaller number of data points.
 
@@
 
-Two different configurations were tested: one that operates strictly on $\omega$ with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
-The second configuration was introduced because the position corrections applied to $\beta_{8}$ in previous PICO-60 research were applied only to broad frequency bands and are thus not practical to apply to a full resolution waveform.
+Two different configurations were tested:
+\begin{enumerate}
+    \item A configuration that operates strictly on $\omega$ with no position corrections of any kind (furthermore $DeepConv(\omega)$).
+    \item A configuration with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
+\end{enumerate}
 
-During early tests with variations on $DeepConv(\omega)$, it was revealed to have an incredible capability to separate run types based on unwanted biases.
+During early tests with variations on $DeepConv(\omega)$, it was revealed to have an incredible capacity to separate run types based on unwanted biases.
 It was able to separate the AmBe calibration run from combined background runs with 96\% accuracy, and to separate all calibration runs from background runs with 91\% accuracy.
-However, during these tests, the white noise at the beginning and hydraulic recompression sounds at the end were not cropped off, and the network was shown to be overfitting on this information.
+This was determined to be caused by the network detecting and fitting on biases in the white noise at the beginning and hydraulic recompression sounds at the end of the raw waveform.
+This extraneous information was subsequently cropped out.
 
 Once the audio was cropped properly, $DeepConv(\omega)$ achieved a maximum of 77\% accuracy on validation data.
-Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ managed 85\%.
+Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ achieved 85\%.
 The maximum obtained on a grid search with position correction was 91\%.
@@
 
-A multi-layer perceptron (a neural network composed of dense layers) was applied to analysis of audio data in the frequency domain.
-Preprocessing by applying a Fourier transform makes it an easier task for the neural network to accomplish in some ways.
+Preprocessing the audio by applying a Fourier transform may be advantageous.
 If the frequency distribution and overall volume are indeed the most important factors for discrimination, the network can gather them straight from the Fourier transform $\beta_{N}$ rather than having to analyze $\omega$ to extract this information.
+
+In addition to the Fourier transform used in the original PICO-60 analysis, higher resolutions were computed by similar means.
+These included $\beta _{5}$, $\beta _{10}$, $\beta _{20}$, and $\beta _{40}$.
+Also, the full-resolution $\beta _{50,001}$ was input into a neural network directly, without any downscaling.
 

### 2018-08-06 visualization/grid_search_visualize_means.py
Created script to graph mean accuracy by configuration

@@ -1,0 +1,30 @@
+#!/usr/bin/env python3
+"""Visualize the mean accuracy values over the runs of a grid search"""
+# Created by Brendon Matusch, August 2018
+
+import os
+import sys
+
+import matplotlib.pyplot as plt
+
+from utilities.verify_arguments import verify_arguments
+
+# A path to an AP similarity log file is expected
+verify_arguments('path to AP similarity log file')
+# Load all lines in the file
+with open(os.path.expanduser(sys.argv[1])) as file:
+    lines = file.readlines()
+# Take only the lines containing the mean disagreement statistics, and extract the numeric values
+mean_disagreements = [float(line.split()[1]) for line in lines if 'Mean' in line]
+# Calculate the corresponding accuracy values, by dividing by the total validation examples and subtracting from 1
+accuracy_values = [1 - (disagreements / 128) for disagreements in mean_disagreements]
+
+# Set the size of the resulting graph (it should be standard across all such graphs)
+plt.figure(figsize=(8, 6))
+# Plot the mean accuracy values by the run number in a bar graph
+plt.bar(range(len(accuracy_values)), accuracy_values)
+# Label the X and Y axes
+plt.xlabel('Grid Search Configuration')
+plt.ylabel('Mean Accuracy in Acoustic Parameter Replication')
+# Display the graph on screen
+plt.show()

### 2018-08-06 statistics/grid_search_optimal_hyperparameters_statistics.py
Created script to find optimal hyperparameters

@@ -1,0 +1,37 @@
+#!/usr/bin/env python3
+"""A script for calculating the optimal value of each hyperparameter in a grid search"""
+# Created by Brendon Matusch, August 2018
+
+import os
+import sys
+
+import numpy as np
+
+from utilities.verify_arguments import verify_arguments
+
+# A path to an AP similarity log file is expected
+verify_arguments('path to AP similarity log file')
+# Load all lines in the file
+with open(os.path.expanduser(sys.argv[1])) as file:
+    lines = file.readlines()
+# Take only the lines containing the mean disagreement statistics, and strip whitespace
+mean_lines = [line.strip() for line in lines if 'Mean' in line]
+# Extract the numeric disagreement values
+mean_disagreements = [float(line.split()[1]) for line in mean_lines]
+# Iterate over each of the hyperparameters that are used
+for hyperparameter in ['dropout', 'l2_lambda', 'initial_examples', 'initial_threshold', 'threshold_multiplier']:
+    # Create a dictionary to add lists of disagreement values according to hyperparameter values
+    disagreement_by_hyperparameter_value = {}
+    # Iterate over each of the mean lines with corresponding disagreement values
+    for line, disagreements in zip(mean_lines, mean_disagreements):
+        # Split the line by the name of the hyperparameter, and then get the hyperparameter value (as a string) up to the next underscore separator
+        hyperparameter_value = line.split(hyperparameter)[1].split('_')[0]
+        # If this value is not already in the dictionary, add a corresponding empty list
+        if hyperparameter_value not in disagreement_by_hyperparameter_value:
+            disagreement_by_hyperparameter_value[hyperparameter_value] = []
+        # Add the corresponding disagreements to the list
+        disagreement_by_hyperparameter_value[hyperparameter_value].append(disagreements)
+    # Calculate the overall mean number of disagreements for each hyperparameter value, and get the corresponding values themselves
+    overall_means, corresponding_values = zip(*((np.mean(disagreement_by_hyperparameter_value[hyper]), hyper) for hyper in disagreement_by_hyperparameter_value))
+    # Find the index of the best number of disagreements, and print out the corresponding optimal hyperparameter value
+    print(f'Optimal {hyperparameter}: {corresponding_values[np.argmin(overall_means)]}')

### 2018-08-06 documentation/paper.tex
Added graph for ICN accuracy by run

@@
 The mean of all epochs within that training run was 98\%.
-Further demonstrating this is a powerful algorithm and that high performance was not a one-off, 35\% of the training runs replicated Acoustic Parameter with a mean accuracy greater than 92\%.
+
+Further demonstrating this is a powerful algorithm and that high performance was not a one-off, 35\% of the training runs replicated Acoustic Parameter with a mean accuracy greater than 92\% (shown in Figure \ref{icn_accuracy_by_run}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{icn_accuracy_by_run}
+    \caption{\label{icn_accuracy_by_run} Mean accuracy of iterative cluster nucleation configurations during grid search.}
+\end{figure}
+
+As seen in Figure \ref{icn_grid_search}, the network learned to replicate Acoustic Parameter very accurately.
+Interestingly, the examples with Acoustic Parameter values between the neutron and alpha particle ranges were also classified as such by the network.
 
@@
 \end{figure}
-
-As seen in Figure \ref{icn_grid_search}, the network learned to replicate Acoustic Parameter very accurately.
-Interestingly, the examples with Acoustic Parameter values between the neutron and alpha particle ranges were also classified as such by the network.
 

### 2018-08-06 visualization/grid_search_visualize_optimal_hyperparameters.py
Bound graph between 75% and 100%

@@ -5,6 +5,7 @@
 import os
 import sys
 
+import matplotlib.pyplot as plt
 import numpy as np
 
 from utilities.verify_arguments import verify_arguments
@@ -18,20 +19,50 @@ with open(os.path.expanduser(sys.argv[1])) as file:
 mean_lines = [line.strip() for line in lines if 'Mean' in line]
 # Extract the numeric disagreement values
 mean_disagreements = [float(line.split()[1]) for line in mean_lines]
-# Iterate over each of the hyperparameters that are used
-for hyperparameter in ['dropout', 'l2_lambda', 'initial_examples', 'initial_threshold', 'threshold_multiplier']:
+# Create a dictionary to add disagreement dictionaries to, based on human-readable names
+disagreement_dictionaries = {}
+# Iterate over each of the hyperparameters that are used, alongside human-readable names
+for hyperparameter, human_readable in [('dropout', 'Dropout'), ('l2_lambda', 'L2 Regularization Lambda'), ('initial_threshold', 'Initial Training Threshold'), ('threshold_multiplier', 'Training Threshold Multiplier')]:
     # Create a dictionary to add lists of disagreement values according to hyperparameter values
     disagreement_by_hyperparameter_value = {}
     # Iterate over each of the mean lines with corresponding disagreement values
     for line, disagreements in zip(mean_lines, mean_disagreements):
-        # Split the line by the name of the hyperparameter, and then get the hyperparameter value (as a string) up to the next underscore separator
-        hyperparameter_value = line.split(hyperparameter)[1].split('_')[0]
+        # Split the line by the name of the hyperparameter, and then get the hyperparameter value (as a floating-point number) up to the next underscore separator
+        hyperparameter_value = float(line.split(hyperparameter)[1].split('_')[0])
         # If this value is not already in the dictionary, add a corresponding empty list
         if hyperparameter_value not in disagreement_by_hyperparameter_value:
             disagreement_by_hyperparameter_value[hyperparameter_value] = []
         # Add the corresponding disagreements to the list
         disagreement_by_hyperparameter_value[hyperparameter_value].append(disagreements)
-    # Calculate the overall mean number of disagreements for each hyperparameter value, and get the corresponding values themselves
-    overall_means, corresponding_values = zip(*((np.mean(disagreement_by_hyperparameter_value[hyper]), hyper) for hyper in disagreement_by_hyperparameter_value))
[…]

### 2018-08-06 documentation/paper.tex
Added data for ICN accuracy as a function of hyperparameters

@@
 
+As seen in Figure \ref{icn_grid_search}, the network learned to replicate Acoustic Parameter very accurately.
+Interestingly, the examples with Acoustic Parameter values between the neutron and alpha particle ranges were also classified as such by the network.
+This may imply the indecisiveness on these examples is fundamental to the data rather than an implementation problem with Acoustic Parameter.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{icn_grid_search}
+    \caption{\label{icn_grid_search} Accuracy of best iterative cluster nucleation training run compared to Acoustic Parameter and original ground truths (100\% accuracy).}
+\end{figure}
+
 Further demonstrating this is a powerful algorithm and that high performance was not a one-off, 35\% of the training runs replicated Acoustic Parameter with a mean accuracy greater than 92\% (shown in Figure \ref{icn_accuracy_by_run}.
@@
 
-As seen in Figure \ref{icn_grid_search}, the network learned to replicate Acoustic Parameter very accurately.
-Interestingly, the examples with Acoustic Parameter values between the neutron and alpha particle ranges were also classified as such by the network.
+Digging deeper into the performance throughout the runs of the grid search, Figure \ref{icn_acc_by_hyper} shows the mean accuracy as a function of each of the hyperparameters that were optimized.
+Dropout has a clear effect, with 25\% shown to be the optimal number of neurons to drop.
+L2 regularization is also significant; all $\lambda$ values greater than 0 make performance worse.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{icn_grid_search}
-    \caption{\label{icn_grid_search} Accuracy of best iterative cluster nucleation training run compared to Acoustic Parameter and original ground truths (100\% accuracy).}
+    \includegraphics[width=\textwidth]{icn_acc_by_hyper}
+    \caption{\label{icn_acc_by_hyper} Mean accuracy of iterative cluster nucleation as a function of hyperparameter values.}
 \end{figure}

### 2018-08-07 documentation/paper.tex
Added gravitational differentiation validation data graph

@@
 
+During a grid search of parameters of the gravitational differentiation algorithm (in addition to the stochastic gradient descent learning rate), the algorithm achieved a maximum of 98\% validation accuracy.
+The mean over that run (also the highest overall) was 96\%.
+
+Figure \ref{grav_opt_validation} shows the highly accurate replication of Acoustic Parameter.
+However, the examples that are classified incorrectly seem to have no correlation with Acoustic Parameter, suggesting that the learned function cues on different properties of the audio.
+While the discriminator may still be accurate, this property implies it may be less predictable when trained on new data sets.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{grav_opt_validation}
+    \caption{\label{grav_opt_validation} Accuracy of optimal gravitational differentiation training run compared to Acoustic Parameter (98\% accuracy).}
+\end{figure}
+
 \subsection{Iterative Cluster Nucleation}
@@
     \includegraphics[width=\textwidth]{icn_grid_search}
-    \caption{\label{icn_grid_search} Accuracy of best iterative cluster nucleation training run compared to Acoustic Parameter and original ground truths (100\% accuracy).}
+    \caption{\label{icn_grid_search} Accuracy of optimal iterative cluster nucleation training run compared to Acoustic Parameter (100\% accuracy).}
 \end{figure}

### 2018-08-07 visualization/grid_search_visualize_means.py
Skip ground truth sets from gravitational differentiation

@@ -14,6 +14,8 @@ verify_arguments('path to AP similarity log file')
 # Load all lines in the file
 with open(os.path.expanduser(sys.argv[1])) as file:
     lines = file.readlines()
+# Remove lines that contain ground truth data from the gravitational differentiation grid search
+lines = [line for line in lines if '_ground_truths' not in line]
 # Take only the lines containing the mean disagreement statistics, and extract the numeric values
 mean_disagreements = [float(line.split()[1]) for line in lines if 'Mean' in line]
 # Calculate the corresponding accuracy values, by dividing by the total validation examples and subtracting from 1

### 2018-08-07 visualization/grid_search_visualize_optimal_hyperparameters.py
Fixed optimal hyperparameter visualization script for gravitational differentiation grid search

@@ -15,6 +15,8 @@ verify_arguments('path to AP similarity log file')
 # Load all lines in the file
 with open(os.path.expanduser(sys.argv[1])) as file:
     lines = file.readlines()
+# Remove lines that contain ground truth data from the gravitational differentiation grid search
+lines = [line for line in lines if '_ground_truths' not in line]
 # Take only the lines containing the mean disagreement statistics, and strip whitespace
 mean_lines = [line.strip() for line in lines if 'Mean' in line]
 # Extract the numeric disagreement values
@@ -22,7 +24,8 @@ mean_disagreements = [float(line.split()[1]) for line in mean_lines]
 # Create a dictionary to add disagreement dictionaries to, based on human-readable names
 disagreement_dictionaries = {}
 # Iterate over each of the hyperparameters that are used, alongside human-readable names
-for hyperparameter, human_readable in [('dropout', 'Dropout'), ('l2_lambda', 'L2 Regularization Lambda'), ('initial_threshold', 'Initial Training Threshold'), ('threshold_multiplier', 'Training Threshold Multiplier')]:
+# `distortion_root` is the old name for distortion power
+for hyperparameter, human_readable in [('gravity_multiplier_increment', 'Gravity Multiplier Increment'), ('learning_rate', 'Learning Rate'), ('distortion_root', 'Distortion Power')]:
     # Create a dictionary to add lists of disagreement values according to hyperparameter values
     disagreement_by_hyperparameter_value = {}
     # Iterate over each of the mean lines with corresponding disagreement values
@@ -53,8 +56,8 @@ for plot_index, human_readable in enumerate(disagreement_dictionaries):
     hyperparameter_values = list(disagreement_dictionaries[human_readable].keys())
     # Calculate the corresponding accuracy values, by dividing by the total validation examples and subtracting from 1
     accuracy_values = [1 - (disagreements / 128) for disagreements in list(disagreement_dictionaries[human_readable].values())]
-    # Calculate the bar width, which should be half of the mean difference between a pair of examples
-    bar_width = (max(hyperparameter_values) - min(hyperparameter_values)) / (2 * (len(hyperparameter_values) - 1))
+    # Calculate the bar width, which should be 1/6 of the mean difference between a pair of examples
+    bar_width = (max(hyperparameter_values) - min(hyperparameter_values)) / (6 * (len(hyperparameter_values) - 1))
[…]

### 2018-08-07 documentation/paper.tex
Added graphs for gravitational accuracy by run and hyperparameters

@@
 \usepackage{wrapfig}
+\usepackage{svg}
 
@@
 
+This was not the only configuration that performed well; as seen in Figure \ref{grav_acc_by_run}, 10\% of the runs during the grid search produced accuracy greater than 93\%.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{grav_acc_by_run}
+    \caption{\label{grav_acc_by_run} Accuracy of gravitational differentiation arranged by training run during the grid search.}
+\end{figure}
+
+As seen in Figure \ref{grav_acc_by_hyper}, all 3 hyperparameters that were optimized during the grid search had a significant effect on accuracy:
+\begin{itemize}
+    \item Smaller gravitational multipliers improved performance; the best value tested was 0.0005.
+This implies that the gravity effect on each unlabeled event should be weighted lightly compared to the labeled examples.
+    \item Higher learning rates improve performance, with the best value tested at a quite high 0.03.
+    \item Larger distortion powers (which raise the confidence threshold for a significant gravitational effect) improved performance, with the best tested at 11.
+\end{itemize}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{grav_acc_by_hyper}
+    \caption{\label{grav_acc_by_hyper} Accuracy of gravitational differentiation as a function of hyperparameter values.}
+\end{figure}
+
 \subsection{Iterative Cluster Nucleation}

### 2018-08-07 documentation/paper.tex
Removed unnecessary SVG package
Completed section on machine learning techniques

@@
 \usepackage{wrapfig}
-\usepackage{svg}
 
@@
     Furthermore, what little data is available to optimize a discriminator very often contains impurities, because whatever background radiation is present during WIMP detection runs is also present during calibration runs.
-These are difficult to separate without already having access to a functional discriminator, creating a ``chicken or the egg'' problem (from which one of the most common ways to escape is to create a time- and labor-intensive simulation).
+These are difficult to separate without already having access to a functional discriminator, creating a ``chicken or the egg'' problem (from which a common way to escape is to create a time- and labor-intensive simulation).
 \end{enumerate}
@@
 
+\subsection{Techniques Considered}
+
+Machine learning techniques are usually divided into two main categories:
+
+\begin{itemize}
+    \item Supervised learning, in which a system is trained on a set of labeled data to classify unseen examples according to the patterns it observes in the training set.
+The training set is very important; biases and errors often slow training and produce worse results.
+    \item Unsupervised learning, in which a system finds patterns or clusters in a set of unlabeled data.
+This can find order in almost any training set, but it is unpredictable and may not find the particular patterns it is expected to find.
+\end{itemize}
+
+A significant challenge in this application is that neither supervised nor unsupervised learning is ideal.
+The calibration sets available for training are not pure, so supervised learning is likely to overfit on biases and produce undesirable results.
+Conversely, unsupervised learning may be able to find clusters in the data, but it is impossible to guarantee it will distinguish between nuclear recoils and alpha particles as opposed to some other binary separation.
+
+For these reasons, a technique is needed that is not as sensitive to problematic training data as supervised learning, but more predictable and controllable than unsupervised learning.
+Semi-supervised learning is a middle ground in this regard.
+It makes use of a labeled set as well as an unlabeled set.
+It uses the labeled set for training, and uses the unlabeled set to further structure the patterns it finds in the training set (which may not be sufficiently large to apply to supervised learning).
+
+Two semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
+Both are discussed in detail in section \ref{semi_supervised}.
+
 \subsection{Hyperparameter Optimization}
@@
 
-Supervised learning was the first step in the optimization procedure.
-It makes use of fully labeled calibration sets, which are used for training.
-Performance of these supervised neural networks is judged based on its accuracy at replication of the provided labels on a validation set (which is set aside and not trained on).
+Supervised learning was the first general technique that was applied and optimized.
+It makes use of fully (but imperfectly) labeled calibration data sets, which are used for training.
+Performance of these supervised neural networks is evaluated based on its accuracy at replication of the provided labels on a validation set (which is set aside and not trained on).
 
@@
 This avoids any destruction of information, and should be possible, in theory, for a neural network to handle.
-For this task, a very deep 1-dimensional fully convolutional neural network was applied.
+For this task, a very deep (20 layers in its shallowest configuration) 1-dimensional fully convolutional neural network was applied.
 The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.
@@
 
-\section{Semi-Supervised Learning}
+\section{Semi-Supervised Learning} \label{semi_supervised}
 

### 2018-08-07 documentation/paper.tex
Clarified wall cuts a bit

@@
 
-A convolutional neural network was used for analyzing the raw waveform $\omega$ directly, without any preprocessing whatsoever, and without application of $WallCut(D)$.
+A convolutional neural network was used for analyzing the raw waveform $\omega$ directly, without any preprocessing whatsoever.
 This avoids any destruction of information, and should be possible, in theory, for a neural network to handle.
@@
 No position input is used (since there are no microphones that would require position correction), so the basic configuration is $ImageConv(\iota)$.
+
+$WallCut(D)$ was not applied to the training set, because while the resonance in the vessel varies greatly depending on the position of the bubble, any visual differences in the bubble should not.
+In addition, applying it reduces the size of the training set considerably.
 

### 2018-08-07 documentation/paper.tex
Reorganization and clarification of purpose of supervised learning

@@
 
-In this study, newly developed methodologies were compared to the following discrimination techniques developed previously for the PICO-60 experiment.
+\subsubsection{Overview}
+
+In this study, newly developed methodologies were compared to the 2 discrimination techniques developed previously for the PICO-60 experiment: Acoustic Parameter and machine learning.
 
@@
 
-Two semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
+2 semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
 Both are discussed in detail in section \ref{semi_supervised}.
 
-\subsection{Hyperparameter Optimization}
+\subsection{Optimization Process}
 
-Learning algorithms in this study are optimized using grid searches, in which many different combinations of hyperparameters are tested.
-The computational work required for a grid search increases exponentially with the number of hyperparameters $n$ and the number of options per hyperparameter $m$; to be exact, its complexity is $O(m^{n})$.
-Thus, it is important to minimize the number of hyperparameters that are optimized at once.
-
-Since the semi-supervised learning algorithms used incorporate neural network models, optimization is a much simpler task if the network's hyperparameters are already optimized.
-Thus, the optimization task is split into two distinct phases:
+Learning algorithms in this study are optimized systematically, with a variety of neural network configurations and input formats.
+To obtain meaningful results, it is important to minimize the number of variables during each individual run.
+Thus, since the 2 semi-supervised algorithms incorporate neural network models, it is important that the input formats and other hyperparameters of these networks are optimized first to some degree.
+This means the task is split into 2 distinct phases:
 
 \begin{enumerate}
-    \item Optimize the hyperparameters of a neural network on validation accuracy, to accurately reproduce the imperfect ground truths it is trained on.
+    \item Develop and verify several different supervised learning systems.
+Their performance is evaluated based on their ability to replicate the ground truths they are trained on.
 While this does not produce an optimal discriminator, it does reveal architectures that are capable, in essence, of learning to accomplish the discrimination task.
-    \item Using neural network hyperparameters that were determined to be effective, optimize the hyperparameters of the semi-supervised learning algorithm to reproduce Acoustic Parameter as closely as possible.
+    \item Using general supervised learning systems and input formats that were determined to be effective, optimize the hyperparameters of the semi-supervised learning algorithm to reproduce Acoustic Parameter as closely as possible.
 Since Acoustic Parameter is not used to optimize the neural network directly at any point, models that excel at this task demonstrate an ability to learn the desired discriminator while ignoring the impurities in the training set.
@@
 
-Supervised learning was the first general technique that was applied and optimized.
-It makes use of fully (but imperfectly) labeled calibration data sets, which are used for training.
-Performance of these supervised neural networks is evaluated based on its accuracy at replication of the provided labels on a validation set (which is set aside and not trained on).
+\subsection{Overview}
+
+Important unknowns to be determined through experimentation with supervised learning included the most effective input format for the network, and the associated general network structure.
+Three major such configurations were applied:
+
+\begin{enumerate}
+    \item A convolutional neural network trained on the raw waveform $\omega$.
+    \item A dense neural network trained on banded Fourier transforms $\beta_{N}$.
+    \item A convolutional neural network trained on the image window data $\iota$.
+\end{enumerate}
+
+In addition, configurations with and without application of $WallCut(D)$ were tested.
+While $WallCut(D)$ was used throughout the original PICO-60 analysis because it is necessary for Acoustic Parameter, it was unknown whether or not it was required for neural networks.
 
 \subsection{Convolutional Neural Network for Raw Waveform Analysis}
+
+\subsubsection{Structure}
 
@@
 \end{enumerate}
+
+\subsubsection{Results}
 
@@
 
+\subsubsection{Structure}
+
 Preprocessing the audio by applying a Fourier transform may be advantageous.
@@
 While there were variety of network architectures tested, most had a small number of layers (on the order of 3), applied dropout and L2 regularization, and used batch normalization once again.
+
+\subsubsection{Results}
 
@@
 
+\subsubsection{Structure}
+
 It is an open question whether or not there is any information in the image data $\iota$ that could be used to distinguish between particle classes.
@@
 
+\subsubsection{Results}
+
 Throughout a variety of different network architectures, the best validation accuracy obtained was 85\%.
@@
 \section{Semi-Supervised Learning} \label{semi_supervised}
+
+\subsection{Overview}
 
@@
 
+\subsubsection{Concept}
+
 Gravitational differentiation is a novel technique for gradient computation in the final layer of a neural network.
 It is intended for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
+
+\subsubsection{Structure}
 
@@
 
+\subsubsection{Concept}
+
 Iterative cluster nucleation is a semi-supervised learning algorithm that takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
+
+\subsubsection{Structure}
+
 The algorithm is initialized as follows:

### 2018-08-07 documentation/paper.tex
Added information on minimization of initial set size

@@
 
-Performance of the supervised learning systems was quantified by how accurately they reproduced the labels provided to them (on a validation set).
-They were observed to learn to separate the calibration runs from background runs, including the percentage of alpha particles present in the calibration runs.
-One of the main intents of semi-supervised learning is to address the impure training data.
-Therefore, their performance is quantified by how well it is able to replicate Acoustic Parameter, rather than simply reproduce the labels provided.
+In addition to improving performance with a labeled set known to be adequate, a goal of this category of techniques is to minimize the amount of labeled data necessary.
+This means less time must be spent on collection of calibration data, which can be a very slow process.
+Regularization techniques, which permit effective training on smaller sets, are applied with this purpose.
+
+\subsection{Performance Analysis}
+
+One of the main purposes of semi-supervised learning is to handle the impure training data effectively, and learn to replicate the imperfectly represented underlying pattern (nuclear recoils versus alpha particles) rather than merely fitting to separate run types.
+
+Therefore, performance is quantified by how accurately a system is able to replicate Acoustic Parameter, rather than simply reproduce the labels provided.
+Accuracy of the neural network in the original PICO-60 paper is also provided as a baseline.
 

### 2018-08-07 visualization/grid_search_visualize_optimal_hyperparameters.py
Reordered code a bit

@@ -37,14 +37,15 @@ for hyperparameter, human_readable in [('gravity_multiplier_increment', 'Gravity
             disagreement_by_hyperparameter_value[hyperparameter_value] = []
         # Add the corresponding disagreements to the list
         disagreement_by_hyperparameter_value[hyperparameter_value].append(disagreements)
-    # Add the dictionary to the dictionary for all hyperparameters
-    disagreement_dictionaries[human_readable] = disagreement_by_hyperparameter_value
     # Replace the lists of means by run with the overall mean for that hyperparameter value
     for hyperparameter_value in disagreement_by_hyperparameter_value:
         disagreement_by_hyperparameter_value[hyperparameter_value] = np.mean(disagreement_by_hyperparameter_value[hyperparameter_value])
+    print(disagreement_by_hyperparameter_value)
     # Find the index of the best number of disagreements, and print out the corresponding optimal hyperparameter value
     # The key and value views must be converted to lists, otherwise argmin will return an incorrect index
     print(f'Optimal {hyperparameter}: {list(disagreement_by_hyperparameter_value.keys())[np.argmin(list(disagreement_by_hyperparameter_value.values()))]}')
+    # Add the dictionary to the dictionary for all hyperparameters
+    disagreement_dictionaries[human_readable] = disagreement_by_hyperparameter_value
 
 # Create a figure with a predefined size
 plt.figure(figsize=(10, 12))

### 2018-08-07 visualization/grid_search_visualize_optimal_hyperparameters.py
Include definitive training examples in analysis

@@ -25,7 +25,7 @@ mean_disagreements = [float(line.split()[1]) for line in mean_lines]
 disagreement_dictionaries = {}
 # Iterate over each of the hyperparameters that are used, alongside human-readable names
 # `distortion_root` is the old name for distortion power
-for hyperparameter, human_readable in [('gravity_multiplier_increment', 'Gravity Multiplier Increment'), ('learning_rate', 'Learning Rate'), ('distortion_root', 'Distortion Power')]:
+for hyperparameter, human_readable in [('gravity_multiplier_increment', 'Gravity Multiplier Increment'), ('learning_rate', 'Learning Rate'), ('distortion_root', 'Distortion Power'), ('definitive_training_examples', 'Definitive Training Examples')]:
     # Create a dictionary to add lists of disagreement values according to hyperparameter values
     disagreement_by_hyperparameter_value = {}
     # Iterate over each of the mean lines with corresponding disagreement values
@@ -40,7 +40,6 @@ for hyperparameter, human_readable in [('gravity_multiplier_increment', 'Gravity
     # Replace the lists of means by run with the overall mean for that hyperparameter value
     for hyperparameter_value in disagreement_by_hyperparameter_value:
         disagreement_by_hyperparameter_value[hyperparameter_value] = np.mean(disagreement_by_hyperparameter_value[hyperparameter_value])
-    print(disagreement_by_hyperparameter_value)
     # Find the index of the best number of disagreements, and print out the corresponding optimal hyperparameter value
     # The key and value views must be converted to lists, otherwise argmin will return an incorrect index
     print(f'Optimal {hyperparameter}: {list(disagreement_by_hyperparameter_value.keys())[np.argmin(list(disagreement_by_hyperparameter_value.values()))]}')

### 2018-08-07 documentation/paper.tex
Added statistic for number of examples on gravitational differentiation

@@
 As seen in Figure \ref{grav_acc_by_hyper}, all 3 hyperparameters that were optimized during the grid search had a significant effect on accuracy:
+
 \begin{itemize}
@@
 \end{figure}
+
+Two different options were tested with regard to size of the labeled data set $\varsigma$.
+Out of a set of 496 total examples available for training, either 128 (approximately $1/4$ of the total set) or 256 (approximately $1/2$) were used as labeled data.
+Over the entire grid search, runs with 128 examples in $\varsigma$ produced on average 64\% more errors on validation data, compared to those with 256 examples.
 

### 2018-08-07 visualization/grid_search_visualize_optimal_hyperparameters.py
Changed back to ICN, with initial examples this time

@@ -25,7 +25,7 @@ mean_disagreements = [float(line.split()[1]) for line in mean_lines]
 disagreement_dictionaries = {}
 # Iterate over each of the hyperparameters that are used, alongside human-readable names
 # `distortion_root` is the old name for distortion power
-for hyperparameter, human_readable in [('gravity_multiplier_increment', 'Gravity Multiplier Increment'), ('learning_rate', 'Learning Rate'), ('distortion_root', 'Distortion Power'), ('definitive_training_examples', 'Definitive Training Examples')]:
+for hyperparameter, human_readable in [('dropout', 'Dropout'), ('l2_lambda', 'L2 Regularization Lambda'), ('initial_threshold', 'Initial Training Threshold'), ('threshold_multiplier', 'Training Threshold Multiplier'), ('initial_examples', 'Initial Examples')]:
     # Create a dictionary to add lists of disagreement values according to hyperparameter values
     disagreement_by_hyperparameter_value = {}
     # Iterate over each of the mean lines with corresponding disagreement values
@@ -51,7 +51,7 @@ plt.figure(figsize=(10, 12))
 # Iterate over the human readable names of the 4 hyperparameters with a corresponding index for the plot
 for plot_index, human_readable in enumerate(disagreement_dictionaries):
     # Select a subplot with the current index (plus 1, since Matplotlib is 1-indexed)
-    plt.subplot(2, 2, plot_index + 1)
+    plt.subplot(2, 3, plot_index + 1)
     # Get the hyperparameter values from the dictionary
     hyperparameter_values = list(disagreement_dictionaries[human_readable].keys())
     # Calculate the corresponding accuracy values, by dividing by the total validation examples and subtracting from 1

### 2018-08-07 documentation/paper.tex
Added data for number of training examples (ICN)

@@
 
+Compared to gravitational differentiation, there is a larger delta between the number of errors on the validation set depending on the number of initial training examples used.
+Those with 128 examples in $\varsigma$ make 85\% more errors on average.
+However, iterative cluster nucleation with 128 examples still makes 19\% fewer errors than gravitational differentiation with 256 examples in $\varsigma$.
+
 \section{Software}

### 2018-08-07 documentation/paper.tex
Documented class-wise standard deviation

@@
 Since DFTs produce sequences of complex numbers, the magnitude of each element was computed.
-The full-resolution Fourier transform $\beta _{50,000}$ consists of a sequence of complex numbers, half the length of the raw waveform $\omega$.
+The full-resolution Fourier transform $\beta _{50,001}$ consists of a sequence of complex numbers, half the length of the raw waveform $\omega$.
 
@@
 While $WallCut(D)$ was used throughout the original PICO-60 analysis because it is necessary for Acoustic Parameter, it was unknown whether or not it was required for neural networks.
+
+\subsection{Performance Analysis}
+
+The performance of supervised learning neural networks was evaluated using 2 metrics.
+One is a simple classification accuracy value, which corresponds to the number of validation examples the network classifies incorrectly.
+This effectively summarizes how often the network is outright correct or incorrect, but fails to capture how decisive it is.
+
+A metric that captures this property is the class-wise standard deviation.
+This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
+It calculates the spread of the network's predictions for each ground truth class.
+This means that a decisive discriminator, which produces a wide separation between the two classes, is preferred over one that produces a nebulous cloud of outputs with a seemingly arbitrary decision boundary.
+
+$S=std(N \cup A)$
+
+$C=(std(N \div S) + std(A \div S)) \div 2$
+
+The first step is to calculate the standard deviation S of the union of the two sets.
+This gives an indication of the scale of the overall distribution.
+When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
+While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, Acoustic Parameter has a significantly wider range.
+Normalization of the union prevents this from creating a bias where Acoustic Parameter would produce a higher standard deviation with a similarly proportioned error.
+
+The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually.
+This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are.
+Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off of those specific values, will produce a low class-wise standard deviation.
 

### 2018-08-07 visualization/graph_binary_predictions.py
Print accuracy in graphing script

@@ -30,12 +30,18 @@ else:
     acoustic_parameters = random_values
     original_neural_network_scores = random_values
 
-    # Iterate over the three criteria standard deviations will be calculated for, and corresponding names
-for criterion_data, criterion_name in zip(
-    copy.deepcopy([network_outputs, acoustic_parameters,
-                   original_neural_network_scores]),
-    ['network outputs', 'acoustic parameters', 'original neural network scores']
+    # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
+for criterion_data, criterion_name, classification_threshold in zip(
+    copy.deepcopy([network_outputs, acoustic_parameters, original_neural_network_scores]),
+    ['network outputs', 'acoustic parameters', 'original neural network scores'],
+    [0.5, 0.25, 0.5]
 ):
+    # Convert the raw outputs to binary classifications, based on whether they are greater than or equal to the threshold
+    classifications = [data_point >= classification_threshold for data_point in criterion_data]
+    # Get the number of these classifications that agree with the ground truths
+    agreements = np.count_nonzero(np.array(classifications) == ground_truths)
+    # Print out the resulting accuracy statistic to the user
+    print(f'Accuracy of {agreements / 128} for {criterion_name}')
     # Divide all of these data points by their overall standard deviation, to normalize their range
     criterion_data /= np.std(criterion_data)
     # Iterate over both possible values of the ground truth, and corresponding names

### 2018-08-07 documentation/paper.tex
Added performance table

@@
 One of the main purposes of semi-supervised learning is to handle the impure training data effectively, and learn to replicate the imperfectly represented underlying pattern (nuclear recoils versus alpha particles) rather than merely fitting to separate run types.
-
 Therefore, performance is quantified by how accurately a system is able to replicate Acoustic Parameter, rather than simply reproduce the labels provided.
-Accuracy of the neural network in the original PICO-60 paper is also provided as a baseline.
 
@@
 
+\subsection{Performance Summary}
+
+The following table directly compares performance of the semi-supervised learning systems, and also includes the neural network used in the original PICO-60 paper as a baseline.
+
+\begin{tabular}{|l|l|l|l|}
+    Technique & Max Accuracy & Max Mean by Run & $dim(\varsigma)=128$ Errors versus 256 \\
+    Gravitational Differentiation & 98\% & 96\% & 64\% Increase \\
+    Iterative Cluster Nucleation & 100\% & 98\% & 85\% Increase \\
+    Original PICO-60 NN & X\% & N/A & N/A
+\end{tabular}
+
 \section{Software}

### 2018-08-07 statistics/disagreement_statistics.py
Compare to original neural network score

@@ -13,14 +13,14 @@ from utilities.verify_arguments import verify_arguments
 verify_arguments('JSON data file')
 # Load the data set from the file
 events, ground_truths, network_outputs = load_test(sys.argv[1])
-# Get the acoustic parameter data from the events
-acoustic_parameters = [event.logarithmic_acoustic_parameter for event in events]
+# Get the original neural network scores from the events
+original_neural_network_scores = [event.original_neural_network_score for event in events]
 # Convert the network's outputs to binary predictions (as a NumPy array)
 network_predictions = np.array([output >= 0.5 for output in network_outputs])
-# Repeat with acoustic parameters, dividing at a specific threshold
-acoustic_predictions = np.array([acoustic_parameter >= 0.22 for acoustic_parameter in acoustic_parameters])
+# Repeat with original neural network scores
+original_predictions = np.array([score >= 0.5 for score in original_neural_network_scores])
 # Get the indices of the bubbles on which the network and the acoustic parameter disagree
-disagreement_indices = np.argwhere(network_predictions != acoustic_predictions)
+disagreement_indices = np.argwhere(network_predictions != original_neural_network_scores)
 # Get the events on which the two techniques disagree (indices are wrapped in single-element arrays)
 disagreement_events = [events[index[0]] for index in disagreement_indices]
 # Iterate over the events with corresponding network predictions and ground truths, printing information
@@ -41,3 +41,5 @@ for event, network_prediction, ground_truth in zip(disagreement_events, network_
     print(correct, 'is correct')
     # Print a blank line for separation
     print()
+# Print the total number of events on which the techniques disagree
+print(f'{len(disagreement_events)} disagreements in total')

### 2018-08-07 documentation/paper.tex
Added horizontal lines to table

@@
 \begin{tabular}{|l|l|l|l|}
+    \hline
     Technique & Max Accuracy & Max Mean by Run & $dim(\varsigma)=128$ Errors versus 256 \\
+    \hline
     Gravitational Differentiation & 98\% & 96\% & 64\% Increase \\
+    \hline
     Iterative Cluster Nucleation & 100\% & 98\% & 85\% Increase \\
-    Original PICO-60 NN & X\% & N/A & N/A
+    \hline
+    Original PICO-60 NN & TODO\% & N/A & N/A \\
+    \hline
 \end{tabular}

### 2018-08-07 documentation/paper.tex
Added information with and without wall cuts

@@
 
-In addition, configurations with and without application of $WallCut(D)$ were tested.
-While $WallCut(D)$ was used throughout the original PICO-60 analysis because it is necessary for Acoustic Parameter, it was unknown whether or not it was required for neural networks.
+Except where specifically mentioned, $WallCut(D)$ was not applied.
+This was done to test the network's ability to handle complex acoustic characteristics including resonance.
+There is a relatively clear-cut discrimination boundary in PICO-60, but this may not be the case in other applications.
 
@@
 
-It is difficult to say with a high degree of confidence whether or not this is an effective discriminator, since Acoustic Parameter is not decisive on data without $WallCut(D)$.
 However, the lack of decisiveness observed even when accuracy was relatively high, as seen in Figure \ref{waveform_best_validation}, implies this is a suboptimal technique.
@@
 
+$WallCut(D)$ was applied to the low-resolution Fourier transforms during early experiments (since they were used for Acoustic Parameter) to minimize the number of variables between the neural network and a proven discriminator.
+
 Once again, configurations with and without position input were tested.
@@
 
-It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
+It became evident very quickly that, without $WallCut(D)$ a high resolution was not required to get a high accuracy relative to the ground truth data.
 $FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy by this metric.
-Increasing the number of bands seemed initially not to provide any improvement; the best accuracy reached with $\beta_{20}$ was 93\%, and with $\beta_{40}$ it was 94\%.
+Increasing the number of bands, even up to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 94\%.
 
@@
 
-In all previous experimentation, $WallCut(D)$ was applied to the training and validation sets.
-The higher resolution demonstrated significant power when $WallCut(D)$ was not applied, and only $QualCut(MultiCut(D))$ was used.
-$FourierMLPWithPos(\beta_{40}, \chi)$ still managed 92\% validation accuracy.
+The higher resolution demonstrated significant power when $WallCut(D)$ was not applied, and only \\ $QualCut(MultiCut(D))$ was used.
+$FourierMLPWithPos(\beta_{20}, \chi)$ achieved only 92\% accuracy (2 percentage points lower than the same configuration without $WallCut(D)$).
+$FourierMLPWithPos(\beta_{40}, \chi)$ also managed 92\% validation accuracy.
 However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%.

### 2018-08-07 documentation/paper.tex
Added class-wise standard deviation information for supervised learning

@@
 Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ achieved 85\%.
-The maximum obtained on a grid search with position correction was 91\%.
+The maximum obtained on a grid search with position input was 91\% (although with a high class-wise standard deviation of 0.61).
 While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
@@
 
-It became evident very quickly that, without $WallCut(D)$ a high resolution was not required to get a high accuracy relative to the ground truth data.
-$FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy by this metric.
-Increasing the number of bands, even up to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 94\%.
+It became evident very quickly that, without $WallCut(D)$, a high resolution was not required to get a high accuracy relative to the ground truth data.
+$FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
+Increasing the number of bands, even up to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.
 
@@
 However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%.
+However, that epoch had a relatively high standard deviation, at 0.47.
 
@@
 
-$WallCut(D)$ was not applied to the training set, because while the resonance in the vessel varies greatly depending on the position of the bubble, any visual differences in the bubble should not.
-In addition, applying it reduces the size of the training set considerably.
-
 \subsubsection{Results}
@@
 Throughout a variety of different network architectures, the best validation accuracy obtained was 85\%.
-(The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.) It is likely fitting on some form of noise within $\iota$, since the network's predictions seem almost random when represented graphically (as seen in Figure \ref{image_window}).
+The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.
+Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random (as is apparent in \ref{image_window}).
+It is likely fitting on some form of noise within $\iota$.
 The fact that throughout many trials (including a grid search), no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.

### 2018-08-07 documentation/paper.tex
Created data table for supervised learning

@@
 
+\subsection{Performance Summary}
+
+The following table compares the accuracy and mean class-wise standard deviation values of various supervised learning configurations:
+
+\begin{tabular}{|l|l|l|}
+    \hline
+    Configuration & Optimal Accuracy & Mean Class-Wise Standard Deviation \\
+    \hline
+    $DeepConv(\omega)$ & 77\% & X \\
+    \hline
+    $DeepConvWithPos(\omega, \chi)$ & 91\% & X \\
+    \hline
+\end{tabular}
+
 \section{Semi-Supervised Learning} \label{semi_supervised}
@@
 
-The following table directly compares performance of the semi-supervised learning systems, and also includes the neural network used in the original PICO-60 paper as a baseline.
+The following table directly compares performance of the semi-supervised learning systems, and also includes the neural network used in the original PICO-60 paper as a baseline:
 

### 2018-08-07 documentation/paper.tex
Lots of cleanup, added conclusion

--- 2018-08-07 documentation/paper.tex
@@
 
+\subsection{Performance Summary}
+
+The following table compares the accuracy and mean class-wise standard deviation values of various supervised learning configurations, with the original PICO-60 neural network provided as a baseline:
+
+\begin{tabular}{|l|l|l|}
+    \hline
+    Configuration & Max Accuracy & Mean Class-Wise Standard Deviation \\
+    \hline
+    $DeepConv(\omega)$ & 77\% & TODO \\
+    \hline
+    $DeepConvWithPos(\omega, \chi)$ & 91\% & TODO \\
+    \hline
+\end{tabular}
+
 \section{Semi-Supervised Learning} \label{semi_supervised}
@@
 
+This system was trained using $FourierMLP(\beta_{8})$ for $NN(x)$, because this was the most successful configuration in supervised learning.
+It produced the highest accuracy as well as the lowest class-wise standard deviation.
+
 \subsubsection{Results}
@@
     \caption{\label{grav_opt_validation} Accuracy of optimal gravitational differentiation training run compared to Acoustic Parameter (98\% accuracy).}
-\end{figure}
-
-This was not the only configuration that performed well; as seen in Figure \ref{grav_acc_by_run}, 10\% of the runs during the grid search produced accuracy greater than 93\%.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{grav_acc_by_run}
-    \caption{\label{grav_acc_by_run} Accuracy of gravitational differentiation arranged by training run during the grid search.}
 \end{figure}
@@
 
-Further demonstrating this is a powerful algorithm and that high performance was not a one-off, 35\% of the training runs replicated Acoustic Parameter with a mean accuracy greater than 92\% (shown in Figure \ref{icn_accuracy_by_run}.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{icn_accuracy_by_run}
-    \caption{\label{icn_accuracy_by_run} Mean accuracy of iterative cluster nucleation configurations during grid search.}
-\end{figure}
-
 Digging deeper into the performance throughout the runs of the grid search, Figure \ref{icn_acc_by_hyper} shows the mean accuracy as a function of each of the hyperparameters that were optimized.
@@
 
-The following table directly compares performance of the semi-supervised learning systems, and also includes the neural network used in the original PICO-60 paper as a baseline.
+The following table directly compares performance of the semi-supervised learning systems, with the existing PICO-60 neural network provided as a baseline once again.
+Note that ``Max Accuracy'' refers to accuracy at replication of Acoustic Parameter, rather than replication of the ground truths.
 
@@
     \hline
-    Technique & Max Accuracy & Max Mean by Run & $dim(\varsigma)=128$ Errors versus 256 \\
+    Technique & Max AP Accuracy & Max Mean by Run & $\varsigma$ Errors: 128 versus 256 \\
     \hline
@@
     \hline
-    Original PICO-60 NN & TODO\% & N/A & N/A \\
-    \hline
 \end{tabular}
+
+\section{Conclusions}
+
+This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like Acoustic Parameter.
+This conclusion has been achieved through 2 rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems and replicate Acoustic Parameter as accurately as possible.
+
+In the realm of supervised learning, it was made clear that the banded Fourier transform is the most effective input format.
+It produced significantly higher accuracy and tighter prediction distributions than neural networks trained on either the raw audio waveform or image window data.
+There is strong evidence that image data in particular does not contain sufficient information to be used for discrimination.
+
+2 semi-supervised learning algorithms were developed, applying the best configuration found for supervised learning.
+While both techniques produced some very effective discriminators, iterative cluster nucleation was found to be the more effective of the 2.
+It replicated Acoustic Parameter with a mean of 98\% accuracy on the best configuration.
+This indicates that the technique shows significant promise, and may be worth consideration as a discrimination technique for future experiments.
 

### 2018-08-07 documentation/paper.tex
Added TODOs for existing NN
Added lots of TODOs

@@
     \hline
+    Existing NN TODO & X & X \\
+    \hline
+    AP TODO & X & X \\
+    \hline
 \end{tabular}
@@
 
-This system was trained using $FourierMLP(\beta_{8})$ for $NN(x)$, because this was the most successful configuration in supervised learning.
+This system was trained using a multi-layer perceptron with an 8-band Fourier transform input for $NN(x)$, because this was one of the most successful configurations in supervised learning.
 It produced the highest accuracy as well as the lowest class-wise standard deviation.
@@
 The following table directly compares performance of the semi-supervised learning systems, with the existing PICO-60 neural network provided as a baseline once again.
-Note that ``Max Accuracy'' refers to accuracy at replication of Acoustic Parameter, rather than replication of the ground truths.
+Note that ``Max AP Accuracy'' refers to accuracy at replication of Acoustic Parameter, rather than replication of the ground truths.
 
-\begin{tabular}{|l|l|l|l|}
+\begin{tabular}{|l|l|l|}
     \hline
-    Technique & Max AP Accuracy & Max Mean by Run & $\varsigma$ Errors: 128 versus 256 \\
+    Technique & Max AP Accuracy & Max Mean by Run \\
     \hline
-    Gravitational Differentiation & 98\% & 96\% & 64\% Increase \\
+    Gravitational Differentiation (256 $\varsigma$) & 98\% & 96\% \\
     \hline
-    Iterative Cluster Nucleation & 100\% & 98\% & 85\% Increase \\
+    Gravitational Differentiation (128 $\varsigma$) & TODO & TODO \\
+    \hline
+    Iterative Cluster Nucleation (256 $\varsigma$) & 100\% & 98\% \\
+    \hline
+    Iterative Cluster Nucleation (128 $\varsigma$) & TODO & TODO \\
+    \hline
+    Existing NN TODO & X & X \\
+    \hline
+    AP TODO & X & X \\
     \hline

### 2018-08-07 documentation/paper.tex
More TODOs
Rearranged graphs

@@
 \usepackage{wrapfig}
+\usepackage{float}
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{wall_event_positions}
+    \caption{\label{} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
+\end{figure}
+
 The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the pressure transducer), and the acoustic cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
@@
 This takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{wall_event_positions}
-    \caption{\label{} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
-\end{figure}
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{image_window}
-    \caption{\label{image_window} Accuracy of $ImageConv(\iota)$ compared to Acoustic Parameter at discriminating between run types (85\% accuracy).}
-\end{figure}
-
 \subsection{Performance Summary}
@@
 \end{tabular}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{image_window}
+    \caption{\label{image_window} Accuracy of $ImageConv(\iota)$ compared to Acoustic Parameter at discriminating between run types (85\% accuracy).}
+\end{figure}
 
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{grav_opt_validation}
+    \caption{\label{grav_opt_validation} Accuracy of best gravitational differentiation training run compared to Acoustic Parameter (98\% accuracy).}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{grav_acc_by_hyper}
+    \caption{\label{grav_acc_by_hyper} Accuracy of gravitational differentiation as a function of hyperparameter values.}
+\end{figure}
+
 \subsubsection{Results}
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{grav_opt_validation}
-    \caption{\label{grav_opt_validation} Accuracy of optimal gravitational differentiation training run compared to Acoustic Parameter (98\% accuracy).}
-\end{figure}
-
 As seen in Figure \ref{grav_acc_by_hyper}, all 3 hyperparameters that were optimized during the grid search had a significant effect on accuracy:
@@
 \end{itemize}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{grav_acc_by_hyper}
-    \caption{\label{grav_acc_by_hyper} Accuracy of gravitational differentiation as a function of hyperparameter values.}
-\end{figure}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \centering
@@
 
-\begin{tabular}{|l|l|l|}
+\begin{tabular}{|l|l|l|l|}
     \hline
-    Technique & Max AP Accuracy & Max Mean by Run \\
+    Technique & Max AP Accuracy & Max Mean by Run & Class-Wise SD \\
     \hline
-    Gravitational Differentiation (256 $\varsigma$) & 98\% & 96\% \\
+    Gravitational Differentiation (256 $\varsigma$) & 98\% & 96\% & TODO \\
     \hline
-    Gravitational Differentiation (128 $\varsigma$) & TODO & TODO \\
+    Gravitational Differentiation (128 $\varsigma$) & TODO & TODO & TODO \\
     \hline
-    Iterative Cluster Nucleation (256 $\varsigma$) & 100\% & 98\% \\
+    Iterative Cluster Nucleation (256 $\varsigma$) & 100\% & 98\% & TODO \\
     \hline
-    Iterative Cluster Nucleation (128 $\varsigma$) & TODO & TODO \\
+    Iterative Cluster Nucleation (128 $\varsigma$) & TODO & TODO & TODO \\
     \hline
-    Existing NN TODO & X & X \\
+    Existing NN TODO & X & X & X \\
     \hline
-    AP TODO & X & X \\
+    AP TODO & X & X & X \\
     \hline

### 2018-08-08 documentation/paper.tex
Cited PICO-60 paper
Filled in data tables for supervised learning

@@
 
-The PICO-60 bubble chamber experiment for detection of weakly interacting massive particles (WIMPs) was completed in 2017.
+The PICO-60 bubble chamber experiment for detection of weakly interacting massive particles (WIMPs) \cite{pico} was completed in 2017.
 It provides an excellent basis on which to develop and test the efficacy of such techniques as a potential tool for event discrimination, relative to the conventional techniques used during initial analysis of the experiment.
@@
 
-$WallCut(D)$ was applied to the low-resolution Fourier transforms during early experiments (since they were used for Acoustic Parameter) to minimize the number of variables between the neural network and a proven discriminator.
+$WallCut(D)$ was applied to the low-resolution Fourier transforms during early experiments (since they were used for Acoustic Parameter) to minimize the number of variables between initial neural networks and a proven discriminator.
 
@@
 This time, position corrections were optionally included when position input was not used.
-The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$, $FourierMLP(PosCor(\beta_{N}))$ and $FourierMLPWithPos(\beta_{N}, \chi)$.
+The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$, $FourierMLP(PosCor(\beta_{N}, \chi))$ and $FourierMLPWithPos(\beta_{N}, \chi)$.
 Several different resolutions $N$ of $\beta_{N}$ were used.
@@
 $FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
-Increasing the number of bands, even up to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.
+Increasing the number of bands to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.
+40 bands reduced the accuracy once again, to 91\%.
+
+Very interestingly, $FourierMLP(\beta_{8})$ performed better than either $FourierMLP(PosCor(\beta{8}), \chi)$ or \\ $FourierMLPWithPos(\beta{8}, \chi)$, which produced accuracy values of 97\% and 95\%, respectively.
 
@@
 $FourierMLPWithPos(\beta_{20}, \chi)$ achieved only 92\% accuracy (2 percentage points lower than the same configuration without $WallCut(D)$).
-$FourierMLPWithPos(\beta_{40}, \chi)$ also managed 92\% validation accuracy.
 However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%.
@@
 
-The following table compares the accuracy and mean class-wise standard deviation values of various supervised learning configurations, with the original PICO-60 neural network provided as a baseline:
+The following tables compare the accuracy and mean class-wise standard deviation values of various supervised learning configurations, with Acoustic Parameter and the original PICO-60 neural network provided as baselines.
+
+With $WallCut(D)$:
 
@@
     \hline
-    $DeepConv(\omega)$ & 77\% & TODO \\
+    $FourierMLP(PosCor(\beta{8}), \chi)$ & 97\% & 0.42 \\
     \hline
-    $DeepConvWithPos(\omega, \chi)$ & 91\% & TODO \\
+    $FourierMLP(\beta{8})$ & 98\% & 0.29 \\
     \hline
-    Existing NN TODO & X & X \\
+    $FourierMLPWithPos(\beta{8}, \chi)$ & 95\% & 0.44 \\
     \hline
-    AP TODO & X & X \\
+    $FourierMLPWithPos(\beta{20}, \chi)$ & 94\% & 0.52 \\
+    \hline
+    $FourierMLPWithPos(\beta{40}, \chi)$ & 91\% & 0.54 \\
+    \hline
+    Original PICO-60 Neural Network & 94\% & 0.99 \\
+    \hline
+    Acoustic Parameter & N/A & 0.56 \\
+    \hline
+\end{tabular}
+
+Without $WallCut(D)$:
+
+\begin{tabular}{|l|l|l|}
+    \hline
+    Configuration & Max Accuracy & Mean Class-Wise Standard Deviation \\
+    \hline
+    $FourierMLPWithPos(\beta_{20}, \chi)$ & 92\% & 0.63 \\
+    \hline
+    $FourierMLPWithPos(\beta{50,001}, \chi)$ & 97\% & 0.47 \\
+    \hline
+    $DeepConv(\omega)$ & 77\% & 0.86 \\
+    \hline
+    $DeepConvWithPos(\omega, \chi)$ & 91\% & 0.61 \\
+    \hline
+    $ImageConv(\iota)$ & 85\% & 0.95 \\
+    \hline
+    Acoustic Parameter & N/A & 0.73 \\
     \hline

### 2018-08-09 training/triplets_neural_network_train.py
Combination of L1 and L2

@@ -7,7 +7,7 @@ import random
 import numpy as np
 from keras.layers import Dense, Dropout, BatchNormalization, InputLayer
 from keras.models import Model, Sequential
-from keras.regularizers import l2
+from keras.regularizers import l1_l2
 
 from data_processing.load_triplet_classification_data import load_triplet_classification_data
 
@@ -26,8 +26,8 @@ inputs = inputs[order]
 outputs = outputs[order]
 
 # Create a neural network model that includes several dense layers with hyperbolic tangent activations, L2 regularization, and batch normalization
-regularizer = l2(0.02)
-dropout = 0.15
+regularizer = l1_l2(l1=0.005, l2=0.02)
+dropout = 0.25
 activation = 'tanh'
 model = Sequential([
     InputLayer(input_shape=(16,)),

### 2018-08-09 training/triplets_neural_network_train.py
Played with regularization; surprisingly consistent 100%

@@ -26,8 +26,8 @@ inputs = inputs[order]
 outputs = outputs[order]
 
 # Create a neural network model that includes several dense layers with hyperbolic tangent activations, L2 regularization, and batch normalization
-regularizer = l1_l2(l1=0.005, l2=0.02)
-dropout = 0.25
+regularizer = l1_l2(l1=0, l2=0.03)
+dropout = 0.4
 activation = 'tanh'
 model = Sequential([
     InputLayer(input_shape=(16,)),
@@ -37,11 +37,11 @@ model = Sequential([
     Dense(9, activation=activation, kernel_regularizer=regularizer),
     Dropout(dropout),
     Dense(6, activation=activation, kernel_regularizer=regularizer),
-    Dropout(dropout),
+    # Dropout(dropout),
     Dense(3, activation=activation, kernel_regularizer=regularizer),
-    Dropout(dropout),
+    # Dropout(dropout),
     Dense(2, activation=activation, kernel_regularizer=regularizer),
-    Dropout(dropout),
+    # Dropout(dropout),
     Dense(1, activation=activation, kernel_regularizer=regularizer)
 ])
 # Output a summary of the model's architecture

### 2018-08-09 training/triplets_neural_network_train.py
Incorporated L1 regularization to apparently improve results somewhat

@@ -26,8 +26,8 @@ inputs = inputs[order]
 outputs = outputs[order]
 
 # Create a neural network model that includes several dense layers with hyperbolic tangent activations, L2 regularization, and batch normalization
-regularizer = l1_l2(l1=0, l2=0.03)
-dropout = 0.4
+regularizer = l1_l2(l1=0.005, l2=0.03)
+dropout = 0.3
 activation = 'tanh'
 model = Sequential([
     InputLayer(input_shape=(16,)),

### 2018-08-09 training/triplets_neural_network_train.py
Tweaked further for more frequent 100%

@@ -26,8 +26,8 @@ inputs = inputs[order]
 outputs = outputs[order]
 
 # Create a neural network model that includes several dense layers with hyperbolic tangent activations, L2 regularization, and batch normalization
-regularizer = l1_l2(l1=0.005, l2=0.03)
-dropout = 0.3
+regularizer = l1_l2(l1=0.0025, l2=0.03)
+dropout = 0.4
 activation = 'tanh'
 model = Sequential([
     InputLayer(input_shape=(16,)),

### 2018-08-09 training/waveform_grid_search.py
Enabled wall cuts for waveform grid search

@@ -23,7 +23,7 @@ event_data_set = EventDataSet(
         RunType.AMERICIUM_BERYLLIUM,
         RunType.CALIFORNIUM
     },
-    use_wall_cuts=False
+    use_wall_cuts=True
 )
 # Load training and validation data as NumPy arrays
 training_inputs, training_ground_truths, validation_inputs, validation_ground_truths = \
@@ -131,7 +131,7 @@ for l2_lambda in [0.003, 0.001, 0.0003]:
                     )
 
                     # Iterate over training and validation for several epochs
-                    for epoch in range(100):
+                    for epoch in range(300):
                         # Train the model on the input and ground truth arrays
                         model.fit(
                             x=training_inputs,

### 2018-08-09 training/waveform_grid_search.py
Reduced epochs to 100

@@ -131,7 +131,7 @@ for l2_lambda in [0.003, 0.001, 0.0003]:
                     )
 
                     # Iterate over training and validation for several epochs
-                    for epoch in range(300):
+                    for epoch in range(100):
                         # Train the model on the input and ground truth arrays
                         model.fit(
                             x=training_inputs,

### 2018-08-09 training/waveform_grid_search.py
Further reduced epochs to 50

@@ -131,7 +131,7 @@ for l2_lambda in [0.003, 0.001, 0.0003]:
                     )
 
                     # Iterate over training and validation for several epochs
-                    for epoch in range(100):
+                    for epoch in range(50):
                         # Train the model on the input and ground truth arrays
                         model.fit(
                             x=training_inputs,

### 2018-08-09 training/triplets_nucleation_grid_search.py
Changed grid search parameters according to lessons learned with supervised NN

@@ -26,8 +26,8 @@ VALIDATION_SET_SIZE = 8
 # Iterate over possible configurations for the number of initial training examples, the initial threshold, the threshold multiplier, the L2 lambda, and dropout regularization
 for initial_threshold in [0.01, 0.02]:
     for threshold_multiplier in [1.025, 1.05]:
-        for l2_lambda in [0, 0.001, 0.003]:
-            for dropout in [0, 0.25, 0.5]:
+        for l2_lambda in [0, 0.001, 0.003, 0.01, 0.03]:
+            for dropout in [0, 0.15, 0.3]:
                 # Test each configuration multiple times so the initial set does not cause a bias
                 for configuration_test_index in range(3):
                     # Print a few blank lines for separation

### 2018-08-09 documentation/paper.tex
Added subscripts to betas
Explained wall cuts more clearly

@@
 
-Except where specifically mentioned, $WallCut(D)$ was not applied.
-This was done to test the network's ability to handle complex acoustic characteristics including resonance.
+Except where specifically mentioned, $WallCut(D)$ was not applied. This was done for 2 reasons:
+
+\begin{enumerate}
+    \item As an exploratory effort, to determine whether or not it was possible to produce an effective discriminator without wall cuts.
+No such thing was attempted during the original PICO-60 analysis.
+    \item To test the network's ability to handle complex acoustic characteristics including resonance.
 There is a relatively clear-cut discrimination boundary in PICO-60, but this may not be the case in other applications.
+\end{enumerate}
 
@@
     \hline
-    $FourierMLP(PosCor(\beta{8}), \chi)$ & 97\% & 0.42 \\
+    $FourierMLP(PosCor(\beta_{8}), \chi)$ & 97\% & 0.42 \\
     \hline
-    $FourierMLP(\beta{8})$ & 98\% & 0.29 \\
+    $FourierMLP(\beta_{8})$ & 98\% & 0.29 \\
     \hline
-    $FourierMLPWithPos(\beta{8}, \chi)$ & 95\% & 0.44 \\
+    $FourierMLPWithPos(\beta_{8}, \chi)$ & 95\% & 0.44 \\
     \hline
-    $FourierMLPWithPos(\beta{20}, \chi)$ & 94\% & 0.52 \\
+    $FourierMLPWithPos(\beta_{20}, \chi)$ & 94\% & 0.52 \\
     \hline
-    $FourierMLPWithPos(\beta{40}, \chi)$ & 91\% & 0.54 \\
+    $FourierMLPWithPos(\beta_{40}, \chi)$ & 91\% & 0.54 \\
     \hline
@@
     \hline
-    $FourierMLPWithPos(\beta{50,001}, \chi)$ & 97\% & 0.47 \\
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 97\% & 0.47 \\
     \hline

### 2018-08-09 training/gravitational_grid_search.py
Separated offsets function into its own file

@@ -18,20 +18,7 @@ from keras.regularizers import l2
 from data_processing.bubble_data_point import RunType
 from data_processing.event_data_set import EventDataSet
 from data_processing.experiment_serialization import save_test
-
-
-def gravitational_ground_truth_offsets(predictions: np.ndarray, distortion_power: float, gravity_multiplier: float) -> np.ndarray:
-    """Get an array of ground truth offsets based on a gravitational model where examples classified very close to one edge will be pulled toward that edge, and examples near the middle will make little difference"""
-    # The function should pass through (0.5, 0), should change very little near that point, and should rapidly asymptote in the negative or positive directions as the prediction comes close to 0 or 1
-    # First, scale the predictions to the range of -1 to 1
-    predictions_scaled = (predictions - 0.5) * 2
-    # Take the hyperbolic tangent so examples in the middle are affected minimally
-    hyperbolic_tangent = np.tanh(predictions_scaled)
-    # Take the Nth power (removing the sign and multiplying it back in after so the negative side is the same as the positive side) of the hyperbolic tangent so that the area around 0 is squashed
-    power_distorted = np.sign(hyperbolic_tangent) * np.power(np.abs(hyperbolic_tangent), distortion_power)
-    # Multiply it by a constant so the gravitational offset does not dominate the training process
-    return power_distorted * gravity_multiplier
-
+from training.gravitational_ground_truth_offsets import gravitational_ground_truth_offsets
 
 # Iterate over different values for the number of definitive training examples, the gravity multiplier increment, the distortion power, and the stochastic gradient descent learning rate
 for definitive_training_examples in [32, 64, 128, 256]:

### 2018-08-09 training/gravitational_ground_truth_offsets.py

@@ -1,0 +1,18 @@
+#!/usr/bin/env python3
+"""A function used to distort a neural network's predictions and produce ground truth offsets that push unlabeled examples further to the side they are already nearest"""
+# Created by Brendon Matusch, July 2018
+
+import numpy as np
+
+
+def gravitational_ground_truth_offsets(predictions: np.ndarray, distortion_power: float, gravity_multiplier: float) -> np.ndarray:
+    """Get an array of ground truth offsets based on a gravitational model where examples classified very close to one edge will be pulled toward that edge, and examples near the middle will make little difference"""
+    # The function should pass through (0.5, 0), should change very little near that point, and should rapidly asymptote in the negative or positive directions as the prediction comes close to 0 or 1
+    # First, scale the predictions to the range of -1 to 1
+    predictions_scaled = (predictions - 0.5) * 2
+    # Take the hyperbolic tangent so examples in the middle are affected minimally
+    hyperbolic_tangent = np.tanh(predictions_scaled)
+    # Take the Nth power (removing the sign and multiplying it back in after so the negative side is the same as the positive side) of the hyperbolic tangent so that the area around 0 is squashed
+    power_distorted = np.sign(hyperbolic_tangent) * np.power(np.abs(hyperbolic_tangent), distortion_power)
+    # Multiply it by a constant so the gravitational offset does not dominate the training process
+    return power_distorted * gravity_multiplier

### 2018-08-09 training/gravitational_train.py

@@ -11,6 +11,7 @@ from data_processing.bubble_data_point import RunType
 from data_processing.event_data_set import EventDataSet
 from data_processing.experiment_serialization import save_test
 from models.banded_frequency_network import create_model
+from training.gravitational_ground_truth_offsets import gravitational_ground_truth_offsets
 
 # The number of training examples for which the ground truth is actually used, and is not dynamically generated
 DEFINITIVE_TRAINING_EXAMPLES = 128
@@ -52,20 +53,6 @@ training_data_set.validation_events = \
 # It must first be converted to floating-point
 training_ground_truths = training_ground_truths.astype(float)
 training_ground_truths[DEFINITIVE_TRAINING_EXAMPLES:] = 0.5
-
-
-def gravitational_ground_truth_offsets(predictions: np.ndarray, distortion_power: float, gravity_multiplier: float) -> np.ndarray:
-    """Get an array of ground truth offsets based on a gravitational model where examples classified very close to one edge will be pulled toward that edge, and examples near the middle will make little difference"""
-    # The function should pass through (0.5, 0), should change very little near that point, and should rapidly asymptote in the negative or positive directions as the prediction comes close to 0 or 1
-    # First, scale the predictions to the range of -1 to 1
-    predictions_scaled = (predictions - 0.5) * 2
-    # Take the hyperbolic tangent so examples in the middle are affected minimally
-    hyperbolic_tangent = np.tanh(predictions_scaled)
-    # Take the Nth power (removing the sign and multiplying it back in after so the negative side is the same as the positive side) of the hyperbolic tangent so that the area around 0 is squashed
-    power_distorted = np.sign(hyperbolic_tangent) * np.power(np.abs(hyperbolic_tangent), distortion_power)
-    # Multiply it by a constant so the gravitational offset does not dominate the training process
-    return power_distorted * gravity_multiplier
-
 
 # The gravity multiplier should start at 0 and is added to every epoch
 gravity_multiplier = 0

### 2018-08-09 training/gravitational_ground_truth_offsets.py
Removed hashbang for function file

@@ -1,4 +1,3 @@
-#!/usr/bin/env python3
 """A function used to distort a neural network's predictions and produce ground truth offsets that push unlabeled examples further to the side they are already nearest"""
 # Created by Brendon Matusch, July 2018
 

### 2018-08-09 visualization/visualize_gravity.py
Created gravity visualization script

@@ -1,0 +1,31 @@
+#!/usr/bin/env python3
+"""A script for creating a 3-dimensional graph of the gravitational ground truth offset generation function, parameterized by the prediction and distortion power"""
+# Created by Brendon Matusch, August 2018
+
+import matplotlib.pyplot as plt
+import numpy as np
+from mpl_toolkits.mplot3d import Axes3D
+
+from training.gravitational_ground_truth_offsets import gravitational_ground_truth_offsets
+
+# Create an array of possible predictions between 0 and 1 which will be graphed
+possible_predictions = np.linspace(0, 1, 100)
+# Create an array several distortion power options (not all of which are integers)
+distortion_powers = np.linspace(3, 11, 100)
+# Create mesh grids of all possible combinations
+possible_predictions_mesh, distortion_powers_mesh = np.meshgrid(possible_predictions, distortion_powers)
+# Vectorize the offset calculation function (inefficiently) so the entire grids can be processed in one call
+offsets_vectorized = np.vectorize(gravitational_ground_truth_offsets)
+# Run the function on the mesh grids, assuming a gravitational multiplier of 1
+offsets_mesh = offsets_vectorized(possible_predictions_mesh, distortion_powers_mesh, 1)
+
+# Create 3D axes in the default figure
+axes = plt.gca(projection='3d')
+# Plot the resulting on the vertical axis, with the predictions and distortion powers on the (X, Y) plane
+axes.plot_surface(X=possible_predictions_mesh, Y=distortion_powers_mesh, Z=offsets_mesh)
+# Label the 3 axes
+axes.set_xlabel('Network Prediction $p$')
+axes.set_ylabel('Distortion Power $\\psi$')
+axes.set_zlabel('$GravDiff(p, \\psi, 1)$')
+# Display the plot on screen
+plt.show()

### 2018-08-09 documentation/paper.tex
Cropped and renamed graph; explained grav diff a little better

@@
 
-\subsubsection{Structure}
+\subsubsection{Algorithm}
 
@@
 
-The essential purpose of $GravDiff(p, \psi, g)$ is to allow confidently classified unlabeled examples to influence the training process, pushing them further towards the classification they are already near (like gravity), while ensuring examples with low-confidence classifications do not have a significant adverse effect.
+The essential purpose of $GravDiff(p, \psi, g)$ is to allow confidently classified (near 0 or 1) unlabeled examples to influence the training process, pushing them further towards the classification they are already near (like gravity), while ensuring examples with low-confidence (near 0.5) classifications do not have a significant effect.
 
-It can be considered a distortion of the hyperbolic tangent that flattens the central range and exaggerates the asymptote on either side.
+It is a distortion of the hyperbolic tangent that flattens the central range and comparatively exaggerates the asymptote on either side.
 The network's prediction $p$ is first transformed so its range is 1 to -1 rather than 0 to 1 ($2(p - 0.5)$).
@@
 This permits use of the full range of curves produced by even and non-integer values of $\psi$.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{grav_diff}
+    \caption{\label{grav_diff} A visualization of $GravDiff(p, \psi, g)$ with respect to $p$ and $\psi$.}
+\end{figure}
+
+\subsubsection{Application}
 

### 2018-08-09 documentation/paper.tex
Clarified grav diff quite a bit

@@
 
-Gravitational differentiation is a novel technique for gradient computation in the final layer of a neural network.
-It is intended for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
+Gravitational differentiation is a novel technique for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
+
+It accomplishes this using a new method for calculating gradients for the last layer of the neural network.
+Based on the current predictions of a trained network on a certain unlabeled training example, a gradient is calculated for the last layer with the following purposes:
+
+\begin{itemize}
+    \item For examples with confident predictions close to 0 or 1: Create a large gradient that pulls the prediction closer to 0 or 1 through gradient descent (like gravity).
+    \item For examples with low-confidence predictions close to 0.5: Create a gradient near 0 that does not affect training significantly.
+\end{itemize}
+
+The gradients produced by this technique are used for training in the same way as gradients calculated with a loss function.
+Backpropagation is used to calculate gradients for previous layers, and a stochastic gradient descent optimizer is used for weight updates.
 
@@
 
-It makes use of the gravitational differentiation function $GravDiff(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$.
+Gradients are calculated using the gravitational differentiation function $GravDiff(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$.
 The function is defined as follows:
@@
 $GravDiff(p, \psi, g) = g \cdot sgn(p) \cdot abs(tanh(2(p - 0.5))) ^ \psi$
-
-The essential purpose of $GravDiff(p, \psi, g)$ is to allow confidently classified (near 0 or 1) unlabeled examples to influence the training process, pushing them further towards the classification they are already near (like gravity), while ensuring examples with low-confidence (near 0.5) classifications do not have a significant effect.
 
@@
 
-However, what is really desired is a shallow slope in the center (so low-confidence network outputs in the range of, for instance, 0.2 to 0.8 will produce very shallow output gradients).
-Applying a relatively large exponent (on the order of 9) accomplishes this, because values less than 1 will rapidly approach 0.
+However, what is really desired is a shallow slope in the center (so low-confidence network outputs near 0.5 will produce very shallow output gradients).
+Applying a relatively large exponent (in the range of 3 to 11) accomplishes this; it squashes values close to 0 closer to 0.
+
 However, this only works with odd integer exponents (because they produce negative outputs for negative inputs), limiting the ability to fine-tune this function by changing $\psi$.
-To resolve this, a piecewise exponential function is used, where the absolute value is taken prior to applying the power and the sign is multiplied in afterwards.
+To resolve this, a piecewise exponential function is used, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards.
 This permits use of the full range of curves produced by even and non-integer values of $\psi$.

### 2018-08-09 documentation/paper.tex
Cleaned up wording

@@
 
-However, this only works with odd integer exponents (because they produce negative outputs for negative inputs), limiting the ability to fine-tune this function by changing $\psi$.
+However, this only works if $p$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$.
 To resolve this, a piecewise exponential function is used, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards.

### 2018-08-09 documentation/paper.tex
Rearranged into enumerate

@@
 
-It is a distortion of the hyperbolic tangent that flattens the central range and comparatively exaggerates the asymptote on either side.
-The network's prediction $p$ is first transformed so its range is 1 to -1 rather than 0 to 1 ($2(p - 0.5)$).
-Next, the sigmoidal hyperbolic tangent is applied, producing a value of 0 in the center and asymptotic slopes to -1 and 1 at the edges.
+It is a distortion of the hyperbolic tangent that flattens the central range and comparatively exaggerates the asymptote on either side. The equation above can be described in the following steps:
 
-However, what is really desired is a shallow slope in the center (so low-confidence network outputs near 0.5 will produce very shallow output gradients).
-Applying a relatively large exponent (in the range of 3 to 11) accomplishes this; it squashes values close to 0 closer to 0.
+\begin{enumerate}
+    \item Transform the network's prediction $p$, so its range is 1 to -1 rather than 0 to 1.
+    \item Apply the sigmoidal hyperbolic tangent, producing a value of 0 in the center and asymptotic slopes to -1 and 1 at the edges.
+    \item \label{exponent} What is really desired is a shallow slope in the center (so low-confidence network outputs near 0.5 will produce very shallow output gradients).
+Apply a relatively large exponent which is $\psi$ (in the range of 3 to 11) to squash values close to 0 closer to 0.
+\end{enumerate}
 
-However, this only works if $p$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$.
-To resolve this, a piecewise exponential function is used, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards.
-This permits use of the full range of curves produced by even and non-integer values of $\psi$.
+(Note that step \ref{exponent} only works if $p$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$.
+To resolve this, a piecewise exponential function is used instead, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards.
+This permits use of the full range of curves produced by even and non-integer values of $\psi$.)
 
@@
     \centering
-    \includegraphics[width=\textwidth]{grav_diff}
+    \includegraphics[width=0.8\textwidth]{grav_diff}
     \caption{\label{grav_diff} A visualization of $GravDiff(p, \psi, g)$ with respect to $p$ and $\psi$.}

### 2018-08-09 documentation/paper.tex
Explained basic concept of ICN

@@
 
-\subsubsection{Structure}
+The basic concept is to start with a relatively small amount of labeled data from 2 classes (nuclear recoils and alpha particles, in this case).
+They can be visualized as clusters.
+A neural network is trained on the clusters, and used to run predictions on unlabeled data.
+
+Those examples that, based on the network's predictions, fall firmly into one cluster or the other, are added to that cluster, and are used for training.
+By repeating this process, the clusters are nucleated (expanded) to gradually include more and more of the data set.
+
+\subsubsection{Algorithm}
 

### 2018-08-10 documentation/paper.tex
Removed meaningless detail on grav diff

@@
 The mean over that run (also the highest overall) was 96\%.
-
 Figure \ref{grav_opt_validation} shows the highly accurate replication of Acoustic Parameter.
-However, the examples that are classified incorrectly seem to have no correlation with Acoustic Parameter, suggesting that the learned function cues on different properties of the audio.
-While the discriminator may still be accurate, this property implies it may be less predictable when trained on new data sets.
 

### 2018-08-10 documentation/paper.tex
Included table of results for waveform without wall cuts

@@
     \item A configuration with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
+This configuration was tested both with and without $WallCut(D)$.
 \end{enumerate}
@@
 Once the audio was cropped properly, $DeepConv(\omega)$ achieved a maximum of 77\% accuracy on validation data.
-Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ achieved 85\%.
-The maximum obtained on a grid search with position input was 91\% (although with a high class-wise standard deviation of 0.61).
+Using the same network architecture (except for the input layer), and still without $WallCut(D)$, $DeepConvWithPos(\omega, \chi)$ achieved 85\%.
+The maximum obtained on a grid search with this configuration was 91\% (although with a high class-wise standard deviation of 0.61).
 While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
 
-However, the lack of decisiveness observed even when accuracy was relatively high, as seen in Figure \ref{waveform_best_validation}, implies this is a suboptimal technique.
+However, the lack of decisiveness observed even when accuracy was relatively high, as seen in Figure \ref{waveform_best_validation}, implies that without $WallCut(D)$, this is a suboptimal technique.
 
@@
 \end{figure}
+
+!!!Include results with wall cuts!!!
+
+The following tables show the results of several $DeepConvWithPos(\omega, \chi)$ discriminators produced throughout testing, with corresponding network hyperparameters.
+
+With $WallCut(D)$:
+
+!!!TODO!!!
+
+Without $WallCut(D)$:
+
+\begin{tabular}{|l|l|l|l|l|l|l|}
+    \hline
+    L2 $\lambda$ & Dropout & Filters & Kernel Size & \# Layers Multiplier & Max Accuracy & Class-Wise Std Dev \\
+    \hline
+    0.003 & 0 & 24 & 3 & 3 & 91\% & 0.61 \\
+    \hline
+    0.003 & 0.5 & 24 & 5 & 3 & 91\% & 0.65 \\
+    \hline
+    0.001 & 0.25 & 24 & 3 & 3 & 91\% & 0.66 \\
+    \hline
+    0.001 & 0.5 & 48 & 5 & 3 & 88\% & 0.73 \\
+    \hline
+    0.0003 & 0 & 48 & 5 & 6 & 78\% & 0.95 \\
+    \hline
+    0.0003 & 0.5 & 48 & 5 & 6 & 74\% & 0.92 \\
+    \hline
+\end{tabular}
 

### 2018-08-10 documentation/paper.tex
Explained table

@@
 Using the same network architecture (except for the input layer), and still without $WallCut(D)$, $DeepConvWithPos(\omega, \chi)$ achieved 85\%.
-The maximum obtained on a grid search with this configuration was 91\% (although with a high class-wise standard deviation of 0.61).
+The maximum obtained on a grid search with this configuration was 91\% (although with a relatively high class-wise standard deviation of 0.61).
 While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
@@
 \end{tabular}
+
+This chart shows that simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
+Larger kernel sizes, more filters, and more layers make performance generally worse, while higher L2 regularization does the opposite.
+Dropout seems to have a weak correlation, possibly reducing accuracy slightly.
 

### 2018-08-10 training/high_resolution_grid_search.py
Enabled wall cuts for high resolution

@@ -2,8 +2,6 @@
 """Training script that trains the banded frequency neural network on several different hyperparameter combinations"""
 # Created by Brendon Matusch, July 2018
 
-import os
-
 from keras.layers import Input, BatchNormalization, Dense, Dropout, concatenate
 from keras.models import Model
 from keras.regularizers import l2
@@ -11,7 +9,6 @@ from keras.regularizers import l2
 from data_processing.event_data_set import EventDataSet
 from data_processing.bubble_data_point import RunType, load_bubble_frequency_domain
 from data_processing.experiment_serialization import save_test
-from models.banded_frequency_network import create_model
 
 # Load the event data set from the file, removing multiple-bubble events, disabling acoustic parameter cuts, and keeping background radiation and calibration runs
 event_data_set = EventDataSet(
@@ -20,7 +17,7 @@ event_data_set = EventDataSet(
         RunType.AMERICIUM_BERYLLIUM,
         RunType.CALIFORNIUM
     },
-    use_wall_cuts=False
+    use_wall_cuts=True
 )
 # Load training and validation data as NumPy arrays, currying the loading function to disable banding
 training_inputs, training_ground_truths, validation_inputs, validation_ground_truths = \

### 2018-08-10 documentation/paper.tex
Added data for banded frequency with wall cuts

@@
 
-Very interestingly, $FourierMLP(\beta_{8})$ performed better than either $FourierMLP(PosCor(\beta{8}), \chi)$ or \\ $FourierMLPWithPos(\beta{8}, \chi)$, which produced accuracy values of 97\% and 95\%, respectively.
+Very interestingly, $FourierMLP(\beta_{8})$ performed better than either $FourierMLP(PosCor(\beta_{8}), \chi)$ or \\ $FourierMLPWithPos(\beta{8}, \chi)$, which produced accuracy values of 97\% and 95\%, respectively.
 
@@
 However, that epoch had a relatively high standard deviation, at 0.47.
+
+The following tables compare various multi-layer perceptron discriminators, with and without wall cuts.
+
+With $WallCut(D)$:
+
+\begin{tabular}{|l|l|l|l|l|l|}
+    \hline
+    Configuration & L2 $\lambda$ & Dropout & Layers & Max Accuracy & Class-Wise Std Dev \\
+    \hline
+    $FourierMLP(PosCor(\beta_{8}))$ & 0 & 0.5 & 2 & 95\% & 0.56 \\
+    \hline
+    $FourierMLP(\beta_{8})$ & 0 & 0.5 & 2 & 98\% & 0.29 \\
+    \hline
+    $FourierMLPWithPos(\beta_{8}, \chi)$ & 0 & 0.5 & 3 & 95\% & 0.44 \\
+    \hline
+    $FourierMLPWithPos(\beta_{8}, \chi)$ & 0 & 0.25 & 3 & 97\% & 0.43 \\
+    \hline
+    $FourierMLPWithPos(\beta_{20}, \chi)$ & 0.003 & 0.25 & 3 & 93\% & 0.54 \\
+    \hline
+    $FourierMLPWithPos(\beta_{20}, \chi)$ & 0 & 0 & 3 & 94\% & 0.52 \\
+    \hline
+\end{tabular}
+
+Without $WallCut(D)$:
 
@@
 
-The following tables compare the accuracy and mean class-wise standard deviation values of various supervised learning configurations, with Acoustic Parameter and the original PICO-60 neural network provided as baselines.
+The following tables compare the accuracy and mean class-wise standard deviation values of various supervised learning configurations with multiple input formats, with Acoustic Parameter and the original PICO-60 neural network provided as baselines.
 

### 2018-08-10 documentation/paper.tex
Filled in table without wall cuts
Described tables

@@
 
-This chart shows that simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
+This table shows that simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
 Larger kernel sizes, more filters, and more layers make performance generally worse, while higher L2 regularization does the opposite.
@@
     \hline
+    $FourierMLPWithPos(\beta_{40}, \chi)$ & 0 & 0 & 3 & 91\% & 0.57 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.0003 & 0 & 4 & 91\% & 0.53 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0.25 & 4 & 91\% & 0.56 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0.5 & 3 & 89\% & 0.64 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.5 & 4 & 89\% & 0.65 \\
+    \hline
 \end{tabular}
 
+It is very clear here that higher resolutions do not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
+
 Without $WallCut(D)$:
+
+\begin{tabular}{|l|l|l|l|l|l|}
+    \hline
+    Configuration & L2 $\lambda$ & Dropout & Layers & Max Accuracy & Class-Wise Std Dev \\
+    \hline
+    $FourierMLPWithPos(\beta_{20}, \chi)$ & 0 & 0 & 3 & 92\% & 0.63 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0 & 0 & 3 & 95\% & 0.49 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0 & 2 & 96\% & 0.47 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0 & 2 & 97\% & 0.47 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.25 & 2 & 95\% & 0.54 \\
+    \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.5 & 2 & 94\% & 0.53 \\
+    \hline
+\end{tabular}
+
+Without wall cuts, the full-resolution Fourier transform produces much more effective discriminators.
+Here, L2 regularization is beneficial, possibly because there is more noise that can be overfit on.
 

### 2018-08-10 documentation/paper.tex
Added results table for ImageConv

@@
 
-Throughout a variety of different network architectures, the best validation accuracy obtained was 85\%.
+Throughout a variety of different network architectures, the best validation accuracy obtained was a deceptively high 85\%.
 The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.
-Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random (as is apparent in \ref{image_window}).
+Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random (as is visually apparent in \ref{image_window}).
 It is likely fitting on some form of noise within $\iota$.
@@
 
+The following table shows performance of various configurations of $ImageConv(\iota)$:
+
+\begin{tabular}{|l|l|l|l|l|l|}
+    \hline
+    L2 $\lambda$ & Dropout & Convolutional Layers & Dense Layers & Max Accuracy & Class-Wise Std Dev \\
+    \hline
+    0 & 0.5 & 4 & 3 & 77\% & 1.01 \\
+    \hline
+    0 & 0.25 & 4 & 3 & 85\% & 0.95 \\
+    \hline
+    0.01 & 0.25 & 4 & 3 & 66\% & 0.97 \\
+    \hline
+    0.003 & 0.25 & 4 & 3 & 69\% & 0.97 \\
+    \hline
+    0.01 & 0.25 & 12 & 3 & 66\% & 0.98 \\
+    \hline
+    0.01 & 0.5 & 12 & 3 & 68\% & 0.99 \\
+    \hline
+\end{tabular}
+
 \subsection{Performance Summary}
 
-The following tables compare the accuracy and mean class-wise standard deviation values of various supervised learning configurations with multiple input formats, with Acoustic Parameter and the original PICO-60 neural network provided as baselines.
+The following tables compare the accuracy and mean class-wise standard deviation values of certain highlight supervised learning configurations with multiple input formats, with Acoustic Parameter and the original PICO-60 neural network provided as baselines.
 

### 2018-08-10 documentation/paper.tex
Updated and described overall table for supervised learning

@@
     \hline
+    $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 91\% & 0.53 \\
+    \hline
     Original PICO-60 Neural Network & 94\% & 0.99 \\
@@
 \end{tabular}
+
+There are 2 key conclusions to be drawn from the above tables:
+
+\begin{enumerate}
+    \item With $WallCut(D)$, the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
+Higher resolutions seem to make performance worse, likely due to overfitting.
+It is significantly better than the original PICO-60 neural network, and is more decisive (as measured by the class-wise standard deviation)
+    \item Without $WallCut(D)$, full-resolution $\beta_{50,001}$ is the most effective, nearing the best performance with wall cuts at separating run types.
+All models are notably less decisive, given that the lowest standard deviation without wall cuts is 0.47, compared to 0.29 with wall cuts.
+\end{enumerate}
 

### 2018-08-11 documentation/paper.tex
Removed highly misleading graph
Rearranged wording of introduction

@@
 
-In experiments for dark matter detection, one of the most important and challenging hurdles to overcome is that of background radiation.
+In the PICO-60 experiment, and in most other experiments for dark matter detection, one of the most important and challenging hurdles to overcome is that of background radiation.
 Unwanted particles from a variety of sources can produce event signatures that appear very similar to those of dark matter candidates.
@@
 
-The objective of this study is to explore a potential methodology that resolves both of these problems using techniques based on machine learning.
-These include supervised learning algorithms that learn to separate calibration data sets, and semi-supervised learning algorithms that simultaneously learn an effective discriminator and improve the purity of their own training data sets.
+The objective of this study is to investigate the potential for machine learning techniques to address both of these issues.
+More specifically, in the context of the PICO-60 experiment, there are 2 objectives:
+
+\begin{enumerate}
+    \item Determine whether machine learning techniques can function comparably to a conventionally derived discriminator.
+Furthermore, develop several neural network architectures and apply them to different input formats, in order to find which are the most effective.
+
+    \item Determine whether these techniques can be extended to develop an effective discriminator function based on incomplete and inaccurate initial information.
+
+    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only calibration data for training.
+While calibration runs are known to be mostly correlated with particle types, the relatively large percentage of impurities has the potential to hinder conventional learning techniques.
+\end{enumerate}
+
+The benefit to experimental science would be to potentially allow the creation of a discriminator in cases where it would be difficult to construct one by conventional means.
+There may be various reasons for this:
+
+\begin{enumerate}
+    \item Limited ability to collect accurate calibration data, and/or imperfect correlation between calibration data and particle types.
+    \item Unknown and/or complex interactions in the experimental apparatus make a conventional discriminator difficult to define, where machine learning techniques may be able to uncover hidden correlations in the input data.
+\end{enumerate}
 
@@
 
+\subsection{Performance Analysis}
+
+Performance of machine learning systems was evaluated using 2 metrics.
+One is a simple classification accuracy value, which corresponds to the number of validation examples the network classifies incorrectly.
+This effectively summarizes how often the network is outright correct or incorrect relative to the baseline provided for it.
+However, this fails to capture how decisive the network is: whether its prediction is confidently high or low, or is only slightly closer to one edge of the spectrum.
+
+A metric that captures this property is the class-wise standard deviation.
+This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
+It calculates the spread of the network's predictions for each ground truth class.
+This means that a decisive discriminator, which produces a wide separation between the two classes, is preferred over one that produces a nebulous cloud of outputs with a seemingly arbitrary decision boundary.
+
+$S=std(N \cup A)$
+
+$C=(std(N \div S) + std(A \div S)) \div 2$
+
+The first step is to calculate the standard deviation S of the union of the two sets.
+This gives an indication of the scale of the overall distribution.
+When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
+While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, Acoustic Parameter has a significantly wider range.
+Normalization of the union prevents this from creating a bias where Acoustic Parameter would produce a higher standard deviation with a similarly proportioned error.
+
+The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually.
+This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are.
+Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off of those specific values, will produce a low class-wise standard deviation.
+
 \subsection{Optimization Process}
@@
 \end{enumerate}
-
-\subsection{Performance Analysis}
-
-The performance of supervised learning neural networks was evaluated using 2 metrics.
-One is a simple classification accuracy value, which corresponds to the number of validation examples the network classifies incorrectly.
-This effectively summarizes how often the network is outright correct or incorrect, but fails to capture how decisive it is.
-
-A metric that captures this property is the class-wise standard deviation.
-This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
-It calculates the spread of the network's predictions for each ground truth class.
-This means that a decisive discriminator, which produces a wide separation between the two classes, is preferred over one that produces a nebulous cloud of outputs with a seemingly arbitrary decision boundary.
-
-$S=std(N \cup A)$
-
-$C=(std(N \div S) + std(A \div S)) \div 2$
-
-The first step is to calculate the standard deviation S of the union of the two sets.
-This gives an indication of the scale of the overall distribution.
-When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
-While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, Acoustic Parameter has a significantly wider range.
-Normalization of the union prevents this from creating a bias where Acoustic Parameter would produce a higher standard deviation with a similarly proportioned error.
-
-The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually.
-This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are.
-Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off of those specific values, will produce a low class-wise standard deviation.
 
@@
 Very interestingly, $FourierMLP(\beta_{8})$ performed better than either $FourierMLP(PosCor(\beta_{8}), \chi)$ or \\ $FourierMLPWithPos(\beta{8}, \chi)$, which produced accuracy values of 97\% and 95\%, respectively.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{fourier_mlp_no_correction}
-    \caption{\label{fourier_mlp_no_correction} Accuracy of $FourierMLP(\beta_{8})$ compared to Acoustic Parameter at discriminating between run types (98\% accuracy).}
-\end{figure}
 
@@
 Regularization techniques, which permit effective training on smaller sets, are applied with this purpose.
-
-\subsection{Performance Analysis}
-
-One of the main purposes of semi-supervised learning is to handle the impure training data effectively, and learn to replicate the imperfectly represented underlying pattern (nuclear recoils versus alpha particles) rather than merely fitting to separate run types.
-Therefore, performance is quantified by how accurately a system is able to replicate Acoustic Parameter, rather than simply reproduce the labels provided.
 

### 2018-08-11 visualization/graph_binary_predictions.py
Scrapped the run type stuff; going to do this analysis the right way

@@ -16,8 +16,8 @@ from utilities.verify_arguments import verify_arguments
 # Verify that a path to the JSON data file is passed
 verify_arguments('JSON data file')
 
-# Load the data set from the file
-events, ground_truths, network_outputs = load_test(sys.argv[1])
+# Load the data set from the file, ignoring the run types
+events, _, network_outputs = load_test(sys.argv[1])
 # Get the acoustic parameter and neural network score data from the events, if they are present (if one is, both will be)
 if hasattr(events[0], 'logarithmic_acoustic_parameter'):
     acoustic_parameters, original_neural_network_scores = zip(
@@ -29,8 +29,10 @@ else:
     random_values = [random.uniform(0, 1) for _ in range(len(network_outputs))]
     acoustic_parameters = random_values
     original_neural_network_scores = random_values
+# Calculate the actual ground truths, which are defined based on whether AP is above or below a certain threshold
+ground_truths = np.array(acoustic_parameters) > 0.25
 
-    # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
+# Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
 for criterion_data, criterion_name, classification_threshold in zip(
     copy.deepcopy([network_outputs, acoustic_parameters, original_neural_network_scores]),
     ['network outputs', 'acoustic parameters', 'original neural network scores'],
@@ -74,8 +76,8 @@ plt.scatter(
     c=point_colors
 )
 # Create patches that describe the 2 differently colored classes
-background_patch = Patch(color='r', label='Background radiation runs')
-calibration_patch = Patch(color='b', label='Neutron calibration runs')
+background_patch = Patch(color='r', label='Alpha Particles')
+calibration_patch = Patch(color='b', label='Neutrons')
 # Display them in a legend
 plt.legend(handles=[background_patch, calibration_patch])
 # Label the X and Y axes

### 2018-08-11 visualization/graph_binary_predictions.py
Revert "Scrapped the run type stuff; going to do this analysis the right way"
This reverts commit 5c5b12dd66d409d6942a63270d517e46e859459f.

@@ -16,8 +16,8 @@ from utilities.verify_arguments import verify_arguments
 # Verify that a path to the JSON data file is passed
 verify_arguments('JSON data file')
 
-# Load the data set from the file, ignoring the run types
-events, _, network_outputs = load_test(sys.argv[1])
+# Load the data set from the file
+events, ground_truths, network_outputs = load_test(sys.argv[1])
 # Get the acoustic parameter and neural network score data from the events, if they are present (if one is, both will be)
 if hasattr(events[0], 'logarithmic_acoustic_parameter'):
     acoustic_parameters, original_neural_network_scores = zip(
@@ -29,10 +29,8 @@ else:
     random_values = [random.uniform(0, 1) for _ in range(len(network_outputs))]
     acoustic_parameters = random_values
     original_neural_network_scores = random_values
-# Calculate the actual ground truths, which are defined based on whether AP is above or below a certain threshold
-ground_truths = np.array(acoustic_parameters) > 0.25
 
-# Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
+    # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
 for criterion_data, criterion_name, classification_threshold in zip(
     copy.deepcopy([network_outputs, acoustic_parameters, original_neural_network_scores]),
     ['network outputs', 'acoustic parameters', 'original neural network scores'],
@@ -76,8 +74,8 @@ plt.scatter(
     c=point_colors
 )
 # Create patches that describe the 2 differently colored classes
-background_patch = Patch(color='r', label='Alpha Particles')
-calibration_patch = Patch(color='b', label='Neutrons')
+background_patch = Patch(color='r', label='Background radiation runs')
+calibration_patch = Patch(color='b', label='Neutron calibration runs')
 # Display them in a legend
 plt.legend(handles=[background_patch, calibration_patch])
 # Label the X and Y axes

### 2018-08-11 documentation/paper.tex
Cleared up description of semi-supervised learning

@@
 \begin{enumerate}
-    \item Determine whether machine learning techniques can function comparably to a conventionally derived discriminator.
+    \item Determine the most effective machine learning architecture to use in development of a discriminator.
 Furthermore, develop several neural network architectures and apply them to different input formats, in order to find which are the most effective.
 
-    \item Determine whether these techniques can be extended to develop an effective discriminator function based on incomplete and inaccurate initial information.
+    \item Determine whether these techniques can be extended, applying novel algorithms based on semi-supervised learning, to develop an effective discriminator function based on incomplete and inaccurate initial information.
 
@@
 
+For all supervised learning configurations, accuracy and class-wise standard deviation statistics were computed using the calibration run type as a base line, as opposed to Acoustic Parameter.
+
+Because supervised learning systems are trained strictly to replicate ground truth data provided to them (the run type), their predictions will tend to incorporate biases similar to those found in the training set.
+Thus, comparison to Acoustic Parameter is not particularly meaningful.
+
 \subsection{Convolutional Neural Network for Raw Waveform Analysis}
@@
 
-During early tests with variations on $DeepConv(\omega)$, it was revealed to have an incredible capacity to separate run types based on unwanted biases.
+During early tests with variations on $DeepConv(\omega)$, it was revealed to have a surprising and undesirable capacity to separate run types based on meaningless information.
 It was able to separate the AmBe calibration run from combined background runs with 96\% accuracy, and to separate all calibration runs from background runs with 91\% accuracy.
@@
 The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.
-Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random (as is visually apparent in \ref{image_window}).
+Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random.
 It is likely fitting on some form of noise within $\iota$.
@@
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{image_window}
-    \caption{\label{image_window} Accuracy of $ImageConv(\iota)$ compared to Acoustic Parameter at discriminating between run types (85\% accuracy).}
-\end{figure}
-
 \section{Semi-Supervised Learning} \label{semi_supervised}
@@
 
-Semi-supervised learning is the idea of training a machine learning model on a set of labeled data in addition to a set of unlabeled data.
-This has the advantage of requiring fewer labels to be collected than conventional supervised learning.
-However, it also has a very powerful advantage when training data is impure and the function to be learned is relatively simple (not simple in that it can be easily calculated by a person, but simple in that it does not require a huge set of training data).
-It allows a neural network to reinforce its own decisions.
+Semi-supervised learning is an uncommon machine learning technique, relative to widely-used supervised and unsupervised learning.
+The concept is to train a machine learning model on a set of labeled data in addition to a set of unlabeled data.
+
+In the context of PICO-60, the labeled sets consist of single-bubble events from the calibration sets, which are known to contain approximately 90\% neutrons, and the background radiation data sets, which contain approximately 99\% alpha particles.
+In general, these labeled sets should have a strong but not perfect correlation with particle types.
+The unlabeled data can consist of any mixture of particle types, as long as they are all represented in the labeled data.
+
+The primary advantage of this technique is clear: it requires fewer labels to be collected than conventional supervised learning.
+However, it also has a very powerful advantage when training data is impure and the function to be learned is relatively simple (not simple in that it can be easily calculated by a person, but simple in that it does not require a huge set of training data): It allows a neural network to reinforce its own decisions.
 
@@
 Regularization techniques, which permit effective training on smaller sets, are applied with this purpose.
+
+Since a goal of these techniques is for the network to reinforce the pattern of particle types which is represented imperfectly in the run types, it is reasonable to compare them to Acoustic Parameter.
+Thus, the accuracy and class-wise standard deviation use Acoustic Parameter as a baseline.
 

### 2018-08-11 documentation/paper.tex
Further clarified semi-supervised stuff, and scrapped misleading graphs

@@
 
-However, the lack of decisiveness observed even when accuracy was relatively high, as seen in Figure \ref{waveform_best_validation}, implies that without $WallCut(D)$, this is a suboptimal technique.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{waveform_best_validation}
-    \caption{\label{waveform_best_validation} Accuracy of $DeepConvWithPos(\omega, \chi)$ compared to Acoustic Parameter at discriminating between run types without $WallCut(D)$ (91\% accuracy).}
-\end{figure}
+However, the lack of decisiveness observed even when accuracy was relatively high implies that, without $WallCut(D)$, this is a suboptimal technique.
 
@@
 
-Gravitational differentiation is a novel technique for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
+Gravitational differentiation is a novel technique, developed during this study, for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
 
-It accomplishes this using a new method for calculating gradients for the last layer of the neural network.
+In general terms, it amplifies the training effect associated with high-confidence predictions (those that most clearly show characteristics of either an alpha particle or a neutron) on unlabeled examples.
+Meanwhile, it minimizes the training impact of low-confidence predictions, such that the network will not be trained to make incorrect predictions.
+
+Analogous to a gravitational effect, events with high-confidence predictions are attracted closer to their corresponding classifications.
+They are then used as training data, allowing the patterns they represent to be used more generally to make predictions on other events.
+
+The system accomplishes this using a new method for calculating gradients for the last layer of the neural network.
 Based on the current predictions of a trained network on a certain unlabeled training example, a gradient is calculated for the last layer with the following purposes:
@@
 \begin{enumerate}
-    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for $\upsilon$.
+    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
     \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $GravDiff(p, \psi, g)$ and record the resulting gradients for the next iteration.
@@
 
+!!!Include table on grid search!!!
+
 \subsection{Iterative Cluster Nucleation}
@@
 
-Iterative cluster nucleation is a semi-supervised learning algorithm that takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
+Iterative cluster nucleation is a second semi-supervised learning algorithm developed as part of this study.
+It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
 
-The basic concept is to start with a relatively small amount of labeled data from 2 classes (nuclear recoils and alpha particles, in this case).
-They can be visualized as clusters.
-A neural network is trained on the clusters, and used to run predictions on unlabeled data.
+The basic concept is similar to that of gravitational differentiation.
+Starting with a relatively small amount of labeled data from 2 classes (nuclear recoils and alpha particles, in this case), predictions are run on a set of unlabeled data, and those predictions are used to produce further training data.
+However, there are 2 key differences:
 
-Those examples that, based on the network's predictions, fall firmly into one cluster or the other, are added to that cluster, and are used for training.
-By repeating this process, the clusters are nucleated (expanded) to gradually include more and more of the data set.
+\begin{enumerate}
+    \item In iterative cluster nucleation, an unlabeled example is not included in the training process at all, until a very confident prediction is made on it.
+Once that occurs, it is added to the training set, expanding 1 of the 2 ``clusters'' of training data.
+    \item Once an example is added to the active training set, it is never removed.
+This ensures that the highly confident predictions made at the very beginning continue to influence the training process.
+    \item The unlabeled examples, once added to the active training set, are weighted just as strongly as examples from the original labeled data set.
+This ensures imperfect labels in the labeled data do not outweigh the confident and more likely correct predictions from later in the training process.
+\end{enumerate}
 

### 2018-08-11 visualization/graph_supervised_prediction_histogram.py
Created script for graphing prediction histogram

@@ -1,0 +1,29 @@
+#!/usr/bin/env python3
+"""Given a saved validation set from a supervised learning system, load it and display a histogram showing the distribution of the network's predictions for both ground truth classes"""
+# Created by Brendon Matusch, August 2018
+
+import sys
+
+import matplotlib.pyplot as plt
+
+from data_processing.experiment_serialization import load_test
+from utilities.verify_arguments import verify_arguments
+
+# Verify that a path to the JSON data file is passed
+verify_arguments('JSON data file')
+
+# Load the data set from the file
+events, ground_truths, network_outputs = load_test(sys.argv[1])
+# Separate the network's outputs based on the value of the corresponding ground truth
+network_outputs_false = [output for output, ground_truth in zip(network_outputs, ground_truths) if not ground_truth]
+network_outputs_true = [output for output, ground_truth in zip(network_outputs, ground_truths) if ground_truth]
+
+# Set the size of the resulting graph (it should be standard across all such graphs)
+plt.figure(figsize=(8, 6))
+# Plot the network's outputs by ground truth in a histogram, labeling the 2 classes
+plt.hist([network_outputs_false, network_outputs_true], label=['Neutron Calibration Data', 'Alpha Background Data'])
+# Label the axes of the graph
+plt.xlabel('Network Prediction')
+plt.ylabel('Count')
+# Display the graph on screen
+plt.show()

### 2018-08-11 visualization/graph_supervised_prediction_histogram.py
Increased number of bins to 20

@@ -21,7 +21,7 @@ network_outputs_true = [output for output, ground_truth in zip(network_outputs, 
 # Set the size of the resulting graph (it should be standard across all such graphs)
 plt.figure(figsize=(8, 6))
 # Plot the network's outputs by ground truth in a histogram, labeling the 2 classes
-plt.hist([network_outputs_false, network_outputs_true], label=['Neutron Calibration Data', 'Alpha Background Data'])
+plt.hist([network_outputs_false, network_outputs_true], bins=20, label=['Neutron Calibration Data', 'Alpha Background Data'])
 # Label the axes of the graph
 plt.xlabel('Network Prediction')
 plt.ylabel('Count')

### 2018-08-11 documentation/paper.tex
Added histograms for supervised learning techniques

@@
 
-However, the lack of decisiveness observed even when accuracy was relatively high implies that, without $WallCut(D)$, this is a suboptimal technique.
+However, there is a lack of decisiveness observed even when accuracy was relatively high, in Figure \ref{waveform_hist}. This implies that, without $WallCut(D)$, this is a suboptimal technique.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{waveform_hist}
+    \caption{\label{waveform_hist} Prediction distribution of the best $DeepConvWithPos(\omega, \chi)$ discriminator.}
+\end{figure}
 
@@
 $FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
+Its very decisive prediction distribution can be observed in Figure \ref{banded_no_pos_input_hist}.
 Increasing the number of bands to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.
 40 bands reduced the accuracy once again, to 91\%.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{banded_no_pos_input_hist}
+    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best $FourierMLP(\beta_{8})$ discriminator, with $WallCut(D)$.}
+\end{figure}
 
@@
 However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%.
-However, that epoch had a relatively high standard deviation, at 0.47.
+However, that epoch had a relatively high class-wise standard deviation, at 0.47, as seen in Figure \ref{high_freq_no_wall_hist}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{high_freq_no_wall_hist}
+    \caption{\label{high_freq_no_wall_hist} Prediction distribution of the best $FourierMLPWithPos(\beta_{50,001}, \chi)$ discriminator, without $WallCut(D)$.}
+\end{figure}
 
@@
 The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.
-Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random.
+Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random (as seen in Figure \ref{image_hist}).
 It is likely fitting on some form of noise within $\iota$.
 The fact that throughout many trials (including a grid search), no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{image_hist}
+    \caption{\label{image_hist} Prediction distribution of the best $ImageConv(\iota)$ discriminator.}
+\end{figure}
 
@@
 
-The following tables compare the accuracy and mean class-wise standard deviation values of certain highlight supervised learning configurations with multiple input formats, with Acoustic Parameter and the original PICO-60 neural network provided as baselines.
+The following tables compare the accuracy and mean class-wise standard deviation values of certain highlight supervised learning configurations with multiple input formats, with the original PICO-60 neural network provided as a baseline.
 
@@
     \hline
-    Acoustic Parameter & N/A & 0.56 \\
-    \hline
 \end{tabular}
@@
     $ImageConv(\iota)$ & 85\% & 0.95 \\
-    \hline
-    Acoustic Parameter & N/A & 0.73 \\
     \hline

### 2018-08-11 visualization/graph_binary_predictions.py
Color by particle type rather than run type

@@ -16,8 +16,8 @@ from utilities.verify_arguments import verify_arguments
 # Verify that a path to the JSON data file is passed
 verify_arguments('JSON data file')
 
-# Load the data set from the file
-events, ground_truths, network_outputs = load_test(sys.argv[1])
+# Load the data set from the file, ignoring the run type ground truths
+events, _, network_outputs = load_test(sys.argv[1])
 # Get the acoustic parameter and neural network score data from the events, if they are present (if one is, both will be)
 if hasattr(events[0], 'logarithmic_acoustic_parameter'):
     acoustic_parameters, original_neural_network_scores = zip(
@@ -29,8 +29,10 @@ else:
     random_values = [random.uniform(0, 1) for _ in range(len(network_outputs))]
     acoustic_parameters = random_values
     original_neural_network_scores = random_values
+# Calculate actual neutron/alpha ground truths based on AP
+ground_truths = np.array(acoustic_parameters) > 0.25
 
-    # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
+# Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
 for criterion_data, criterion_name, classification_threshold in zip(
     copy.deepcopy([network_outputs, acoustic_parameters, original_neural_network_scores]),
     ['network outputs', 'acoustic parameters', 'original neural network scores'],
@@ -74,8 +76,8 @@ plt.scatter(
     c=point_colors
 )
 # Create patches that describe the 2 differently colored classes
-background_patch = Patch(color='r', label='Background radiation runs')
-calibration_patch = Patch(color='b', label='Neutron calibration runs')
+background_patch = Patch(color='r', label='Alpha Particles')
+calibration_patch = Patch(color='b', label='Neutrons')
 # Display them in a legend
 plt.legend(handles=[background_patch, calibration_patch])
 # Label the X and Y axes

### 2018-08-11 documentation/paper.tex
Switched out graphs to highlight particle types rather than run types

@@
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{grav_opt_validation}
-    \caption{\label{grav_opt_validation} Accuracy of best gravitational differentiation training run compared to Acoustic Parameter (98\% accuracy).}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{grav_acc_by_hyper}
-    \caption{\label{grav_acc_by_hyper} Accuracy of gravitational differentiation as a function of hyperparameter values.}
-\end{figure}
-
 \subsubsection{Results}
@@
 The mean over that run (also the highest overall) was 96\%.
-Figure \ref{grav_opt_validation} shows the highly accurate replication of Acoustic Parameter.
+Figure \ref{grav_grid_search} shows the highly accurate replication of Acoustic Parameter.
 
@@
 !!!Include table on grid search!!!
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{grav_grid_search}
+    \caption{\label{grav_grid_search} Accuracy of best gravitational differentiation training run compared to Acoustic Parameter (98\% accuracy).}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{grav_acc_by_hyper}
+    \caption{\label{grav_acc_by_hyper} Accuracy of gravitational differentiation as a function of hyperparameter values.}
+\end{figure}
 
@@
 
+Compared to gravitational differentiation, there is a larger delta between the number of errors on the validation set depending on the number of initial training examples used.
+Those with 128 examples in $\varsigma$ make 85\% more errors on average.
+However, iterative cluster nucleation with 128 examples still makes 19\% fewer errors than gravitational differentiation with 256 examples in $\varsigma$.
+
 \begin{figure}[H]
@@
 \end{figure}
-
-Compared to gravitational differentiation, there is a larger delta between the number of errors on the validation set depending on the number of initial training examples used.
-Those with 128 examples in $\varsigma$ make 85\% more errors on average.
-However, iterative cluster nucleation with 128 examples still makes 19\% fewer errors than gravitational differentiation with 256 examples in $\varsigma$.
 

### 2018-08-11 visualization/graph_supervised_prediction_histogram.py
Add legend to graphs

@@ -25,5 +25,7 @@ plt.hist([network_outputs_false, network_outputs_true], bins=20, label=['Neutron
 # Label the axes of the graph
 plt.xlabel('Network Prediction')
 plt.ylabel('Count')
+# Include a legend in the graph, explaining the colors
+plt.legend()
 # Display the graph on screen
 plt.show()

### 2018-08-11 documentation/paper.tex
Cleared up wording for supervised learning introduction

@@
 
-These conversion, preprocessing, and data cutting steps were reapplied during this study, and are described in section \ref{data_formats}.
+These conversion, preprocessing, and data cutting steps were reapplied during this study, and are described in Section \ref{data_formats}.
 
@@
 2 semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
-Both are discussed in detail in section \ref{semi_supervised}.
+Both are discussed in detail in Section \ref{semi_supervised}.
 
@@
 
-Important unknowns to be determined through experimentation with supervised learning included the most effective input format for the network, and the associated general network structure.
-Three major such configurations were applied:
+Supervised learning is the most common machine learning paradigm, in which a system is trained on a training set of predefined inputs and ground truths.
+
+As overviewed in Section \ref{data_formats}, there are many possible formats for the input.
+In addition, different network architectures are necessitated to process these inputs.
+Three major such configurations were tested:
 

### 2018-08-11 data_processing/experiment_serialization.py
Switched to data drive for saved validation sets because EXT4 is special

@@ -45,8 +45,7 @@ def save_test(event_data_set: EventDataSet, validation_ground_truths: np.ndarray
         # Add the processed dictionary to the list
         output_list.append(bubble_information)
     # Create a JSON file in the temporary folder, named with the prefix, current Unix time, and epoch number, save the data in it, and notify the user
-    json_file_path = os.path.expanduser(
-        f'~/{prefix}time{int(time.time())}_epoch{epoch}.json')
+    json_file_path = os.path.expanduser(f'/media/brendonm/Data/{prefix}time{int(time.time())}_epoch{epoch}.json')
     with open(json_file_path, 'w') as output_file:
         json.dump(output_list, output_file)
     print('Data saved at', json_file_path)

### 2018-08-12 documentation/paper.tex
Made a ton of edits, mostly for readability

@@
 The objective of this study is to investigate the potential for machine learning techniques to address both of these issues.
-More specifically, in the context of the PICO-60 experiment, there are 2 objectives:
+More specifically, in the context of the PICO-60 experiment, there are two objectives:
 
@@
 
-In this study, newly developed methodologies were compared to the 2 discrimination techniques developed previously for the PICO-60 experiment: Acoustic Parameter and machine learning.
+In this study, newly developed methodologies were compared to the two discrimination techniques developed previously for the PICO-60 experiment: Acoustic Parameter and machine learning.
 
@@
 
-Acoustic Parameter is the technique used for event discrimination in the original PICO-60 analysis.
+Acoustic Parameter is the technique that was used for event discrimination in the original PICO-60 analysis.
 It is a function that discriminates between alpha particles and nuclear recoils based on audio data.
@@
 The lowest-level data derived from the piezos is the raw waveform $\omega$.
-$\omega$ consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
+This consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
 They are each represented as a 16-bit integer.
@@
 
-The raw waveform $\omega$ was converted into the frequency domain by means of a 1-dimensional Discrete Fourier Transform for real input.
+The raw waveform $\omega$ was converted into the frequency domain by means of a one-dimensional Discrete Fourier Transform for real input.
 Since DFTs produce sequences of complex numbers, the magnitude of each element was computed.
@@
 
-For each event that is captured, the 4 cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected.
+For each event that is captured, the four cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected.
 These raw images contain a large amount of extraneous information; they encompass the entire vessel.
@@
 The 71 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
-$\iota$ includes 5 frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
+$\iota$ includes five frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
 
@@
 The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the pressure transducer), and the acoustic cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
-
-\textit{Fiducial Cut}
-
-The fiducial cut $FidWallCut(D)$ defines an spatial area along the walls of the vessel within which no events are accepted.
-It makes use of the bubble position $\chi$, the distance from the bubble to the center of the vessel $R$, and the distance from the bubble to the nearest wall, which is defined as $min(\texttt{Dwall}, \texttt{Dwall\_horiz})$.
-It is defined as follows:
+Those cuts restrict data as follows:
 
 \begin{itemize}
-    \item $Z \leq 523$
-    \item Any of the following 4 restrictions are true:
-    \begin{itemize}
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (0 < \chi_{Z} \leq 400)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 6) \land (\chi_{Z} \leq 0) \land (R \leq 100)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (\chi_{Z} \leq 0) \land (R > 100)$
-        \item $(min(\texttt{Dwall}, \texttt{Dwall\_horiz}) > 13) \land (\chi_{Z} > 400)$
+    \item The fiducial cut $FidWallCut(D)$ defines a spatial area along the walls of the vessel within which no events are accepted.
+It determines whether events fall within this area based on the visually calculated position $\chi$.
+    \item The pressure cut $PresWallCut(D)$ restricts the pressure detected by the pressure transducer, without any position corrections, to be close to 1 ($0.7<\texttt{dytranC}<1.3$).
+    \item The acoustic cut $AcWallCut(D)$ is defined using the banded Fourier transform $\beta_{8}$, and takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
 \end{itemize}
-\end{itemize}
-
-\textit{Pressure Cut}
-
-The pressure cut $PresWallCut(D)$ restricts the pressure detected by the pressure transducer, without any position corrections, to be close to 1 ($0.7<\texttt{dytranC}<1.3$).
-In combination with the acoustic cut $AcWallCut(D)$, this acts as a backup to the fiducial cut.
-
-\textit{Acoustic Cut}
-
-The acoustic cut $AcWallCut(D)$ is defined using the banded Fourier transform $\beta_{8}$ as follows:
-
-$45 < mean(\beta_{8}[(0, 2), (0, 1)]) < 300$
-
-This takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
 
@@
 \begin{itemize}
-    \item Supervised learning, in which a system is trained on a set of labeled data to classify unseen examples according to the patterns it observes in the training set.
-The training set is very important; biases and errors often slow training and produce worse results.
+    \item Supervised learning, in which a system is trained on a set of fully-labeled data to classify unseen examples according to the patterns it observes in the training set.
     \item Unsupervised learning, in which a system finds patterns or clusters in a set of unlabeled data.
-This can find order in almost any training set, but it is unpredictable and may not find the particular patterns it is expected to find.
+This can find order in almost any training set, but it is unpredictable and may not find the particular patterns desired.
 \end{itemize}
@@
 
-2 semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
+Two semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
 Both are discussed in detail in Section \ref{semi_supervised}.
@@
 
-Performance of machine learning systems was evaluated using 2 metrics.
-One is a simple classification accuracy value, which corresponds to the number of validation examples the network classifies incorrectly.
-This effectively summarizes how often the network is outright correct or incorrect relative to the baseline provided for it.
-However, this fails to capture how decisive the network is: whether its prediction is confidently high or low, or is only slightly closer to one edge of the spectrum.
+Performance of machine learning systems was evaluated using two metrics: classification accuracy and class-wise standard deviation.
 
-A metric that captures this property is the class-wise standard deviation.
-This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
+Classification accuracy is the ability of the network to separate the events into the two desired classes.
+Numerically, this corresponds to the number of validation examples the network classifies correctly.
+
+Class-wise standard deviation captures how decisive the network is: whether its prediction is confidently high or low, or is only slightly closer to one edge of the spectrum.
+
+Numerically, This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
 It calculates the spread of the network's predictions for each ground truth class.
@@
 This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are.
-Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off of those specific values, will produce a low class-wise standard deviation.
+Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off those specific values, will produce a low class-wise standard deviation.
 
@@
 
-Learning algorithms in this study are optimized systematically, with a variety of neural network configurations and input formats.
-To obtain meaningful results, it is important to minimize the number of variables during each individual run.
-Thus, since the 2 semi-supervised algorithms incorporate neural network models, it is important that the input formats and other hyperparameters of these networks are optimized first to some degree.
-This means the task is split into 2 distinct phases:
+Each of the learning algorithms used in this study applies a neural network.
+For each of the general configurations (multi-layer perceptron, one-dimensional convolutional neural network, et cetera) there are many possible specific architectures with different hyperparameters, which have to be optimized to improve performance.
+Some examples of hyperparameters are:
 
-\begin{enumerate}
-    \item Develop and verify several different supervised learning systems.
-Their performance is evaluated based on their ability to replicate the ground truths they are trained on.
-While this does not produce an optimal discriminator, it does reveal architectures that are capable, in essence, of learning to accomplish the discrimination task.
-    \item Using general supervised learning systems and input formats that were determined to be effective, optimize the hyperparameters of the semi-supervised learning algorithm to reproduce Acoustic Parameter as closely as possible.
-Since Acoustic Parameter is not used to optimize the neural network directly at any point, models that excel at this task demonstrate an ability to learn the desired discriminator while ignoring the impurities in the training set.
-\end{enumerate}
+\begin{itemize}
+    \item Number of layers of each type (dense, convolutional)
+    \item Number of neurons in dense layers
+    \item Number of filters, kernel size, and stride in convolutional layers
+    \item Types and amounts of regularization to apply (L2, dropout \cite{dropout})
+\end{itemize}
+
+Several of these hyperparameters are optimized at a time, using a grid search.
+Given $n$ different hyperparameters to optimize and $m$ different possible values for each, $m^n$ different networks are trained (using every possible combination of hyperparameters) and tested on a validation set.
 
@@
 
-Supervised learning is the most common machine learning paradigm, in which a system is trained on a training set of predefined inputs and ground truths.
-
-As overviewed in Section \ref{data_formats}, there are many possible formats for the input.
-In addition, different network architectures are necessitated to process these inputs.
-Three major such configurations were tested:
+The primary objective, in experimentation with supervised learning, is to determine the most effective combination of input format and neural network architecture to use in replicating a discriminator function.
+Three major configurations were tested:
 
@@
 \begin{enumerate}
-    \item As an exploratory effort, to determine whether or not it was possible to produce an effective discriminator without wall cuts.
-No such thing was attempted during the original PICO-60 analysis.
+    \item As an exploratory effort, to determine whether or not it was possible to produce an effective discriminator without wall cuts. No such attempt was made during the original PICO-60 analysis.
     \item To test the network's ability to handle complex acoustic characteristics including resonance.
@@
 
-For all supervised learning configurations, accuracy and class-wise standard deviation statistics were computed using the calibration run type as a base line, as opposed to Acoustic Parameter.
+All supervised learning configurations were trained and evaluated based on their ability to discriminate calibration runs (predominantly neutrons) from background runs (predominantly alpha particles).
+This was used instead of Acoustic Parameter for two reasons:
 
-Because supervised learning systems are trained strictly to replicate ground truth data provided to them (the run type), their predictions will tend to incorporate biases similar to those found in the training set.
-Thus, comparison to Acoustic Parameter is not particularly meaningful.
+\begin{enumerate}
+    \item In future experiments, when no Acoustic Parameter is available, impure data is the only information available for training. The network structure should be selected to accommodate this.
+    \item In general, supervised learning systems learn to replicate biases present in the training data set.
+Thus, to evaluate the network's effectiveness at processing the input data, the run type (which is used for training) must be used as a baseline instead of Acoustic Parameter.
+\end{enumerate}
 
@@
 A convolutional neural network was used for analyzing the raw waveform $\omega$ directly, without any preprocessing whatsoever.
-This avoids any destruction of information, and should be possible, in theory, for a neural network to handle.
-For this task, a very deep (20 layers in its shallowest configuration) 1-dimensional fully convolutional neural network was applied.
+This avoids any destruction of information; theoretically, a sufficiently complex neural network should be capable of deriving any characteristics needed.
+
+For this task, a very deep (20 layers in its shallowest configuration) one-dimensional fully convolutional neural network was applied.
 The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.
-L2 regularization was used to alleviate overfitting, and batch normalization was applied to the input.
-(Its hyperparameters were modified significantly through manual empirical testing as well as grid searches.)
+L2 regularization was used to alleviate overfitting, and batch normalization was applied to the input to stabilize training.
+The network's hyperparameters were modified significantly through manual empirical testing as well as grid searches.
 
@@
 Several different resolutions $N$ of $\beta_{N}$ were used.
-While there were variety of network architectures tested, most had a small number of layers (on the order of 3), applied dropout and L2 regularization, and used batch normalization once again.
+While a variety of network architectures were tested, most had a small number of layers (on the order of three), applied dropout and L2 regularization, and used batch normalization once again.
 
@@
 Increasing the number of bands to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.
-40 bands reduced the accuracy once again, to 91\%.
+Using 40 bands reduced the accuracy once again, to 91\%.
 
@@
 
-In an effort to resolve this, a 2-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$.
-The network architecture consists of a moderate number of convolutional layers (on the scale of 9) and 3 dense layers at the end.
+In an effort to resolve this, a two-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$.
+The network architecture consists of a moderate number of convolutional layers (on the scale of nine) and three dense layers at the end.
 L2 regularization was used on all layers, and dropout was additionally used on the dense layers.
@@
 Throughout a variety of different network architectures, the best validation accuracy obtained was a deceptively high 85\%.
-The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.
+The corresponding training accuracy of 100\% indicates that severe overfitting is taking place (not unexpected, given the relatively small image data set).
 Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random (as seen in Figure \ref{image_hist}).
@@
 \end{tabular}
+
+The high standard deviations across the board are further evidence for the hypothesis that there is insufficient information in the image data.
 
@@
 
+Based on these results, experimentation with semi-supervised learning proceeded using \\ $FourierMLP(PosCor(\beta_{8}))$, applying a 3-layer perceptron architecture, and using $WallCut(D)$.
+There were three main reasons for this:
+
+\begin{enumerate}
+    \item This was a highly effective configuration with wall cuts, producing 97\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
+    \item The use of the exact same input format as Acoustic Parameter makes the algorithms more directly comparable.
+\end{enumerate}
+
 \section{Semi-Supervised Learning} \label{semi_supervised}
@@
 Semi-supervised learning is an uncommon machine learning technique, relative to widely-used supervised and unsupervised learning.
-The concept is to train a machine learning model on a set of labeled data in addition to a set of unlabeled data.
+The concept is to train a machine learning model on a set of labeled (classified) data in addition to a set of unlabeled (unclasified) data.
 
-In the context of PICO-60, the labeled sets consist of single-bubble events from the calibration sets, which are known to contain approximately 90\% neutrons, and the background radiation data sets, which contain approximately 99\% alpha particles.
-In general, these labeled sets should have a strong but not perfect correlation with particle types.
-The unlabeled data can consist of any mixture of particle types, as long as they are all represented in the labeled data.
+In the context of PICO-60, the labeled sets consist of:
 
-The primary advantage of this technique is clear: it requires fewer labels to be collected than conventional supervised learning.
-However, it also has a very powerful advantage when training data is impure and the function to be learned is relatively simple (not simple in that it can be easily calculated by a person, but simple in that it does not require a huge set of training data): It allows a neural network to reinforce its own decisions.
+\begin{enumerate}
+    \item Single-bubble events from the calibration sets, which are known to contain approximately 90\% neutrons
+    \item Single-bubble events from the background radiation data sets, which contain approximately 99\% alpha particles
+\end{enumerate}
 
-In both of the semi-supervised learning algorithms applied, to arrive at an effective discriminator, the network first trains on a smaller amount of imperfect data.
-Because the ground truth data is imperfect, training on a relatively small set for a short period of time (potentially using regularization, in addition) ensures that the network will be less likely to overfit and its learned function will be simpler.
-This is in line with the objective, since, assuming the training inputs are normalized correctly, it should be simpler to discriminate between classes of particles than calibration runs.
+In general, these labeled sets should have a strong but not perfect correlation with particle types. The unlabeled data can consist of any mixture of particle types.
 
-The essence of these 2 algorithms, then, is a positive feedback loop in which, by using the network's most confident predictions in the training process, the desired (simpler) function is reinforced, when a conventional supervised learning system would have learned the more complex function that is taught by the ground truths.
+The primary advantage of semi-supervised learning is clear: it requires fewer labels to be collected than conventional supervised learning.
+A less obvious but powerful secondary advantage is that, since the learning process allows the network to reinforce its own decisions, the negative effect of impure training data is minimized, potentially allowing the resulting network to perform better than a network trained with supervised learning.
 
-In addition to improving performance with a labeled set known to be adequate, a goal of this category of techniques is to minimize the amount of labeled data necessary.
-This means less time must be spent on collection of calibration data, which can be a very slow process.
-Regularization techniques, which permit effective training on smaller sets, are applied with this purpose.
+Two techniques for semi-supervised learning have been developed for this study: gravitational differentiation and iterative cluster nucleation.
+The essence of both algorithms is a positive feedback loop.
+First, the network is trained on a smaller amount of imperfect data.
+Throughout the training process, the network runs predictions on unlabeled data.
+By using its most confident predictions are used in the training process, the desired function (separating particle types) is continuously reinforced, while less confident predictions are de-emphasized.
+This effectively draws new, unlabeled data into the training set.
 
-Since a goal of these techniques is for the network to reinforce the pattern of particle types which is represented imperfectly in the run types, it is reasonable to compare them to Acoustic Parameter.
-Thus, the accuracy and class-wise standard deviation use Acoustic Parameter as a baseline.
+Since the intent of these techniques is to separate particle types (the same as Acoustic Parameter), performance evaluations (accuracy and class-wise standard deviation) use Acoustic Parameter as a baseline.
 
@@
 
-As seen in Figure \ref{grav_acc_by_hyper}, all 3 hyperparameters that were optimized during the grid search had a significant effect on accuracy:
+As seen in Figure \ref{grav_acc_by_hyper}, all three hyperparameters that were optimized during the grid search had a significant effect on accuracy:
 
@@
 The basic concept is similar to that of gravitational differentiation.
-Starting with a relatively small amount of labeled data from 2 classes (nuclear recoils and alpha particles, in this case), predictions are run on a set of unlabeled data, and those predictions are used to produce further training data.
-However, there are 2 key differences:
+Starting with a relatively small amount of labeled data from two classes (nuclear recoils and alpha particles, in this case), predictions are run on a set of unlabeled data, and those predictions are used to produce further training data.
+However, there are two key differences:
 
@@
     \item In iterative cluster nucleation, an unlabeled example is not included in the training process at all, until a very confident prediction is made on it.
-Once that occurs, it is added to the training set, expanding 1 of the 2 ``clusters'' of training data.
+Once that occurs, it is added to the training set, expanding one of the two ``clusters'' of training data.
     \item Once an example is added to the active training set, it is never removed.
@@
 This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like Acoustic Parameter.
-This conclusion has been achieved through 2 rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems and replicate Acoustic Parameter as accurately as possible.
+This conclusion has been achieved through two rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems and replicate Acoustic Parameter as accurately as possible.
 
@@
 
-2 semi-supervised learning algorithms were developed, applying the best configuration found for supervised learning.
-While both techniques produced some very effective discriminators, iterative cluster nucleation was found to be the more effective of the 2.
+Two semi-supervised learning algorithms were developed, applying the best configuration found for supervised learning.
+While both techniques produced some very effective discriminators, iterative cluster nucleation was found to be the more effective of the two.
 It replicated Acoustic Parameter with a mean of 98\% accuracy on the best configuration.

### 2018-08-12 visualization/graph_supervised_prediction_histogram.py
Clarified labels in graphing script

@@ -23,8 +23,8 @@ plt.figure(figsize=(8, 6))
 # Plot the network's outputs by ground truth in a histogram, labeling the 2 classes
 plt.hist([network_outputs_false, network_outputs_true], bins=20, label=['Neutron Calibration Data', 'Alpha Background Data'])
 # Label the axes of the graph
-plt.xlabel('Network Prediction')
-plt.ylabel('Count')
+plt.xlabel('Network Prediction (0 $\Rightarrow$ Neutron Calibration, 1 $\Rightarrow$ Alpha Background)')
+plt.ylabel('Validation Event Count')
 # Include a legend in the graph, explaining the colors
 plt.legend()
 # Display the graph on screen

### 2018-08-12 documentation/paper.tex
Added placeholder for abstract; minor fixes

@@
 \begin{document}
+
+\title{Developing a Bubble Chamber Particle Discriminator Using Semi-Supervised Learning}
+\author{Brendon Matusch, X, Y, and Z}
+\date{August 2018}
+\maketitle
+
+\begin{abstract}
+    TODO, once content is affirmed.
+\end{abstract}
 
@@
 
-Numerically, This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
+Numerically, this is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
 It calculates the spread of the network's predictions for each ground truth class.
@@
 Throughout the training process, the network runs predictions on unlabeled data.
-By using its most confident predictions are used in the training process, the desired function (separating particle types) is continuously reinforced, while less confident predictions are de-emphasized.
+By using its most confident predictions in the training process, the desired function (separating particle types) is continuously reinforced, while less confident predictions are de-emphasized.
 This effectively draws new, unlabeled data into the training set.

### 2018-08-13 documentation/paper.tex
Added data table for waveform with wall cuts
Added and explained table for waveform with wall cut

@@
 
-During early tests with variations on $DeepConv(\omega)$, it was revealed to have a surprising and undesirable capacity to separate run types based on meaningless information.
+During early tests with variations on $DeepConv(\omega)$ without $WallCut(D)$, it was revealed to have a surprising and undesirable capacity to separate run types based on meaningless information.
 It was able to separate the AmBe calibration run from combined background runs with 96\% accuracy, and to separate all calibration runs from background runs with 91\% accuracy.
@@
 
-!!!Include results with wall cuts!!!
+When $WallCut(D)$ was included, both accuracy and class-wise standard deviation improved drastically.
+The best configuration during a grid search was 96\% accurate and had a standard deviation of 0.41.
+This is better by both metrics than the multi-layer perceptron used in the original PICO-60 analysis.
 
@@
 
-!!!TODO!!!
+\begin{tabular}{|l|l|l|l|l|l|l|}
+    \hline
+    L2 $\lambda$ & Dropout & Filters & Kernel Size & \# Layers Multiplier & Max Accuracy & Class-Wise Std Dev \\
+    \hline
+    0.003 & 0 & 24 & 3 & 3 & 95\% & 0.44 \\
+    \hline
+    0.003 & 0 & 48 & 5 & 3 & 96\% & 0.41 \\
+    \hline
+    0.003 & 0.5 & 24 & 5 & 3 & 93\% & 0.50 \\
+    \hline
+    0.001 & 0 & 48 & 5 & 3 & 95\% & 0.48 \\
+    \hline
+    0.001 & 0.25 & 48 & 3 & 6 & 93\% & 0.62 \\
+    \hline
+    0.0003 & 0.5 & 24 & 5 & 6 & 95\% & 0.49 \\
+    \hline
+\end{tabular}
 
@@
 
-This table shows that simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
+These tables show that, both with and without $WallCut(D)$, simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
 Larger kernel sizes, more filters, and more layers make performance generally worse, while higher L2 regularization does the opposite.

### 2018-08-13 documentation/paper.tex
Wrote initial abstract

@@
 \begin{abstract}
-    TODO, once content is affirmed.
+    Background radiation is a major hurdle to overcome for bubble chamber dark matter experiments such as PICO-60.
+The current practice of manually developing a discriminator function to eliminate background data is difficult when calibration data is frequently impure and present only in small quantities.
+In this study, several different input formats and neural network architectures are applied to the task.
+First, they are optimized in a supervised learning context.
+Next, novel semi-supervised learning algorithms are trained, and found to replicate the Acoustic Parameter discriminator previously used in PICO-60 with up to 100\% accuracy.
 \end{abstract}

### 2018-08-13 documentation/paper.tex
Added histogram for waveform with wall cut

@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \centering
     \includegraphics[width=\textwidth]{waveform_hist}
-    \caption{\label{waveform_hist} Prediction distribution of the best $DeepConvWithPos(\omega, \chi)$ discriminator.}
+    \caption{\label{waveform_hist} Prediction distribution of the best $DeepConvWithPos(\omega, \chi)$ discriminator without $WallCut(D)$.}
 \end{figure}
@@
 When $WallCut(D)$ was included, both accuracy and class-wise standard deviation improved drastically.
-The best configuration during a grid search was 96\% accurate and had a standard deviation of 0.41.
+The best configuration during a grid search was 96\% accurate and had a standard deviation of 0.41 (seen in Figure \ref{waveform_wall_cut_hist}).
 This is better by both metrics than the multi-layer perceptron used in the original PICO-60 analysis.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{waveform_wall_cut_hist}
+    \caption{\label{waveform_wall_cut_hist} Prediction distribution of the best $DeepConvWithPos(\omega, \chi)$ discriminator with $WallCut(D)$.}
+\end{figure}
 

### 2018-08-20 documentation/paper.tex
Lots of fairly minor edits

@@
 \usepackage{float}
+\usepackage{caption}
 
@@
 \begin{abstract}
-    Background radiation is a major hurdle to overcome for bubble chamber dark matter experiments such as PICO-60.
-The current practice of manually developing a discriminator function to eliminate background data is difficult when calibration data is frequently impure and present only in small quantities.
+    The identification of non-signal events is a major hurdle to overcome for bubble chamber dark matter experiments such as PICO-60.
+The current practice of manually developing a discriminator function to eliminate background events is difficult when calibration data is frequently impure and present only in small quantities.
 In this study, several different input formats and neural network architectures are applied to the task.
@@
 
-In the PICO-60 experiment, and in most other experiments for dark matter detection, one of the most important and challenging hurdles to overcome is that of background radiation.
-Unwanted particles from a variety of sources can produce event signatures that appear very similar to those of dark matter candidates.
+In the PICO-60 experiment, and in most other experiments striving to directly detect dark matter, one of the most important and challenging hurdles to overcome is that of background events.
+Unwanted particles from a variety of sources can produce event signatures that appear very similar to those which are expected to be created by dark matter candidates.
 
-To resolve this, physicists must determine a discriminator function that can, based on some sensory data, separate dark matter candidates from background radiation.
+To resolve this, physicists must determine a discriminator function that can, based on experimental data, separate events generated by possible dark matter candidates from those due to background radiation.
 Two general problems stand in the way:
@@
     \item A detailed model of the physical environment is often used to produce an accurate discriminator.
-This takes a long time even for a team of experienced physicists, and will have to be updated whenever the physical variables of the experimental apparatus change.
+This takes a long time, and will have to be updated whenever the physical variables of the experimental apparatus change.
 
-    \item The primary intention of many dark matter experiments, which is to reduce background radiation to the lowest level possible, means that there will be a very small number of background events detected.
-This places a severe limit on the amount of data that is available to optimize such a discriminator.
+    \item The primary intention of many dark matter experiments, which is to reduce the number of background events to the lowest level possible, means that there will be a very small number detected.
+This places a significant constraint on the amount of data that is available to optimize such a discriminator.
 
@@
 
-The benefit to experimental science would be to potentially allow the creation of a discriminator in cases where it would be difficult to construct one by conventional means.
-There may be various reasons for this:
+The architecture described may have wider applications to several similar cases, such as those that require a discriminator which is difficult to construct using conventional techniques.
 
@@
 
-The PICO-60 bubble chamber experiment for detection of weakly interacting massive particles (WIMPs) \cite{pico} was completed in 2017.
+The PICO-60 bubble chamber experiment, created for the detection of weakly interacting massive particles (WIMPs) \cite{pico}, was completed in 2017.
 It provides an excellent basis on which to develop and test the efficacy of such techniques as a potential tool for event discrimination, relative to the conventional techniques used during initial analysis of the experiment.
 
-WIMP candidates in the cross section and mass ranges that the PICO-60 detector is sensitive to are predicted to produce nuclear recoils with the same properties as those created by neutrons.
-The key difference is that neutrons frequently (but not always) scatter and produce multiple bubbles, where the extremely small predicted cross sections of WIMP candidates mean that they should almost invariably create just one bubble.
+WIMP candidates in the cross-section and mass ranges to which the PICO-60 detector is sensitive are predicted to produce nuclear recoils with the same properties as those induced by neutrons.
+The key difference is that neutrons frequently (but not always) scatter several times and produce multiple bubbles, where the extremely small predicted cross-sections of WIMP candidates mean that they should almost invariably create just one bubble.
 This means single-bubble neutron events can be used to optimize a discriminator to detect WIMP events.
@@
 
-Acoustic Parameter is the technique that was used for event discrimination in the original PICO-60 analysis.
-It is a function that discriminates between alpha particles and nuclear recoils based on audio data.
-It was derived based on a physical model of the bubble chamber, and was tuned using calibration runs using neutron sources to produce events in the PICO-60 apparatus.
+The Acoustic Parameter method (abbreviated AP) is the technique that was used for event discrimination in the original PICO-60 analysis.
+A function was defined that discriminates between alpha particles and nuclear recoils based on audio data.
+It was derived based on a physical model of the bubble chamber, and was tuned using neutron calibration sources to produce events in the PICO-60 apparatus.
 
-Before Acoustic Parameter can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$).
+Before the Acoustic Parameter can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$).
 It must also be preprocessed with the position correction function $PosCor(\beta_{8}, \chi)$, which adjusts the overall volume of an audio recording according to its position in the vessel.
-This is done because the resonance characteristics differ significantly depending on the position of an event.
+This is done because the acoustic characteristics differ significantly depending on the position of an event in the vessel.
 
@@
 The lowest-level data derived from the piezos is the raw waveform $\omega$.
-This consists of a series of samples collected every 400,000\textsuperscript{th} of a second.
+This consists of a series of samples collected every $2.5 \times 10^{-6}$ seconds.
 They are each represented as a 16-bit integer.
-Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because before this there is background noise, and afterwards there is a clipped signal produced by hydraulics repressurizing the vessel.
+Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because there is only background noise prior to this window, and a clipped signal produced by hydraulics repressurizing the vessel after.
 (These sections contain no information and can be undesirably fit on by neural networks.)
@@
 
-The raw waveform $\omega$ was converted into the frequency domain by means of a one-dimensional Discrete Fourier Transform for real input.
+The raw waveform $\omega$ was converted into the frequency domain by means of a one-dimensional Discrete Fourier Transform (DFT) for real input.
 Since DFTs produce sequences of complex numbers, the magnitude of each element was computed.
-The full-resolution Fourier transform $\beta _{50,001}$ consists of a sequence of complex numbers, half the length of the raw waveform $\omega$.
+The direct output from such a DFT is the full-resolution Fourier transform $\beta _{50,001}$, which consists of a sequence of numbers half the length of the raw waveform $\omega$.
+(It is half the length because the output of a DFT of real inputs is Hermitian, meaning there is no new information contained in the second half of the output.)
 
 Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to Acoustic Parameter and the original neural analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands.
-This can be thought of as a method of signal downscaling, compressing the signal into a smaller number of data points.
+This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.
+
+This process was generalized to several other banding resolutions. Higher-resolution counterparts of $\beta_{8}$ included $\beta_{20}$ and $\beta_{40}$.
 
@@
 
-For each event that is captured, the four cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected.
+For each event that is captured, the four cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected using the cameras and an entropy-based trigger.
 These raw images contain a large amount of extraneous information; they encompass the entire vessel.
@@
 The 71 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
-$\iota$ includes five frames from before the recording trigger, because the bubble does not cause a trigger until it is already at a significant size.
+Five frames from before the recording trigger are included in $\iota$, because the bubble does not cause a trigger until it is already of a significant size.
 
@@
 \begin{itemize}
-    \item The run was not collected during engineering or testing ($\texttt{run\_type}\neq99$)
-    \item Recording was triggered by the camera ($\texttt{trigger\_main}=0$)
-    \item Acoustic Parameter is not erroneously large and negative ($log_{10}(\texttt{acoustic\_bubnum})>-100$)
-    \item Recorded more than 25 seconds after reaching target pressure ($\texttt{te}>25$)
+    \item The run was not collected during engineering or testing
+    \item Recording was triggered by the camera
+    \item Acoustic Parameter is not erroneously large and negative ($log_{10}(AP)>-100$)
+    \item Recorded more than 25 seconds after reaching target pressure
     \item The bubble position $\chi$ was successfully calculated ($[\chi_{X}, \chi_{Y}, \chi_{Z}]\neq[-100, -100, -100]$)
@@
 \begin{itemize}
-    \item Either 0 or 1 bubbles are detected based on images from the camera ($\texttt{nbub}<2$)
-    \item The number of bubbles approximated using the pressure transducer is close to 1 ($0.7<\texttt{dytranCZ}<1.3$)
+    \item Either 0 or 1 bubbles are detected based on images from the camera
+    \item The number of bubbles approximated using the pressure transducer is close to 1
 \end{itemize}
@@
 It determines whether events fall within this area based on the visually calculated position $\chi$.
-    \item The pressure cut $PresWallCut(D)$ restricts the pressure detected by the pressure transducer, without any position corrections, to be close to 1 ($0.7<\texttt{dytranC}<1.3$).
+    \item The pressure cut $PresWallCut(D)$ restricts the pressure detected by the pressure transducer, without any position corrections, to be within a range of 0.3 of 1
     \item The acoustic cut $AcWallCut(D)$ is defined using the banded Fourier transform $\beta_{8}$, and takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
@@
 \begin{itemize}
-    \item Supervised learning, in which a system is trained on a set of fully-labeled data to classify unseen examples according to the patterns it observes in the training set.
+    \item Supervised learning, in which a system is trained on a set of fully-labeled data to classify unseen examples according to the patterns it observes in the training set, and
     \item Unsupervised learning, in which a system finds patterns or clusters in a set of unlabeled data.
@@
 The calibration sets available for training are not pure, so supervised learning is likely to overfit on biases and produce undesirable results.
-Conversely, unsupervised learning may be able to find clusters in the data, but it is impossible to guarantee it will distinguish between nuclear recoils and alpha particles as opposed to some other binary separation.
+Conversely, unsupervised learning may be able to find clusters in the data, but it is impossible to guarantee that it will distinguish between nuclear recoils and alpha particles as opposed to some other binary separation.
 
@@
 For each of the general configurations (multi-layer perceptron, one-dimensional convolutional neural network, et cetera) there are many possible specific architectures with different hyperparameters, which have to be optimized to improve performance.
-Some examples of hyperparameters are:
+Specific hyperparameters that were optimized include:
 
@@
     \item Number of neurons in dense layers
-    \item Number of filters, kernel size, and stride in convolutional layers
-    \item Types and amounts of regularization to apply (L2, dropout \cite{dropout})
+    \item Number of filters in convolutional layers (depth of output tensor)
+    \item Kernel size in convolutional layers (spatial area that a single filter covers)
+    \item Stride in convolutional layers (spacing between kernel positions during convolution)
+    \item Dropout regularization parameter (proportion of neurons to randomly remove for any given training example)
+    \item L2 regularization $\lambda$ (multiplier for squared weights before adding to loss function)
 \end{itemize}
@@
 
-Except where specifically mentioned, $WallCut(D)$ was not applied. This was done for 2 reasons:
+For some experimental runs, $WallCut(D)$ was not applied. This was done for 2 reasons:
 
@@
 
-All supervised learning configurations were trained and evaluated based on their ability to discriminate calibration runs (predominantly neutrons) from background runs (predominantly alpha particles).
-This was used instead of Acoustic Parameter for two reasons:
+All supervised learning configurations were trained and evaluated based on their ability to determine the origin of events: whether they are from calibration runs (predominantly neutrons) or background runs (predominantly alpha particles).
+This metric was used for training and testing instead of Acoustic Parameter for two reasons:
 
@@
     \includegraphics[width=\textwidth]{waveform_hist}
-    \caption{\label{waveform_hist} Prediction distribution of the best $DeepConvWithPos(\omega, \chi)$ discriminator without $WallCut(D)$.}
+    \caption{\label{waveform_hist} Prediction distribution of the best $DeepConvWithPos(\omega, \chi)$ discriminator without $WallCut(D)$.
+(Validation Event Count is the number of events in the validation set that fall within a certain network prediction band.)}
 \end{figure}
@@
 
-With $WallCut(D)$:
-
+\begin{center}
+    \captionof{table}{$DeepConvWithPos(\omega, \chi)$ with $WallCut(D)$.}
     \begin{tabular}{|l|l|l|l|l|l|l|}
@@
 
-Without $WallCut(D)$:
-
+    \captionof{table}{$DeepConvWithPos(\omega, \chi)$ without $WallCut(D)$.}
     \begin{tabular}{|l|l|l|l|l|l|l|}
@@
     \end{tabular}
+\end{center}
 
@@
 In addition to the Fourier transform used in the original PICO-60 analysis, higher resolutions were computed by similar means.
-These included $\beta _{5}$, $\beta _{10}$, $\beta _{20}$, and $\beta _{40}$.
+These included $\beta _{20}$ and $\beta _{40}$.
 Also, the full-resolution $\beta _{50,001}$ was input into a neural network directly, without any downscaling.
@@
 
-With $WallCut(D)$:
-
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Neural networks trained on Fourier transforms with $WallCut(D)$.} \label{fft_wall_cuts}
         \begin{tabular}{|l|l|l|l|l|l|}
@@
         \end{tabular}
+    \end{center}
+\end{minipage}
 
-It is very clear here that higher resolutions do not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
-
-Without $WallCut(D)$:
-
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{$FourierMLPWithPos(\beta{N})$ trained without $WallCut(D)$.}
         \begin{tabular}{|l|l|l|l|l|l|}
@@
         \end{tabular}
+    \end{center}
+\end{minipage}
+
+It is very clear in Table \ref{fft_wall_cuts} that higher resolutions do not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
 
@@
 
-The following table shows performance of various configurations of $ImageConv(\iota)$:
-
+\begin{center}
+    \captionof{table}{Performance of $ImageConv(\iota)$ configurations, trained without $WallCut(D)$.}
     \begin{tabular}{|l|l|l|l|l|l|}
@@
 
-With $WallCut(D)$:
-
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Summary of supervised learning configurations, trained with $WallCut(D)$.}
         \begin{tabular}{|l|l|l|}
@@
         \end{tabular}
+    \end{center}
+\end{minipage}
 
-Without $WallCut(D)$:
-
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Summary of supervised learning configurations, trained without $WallCut(D)$.}
         \begin{tabular}{|l|l|l|}
@@
         \end{tabular}
+    \end{center}
+\end{minipage}
 

### 2018-08-20 documentation/paper.tex
Renamed Acoustic Parameter to AP

@@
 
-The Acoustic Parameter method (abbreviated AP) is the technique that was used for event discrimination in the original PICO-60 analysis.
+The Acoustic Parameter method (furthermore AP) is the technique that was used for event discrimination in the original PICO-60 analysis.
 A function was defined that discriminates between alpha particles and nuclear recoils based on audio data.
@@
 
-Before the Acoustic Parameter can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$).
+Before AP can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$).
 It must also be preprocessed with the position correction function $PosCor(\beta_{8}, \chi)$, which adjusts the overall volume of an audio recording according to its position in the vessel.
@@
 
-Finally, a set of data cuts must be applied to remove certain classes of events which are not accurately classified by Acoustic Parameter. These classes include wall event and multiple-bubble events.
+Finally, a set of data cuts must be applied to remove certain classes of events which are not accurately classified using AP. These classes include wall event and multiple-bubble events.
 
@@
 
-Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to Acoustic Parameter and the original neural analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands.
+Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original neural analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands.
 This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.
@@
     \item Recording was triggered by the camera
-    \item Acoustic Parameter is not erroneously large and negative ($log_{10}(AP)>-100$)
+    \item AP is not erroneously large and negative ($log_{10}(AP)>-100$)
     \item Recorded more than 25 seconds after reaching target pressure
@@
 Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel.
-It can be desirable for a discriminator to handle these events correctly; however, Acoustic Parameter does not, and neither does the neural network used in the previous PICO-60 paper.
+It can be desirable for a discriminator to handle these events correctly; however, AP does not, and neither does the neural network used in the previous PICO-60 paper.
 Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.
@@
 
-One of the key benefits of machine learning systems is that they have the capacity to approximate arbitrary functions (such as Acoustic Parameter) without human intervention.
+One of the key benefits of machine learning systems is that they have the capacity to approximate arbitrary functions (such as AP) without human intervention.
 This greatly alleviates the need for human programmers to define and optimize the exact algorithm used.
@@
 When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
-While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, Acoustic Parameter has a significantly wider range.
-Normalization of the union prevents this from creating a bias where Acoustic Parameter would produce a higher standard deviation with a similarly proportioned error.
+While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, AP has a significantly wider range.
+Normalization of the union prevents this from creating a bias where AP would produce a higher standard deviation with a similarly proportioned error.
 
@@
 All supervised learning configurations were trained and evaluated based on their ability to determine the origin of events: whether they are from calibration runs (predominantly neutrons) or background runs (predominantly alpha particles).
-This metric was used for training and testing instead of Acoustic Parameter for two reasons:
+This metric was used for training and testing instead of AP for two reasons:
 
 \begin{enumerate}
-    \item In future experiments, when no Acoustic Parameter is available, impure data is the only information available for training. The network structure should be selected to accommodate this.
+    \item In future experiments, when no AP equivalent is available, impure data is the only information available for training. The network structure should be selected to accommodate this.
     \item In general, supervised learning systems learn to replicate biases present in the training data set.
-Thus, to evaluate the network's effectiveness at processing the input data, the run type (which is used for training) must be used as a baseline instead of Acoustic Parameter.
+Thus, to evaluate the network's effectiveness at processing the input data, the run type (which is used for training) must be used as a baseline instead of AP.
 \end{enumerate}
@@
 
-$WallCut(D)$ was applied to the low-resolution Fourier transforms during early experiments (since they were used for Acoustic Parameter) to minimize the number of variables between initial neural networks and a proven discriminator.
+$WallCut(D)$ was applied to the low-resolution Fourier transforms during early experiments (since they were used for AP) to minimize the number of variables between initial neural networks and a proven discriminator.
 
@@
     \item This was a highly effective configuration with wall cuts, producing 97\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
-    \item The use of the exact same input format as Acoustic Parameter makes the algorithms more directly comparable.
+    \item The use of the exact same input format as AP makes the algorithms more directly comparable.
 \end{enumerate}
@@
 Semi-supervised learning is an uncommon machine learning technique, relative to widely-used supervised and unsupervised learning.
-The concept is to train a machine learning model on a set of labeled (classified) data in addition to a set of unlabeled (unclasified) data.
+The concept is to train a machine learning model on a set of labeled (classified) data in addition to a set of unlabeled (unclassified) data.
 
@@
 
-Since the intent of these techniques is to separate particle types (the same as Acoustic Parameter), performance evaluations (accuracy and class-wise standard deviation) use Acoustic Parameter as a baseline.
+Since the intent of these techniques is to separate particle types (the same as AP), performance evaluations (accuracy and class-wise standard deviation) use AP as a baseline.
+Note that $WallCut(D)$ was once again applied for semi-supervised learning, because $FourierMLP(PosCor(\beta_{8}))$ was used for a neural network configuration, and because AP requires application of $WallCut(D)$ to be meaningfully comparable to any neural network.
 
@@
 The mean over that run (also the highest overall) was 96\%.
-Figure \ref{grav_grid_search} shows the highly accurate replication of Acoustic Parameter.
+Figure \ref{grav_grid_search} shows the highly accurate replication of AP.
 
@@
     \includegraphics[width=\textwidth]{grav_grid_search}
-    \caption{\label{grav_grid_search} Accuracy of best gravitational differentiation training run compared to Acoustic Parameter (98\% accuracy).}
+    \caption{\label{grav_grid_search} Accuracy of best gravitational differentiation training run compared to AP (98\% accuracy).}
 \end{figure}
@@
 
-During a grid search of these parameters (in addition to regularization parameters of the network), the best accuracy achieved for replication of Acoustic Parameter was exactly 100\%.
+During a grid search of these parameters (in addition to regularization parameters of the network), the best accuracy achieved for replication of AP was exactly 100\%.
 The mean of all epochs within that training run was 98\%.
 
-As seen in Figure \ref{icn_grid_search}, the network learned to replicate Acoustic Parameter very accurately.
-Interestingly, the examples with Acoustic Parameter values between the neutron and alpha particle ranges were also classified as such by the network.
-This may imply the indecisiveness on these examples is fundamental to the data rather than an implementation problem with Acoustic Parameter.
+As seen in Figure \ref{icn_grid_search}, the network learned to replicate AP very accurately.
+Interestingly, the examples with AP values between the neutron and alpha particle ranges were also classified as such by the network.
+This may imply the indecisiveness on these examples is fundamental to the data rather than an implementation problem with AP.
 
@@
     \includegraphics[width=\textwidth]{icn_grid_search}
-    \caption{\label{icn_grid_search} Accuracy of optimal iterative cluster nucleation training run compared to Acoustic Parameter (100\% accuracy).}
+    \caption{\label{icn_grid_search} Accuracy of optimal iterative cluster nucleation training run compared to AP (100\% accuracy).}
 \end{figure}
@@
 The following table directly compares performance of the semi-supervised learning systems, with the existing PICO-60 neural network provided as a baseline once again.
-Note that ``Max AP Accuracy'' refers to accuracy at replication of Acoustic Parameter, rather than replication of the ground truths.
+Note that ``Max AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.
 
@@
 
-This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like Acoustic Parameter.
-This conclusion has been achieved through two rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems and replicate Acoustic Parameter as accurately as possible.
+This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like the Acoustic Parameter technique.
+This conclusion has been achieved through two rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems and replicate AP as accurately as possible.
 
@@
 While both techniques produced some very effective discriminators, iterative cluster nucleation was found to be the more effective of the two.
-It replicated Acoustic Parameter with a mean of 98\% accuracy on the best configuration.
+It replicated AP with a mean of 98\% accuracy on the best configuration.
 This indicates that the technique shows significant promise, and may be worth consideration as a discrimination technique for future experiments.

### 2018-08-21 documentation/paper.tex
Minor edits throughout; added paragraph at end

@@
     The identification of non-signal events is a major hurdle to overcome for bubble chamber dark matter experiments such as PICO-60.
-The current practice of manually developing a discriminator function to eliminate background events is difficult when calibration data is frequently impure and present only in small quantities.
+The current practice of manually developing a discriminator function to eliminate background events is difficult when available calibration data is frequently impure and present only in small quantities.
 In this study, several different input formats and neural network architectures are applied to the task.
 First, they are optimized in a supervised learning context.
-Next, novel semi-supervised learning algorithms are trained, and found to replicate the Acoustic Parameter discriminator previously used in PICO-60 with up to 100\% accuracy.
+Next, two novel semi-supervised learning algorithms are trained, and found to replicate the Acoustic Parameter discriminator previously used in PICO-60 with up to 100\% accuracy.
 \end{abstract}
@@
 
-    \item Determine whether these techniques can be extended, applying novel algorithms based on semi-supervised learning, to develop an effective discriminator function based on incomplete and inaccurate initial information.
+    \item Determine whether these techniques can be extended, applying two original algorithms based on semi-supervised learning, to develop an effective discriminator function based on incomplete and inaccurate initial information.
 
-    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only calibration data for training.
-While calibration runs are known to be mostly correlated with particle types, the relatively large percentage of impurities has the potential to hinder conventional learning techniques.
+    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only run types as training data.
+While run types are known to be mostly correlated with particle types, the relatively large percentage of impurities has the potential to hinder conventional learning techniques.
+Specifically, the run types available for training are:
+
+    \begin{enumerate}
+        \item Neutron calibration sets, which are known to contain approximately 90\% neutrons, and
+        \item Background radiation runs, which contain approximately 99\% alpha particles
+    \end{enumerate}
 \end{enumerate}
@@
 A significant challenge in this application is that neither supervised nor unsupervised learning is ideal.
-The calibration sets available for training are not pure, so supervised learning is likely to overfit on biases and produce undesirable results.
+The calibration and background runs available for training are not pure, so supervised learning is likely to overfit on biases and produce undesirable results.
 Conversely, unsupervised learning may be able to find clusters in the data, but it is impossible to guarantee that it will distinguish between nuclear recoils and alpha particles as opposed to some other binary separation.
@@
 
-Two semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
+Two original semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
 Both are discussed in detail in Section \ref{semi_supervised}.
@@
 Numerically, this corresponds to the number of validation examples the network classifies correctly.
+The baseline for correctness is the run type during supervised learning experiments, and the AP prediction for semi-supervised experiments.
 
@@
 
-Numerically, this is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
+Numerically, this is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively (or calibration and background sets respectively, depending on which is used as ground truth data).
 It calculates the spread of the network's predictions for each ground truth class.
@@
 
-The first step is to calculate the standard deviation S of the union of the two sets.
+The first step is to calculate the standard deviation $S$ of the union of $N$ and $A$.
 This gives an indication of the scale of the overall distribution.
@@
 The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually.
-This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are.
+This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are for each class.
 Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off those specific values, will produce a low class-wise standard deviation.
@@
 
-All supervised learning configurations were trained and evaluated based on their ability to determine the origin of events: whether they are from calibration runs (predominantly neutrons) or background runs (predominantly alpha particles).
+All supervised learning configurations were trained and evaluated based on their ability to determine the origin of events: whether they are from neutron calibration source runs (predominantly neutrons) or background radiation runs (predominantly alpha particles).
 This metric was used for training and testing instead of AP for two reasons:
@@
 During early tests with variations on $DeepConv(\omega)$ without $WallCut(D)$, it was revealed to have a surprising and undesirable capacity to separate run types based on meaningless information.
-It was able to separate the AmBe calibration run from combined background runs with 96\% accuracy, and to separate all calibration runs from background runs with 91\% accuracy.
+It was able to discriminate between events from the AmBe neutron calibration runs and events from alpha background runs with 96\% accuracy, and to separate all combined neutron calibration runs from alpha background runs with 91\% accuracy.
 This was determined to be caused by the network detecting and fitting on biases in the white noise at the beginning and hydraulic recompression sounds at the end of the raw waveform.
@@
 
-In the context of PICO-60, the labeled sets consist of:
-
-\begin{enumerate}
-    \item Single-bubble events from the calibration sets, which are known to contain approximately 90\% neutrons
-    \item Single-bubble events from the background radiation data sets, which contain approximately 99\% alpha particles
-\end{enumerate}
-
-In general, these labeled sets should have a strong but not perfect correlation with particle types. The unlabeled data can consist of any mixture of particle types.
+In the context of PICO-60, the labeled sets consist of the neutron calibration runs, and the background radiation runs (which consist predominantly of alpha particles).
+In general, these labeled sets should have a strong but not perfect correlation with particle types.
+The unlabeled data can consist of any mixture of particle types.
 
@@
 
-Iterative cluster nucleation is a second semi-supervised learning algorithm developed as part of this study.
+Iterative cluster nucleation is a second original semi-supervised learning algorithm developed as part of this study.
 It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
@@
 This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like the Acoustic Parameter technique.
-This conclusion has been achieved through two rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems and replicate AP as accurately as possible.
+This conclusion has been achieved through two rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems with the goal of replicating AP as accurately as possible.
 
@@
 It replicated AP with a mean of 98\% accuracy on the best configuration.
-This indicates that the technique shows significant promise, and may be worth consideration as a discrimination technique for future experiments.
+This indicates that the technique shows significant promise.
 
-\section{Software}
+While there is still of course a place for conventionally developed discriminators such as AP, they are difficult to develop until a large quantity of calibration data has been collected and the physical principles of the experiment are understood well.
+The innovations developed in this study suggest that semi-supervised learning can facilitate very quick development of discriminators immediately after the first calibration runs are conducted on a new experimental apparatus.
+This has the potential to allow the teams working on these experiments to iterate more quickly and better understand intermediate results, before all data has been analyzed in depth.
+
+\section{Technologies}
 

### 2018-08-25 visualization/graph_binary_predictions.py
Print precision and recall individually

@@ -39,11 +39,20 @@ for criterion_data, criterion_name, classification_threshold in zip(
     [0.5, 0.25, 0.5]
 ):
     # Convert the raw outputs to binary classifications, based on whether they are greater than or equal to the threshold
-    classifications = [data_point >= classification_threshold for data_point in criterion_data]
+    classifications = np.array([data_point >= classification_threshold for data_point in criterion_data])
     # Get the number of these classifications that agree with the ground truths
-    agreements = np.count_nonzero(np.array(classifications) == ground_truths)
+    agreements = np.count_nonzero(classifications == ground_truths)
     # Print out the resulting accuracy statistic to the user
     print(f'Accuracy of {agreements / 128} for {criterion_name}')
+    # Repeat this process for precision (erroneous recoil predictions) and recall (erroneous alpha predictions) individually
+    # First, count the number of events predicted as recoils, and then divide the number that are incorrectly predicted to be recoils by that
+    predicted_recoils = np.count_nonzero(classifications == 0)
+    erroneous_recoils = np.count_nonzero(np.logical_and(classifications == 0, ground_truths == 1))
+    print(f'Precision of {(predicted_recoils - erroneous_recoils) / predicted_recoils} for {criterion_name}')
+    # Repeat for the number of events falsely predicted as alphas, which are actually nuclear recoils (possibly representing WIMP candidates in real data)
+    predicted_alphas = np.count_nonzero(classifications == 1)
+    erroneous_alphas = np.count_nonzero(np.logical_and(classifications == 1, ground_truths == 0))
+    print(f'Recall of {(predicted_alphas - erroneous_alphas) / predicted_alphas} for {criterion_name}')
     # Divide all of these data points by their overall standard deviation, to normalize their range
     criterion_data /= np.std(criterion_data)
     # Iterate over both possible values of the ground truth, and corresponding names

### 2018-08-25 visualization/graph_binary_predictions.py
Revert "Reverted visualization script to use AP for ground truths"
This reverts commit 52418a32d213d86fe0f769e84bba735560837a97.

@@ -16,8 +16,8 @@ from utilities.verify_arguments import verify_arguments
 # Verify that a path to the JSON data file is passed
 verify_arguments('JSON data file')
 
-# Load the data set from the file, ignoring the run type ground truths
-events, _, network_outputs = load_test(sys.argv[1])
+# Load the data set from the file
+events, ground_truths, network_outputs = load_test(sys.argv[1])
 # Get the acoustic parameter and neural network score data from the events, if they are present (if one is, both will be)
 if hasattr(events[0], 'logarithmic_acoustic_parameter'):
     acoustic_parameters, original_neural_network_scores = zip(
@@ -29,8 +29,8 @@ else:
     random_values = [random.uniform(0, 1) for _ in range(len(network_outputs))]
     acoustic_parameters = random_values
     original_neural_network_scores = random_values
-# Calculate actual neutron/alpha ground truths based on AP
-ground_truths = np.array(acoustic_parameters) > 0.25
+# # Calculate actual neutron/alpha ground truths based on AP
+# ground_truths = np.array(acoustic_parameters) > 0.25
 
 # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
 for criterion_data, criterion_name, classification_threshold in zip(

### 2018-08-25 visualization/graph_binary_predictions.py
Print mean class-wise standard deviation

@@ -55,6 +55,8 @@ for criterion_data, criterion_name, classification_threshold in zip(
     print(f'Recall of {(predicted_alphas - erroneous_alphas) / predicted_alphas} for {criterion_name}')
     # Divide all of these data points by their overall standard deviation, to normalize their range
     criterion_data /= np.std(criterion_data)
+    # Create a list for the 2 standard deviations
+    standard_deviations = []
     # Iterate over both possible values of the ground truth, and corresponding names
     for ground_truth_value, ground_truth_name in zip([True, False], ['alpha particles', 'neutrons']):
         # Sort out only the data points for this criterion which correspond to ground truths of this value
@@ -70,6 +72,11 @@ for criterion_data, criterion_name, classification_threshold in zip(
             f'for {ground_truth_name}',
             f'is {standard_deviation}'
         )
+        # Add it to the list so the mean can be calculated
+        standard_deviations.append(standard_deviation)
+    # Calculate and print the mean class-wise standard deviation
+    print(f'Mean class-wise standard deviation for {criterion_name} is {np.mean(standard_deviations)}')
+
 
 # Convert the binary ground truth values to colors (red and blue) for graphing
 point_colors = [

### 2018-08-25 visualization/graph_binary_predictions.py
Clarified/corrected recall calculation (I think it's equivalent, but clearer anyway)

@@ -49,10 +49,10 @@ for criterion_data, criterion_name, classification_threshold in zip(
     predicted_recoils = np.count_nonzero(classifications == 0)
     erroneous_recoils = np.count_nonzero(np.logical_and(classifications == 0, ground_truths == 1))
     print(f'Precision of {(predicted_recoils - erroneous_recoils) / predicted_recoils} for {criterion_name}')
-    # Repeat for the number of events falsely predicted as alphas, which are actually nuclear recoils (possibly representing WIMP candidates in real data)
-    predicted_alphas = np.count_nonzero(classifications == 1)
+    # Repeat for the proportion of neutron calibration events that are correctly predicted as recoils; this represents not the purity, but the number missed
+    actual_recoils = np.count_nonzero(classifications == 1)
     erroneous_alphas = np.count_nonzero(np.logical_and(classifications == 1, ground_truths == 0))
-    print(f'Recall of {(predicted_alphas - erroneous_alphas) / predicted_alphas} for {criterion_name}')
+    print(f'Recall of {(actual_recoils - erroneous_alphas) / actual_recoils} for {criterion_name}')
     # Divide all of these data points by their overall standard deviation, to normalize their range
     criterion_data /= np.std(criterion_data)
     # Create a list for the 2 standard deviations

### 2018-08-25 documentation/paper.tex
Explained precision/recall and completed one table accordingly

@@
 
-Class-wise standard deviation captures how decisive the network is: whether its prediction is confidently high or low, or is only slightly closer to one edge of the spectrum.
+The correctness of the network's predictions is further detailed in the precision and recall statistics.
+Precision represents the proportion, of those that the network predicts to be nuclear recoils, that are actually nuclear recoils (or part of the corresponding calibration set).
+It is an indicator of how pure the network's predictions are for the recoil class.
+Recall represents the proportion, of those that are actually nuclear recoils, that the network predicts to be nuclear recoils.
+It indicates how many examples the network misses (i.e., those that would erroneously be removed as background radiation).
 
-Numerically, this is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively (or calibration and background sets respectively, depending on which is used as ground truth data).
+Class-wise standard deviation, abbreviated CWSD, captures how decisive the network is: whether its prediction is confidently high or low, or is only slightly closer to one edge of the spectrum.
+
+Numerically, this is a variable defined below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively (or calibration and background sets respectively, depending on which is used as ground truth data).
 It calculates the spread of the network's predictions for each ground truth class.
@@
 
-$C=(std(N \div S) + std(A \div S)) \div 2$
+$CWSD=(std(N \div S) + std(A \div S)) \div 2$
 
@@
     \captionof{table}{$DeepConvWithPos(\omega, \chi)$ with $WallCut(D)$.}
-    \begin{tabular}{|l|l|l|l|l|l|l|}
+    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
         \hline
-        L2 $\lambda$ & Dropout & Filters & Kernel Size & \# Layers Multiplier & Max Accuracy & Class-Wise Std Dev \\
+        L2 $\lambda$ & Dropout & Filters & Kernel & Layers $\times$ & Max Accuracy & Precision & Recall & CWSD \\
         \hline
-        0.003 & 0 & 24 & 3 & 3 & 95\% & 0.44 \\
+        0.003 & 0 & 24 & 3 & 3 & 95\% & 99\% & 85\% & 0.44 \\
+        % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1533824431_epoch30.json
         \hline
-        0.003 & 0 & 48 & 5 & 3 & 96\% & 0.41 \\
+        0.003 & 0 & 48 & 5 & 3 & 96\% & 99\% & 89\% & 0.41 \\
+        % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1533825716_epoch43.json
         \hline
-        0.003 & 0.5 & 24 & 5 & 3 & 93\% & 0.50 \\
+        0.003 & 0.5 & 24 & 5 & 3 & 93\% & 98\% & 83\% & 0.50 \\
+        % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group3time1533828588_epoch27.json
         \hline
-        0.001 & 0 & 48 & 5 & 3 & 95\% & 0.48 \\
+        0.001 & 0 & 48 & 5 & 3 & 95\% & 98\% & 87\% & 0.48 \\
+        % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.001_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1533832229_epoch16.json
         \hline
-        0.001 & 0.25 & 48 & 3 & 6 & 93\% & 0.62 \\
+        0.001 & 0.25 & 48 & 3 & 6 & 93\% & 94\% & 91\% & 0.62 \\
+        % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.001_dense_dropout0.25_first_layer_filters48_kernel_size3_convolutional_layers_per_group6time1533834812_epoch30.json
         \hline
-        0.0003 & 0.5 & 24 & 5 & 6 & 95\% & 0.49 \\
+        0.0003 & 0.5 & 24 & 5 & 6 & 95\% & 97\% & 91\% & 0.49 \\
+        % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.0003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group6time1533849077_epoch45.json
         \hline

### 2018-08-25 visualization/graph_binary_predictions.py
Fixed dumb mistake with recall calculation

@@ -50,7 +50,7 @@ for criterion_data, criterion_name, classification_threshold in zip(
     erroneous_recoils = np.count_nonzero(np.logical_and(classifications == 0, ground_truths == 1))
     print(f'Precision of {(predicted_recoils - erroneous_recoils) / predicted_recoils} for {criterion_name}')
     # Repeat for the proportion of neutron calibration events that are correctly predicted as recoils; this represents not the purity, but the number missed
-    actual_recoils = np.count_nonzero(classifications == 1)
+    actual_recoils = np.count_nonzero(ground_truths == 0)
     erroneous_alphas = np.count_nonzero(np.logical_and(classifications == 1, ground_truths == 0))
     print(f'Recall of {(actual_recoils - erroneous_alphas) / actual_recoils} for {criterion_name}')
     # Divide all of these data points by their overall standard deviation, to normalize their range

### 2018-08-25 documentation/paper.tex
Updated all supervised learning tables except summary

@@
 
+\begin{minipage}{\textwidth}
     \begin{center}
@@
             \hline
-        0.003 & 0 & 24 & 3 & 3 & 95\% & 99\% & 85\% & 0.44 \\
+            0.003 & 0 & 24 & 3 & 3 & 95\% & 99\% & 94\% & 0.44 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1533824431_epoch30.json
             \hline
-        0.003 & 0 & 48 & 5 & 3 & 96\% & 99\% & 89\% & 0.41 \\
+            0.003 & 0 & 48 & 5 & 3 & 96\% & 99\% & 96\% & 0.41 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1533825716_epoch43.json
             \hline
-        0.003 & 0.5 & 24 & 5 & 3 & 93\% & 98\% & 83\% & 0.50 \\
+            0.003 & 0.5 & 24 & 5 & 3 & 93\% & 98\% & 92\% & 0.50 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group3time1533828588_epoch27.json
             \hline
-        0.001 & 0 & 48 & 5 & 3 & 95\% & 98\% & 87\% & 0.48 \\
+            0.001 & 0 & 48 & 5 & 3 & 95\% & 98\% & 95\% & 0.48 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.001_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1533832229_epoch16.json
             \hline
-        0.001 & 0.25 & 48 & 3 & 6 & 93\% & 94\% & 91\% & 0.62 \\
+            0.001 & 0.25 & 48 & 3 & 6 & 93\% & 94\% & 97\% & 0.62 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.001_dense_dropout0.25_first_layer_filters48_kernel_size3_convolutional_layers_per_group6time1533834812_epoch30.json
             \hline
-        0.0003 & 0.5 & 24 & 5 & 6 & 95\% & 97\% & 91\% & 0.49 \\
+            0.0003 & 0.5 & 24 & 5 & 6 & 95\% & 97\% & 97\% & 0.49 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.0003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group6time1533849077_epoch45.json
@@
         \end{tabular}
+    \end{center}
+\end{minipage}
 
+\begin{minipage}{\textwidth}
+    \begin{center}
         \captionof{table}{$DeepConvWithPos(\omega, \chi)$ without $WallCut(D)$.}
-    \begin{tabular}{|l|l|l|l|l|l|l|}
+        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
             \hline
-        L2 $\lambda$ & Dropout & Filters & Kernel Size & \# Layers Multiplier & Max Accuracy & Class-Wise Std Dev \\
+            L2 $\lambda$ & Dropout & Filters & Kernel & Layers $\times$ & Max Accuracy & Precision & Recall & CWSD \\
             \hline
-        0.003 & 0 & 24 & 3 & 3 & 91\% & 0.61 \\
+            0.003 & 0 & 24 & 3 & 3 & 91\% & 95\% & 78\% & 0.61 \\
+            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1532800009_epoch78.json
             \hline
-        0.003 & 0.5 & 24 & 5 & 3 & 91\% & 0.65 \\
+            0.003 & 0.5 & 24 & 5 & 3 & 91\% & 95\% & 78\% & 0.65 \\
+            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group3time1532813246_epoch49.json
             \hline
-        0.001 & 0.25 & 24 & 3 & 3 & 91\% & 0.66 \\
+            0.001 & 0.25 & 24 & 3 & 3 & 91\% & 88\% & 84\% & 0.66 \\
+            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.001_dense_dropout0.25_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1532825383_epoch88.json
             \hline
-        0.001 & 0.5 & 48 & 5 & 3 & 88\% & 0.73 \\
+            0.001 & 0.5 & 48 & 5 & 3 & 88\% & 92\% & 73\% & 0.73 \\
+            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.001_dense_dropout0.5_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1532837124_epoch0.json
             \hline
-        0.0003 & 0 & 48 & 5 & 6 & 78\% & 0.95 \\
+            0.0003 & 0 & 48 & 5 & 6 & 78\% & 76\% & 56\% & 0.95 \\
+            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.0003_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group6time1532846835_epoch63.json
             \hline
-        0.0003 & 0.5 & 48 & 5 & 6 & 74\% & 0.92 \\
+            0.0003 & 0.5 & 48 & 5 & 6 & 74\% & 67\% & 53\% & 0.92 \\
+            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.0003_dense_dropout0.5_first_layer_filters48_kernel_size5_convolutional_layers_per_group6time1532863869_epoch10.json
             \hline
@@
     \end{center}
+\end{minipage}
 
@@
 Dropout seems to have a weak correlation, possibly reducing accuracy slightly.
+
+It is notable that recall is much, much poorer when wall cuts are applied.
+It seems there are more neutron events that show alpha-like audio signatures in the events near the walls, possibly due to resonance.
 
@@
 
-Very interestingly, $FourierMLP(\beta_{8})$ performed better than either $FourierMLP(PosCor(\beta_{8}), \chi)$ or \\ $FourierMLPWithPos(\beta{8}, \chi)$, which produced accuracy values of 97\% and 95\%, respectively.
+Very interestingly, $FourierMLP(\beta_{8})$ performed better than either $FourierMLP(PosCor(\beta_{8}), \chi)$ or \\ $FourierMLPWithPos(\beta{8}, \chi)$, which produced accuracy values of 96\% and 97\%, respectively.
 
@@
         \captionof{table}{Neural networks trained on Fourier transforms with $WallCut(D)$.} \label{fft_wall_cuts}
-        \begin{tabular}{|l|l|l|l|l|l|}
+        \begin{tabular}{|l|l|l|l|l|l|l|l|}
             \hline
-            Configuration & L2 $\lambda$ & Dropout & Layers & Max Accuracy & Class-Wise Std Dev \\
+            Configuration & L2 $\lambda$ & Dropout & Layers & Max Acc & Prec & Recall & CWSD \\
             \hline
-            $FourierMLP(PosCor(\beta_{8}))$ & 0 & 0.5 & 2 & 95\% & 0.56 \\
+            $FourierMLP(PosCor(\beta_{8}))$ & 0 & 0.5 & 2 & 96\% & 98\% & 97\% & 0.42 \\
+            % ../experimental_data/validation_results/2018_07_16_banded_frequency_validation_passing_weighted_3_times.json
             \hline
-            $FourierMLP(\beta_{8})$ & 0 & 0.5 & 2 & 98\% & 0.29 \\
+            $FourierMLP(\beta_{8})$ & 0 & 0.5 & 2 & 98\% & 100\% & 98\% & 0.29 \\
+            % ../experimental_data/validation_results/2018_07_17_banded_frequency_not_position_corrected/banded_low_resolution_time1531831056_epoch249.json
             \hline
-            $FourierMLPWithPos(\beta_{8}, \chi)$ & 0 & 0.5 & 3 & 95\% & 0.44 \\
+            $FourierMLPWithPos(\beta_{8}, \chi)$ & 0 & 0.5 & 3 & 95\% & 100\% & 92\% & 0.44 \\
+            % ../experimental_data/validation_results/2018_07_17_banded_frequency_not_position_corrected_with_position_input/banded_low_resolution_time1531837503_epoch18.json
             \hline
-            $FourierMLPWithPos(\beta_{8}, \chi)$ & 0 & 0.25 & 3 & 97\% & 0.43 \\
+            $FourierMLPWithPos(\beta_{8}, \chi)$ & 0 & 0.25 & 3 & 97\% & 98\% & 98\% & 0.43 \\
+            % ../experimental_data/validation_results/2018_07_18_banded_frequency_25\%_dropout/banded_low_resolution_time1531922044_epoch154.json
             \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0.003 & 0.25 & 3 & 93\% & 0.54 \\
+            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0.003 & 0.25 & 3 & 93\% & 99\% & 92\% & 0.54 \\
+            % ../experimental_data/validation_results/2018_07_20_20_bands/high_resolution_frequency_time1532105605_epoch90.json
             \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0 & 0 & 3 & 94\% & 0.52 \\
+            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0 & 0 & 3 & 94\% & 99\% & 93\% & 0.52 \\
+            % ../experimental_data/validation_results/2018_07_20_20_bands_no_regularization/high_resolution_frequency_time1532113464_epoch172.json
             \hline
-            $FourierMLPWithPos(\beta_{40}, \chi)$ & 0 & 0 & 3 & 91\% & 0.57 \\
+            $FourierMLPWithPos(\beta_{40}, \chi)$ & 0 & 0 & 3 & 91\% & 98\% & 90\% & 0.57 \\
+            % ../experimental_data/validation_results/2018_07_20_40_bands/high_resolution_frequency_time1532107386_epoch101.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.0003 & 0 & 4 & 91\% & 0.53 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.0003 & 0 & 4 & 91\% & 94\% & 93\% & 0.53 \\
+            % ../experimental_data/validation_results/2018_08_10_high_resolution_grid_search_with_wall_cuts/high_res_grid_search_dropout0_l2_lambda0.0003_hidden_layers3_time1533922958_epoch133.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0.25 & 4 & 91\% & 0.56 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0.25 & 4 & 91\% & 96\% & 89\% & 0.56 \\
+            % ../experimental_data/validation_results/2018_08_10_high_resolution_grid_search_with_wall_cuts/high_res_grid_search_dropout0.25_l2_lambda0.001_hidden_layers3_time1533924257_epoch115.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0.5 & 3 & 89\% & 0.64 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0.5 & 3 & 89\% & 91\% & 91\% & 0.64 \\
+            % ../experimental_data/validation_results/2018_08_10_high_resolution_grid_search_with_wall_cuts/high_res_grid_search_dropout0.5_l2_lambda0.001_hidden_layers2_time1533925299_epoch169.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.5 & 4 & 89\% & 0.65 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.5 & 4 & 89\% & 90\% & 93\% & 0.65 \\
+            % ../experimental_data/validation_results/2018_08_10_high_resolution_grid_search_with_wall_cuts/high_res_grid_search_dropout0.5_l2_lambda0.01_hidden_layers3_time1533925799_epoch91.json
             \hline
@@
         \captionof{table}{$FourierMLPWithPos(\beta{N})$ trained without $WallCut(D)$.}
-        \begin{tabular}{|l|l|l|l|l|l|}
+        \begin{tabular}{|l|l|l|l|l|l|l|l|}
             \hline
-            Configuration & L2 $\lambda$ & Dropout & Layers & Max Accuracy & Class-Wise Std Dev \\
+            Configuration & L2 $\lambda$ & Dropout & Layers & Max Acc & Prec & Recall & CWSD \\
             \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0 & 0 & 3 & 92\% & 0.63 \\
+            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0 & 0 & 3 & 92\% & 93\% & 77\% & 0.63 \\
+            % ../experimental_data/validation_results/2018_07_20_no_wall_cuts/high_resolution_frequency_time1532116206_epoch111.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0 & 0 & 3 & 95\% & 0.49 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0 & 0 & 3 & 95\% & 89\% & 95\% & 0.49 \\
+            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search/high_res_grid_search_dropout0_l2_lambda0_hidden_layers2_time1532446602_epoch210.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0 & 2 & 96\% & 0.47 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0 & 2 & 96\% & 95\% & 93\% & 0.47 \\
+            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search/high_res_grid_search_dropout0_l2_lambda0.001_hidden_layers1_time1532448014_epoch149.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0 & 2 & 97\% & 0.47 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0 & 2 & 97\% & 93\% & 98\% & 0.47 \\
+            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search/high_res_grid_search_dropout0_l2_lambda0.01_hidden_layers1_time1532449765_epoch108.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.25 & 2 & 95\% & 0.54 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.25 & 2 & 95\% & 97\% & 86\% & 0.54 \\
+            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search/high_res_grid_search_dropout0.25_l2_lambda0.01_hidden_layers1_time1532454347_epoch122.json
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.5 & 2 & 94\% & 0.53 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.5 & 2 & 94\% & 95\% & 86\% & 0.53 \\
+            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search//high_res_grid_search_dropout0.5_l2_lambda0.01_hidden_layers1_time1532459365_epoch243.json
             \hline
@@
 Here, L2 regularization is beneficial, possibly because there is more noise that can be overfit on.
+
+Unlike with the raw waveform, recall varies relatively little when wall cuts are applied.
 
@@
 
+\begin{minipage}{\textwidth}
     \begin{center}
         \captionof{table}{Performance of $ImageConv(\iota)$ configurations, trained without $WallCut(D)$.}
-    \begin{tabular}{|l|l|l|l|l|l|}
+        \begin{tabular}{|l|l|l|l|l|l|l|l|}
             \hline
-        L2 $\lambda$ & Dropout & Convolutional Layers & Dense Layers & Max Accuracy & Class-Wise Std Dev \\
+            L2 $\lambda$ & Dropout & Convolutional Layers & Dense Layers & Max Accuracy & Precision & Recall & CWSD \\
             \hline
-        0 & 0.5 & 4 & 3 & 77\% & 1.01 \\
+            0 & 0.5 & 4 & 3 & 77\% & 52\% & 31\% & 1.01 \\
+            % ../experimental_data/validation_results/2018_07_26_image_classification_fixed_loading/image_time1532640532_epoch8.json
             \hline
-        0 & 0.25 & 4 & 3 & 85\% & 0.95 \\
+            0 & 0.25 & 4 & 3 & 85\% & 40\% & 38\% & 0.95 \\
+            % ../experimental_data/validation_results/2018_07_26_image_classification_less_dropout_more_epochs/image_time1532615758_epoch164.json
             \hline
-        0.01 & 0.25 & 4 & 3 & 66\% & 0.97 \\
+            0.01 & 0.25 & 6 & 3 & 68\% & 68\% & 100\% & 0.98 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group2time1532895825_epoch2.json
             \hline
-        0.003 & 0.25 & 4 & 3 & 69\% & 0.97 \\
+            0.003 & 0.25 & 9 & 3 & 68\% & 68\% & 100\% & 0.98 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.003_dropout0.25_convolutional_layers_per_group3time1532895118_epoch3.json
             \hline
-        0.01 & 0.25 & 12 & 3 & 66\% & 0.98 \\
+            0.01 & 0.25 & 12 & 3 & 68\% & 68\% & 100\% & 0.97 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group4time1532896126_epoch0.json
             \hline
-        0.01 & 0.5 & 12 & 3 & 68\% & 0.99 \\
+            0.01 & 0.5 & 12 & 3 & 68\% & 68\% & 100\% & 0.99 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.5_convolutional_layers_per_group4time1532898944_epoch1.json
             \hline
@@
     \end{center}
+\end{minipage}
 
 The high standard deviations across the board are further evidence for the hypothesis that there is insufficient information in the image data.
+
+The 100\% recall statistics are not due to high classification accuracy, but rather due to the network making predictions less than 0.5 for all examples of both classes.
 

### 2018-08-26 documentation/paper.tex
Finished filling out supervised learning tables

@@
 Once the audio was cropped properly, $DeepConv(\omega)$ achieved a maximum of 77\% accuracy on validation data.
+Errors were not balanced at all; precision was 76\% and recall was 94\%, indicating that the network was overzealous to classify particles as neutrons.
 Using the same network architecture (except for the input layer), and still without $WallCut(D)$, $DeepConvWithPos(\omega, \chi)$ achieved 85\%.
-The maximum obtained on a grid search with this configuration was 91\% (although with a relatively high class-wise standard deviation of 0.61).
+The maximum obtained on a grid search with this configuration was 91\%, although with a relatively high class-wise standard deviation of 0.61.
+Precision improved greatly, reaching 95\%, while recall was reduced, at a maximum of 84\%.
 While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
@@
             \hline
-            L2 $\lambda$ & Dropout & Filters & Kernel & Layers $\times$ & Max Accuracy & Precision & Recall & CWSD \\
+            L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Max Accuracy & Precision & Recall & CWSD \\
             \hline
-            0.003 & 0 & 24 & 3 & 3 & 95\% & 99\% & 94\% & 0.44 \\
+            0.003 & 0 & 24 & 3 & 17 & 95\% & 99\% & 94\% & 0.44 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1533824431_epoch30.json
             \hline
-            0.003 & 0 & 48 & 5 & 3 & 96\% & 99\% & 96\% & 0.41 \\
+            0.003 & 0 & 48 & 5 & 17 & 96\% & 99\% & 96\% & 0.41 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1533825716_epoch43.json
             \hline
-            0.003 & 0.5 & 24 & 5 & 3 & 93\% & 98\% & 92\% & 0.50 \\
+            0.003 & 0.5 & 24 & 5 & 17 & 93\% & 98\% & 92\% & 0.50 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group3time1533828588_epoch27.json
             \hline
-            0.001 & 0 & 48 & 5 & 3 & 95\% & 98\% & 95\% & 0.48 \\
+            0.001 & 0 & 48 & 5 & 17 & 95\% & 98\% & 95\% & 0.48 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.001_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1533832229_epoch16.json
             \hline
-            0.001 & 0.25 & 48 & 3 & 6 & 93\% & 94\% & 97\% & 0.62 \\
+            0.001 & 0.25 & 48 & 3 & 32 & 93\% & 94\% & 97\% & 0.62 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.001_dense_dropout0.25_first_layer_filters48_kernel_size3_convolutional_layers_per_group6time1533834812_epoch30.json
             \hline
-            0.0003 & 0.5 & 24 & 5 & 6 & 95\% & 97\% & 97\% & 0.49 \\
+            0.0003 & 0.5 & 24 & 5 & 32 & 95\% & 97\% & 97\% & 0.49 \\
             % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.0003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group6time1533849077_epoch45.json
@@
             \hline
-            L2 $\lambda$ & Dropout & Filters & Kernel & Layers $\times$ & Max Accuracy & Precision & Recall & CWSD \\
+            L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Max Accuracy & Precision & Recall & CWSD \\
             \hline
-            0.003 & 0 & 24 & 3 & 3 & 91\% & 95\% & 78\% & 0.61 \\
+            0.003 & 0 & 24 & 3 & 17 & 91\% & 95\% & 78\% & 0.61 \\
             % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1532800009_epoch78.json
             \hline
-            0.003 & 0.5 & 24 & 5 & 3 & 91\% & 95\% & 78\% & 0.65 \\
+            0.003 & 0.5 & 24 & 5 & 17 & 91\% & 95\% & 78\% & 0.65 \\
             % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group3time1532813246_epoch49.json
             \hline
-            0.001 & 0.25 & 24 & 3 & 3 & 91\% & 88\% & 84\% & 0.66 \\
+            0.001 & 0.25 & 24 & 3 & 17 & 91\% & 88\% & 84\% & 0.66 \\
             % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.001_dense_dropout0.25_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1532825383_epoch88.json
             \hline
-            0.001 & 0.5 & 48 & 5 & 3 & 88\% & 92\% & 73\% & 0.73 \\
+            0.001 & 0.5 & 48 & 5 & 17 & 88\% & 92\% & 73\% & 0.73 \\
             % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.001_dense_dropout0.5_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1532837124_epoch0.json
             \hline
-            0.0003 & 0 & 48 & 5 & 6 & 78\% & 76\% & 56\% & 0.95 \\
+            0.0003 & 0 & 48 & 5 & 32 & 78\% & 76\% & 56\% & 0.95 \\
             % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.0003_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group6time1532846835_epoch63.json
             \hline
-            0.0003 & 0.5 & 48 & 5 & 6 & 74\% & 67\% & 53\% & 0.92 \\
+            0.0003 & 0.5 & 48 & 5 & 32 & 74\% & 67\% & 53\% & 0.92 \\
             % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.0003_dense_dropout0.5_first_layer_filters48_kernel_size5_convolutional_layers_per_group6time1532863869_epoch10.json
@@
         \captionof{table}{Summary of supervised learning configurations, trained with $WallCut(D)$.}
-        \begin{tabular}{|l|l|l|}
+        \begin{tabular}{|l|l|l|l|l|}
             \hline
-            Configuration & Max Accuracy & Mean Class-Wise Standard Deviation \\
+            Configuration & Max Accuracy & Precision & Recall & CWSD \\
             \hline
-            $FourierMLP(PosCor(\beta_{8}), \chi)$ & 97\% & 0.42 \\
+            $DeepConvWithPos(\omega, \chi)$ & 96\% & 99\% & 96\% & 0.41 \\
             \hline
-            $FourierMLP(\beta_{8})$ & 98\% & 0.29 \\
+            $FourierMLP(PosCor(\beta_{8}))$ & 96\% & 98\% & 97\% & 0.42 \\
             \hline
-            $FourierMLPWithPos(\beta_{8}, \chi)$ & 95\% & 0.44 \\
+            $FourierMLP(\beta_{8})$ & 98\% & 100\% & 98\% & 0.29 \\
             \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 94\% & 0.52 \\
+            $FourierMLPWithPos(\beta_{8}, \chi)$ & 97\% & 98\% & 98\% & 0.43 \\
             \hline
-            $FourierMLPWithPos(\beta_{40}, \chi)$ & 91\% & 0.54 \\
+            $FourierMLPWithPos(\beta_{20}, \chi)$ & 94\% & 99\% & 93\% & 0.52 \\
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 91\% & 0.53 \\
+            $FourierMLPWithPos(\beta_{40}, \chi)$ & 91\% & 98\% & 90\% & 0.57 \\
             \hline
-            Original PICO-60 Neural Network & 94\% & 0.99 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 91\% & 94\% & 93\% & 0.53 \\
+            \hline
+            Original PICO-60 Neural Network & 83\% & 83\% & 95\% & 0.83 \\
             \hline
@@
         \captionof{table}{Summary of supervised learning configurations, trained without $WallCut(D)$.}
-        \begin{tabular}{|l|l|l|}
+        \begin{tabular}{|l|l|l|l|l|}
             \hline
-            Configuration & Max Accuracy & Mean Class-Wise Standard Deviation \\
+            Configuration & Max Accuracy & Precision & Recall & CWSD \\
             \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 92\% & 0.63 \\
+            $FourierMLPWithPos(\beta_{20}, \chi)$ & 92\% & 93\% & 77\% & 0.63 \\
             \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 97\% & 0.47 \\
+            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 97\% & 93\% & 98\% & 0.47 \\
             \hline
-            $DeepConv(\omega)$ & 77\% & 0.86 \\
+            $DeepConv(\omega)$ & 77\% & 76\% & 94\% & 0.86 \\
+            % ../experimental_data/validation_results/2018_07_16_waveform_with_new_cuts/waveform_time1531762243_epoch14.json
             \hline
-            $DeepConvWithPos(\omega, \chi)$ & 91\% & 0.61 \\
+            $DeepConvWithPos(\omega, \chi)$ & 91\% & 95\% & 78\% & 0.61 \\
             \hline
-            $ImageConv(\iota)$ & 85\% & 0.95 \\
+            $ImageConv(\iota)$ & 85\% & 40\% & 38\% & 0.95 \\
             \hline

### 2018-08-28 visualization/graph_binary_predictions.py
Revert "Revert "Reverted visualization script to use AP for ground truths""
This reverts commit 7ae76cf4663b8ed8799ed11eb9b9b1ffc7b22ad6.

@@ -16,8 +16,8 @@ from utilities.verify_arguments import verify_arguments
 # Verify that a path to the JSON data file is passed
 verify_arguments('JSON data file')
 
-# Load the data set from the file
-events, ground_truths, network_outputs = load_test(sys.argv[1])
+# Load the data set from the file, ignoring the run type ground truths
+events, _, network_outputs = load_test(sys.argv[1])
 # Get the acoustic parameter and neural network score data from the events, if they are present (if one is, both will be)
 if hasattr(events[0], 'logarithmic_acoustic_parameter'):
     acoustic_parameters, original_neural_network_scores = zip(
@@ -29,8 +29,8 @@ else:
     random_values = [random.uniform(0, 1) for _ in range(len(network_outputs))]
     acoustic_parameters = random_values
     original_neural_network_scores = random_values
-# # Calculate actual neutron/alpha ground truths based on AP
-# ground_truths = np.array(acoustic_parameters) > 0.25
+# Calculate actual neutron/alpha ground truths based on AP
+ground_truths = np.array(acoustic_parameters) > 0.25
 
 # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
 for criterion_data, criterion_name, classification_threshold in zip(

### 2018-08-28 statistics/grid_search_ap_similarity_statistics.py
Parallelized JSON loading to run on 10 threads

@@ -3,8 +3,10 @@
 # Created by Brendon Matusch, August 2018
 
 import glob
+import multiprocessing
 import os
 import sys
+from typing import Tuple
 
 import numpy as np
 
@@ -12,12 +14,8 @@ from data_processing.experiment_serialization import load_test
 from utilities.verify_arguments import verify_arguments
 
 
-# Create a dictionary to hold dictionaries of numbers of disagreements, arranged by a particular set of hyperparameters
-disagreements_by_hyperparameters = {}
-
-
-def load_disagreements(file_path: str) -> None:
-    """Load a saved validation set and print the number of disagreements if it is below a certain threshold"""
+def load_disagreements(file_path: str) -> Tuple[int, str]:
+    """Load a saved validation set and return the run identifier along with the number of disagreements"""
     # Load the validation events and network outputs from the JSON file (ignoring the ground truths)
     events, _, network_outputs = load_test(file_path)
     # Convert the network outputs to binary predictions
@@ -28,11 +26,8 @@ def load_disagreements(file_path: str) -> None:
     disagreements = np.count_nonzero(network_predictions != ap_predictions)
     # Get the full identifier of the run except for the timestamp
     run_identifier = file_path.split('time')[0]
-    # If the identifier is not already in the dictionary, create a sub-dictionary
-    if run_identifier not in disagreements_by_hyperparameters:
-        disagreements_by_hyperparameters[run_identifier] = {}
-    # In the sub-dictionary, add the number of disagreements, referenced by the specific path
-    disagreements_by_hyperparameters[run_identifier][file_path] = disagreements
+    # Return the run identifier, the number of disagreements, and the path (passed directly through to make iteration easier)
+    return run_identifier, disagreements, file_path
 
 
 # An expandable list of files using a wildcard should be provided
@@ -41,13 +36,21 @@ verify_arguments('saved validation sets using wildcard')
 file_paths = glob.glob(os.path.expanduser(sys.argv[1]), recursive=True)
 # Get the length of the list of file paths once, so it does not have to be calculated again
 file_count = len(file_paths)
-# Iterate over the files one by one with corresponding indices; these operations are not thread safe
-for file_index, file_path in enumerate(file_paths):
-    # Load and store information on the file
-    load_disagreements(file_path)
-    # Regularly print the number of files that have been loaded
+# Create a dictionary to hold dictionaries of numbers of disagreements, arranged by a particular set of hyperparameters
[…]

### 2018-08-28 statistics/grid_search_ap_similarity_statistics.py
Calculate class-wise standard deviation in AP similarity script

@@ -14,8 +14,8 @@ from data_processing.experiment_serialization import load_test
 from utilities.verify_arguments import verify_arguments
 
 
-def load_disagreements(file_path: str) -> Tuple[int, str]:
-    """Load a saved validation set and return the run identifier along with the number of disagreements"""
+def load_disagreements(file_path: str) -> Tuple[str, int, float, str]:
+    """Load a saved validation set and return the run identifier, the number of disagreements, the class-wise standard deviation, and the input file path"""
     # Load the validation events and network outputs from the JSON file (ignoring the ground truths)
     events, _, network_outputs = load_test(file_path)
     # Convert the network outputs to binary predictions
@@ -24,10 +24,17 @@ def load_disagreements(file_path: str) -> Tuple[int, str]:
     ap_predictions = np.array([event.logarithmic_acoustic_parameter > 0.25 for event in events])
     # Calculate the number of events on which AP and the network disagree
     disagreements = np.count_nonzero(network_predictions != ap_predictions)
+    # Get indices for the events that are considered neutrons and alphas by AP
+    alpha_indices = np.where(ap_predictions)
+    neutron_indices = np.where(np.logical_not(ap_predictions))
+    # Divide the network's outputs by the standard deviation of the full set to normalize them to the same range as that of AP
+    normalized_outputs = network_outputs / np.std(network_outputs)
+    # Calculate the class-wise standard deviation, using the alphas and neutrons individually
+    class_wise_standard_deviation = np.mean([np.std(normalized_outputs[indices]) for indices in [alpha_indices, neutron_indices]])
     # Get the full identifier of the run except for the timestamp
     run_identifier = file_path.split('time')[0]
-    # Return the run identifier, the number of disagreements, and the path (passed directly through to make iteration easier)
-    return run_identifier, disagreements, file_path
+    # Return the relevant outputs alongside the file path (passed directly through to make iteration easier)
+    return run_identifier, disagreements, class_wise_standard_deviation, file_path
 
 
 # An expandable list of files using a wildcard should be provided
@@ -41,23 +48,24 @@ disagreements_by_hyperparameters = {}
 # Get the run identifiers and corresponding numbers of disagreements using the file paths and corresponding indices, loading and processing files in parallel with a process pool
 pool = multiprocessing.Pool(processes=10)
 # Get a corresponding completion index for each file we iterate over
[…]

### 2018-08-28 statistics/mean_ap_accuracy_across_configurations.py
Updated mean script to calculate CWSD as well as accuracy

@@ -8,14 +8,14 @@ import numpy as np
 
 # Read all lines from standard input, and strip whitespace
 input_lines = [line.strip() for line in sys.stdin.readlines()]
-# Take only the lines that contain mean disagreement statistics
-mean_lines = [line for line in input_lines if line.startswith('Mean')]
+# Take the lines that contain mean disagreement statistics
+mean_disagreement_lines = [line for line in input_lines if line.startswith('Mean disagreements')]
 # Create a dictionary to hold the numbers of disagreements corresponding to each configuration (not including the run index)
 disagreement_counts = {}
-# Iterate over the mean lines, adding to the dictionary
-for line in mean_lines:
+# Iterate over the mean disagreement lines, adding to the dictionary
+for line in mean_disagreement_lines:
     # Take the number of disagreements (for this specific run index) from the line
-    disagreements = float(line.split()[1])
+    disagreements = float(line.split()[3])
     # Take the full path from the line, removing the run index component to get the configuration
     configuration = line.split()[-1].split('configuration_test')[0]
     # If this configuration is not in the dictionary, create an empty list for it
@@ -23,7 +23,24 @@ for line in mean_lines:
         disagreement_counts[configuration] = []
     # Add the disagreement count to the corresponding list in the dictionary
     disagreement_counts[configuration].append(disagreements)
-# Iterate over the configuration keys of the dictionary
+
+# Take the lines that contain mean class-wise standard deviation statistics
+mean_standard_deviation_lines = [line for line in input_lines if line.startswith('Mean class-wise')]
+# Create a dictionary to hold the standard deviations corresponding to each configuration (not including the run index)
+standard_deviation_values = {}
+# Iterate over the mean standard deviation lines, adding to the dictionary
+for line in mean_standard_deviation_lines:
+    # Take the standard deviation (for this specific run index) from the line
+    standard_deviation = float(line.split()[5])
+    # Take the full path from the line, removing the run index component to get the configuration
+    configuration = line.split()[-1].split('configuration_test')[0]
+    # If this configuration is not in the dictionary, create an empty list for it
+    if configuration not in standard_deviation_values:
+        standard_deviation_values[configuration] = []
+    # Add the standard deviation to the corresponding list in the dictionary
+    standard_deviation_values[configuration].append(standard_deviation)
+
+# Iterate over the configuration keys of both dictionaries (the keys should be the same)
 for configuration in disagreement_counts:
[…]

### 2018-08-28 statistics/grid_search_ap_similarity_statistics.py
Unwrap values before passing to zip

@@ -61,7 +61,7 @@ for file_index, (run_identifier, disagreements, class_wise_standard_deviation, f
 # Iterate over the run identifiers in the dictionary
 for run_identifier in disagreements_by_hyperparameters:
     # Get lists of disagreements and class-wise standard deviations from the dictionary
-    disagreement_values, class_wise_standard_deviations = zip(disagreements_by_hyperparameters[run_identifier].values())
+    disagreement_values, class_wise_standard_deviations = zip(*disagreements_by_hyperparameters[run_identifier].values())
     # Print the mean class-wise standard deviation throughout the entire run
     print(f'Mean class-wise standard deviation of {np.mean(class_wise_standard_deviations)} in run {run_identifier}')
     # Repeat for the mean number of disagreements between the network and AP

### 2018-08-28 statistics/grid_search_ap_similarity_statistics.py
Fixed ordering and indexing; time for another 15 minute wait

@@ -53,7 +53,7 @@ for file_index, (run_identifier, disagreements, class_wise_standard_deviation, f
     if run_identifier not in disagreements_by_hyperparameters:
         disagreements_by_hyperparameters[run_identifier] = {}
     # In the sub-dictionary, add the number of disagreements and the class-wise standard deviation, referenced by the specific path
-    disagreements_by_hyperparameters[run_identifier][file_path] = (class_wise_standard_deviation, disagreements)
+    disagreements_by_hyperparameters[run_identifier][file_path] = (disagreements, class_wise_standard_deviation)
     # Regularly print the index of the latest file that has been loaded
     if file_index % 100 == 0:
         print(f'Loaded file {file_index} of {file_count}')
@@ -69,7 +69,7 @@ for run_identifier in disagreements_by_hyperparameters:
     # Iterate over the specific file paths in this run
     for file_path in disagreements_by_hyperparameters[run_identifier]:
         # If there is a small number of disagreements, print out the path and number
-        disagreements = disagreements_by_hyperparameters[run_identifier][file_path]
+        disagreements = disagreements_by_hyperparameters[run_identifier][file_path][0]
         if disagreements < 10:
             print(f'{disagreements} disagreements in file {file_path}')
     # Print a few blank lines for separation

### 2018-08-28 statistics/mean_ap_accuracy_across_configurations.py
Print blank line for separation

@@ -50,3 +50,5 @@ for configuration in disagreement_counts:
     print(f'Mean accuracy for configuration {configuration} is {mean_accuracy}')
     # Calculate and output the mean class-wise standard deviation to the user
     print(f'Mean class-wise standard deviation for configuration {configuration} is {np.mean(standard_deviation_values[configuration])}')
+    # Print a blank line for separation
+    print()

### 2018-08-28 statistics/grid_search_ap_similarity_statistics.py
Calculate precision and recall in AP similarity script

@@ -14,8 +14,8 @@ from data_processing.experiment_serialization import load_test
 from utilities.verify_arguments import verify_arguments
 
 
-def load_disagreements(file_path: str) -> Tuple[str, int, float, str]:
-    """Load a saved validation set and return the run identifier, the number of disagreements, the class-wise standard deviation, and the input file path"""
+def load_disagreements(file_path: str) -> Tuple[str, int, float, float, float, str]:
+    """Load a saved validation set and return the run identifier, the number of disagreements, the precision and recall, the class-wise standard deviation, and the input file path"""
     # Load the validation events and network outputs from the JSON file (ignoring the ground truths)
     events, _, network_outputs = load_test(file_path)
     # Convert the network outputs to binary predictions
@@ -24,6 +24,15 @@ def load_disagreements(file_path: str) -> Tuple[str, int, float, str]:
     ap_predictions = np.array([event.logarithmic_acoustic_parameter > 0.25 for event in events])
     # Calculate the number of events on which AP and the network disagree
     disagreements = np.count_nonzero(network_predictions != ap_predictions)
+    # Calculate precision (erroneous recoil predictions) and recall (erroneous alpha predictions) individually
+    # First, count the number of events predicted as recoils, and then divide the number that are incorrectly predicted to be recoils by that
+    predicted_recoils = np.count_nonzero(network_predictions == 0)
+    erroneous_recoils = np.count_nonzero(np.logical_and(network_predictions == 0, ap_predictions == 1))
+    precision = (predicted_recoils - erroneous_recoils) / predicted_recoils
+    # Repeat for the proportion of neutron calibration events that are correctly predicted as recoils; this represents not the purity, but the number missed
+    actual_recoils = np.count_nonzero(ap_predictions == 0)
+    erroneous_alphas = np.count_nonzero(np.logical_and(network_predictions == 1, ap_predictions == 0))
+    recall = (actual_recoils - erroneous_alphas) / actual_recoils
     # Get indices for the events that are considered neutrons and alphas by AP
     alpha_indices = np.where(ap_predictions)
     neutron_indices = np.where(np.logical_not(ap_predictions))
@@ -34,7 +43,7 @@ def load_disagreements(file_path: str) -> Tuple[str, int, float, str]:
     # Get the full identifier of the run except for the timestamp
     run_identifier = file_path.split('time')[0]
     # Return the relevant outputs alongside the file path (passed directly through to make iteration easier)
[…]

### 2018-08-28 statistics/grid_search_ap_similarity_statistics.py
Removed multiprocessing stuff to see if it changes the weird-looking precision and recall

@@ -3,7 +3,6 @@
 # Created by Brendon Matusch, August 2018
 
 import glob
-import multiprocessing
 import os
 import sys
 from typing import Tuple
@@ -54,10 +53,10 @@ file_paths = glob.glob(os.path.expanduser(sys.argv[1]), recursive=True)
 file_count = len(file_paths)
 # Create a dictionary to hold dictionaries of numbers of disagreements, arranged by a particular set of hyperparameters
 disagreements_by_hyperparameters = {}
-# Get the run identifiers and corresponding numbers of disagreements using the file paths and corresponding indices, loading and processing files in parallel with a process pool
-pool = multiprocessing.Pool(processes=10)
 # Get a corresponding completion index for each file we iterate over
-for file_index, (run_identifier, disagreements, precision, recall, class_wise_standard_deviation, file_path) in enumerate(pool.imap_unordered(load_disagreements, file_paths)):
+for file_index, (run_identifier, disagreements, precision, recall, class_wise_standard_deviation, file_path) in enumerate(load_disagreements(file_path) for file_path in file_paths):
+    # Print the recall for debugging purposes
+    print(f'Recall of {recall} for file {file_path}')
     # If the run identifier is not already in the dictionary, create a sub-dictionary
     if run_identifier not in disagreements_by_hyperparameters:
         disagreements_by_hyperparameters[run_identifier] = {}

### 2018-08-29 documentation/paper.tex
Started adding the report for ICN grid search
Added single example row for best run
Added table and updated one graph for new ICN grid search; gravitational analysis is in progress

@@
 
+To ensure that accuracy statistics are as consistent and reliable as possible, every configuration was trained three times, with a different randomly selected validation set each time.
+Accuracy, precision, recall, and class-wise standard deviation statistics are the mean over those 3 tests.
+
 \subsection{Gravitational Differentiation}
@@
 
-Gravitational differentiation is a novel technique, developed during this study, for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
+Gravitational differentiation (GD) is a novel technique, developed during this study, for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
 
@@
 
-Iterative cluster nucleation is a second original semi-supervised learning algorithm developed as part of this study.
+Iterative cluster nucleation (ICN) is a second original semi-supervised learning algorithm developed as part of this study.
 It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
@@
 
-During a grid search of these parameters (in addition to regularization parameters of the network), the best accuracy achieved for replication of AP was exactly 100\%.
-The mean of all epochs within that training run was 98\%.
+During a grid search of these parameters (in addition to regularization parameters of the network), the best mean accuracy over three runs of an individual configuration was 99\%, using 256 initial labeled examples in $\varsigma$.
+With 128 examples, the maximum was 98\%; with 64 it was 97\%, and with 32 95\%.
+From this, it is evident that it is possible to achieve very high accuracy using this technique.
+It is also clear that, in general, larger initial sets produce higher performance.
 
-As seen in Figure \ref{icn_grid_search}, the network learned to replicate AP very accurately.
-Interestingly, the examples with AP values between the neutron and alpha particle ranges were also classified as such by the network.
-This may imply the indecisiveness on these examples is fundamental to the data rather than an implementation problem with AP.
+The network's highly accurate and decisive predictions can be observed in Figure \ref{icn_grid_search}.
 
@@
 
+THIS NEEDS TO BE EDITED VVVV
+
 Digging deeper into the performance throughout the runs of the grid search, Figure \ref{icn_acc_by_hyper} shows the mean accuracy as a function of each of the hyperparameters that were optimized.
@@
 L2 regularization is also significant; all $\lambda$ values greater than 0 make performance worse.
-
-Compared to gravitational differentiation, there is a larger delta between the number of errors on the validation set depending on the number of initial training examples used.
-Those with 128 examples in $\varsigma$ make 85\% more errors on average.
-However, iterative cluster nucleation with 128 examples still makes 19\% fewer errors than gravitational differentiation with 256 examples in $\varsigma$.
 
@@
 
+The following table represents the performance of a variety of iterative cluster nucleation configurations tested during the grid search.
+As with gravitational differentiation, the accuracy statistic is the mean over three tests of the same configuration.
+
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Performance of iterative cluster nucleation configurations.}
+        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
+            \hline
+            Initial $\varsigma$ Size & L2 $\lambda$ & Dropout & Initial $j$ & $k$ & Mean AP Accuracy & Precision & Recall & CWSD \\
+            \hline
+            32 & 0 & 0.25 & 0.01 & 1.025 & 80\% & 79\% & 99\% & 0.60 \\
+            \hline
+            32 & 0 & 0 & 0.01 & 1.025 & 95\% & 94\% & 99\% & 0.38 \\
+            \hline
+            64 & 0.001 & 0 & 0.01 & 1.05 & 97\% & 96\% & 99\% & 0.28 \\
+            \hline
+            64 & 0 & 0.25 & 0.02 & 1.025 & 97\% & 95\% & 100\% & 0.27 \\
+            \hline
+            128 & 0 & 0.5 & 0.01 & 1.05 & 92\% & 90\% & 100\% & 0.35 \\
+            \hline
+            128 & 0 & 0.5 & 0.02 & 1.05 & 98\% & 98\% & 100\% & 0.21 \\
+            \hline
+            256 & 0 & 0.5 & 0.01 & 1.05 & 99\% & 99\% & 100\% & 0.15 \\
+            \hline
+            256 & 0.003 & 0.5 & 0.01 & 1.05 & 98\% & 98\% & 98\% & 0.26 \\
+            \hline
+        \end{tabular}
+    \end{center}
+\end{minipage}
+
 \subsection{Performance Summary}
 
-The following table directly compares performance of the semi-supervised learning systems, with the existing PICO-60 neural network provided as a baseline once again.
-Note that ``Max AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.
+The following table directly compares performance of certain successful semi-supervised learning systems, with AP provided as a baseline for comparison.
+Remember that ``Mean AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.
 
-\begin{tabular}{|l|l|l|l|}
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Summary of semi-supervised learning systems.}
+        \begin{tabular}{|l|l|l|l|l|l|}
             \hline
-    Technique & Max AP Accuracy & Max Mean by Run & Class-Wise SD \\
+            Initial $\varsigma$ Size & Technique & Mean AP Accuracy & Mean Precision & Mean Recall & Mean CWSD \\
             \hline
-    Gravitational Differentiation (256 $\varsigma$) & 98\% & 96\% & TODO \\
+            32 & ICN & 95\% & 94\% & 99\% & 0.38 \\
             \hline
-    Gravitational Differentiation (128 $\varsigma$) & TODO & TODO & TODO \\
+            64 & ICN & 97\% & 96\% & 99\% & 0.28 \\
             \hline
-    Iterative Cluster Nucleation (256 $\varsigma$) & 100\% & 98\% & TODO \\
+            128 & ICN & 98\% & 98\% & 100\% & 0.21 \\
             \hline
-    Iterative Cluster Nucleation (128 $\varsigma$) & TODO & TODO & TODO \\
+            256 & ICN & 99\% & 99\% & 100\% & 0.15 \\
             \hline
-    Existing NN TODO & X & X & X \\
-    \hline
-    AP TODO & X & X & X \\
+            $\infty$ & AP & 100\% & 100\% & 100\% & TODO \\
             \hline
         \end{tabular}
+    \end{center}
+\end{minipage}
 

### 2018-08-30 visualization/semi_supervised_visualize_optimal_hyperparameters.py
Created optimal hyperparameter visualization script for semi-supervised grid searches

@@ -1,0 +1,62 @@
+#!/usr/bin/env python3
+"""A script for visualizing the optimal value of each hyperparameter in a semi-supervised learning grid search"""
+# Created by Brendon Matusch, August 2018
+
+import sys
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Read all lines provided to standard input (from the script that calculates mean by configuration) and strip whitespace
+lines = [line.strip() for line in sys.stdin.readlines()]
+# Get the configuration identifier from each line
+configurations = [line.split()[1] for line in lines]
+# Extract the accuracy value from each line
+accuracy_values = [float(line.split()[3]) for line in lines]
+# Create a dictionary to add accuracy dictionaries to, based on human-readable names
+accuracy_dictionaries = {}
+# Iterate over each of the hyperparameters that are used, alongside human-readable names
+for hyperparameter, human_readable in [('dropout', 'Dropout'), ('l2_lambda', 'L2 Regularization Lambda'), ('initial_threshold', 'Initial Training Threshold'), ('threshold_multiplier', 'Training Threshold Multiplier'), ('initial_examples', 'Initial Examples')]:
+    # Create a dictionary to add lists of accuracy values according to hyperparameter values
+    accuracy_by_hyperparameter_value = {}
+    # Iterate over each of the configurations with corresponding accuracy values
+    for configuration, accuracy in zip(configurations, accuracy_values):
+        # Split the line by the name of the hyperparameter, and then get the hyperparameter value (as a floating-point number) up to the next slash separator
+        hyperparameter_value = float(configuration.split(hyperparameter)[1].split('/')[0])
+        # If this value is not already in the dictionary, add a corresponding empty list
+        if hyperparameter_value not in accuracy_by_hyperparameter_value:
+            accuracy_by_hyperparameter_value[hyperparameter_value] = []
+        # Add the corresponding accuracy to the list
+        accuracy_by_hyperparameter_value[hyperparameter_value].append(accuracy)
+    # Replace the lists of means by run with the overall mean for that hyperparameter value
+    for hyperparameter_value in accuracy_by_hyperparameter_value:
+        accuracy_by_hyperparameter_value[hyperparameter_value] = np.mean(accuracy_by_hyperparameter_value[hyperparameter_value])
+    # Find the index of the best accuracy value, and print out the corresponding optimal hyperparameter value
[…]

### 2018-08-30 documentation/paper.tex
Removed graphs with accuracy by hyper

@@
 
-As seen in Figure \ref{grav_acc_by_hyper}, all three hyperparameters that were optimized during the grid search had a significant effect on accuracy:
-
-\begin{itemize}
-    \item Smaller gravitational multipliers improved performance; the best value tested was 0.0005.
-This implies that the gravity effect on each unlabeled event should be weighted lightly compared to the labeled examples.
-    \item Higher learning rates improve performance, with the best value tested at a quite high 0.03.
-    \item Larger distortion powers (which raise the confidence threshold for a significant gravitational effect) improved performance, with the best tested at 11.
-\end{itemize}
-
 Two different options were tested with regard to size of the labeled data set $\varsigma$.
@@
     \caption{\label{grav_grid_search} Accuracy of best gravitational differentiation training run compared to AP (98\% accuracy).}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{grav_acc_by_hyper}
-    \caption{\label{grav_acc_by_hyper} Accuracy of gravitational differentiation as a function of hyperparameter values.}
 \end{figure}
@@
 During a grid search of these parameters (in addition to regularization parameters of the network), the best mean accuracy over three runs of an individual configuration was 99\%, using 256 initial labeled examples in $\varsigma$.
-With 128 examples, the maximum was 98\%; with 64 it was 97\%, and with 32 95\%.
+With 128 examples, the maximum was 98\%; with 64 it was 97\%; and with 32, 95\%.
 From this, it is evident that it is possible to achieve very high accuracy using this technique.
@@
     \caption{\label{icn_grid_search} Accuracy of optimal iterative cluster nucleation training run compared to AP (100\% accuracy).}
-\end{figure}
-
-THIS NEEDS TO BE EDITED VVVV
-
-Digging deeper into the performance throughout the runs of the grid search, Figure \ref{icn_acc_by_hyper} shows the mean accuracy as a function of each of the hyperparameters that were optimized.
-Dropout has a clear effect, with 25\% shown to be the optimal number of neurons to drop.
-L2 regularization is also significant; all $\lambda$ values greater than 0 make performance worse.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{icn_acc_by_hyper}
-    \caption{\label{icn_acc_by_hyper} Mean accuracy of iterative cluster nucleation as a function of hyperparameter values.}
 \end{figure}

### 2018-08-30 training/hyperparameter_correlation_linear_regression.py
Created script for calculating correlation using linear regression

@@ -1,0 +1,33 @@
+#!/usr/bin/env python3
+"""Given mean accuracy values by configuration for a semi-supervised learning grid search, run a linear regression to predict accuracy based on the hyperparameter values and determine how well they are correlated"""
+# Created by Brendon Matusch, Augsut 2018
+
+import sys
+
+import numpy as np
+from sklearn.linear_model import LinearRegression
+
+# The hyperparameters that we are concerned with using to predict accuracy
+HYPERPARAMETERS = ['dropout', 'l2_lambda', 'initial_threshold', 'threshold_multiplier']
+
+# Read all lines provided to standard input (from the script that calculates mean by configuration) and strip whitespace
+lines = [line.strip() for line in sys.stdin.readlines()]
+# Get the configuration identifier from each line
+configurations = [line.split()[1] for line in lines]
+# Create a NumPy array that will contain values of the hyperparameters
+# It is okay if not all of the configurations are unique because of disregarded hyperparameters; they can be used for training individually
+hyperparameter_values = np.empty((len(configurations), len(HYPERPARAMETERS)))
+# Iterate over the configurations with corresponding indices, extracting arrays of hyperparameters
+for configuration_index, configuration in enumerate(configurations):
+    # Iterate over the hyperparameters with corresponding indices, extracting them one by one
+    for hyperparameter_index, hyperparameter in enumerate(HYPERPARAMETERS):
+        # Split the line by the name of the hyperparameter, and then get the hyperparameter value (as a floating-point number) up to the next slash separator
+        hyperparameter_values[configuration_index, hyperparameter_index] = float(configuration.split(hyperparameter)[1].split('/')[0])
+# Extract the accuracy value from each line
+accuracy_values = [float(line.split()[3]) for line in lines]
+# Create a linear regression classifier and train it to predict the accuracy values based on the hyperparameters
+linear_regression = LinearRegression()
+linear_regression.fit(hyperparameter_values, accuracy_values)
+# Print the classifier's score on its training set to get an idea of the correlation between the hyperparameters and accuracy
+score = linear_regression.score(hyperparameter_values, accuracy_values)
+print('Score:', score)

### 2018-08-31 training/hyperparameter_correlation_linear_regression.py
Added polynomials to regression

@@ -6,6 +6,7 @@ import sys
 
 import numpy as np
 from sklearn.linear_model import LinearRegression
+from sklearn.preprocessing import PolynomialFeatures
 
 # The hyperparameters that we are concerned with using to predict accuracy
 HYPERPARAMETERS = ['dropout', 'l2_lambda', 'initial_threshold', 'threshold_multiplier']
@@ -23,11 +24,13 @@ for configuration_index, configuration in enumerate(configurations):
     for hyperparameter_index, hyperparameter in enumerate(HYPERPARAMETERS):
         # Split the line by the name of the hyperparameter, and then get the hyperparameter value (as a floating-point number) up to the next slash separator
         hyperparameter_values[configuration_index, hyperparameter_index] = float(configuration.split(hyperparameter)[1].split('/')[0])
+# Add polynomial features up to degree 3 to the hyperparameter matrix
+hyperparameter_polynomials = PolynomialFeatures(degree=3).fit_transform(hyperparameter_values)
 # Extract the accuracy value from each line
 accuracy_values = [float(line.split()[3]) for line in lines]
-# Create a linear regression classifier and train it to predict the accuracy values based on the hyperparameters
+# Create a linear regression classifier and train it to predict the accuracy values based on the hyperparameters including polynomials
 linear_regression = LinearRegression()
-linear_regression.fit(hyperparameter_values, accuracy_values)
+linear_regression.fit(hyperparameter_polynomials, accuracy_values)
 # Print the classifier's score on its training set to get an idea of the correlation between the hyperparameters and accuracy
-score = linear_regression.score(hyperparameter_values, accuracy_values)
+score = linear_regression.score(hyperparameter_polynomials, accuracy_values)
 print('Score:', score)

### 2018-08-31 training/hyperparameter_correlation_linear_regression.py
Weird hack to print out the mean accuracy for each number of initial examples

@@ -9,7 +9,7 @@ from sklearn.linear_model import LinearRegression
 from sklearn.preprocessing import PolynomialFeatures
 
 # The hyperparameters that we are concerned with using to predict accuracy
-HYPERPARAMETERS = ['dropout', 'l2_lambda', 'initial_threshold', 'threshold_multiplier']
+HYPERPARAMETERS = ['dropout', 'l2_lambda', 'initial_threshold', 'threshold_multiplier', 'initial_examples']
 
 # Read all lines provided to standard input (from the script that calculates mean by configuration) and strip whitespace
 lines = [line.strip() for line in sys.stdin.readlines()]
@@ -18,16 +18,22 @@ configurations = [line.split()[1] for line in lines]
 # Create a NumPy array that will contain values of the hyperparameters
 # It is okay if not all of the configurations are unique because of disregarded hyperparameters; they can be used for training individually
 hyperparameter_values = np.empty((len(configurations), len(HYPERPARAMETERS)))
+indices = []
 # Iterate over the configurations with corresponding indices, extracting arrays of hyperparameters
 for configuration_index, configuration in enumerate(configurations):
     # Iterate over the hyperparameters with corresponding indices, extracting them one by one
     for hyperparameter_index, hyperparameter in enumerate(HYPERPARAMETERS):
+        if hyperparameter == 'initial_examples' and float(configuration.split(hyperparameter)[1].split('/')[0]) == 256:
+            indices.append(configuration_index)
         # Split the line by the name of the hyperparameter, and then get the hyperparameter value (as a floating-point number) up to the next slash separator
         hyperparameter_values[configuration_index, hyperparameter_index] = float(configuration.split(hyperparameter)[1].split('/')[0])
+hyperparameter_values = hyperparameter_values[indices]
 # Add polynomial features up to degree 3 to the hyperparameter matrix
 hyperparameter_polynomials = PolynomialFeatures(degree=3).fit_transform(hyperparameter_values)
 # Extract the accuracy value from each line
-accuracy_values = [float(line.split()[3]) for line in lines]
+accuracy_values = np.array([float(line.split()[3]) for line in lines])
+accuracy_values = accuracy_values[indices]
+print(np.mean(accuracy_values))
 # Create a linear regression classifier and train it to predict the accuracy values based on the hyperparameters including polynomials
 linear_regression = LinearRegression()
 linear_regression.fit(hyperparameter_polynomials, accuracy_values)

### 2018-08-31 documentation/paper.tex
Explained accuracy based on hyperparameters

@@
 
+Interestingly, the four hyperparameters optimized during the grid search were observed to have a weak polynomial correlation with the network's accuracy.
+This makes the mean accuracy over all configurations a meaningful statistic, since the accuracy for an individual configuration is difficult to confidently predict.
+
+With 32 and 64 examples it was only 75\% and 84\%, respectively, implying that the high maximum accuracy statistics were subject to a high degree of randomness.
+However, for 128 and 256 initial examples, the mean accuracy statistics over all configurations were 95\% and 97\%, respectively, indicating that the training is much more reliable with a larger initial set.
+
+Note that there are several hyperparameters not specifically optimized in the context of iterative cluster nucleation, including the number of layers and numbers of neurons within those layers.
+The correlation between those hyperparameters and accuracy is subject to future research.
+
 The network's highly accurate and decisive predictions can be observed in Figure \ref{icn_grid_search}.
@@
 
-The following table represents the performance of a variety of iterative cluster nucleation configurations tested during the grid search.
+The following table represents the performance of a variety of high-accuracy iterative cluster nucleation configurations tested during the grid search.
 As with gravitational differentiation, the accuracy statistic is the mean over three tests of the same configuration.
@@
             \hline
-            $\infty$ & AP & 100\% & 100\% & 100\% & TODO \\
+            $\infty$ & AP & 100\% & 100\% & 100\% & 0.41 \\
             \hline

### 2018-08-31 training/hyperparameter_correlation_linear_regression.py
Revert "Weird hack to print out the mean accuracy for each number of initial examples"
This reverts commit 94525024f44740bf7f7864a822b847d3ffd8cf84.

@@ -9,7 +9,7 @@ from sklearn.linear_model import LinearRegression
 from sklearn.preprocessing import PolynomialFeatures
 
 # The hyperparameters that we are concerned with using to predict accuracy
-HYPERPARAMETERS = ['dropout', 'l2_lambda', 'initial_threshold', 'threshold_multiplier', 'initial_examples']
+HYPERPARAMETERS = ['dropout', 'l2_lambda', 'initial_threshold', 'threshold_multiplier']
 
 # Read all lines provided to standard input (from the script that calculates mean by configuration) and strip whitespace
 lines = [line.strip() for line in sys.stdin.readlines()]
@@ -18,22 +18,16 @@ configurations = [line.split()[1] for line in lines]
 # Create a NumPy array that will contain values of the hyperparameters
 # It is okay if not all of the configurations are unique because of disregarded hyperparameters; they can be used for training individually
 hyperparameter_values = np.empty((len(configurations), len(HYPERPARAMETERS)))
-indices = []
 # Iterate over the configurations with corresponding indices, extracting arrays of hyperparameters
 for configuration_index, configuration in enumerate(configurations):
     # Iterate over the hyperparameters with corresponding indices, extracting them one by one
     for hyperparameter_index, hyperparameter in enumerate(HYPERPARAMETERS):
-        if hyperparameter == 'initial_examples' and float(configuration.split(hyperparameter)[1].split('/')[0]) == 256:
-            indices.append(configuration_index)
         # Split the line by the name of the hyperparameter, and then get the hyperparameter value (as a floating-point number) up to the next slash separator
         hyperparameter_values[configuration_index, hyperparameter_index] = float(configuration.split(hyperparameter)[1].split('/')[0])
-hyperparameter_values = hyperparameter_values[indices]
 # Add polynomial features up to degree 3 to the hyperparameter matrix
 hyperparameter_polynomials = PolynomialFeatures(degree=3).fit_transform(hyperparameter_values)
 # Extract the accuracy value from each line
-accuracy_values = np.array([float(line.split()[3]) for line in lines])
-accuracy_values = accuracy_values[indices]
-print(np.mean(accuracy_values))
+accuracy_values = [float(line.split()[3]) for line in lines]
 # Create a linear regression classifier and train it to predict the accuracy values based on the hyperparameters including polynomials
 linear_regression = LinearRegression()
 linear_regression.fit(hyperparameter_polynomials, accuracy_values)

### 2018-09-02 documentation/paper.tex
Added table for gravitational differentiation; did associated writeup

--- 2018-08-30 documentation/paper.tex
@@
 First, they are optimized in a supervised learning context.
-Next, two novel semi-supervised learning algorithms are trained, and found to replicate the Acoustic Parameter discriminator previously used in PICO-60 with up to 100\% accuracy.
+Next, two novel semi-supervised learning algorithms are trained, and found to replicate the Acoustic Parameter discriminator previously used in PICO-60 with a mean of 97\% accuracy.
 \end{abstract}
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
     \centering
@@
 
-During a grid search of parameters of the gravitational differentiation algorithm (in addition to the stochastic gradient descent learning rate), the algorithm achieved a maximum of 98\% validation accuracy.
-The mean over that run (also the highest overall) was 96\%.
-Figure \ref{grav_grid_search} shows the highly accurate replication of AP.
-
-Two different options were tested with regard to size of the labeled data set $\varsigma$.
-Out of a set of 496 total examples available for training, either 128 (approximately $1/4$ of the total set) or 256 (approximately $1/2$) were used as labeled data.
-Over the entire grid search, runs with 128 examples in $\varsigma$ produced on average 64\% more errors on validation data, compared to those with 256 examples.
-
-!!!Include table on grid search!!!
+During a grid search of parameters of the gravitational differentiation algorithm (in addition to the stochastic gradient descent learning rate), the highest mean accuracy over three runs of an individual configuration was 99\%.
+The highest accuracy was obtained with 256 labeled examples in the set $\varsigma$ (out of 496 total training examples).
+High accuracy statistics in this range were also obtained using 128 examples (99\% accuracy, once again) and 64 examples (97\% accuracy).
+32 examples produced a much lower maximum accuracy, at only 83\%.
+From this, it is evident that it is possible to achieve very high accuracy using this technique.
+It is also clear that, in general, larger initial sets produce higher performance.
+
+However, the four hyperparameters optimized during the grid search were observed to have a weak polynomial correlation with the network's accuracy.
+This implies that the training process is generally insensitive to variations in those particular network parameters.
+
+This lack of correlation also means that the mean accuracy over all configurations with a given $\varsigma$ size is a meaningful statistic, since the validation accuracy for an individual configuration is difficult to confidently predict.
+These statistics, shown in Table \ref{grav_accuracy_by_varsigma}, show that for all four set sizes, there is a fairly wide spread between maximum and mean accuracy.
+
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Gravitational differentiation accuracy statistics by size of $\varsigma$.} \label{grav_accuracy_by_varsigma}
+        \begin{tabular}{|l|l|l|}
+            \hline
+            Initial $\varsigma$ Size & Max Accuracy & Mean Accuracy \\
+            \hline
+            32 & 83\% & 69\% \\
+            \hline
+            64 & 97\% & 73\% \\
+            \hline
+            128 & 99\% & 84\% \\
+            \hline
+            256 & 99\% & 94\% \\
+            \hline
+        \end{tabular}
+    \end{center}
+\end{minipage}
+
+The following table represents the performance of several high-accuracy gravitational differentiation configurations tested during the grid search.
+The accuracy statistic is the mean over three tests of the same configuration.
+
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Performance of gravitational differentiation configurations.}
+        \begin{tabular}{|l|l|l|l|l|l|l|l|}
+            \hline
+            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Mean AP Accuracy & Precision & Recall & CWSD \\
+            \hline
+            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% 90\% & 0.78 \\
+            \hline
+            32 & 0.0005 & 0.003 & 9 & 83\% & 80\% & 99\% & 0.63 \\
+            \hline
+            64 & 0.0005 & 0.003 & 9 & 77\% & 74\% & 97\% & 0.71 \\
+            \hline
+            64 & 0.003 & 0.03 & 7 & 97\% & 96\% & 100\% & 0.18 \\
+            \hline
+            128 & 0.008 & 0.001 & 9 & 74\% & 71\% & 95\% & 0.70 \\
+            \hline
+            128 & 0.0005 & 0.03 & 11 & 99\% & 99\% & 99\% & 0.17 \\
+            \hline
+            256 & 0.005 & 0.01 & 11 & 97\% & 96\% & 99\% & 0.25 \\
+            \hline
+            256 & 0.001 & 0.03 & 3 & 99\% & 99\% & 100\% & 0.10 \\
+            \hline
+        \end{tabular}
+    \end{center}
+\end{minipage}
 
@@
     \includegraphics[width=\textwidth]{grav_grid_search}
-    \caption{\label{grav_grid_search} Accuracy of best gravitational differentiation training run compared to AP (98\% accuracy).}
+    \caption{\label{grav_grid_search} Accuracy of best gravitational differentiation training epoch compared to AP.}
 \end{figure}
@@
 During a grid search of these parameters (in addition to regularization parameters of the network), the best mean accuracy over three runs of an individual configuration was 99\%, using 256 initial labeled examples in $\varsigma$.
-With 128 examples, the maximum was 98\%; with 64 it was 97\%; and with 32, 95\%.
-From this, it is evident that it is possible to achieve very high accuracy using this technique.
-It is also clear that, in general, larger initial sets produce higher performance.
-
-The network's highly accurate and decisive predictions can be observed in Figure \ref{icn_grid_search}.
+(The network's predictions on the best performing epoch can be observed in Figure \ref{icn_grid_search}.) With 128 examples, the maximum was 98\%; with 64 it was 97\%; and with 32, 95\%.
+As with gravitational differentiation, larger set sizes improve accuracy significantly.
 
@@
     \includegraphics[width=\textwidth]{icn_grid_search}
-    \caption{\label{icn_grid_search} Accuracy of optimal iterative cluster nucleation training run compared to AP (100\% accuracy).}
+    \caption{\label{icn_grid_search} Accuracy of best iterative cluster nucleation training epoch compared to AP.}
 \end{figure}
 
-The following table represents the performance of a variety of iterative cluster nucleation configurations tested during the grid search.
+Once again, the parameters inherent to the ICN algorithm, as well as dropout and L2 regularization, do not have a meaningful polynomial correlation with accuracy.
+Note that there are several hyperparameters not specifically optimized in the context of iterative cluster nucleation, including the number of layers and numbers of neurons within those layers.
+The correlation between those hyperparameters and accuracy is subject to future research.
+
+The mean accuracy for a given initial $\varsigma$ is again important.
+In Table \ref{icn_accuracy_by_varsigma}, it is clear that with small initial $\varsigma$ sets, the training process is very unreliable, much like gravitational differentiation.
+However, the mean accuracy is much closer to the maximum for initial $\varsigma$ sets of 128 and 256, indicating that the training process of ICN is significantly more consistent than gravitational differentiation.
+
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Iterative cluster nucleation accuracy statistics by initial size of $\varsigma$.} \label{icn_accuracy_by_varsigma}
+        \begin{tabular}{|l|l|l|}
+            \hline
+            Initial $\varsigma$ Size & Max Accuracy & Mean Accuracy \\
+            \hline
+            32 & 95\% & 75\% \\
+            \hline
+            64 & 97\% & 84\% \\
+            \hline
+            128 & 98\% & 95\% \\
+            \hline
+            256 & 99\% & 97\% \\
+            \hline
+        \end{tabular}
+    \end{center}
+\end{minipage}
+
+The following table represents the performance of a variety of high-accuracy iterative cluster nucleation configurations tested during the grid search.
 As with gravitational differentiation, the accuracy statistic is the mean over three tests of the same configuration.
@@
 
-The following table directly compares performance of certain successful semi-supervised learning systems, with AP provided as a baseline for comparison.
+The following table directly compares performance of the most successful semi-supervised learning configurations, with AP provided as a baseline for comparison.
 Remember that ``Mean AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.
@@
     \begin{center}
-        \captionof{table}{Summary of semi-supervised learning systems.}
+        \captionof{table}{Summary of the most successful semi-supervised learning configuration.}
         \begin{tabular}{|l|l|l|l|l|l|}
@@
             \hline
+            32 & GD & 83\% & 80\% & 99\% & 0.63 \\
+            \hline
             32 & ICN & 95\% & 94\% & 99\% & 0.38 \\
             \hline
+            64 & GD & 97\% & 96\% & 100\% & 0.18 \\
+            \hline
             64 & ICN & 97\% & 96\% & 99\% & 0.28 \\
             \hline
+            128 & GD & 99\% & 99\% & 99\% & 0.17 \\
+            \hline
             128 & ICN & 98\% & 98\% & 100\% & 0.21 \\
             \hline
+            256 & GD & 99\% & 99\% & 100\% & 0.10 \\
+            \hline
             256 & ICN & 99\% & 99\% & 100\% & 0.15 \\
             \hline
-            $\infty$ & AP & 100\% & 100\% & 100\% & TODO \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
+            $\infty$ & AP & 100\% & 100\% & 100\% & 0.41 \\
+            \hline
+        \end{tabular}
+    \end{center}
+\end{minipage}
+
+Note that in addition to achieving similar accuracy, configurations from both semi-supervised learning algorithms produce significantly more decisive predictions compared to AP, as measured by the class-wise standard deviation.
 
@@
 While both techniques produced some very effective discriminators, iterative cluster nucleation was found to be the more effective of the two.
-It replicated AP with a mean of 98\% accuracy on the best configuration.
+It replicated AP with a mean of 97\% accuracy, using 256 labeled training examples.
 This indicates that the technique shows significant promise.

### 2018-09-25 documentation/authors.tex
Added LaTeX file containing authors

\author{B.~Matusch}
\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author{C.~Amole}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{M.~Ardid}
\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}

\author{I.~J.~Arnquist}
\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author{D.~M.~Asner}
%\affil{Brookhaven National Laboratory, Upton, New York 11973, USA}
\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author{D.~Baxter}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
\affil{Enrico Fermi Institute, KICP and Department of Physics, 
University of Chicago, Chicago, Illinois 60637, USA}
%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{E.~Behnke}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{P.~Bhattacharjee}
%\affil{Astroparticle Physics and Cosmology Division, Saha Institute of Nuclear Physics, Kolkata, India}

%\author{H.~Borsodi}
%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{M.~Bou-Cabo}
%\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}

\author{M.~Bressler}
\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}

\author{B.~Broerman}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

%\author{P.~Campion}
%\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}

\author{G.~Cao}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{C.~J.~Chen}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}

\author{U.~Chowdhury}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{K.~Clark}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
%\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
%\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author{J.~I.~Collar}
\affil{Enrico Fermi Institute, KICP and Department of Physics,
University of Chicago, Chicago, Illinois 60637, USA}

\author{P.~S.~Cooper}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{M.~Crisler}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author{G.~Crowder}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{N.A.~Cruz-Venegas}
\affil{Instituto de F\'isica, Universidad Nacional Aut\'onoma de M\'exico, M\'exico D.\:F. 01000, M\'exico}

\author{C.~E.~Dahl}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{M.~Das}
\affil{Astroparticle Physics and Cosmology Division, Saha Institute
of Nuclear Physics, Kolkata, India}

\author{S.~Fallows}
\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author{J.~Farine}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{I.~Felis}
\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}

\author{R.~Filgas}
\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}

\author{F.~Girard}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{G.~Giroux}
% \email{ggiroux@owl.phy.queensu.ca}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{J.~Hall}
%\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author{C.~Hardy}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{O.~Harris}
\affil{Northeastern Illinois University, Chicago, Illinois 60625, USA}

\author{E.~W.~Hoppe}
\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}

\author{M.~Jin}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}

\author{L.~Klopfenstein}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author{C.~B.~Krauss}
\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author{M.~Laurin}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{I.~Lawson}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}

\author{A.~Leblanc}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{I.~Levine}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author{W.~H.~Lippincott}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{F.~Mamedov}
\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}

\author{D.~Maurya}
\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}

\author{P.~Mitra}
\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author{C.~Moore}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{T.~Nania}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author{R.~Neilson}
\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}

\author{A.~J.~Noble}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{P.~Oedekerk}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{S.~Olson}
%\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{A.~Ortega}
\affil{Enrico Fermi Institute, KICP and Department of Physics,
University of Chicago, Chicago, Illinois 60637, USA}

\author{M.-C.~Piro}
\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}

\author{A.~Plante}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{R.~Podviyanuk}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{S.~Priya}
\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}

\author{A.~E.~Robinson}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

%\author{A.~Roeder}
%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{R.~Rucinski}
%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{S.~Sahoo}
\affil{Astroparticle Physics and Cosmology Division, Saha Institute
of Nuclear Physics, Kolkata, India}

\author{O.~Scallon}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{S.~Seth}
\affil{Astroparticle Physics and Cosmology Division, Saha Institute
of Nuclear Physics, Kolkata, India}

\author{A.~Sonnenschein}
\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}

\author{N.~Starinski}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{I.~\v{S}tekl}
\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}

\author{T.~Sullivan}
\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}

\author{F.~Tardif}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{E.~V\'azquez-J\'auregui}
\affil{Instituto de F\'isica, Universidad Nacional Aut\'onoma de M\'exico, M\'exico D.\:F. 01000, M\'exico}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{N.~Walkowski}
\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

%\author{J.~Wells}
%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}

\author{U.~Wichoski}
\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}

\author{Y.~Yan}
\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}

\author{V.~Zacek}
\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}

\author{J.~Zhang}
\affil{
Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}

### 2018-09-26 documentation/paper.tex
Added LaTeX file containing authors
Made some of the edits to get ready for collaboration review

@@
 \usepackage{amsmath}
+\usepackage{authblk}
 \usepackage[backend=biber, style=alphabetic, sorting=ynt]{biblatex}
@@
 
+\renewcommand*{\Affilfont}{\footnotesize}
+
 \graphicspath{{./images/}}
@@
 \title{Developing a Bubble Chamber Particle Discriminator Using Semi-Supervised Learning}
-\author{Brendon Matusch, X, Y, and Z}
+\input{authors.tex}
 \date{August 2018}
@@
 The current practice of manually developing a discriminator function to eliminate background events is difficult when available calibration data is frequently impure and present only in small quantities.
-In this study, several different input formats and neural network architectures are applied to the task.
+In this study, several different discriminator input formats and neural network architectures are applied to the task.
 First, they are optimized in a supervised learning context.
-Next, two novel semi-supervised learning algorithms are trained, and found to replicate the Acoustic Parameter discriminator previously used in PICO-60 with a mean of 97\% accuracy.
+Next, two novel semi-supervised learning algorithms are trained, and found to replicate the Acoustic Parameter (AP) discriminator previously used in PICO-60 with a mean of 97\% accuracy.
 \end{abstract}
@@
 
-    \item The primary intention of many dark matter experiments, which is to reduce the number of background events to the lowest level possible, means that there will be a very small number detected.
+    \item A major intent of many dark matter experiments, which is to reduce the number of background events to the lowest level possible, means that there will be a very small number detected.
 This places a significant constraint on the amount of data that is available to optimize such a discriminator.
@@
 The PICO-60 bubble chamber experiment, created for the detection of weakly interacting massive particles (WIMPs) \cite{pico}, was completed in 2017.
-It provides an excellent basis on which to develop and test the efficacy of such techniques as a potential tool for event discrimination, relative to the conventional techniques used during initial analysis of the experiment.
+It provides an excellent basis on which to develop and test the efficacy of new machine learning discriminators, relative to the techniques used during initial analysis of the experiment.
 
@@
 
-In this study, newly developed methodologies were compared to the two discrimination techniques developed previously for the PICO-60 experiment: Acoustic Parameter and machine learning.
+In this study, all newly developed discriminators were compared to the two discrimination techniques used previously for the PICO-60 experiment: Acoustic Parameter and machine learning.
 
@@
 It must also be preprocessed with the position correction function $PosCor(\beta_{8}, \chi)$, which adjusts the overall volume of an audio recording according to its position in the vessel.
-This is done because the acoustic characteristics differ significantly depending on the position of an event in the vessel.
-
-Finally, a set of data cuts must be applied to remove certain classes of events which are not accurately classified using AP. These classes include wall event and multiple-bubble events.
+This is done because the acoustic characteristics differ significantly depending on the position of the event in the vessel.
+
+Finally, a set of data cuts must be applied to remove classes of events which are not accurately classified using AP alone. These classes include wall event and multiple-bubble events.
 
@@
 The direct output from such a DFT is the full-resolution Fourier transform $\beta _{50,001}$, which consists of a sequence of numbers half the length of the raw waveform $\omega$.
-(It is half the length because the output of a DFT of real inputs is Hermitian, meaning there is no new information contained in the second half of the output.)
 
@@
 
-For each event that is captured, the four cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected using the cameras and an entropy-based trigger.
+For each event that is captured, four cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected using the cameras and an image-based trigger.
 These raw images contain a large amount of extraneous information; they encompass the entire vessel.
@@
 This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
+It is also used for $FidWallCut(D)$.
 
@@
     \item Either 0 or 1 bubbles are detected based on images from the camera
-    \item The number of bubbles approximated using the pressure transducer is close to 1
+    \item The number of bubbles approximated using the fast pressure transducer is close to 1
 \end{itemize}
 
-\subsubsection{Wall Cut}
+\subsubsection{Wall Cut} \label{wall_cut}
 
@@
 
-The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the pressure transducer), and the acoustic cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
+The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the fast pressure transducer), and the acoustic cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
 Those cuts restrict data as follows:
@@
 It determines whether events fall within this area based on the visually calculated position $\chi$.
-    \item The pressure cut $PresWallCut(D)$ restricts the pressure detected by the pressure transducer, without any position corrections, to be within a range of 0.3 of 1
+    \item The pressure cut $PresWallCut(D)$ restricts the pressure detected by the fast pressure transducer, without any position corrections, to be within a range of 0.3 of 1
     \item The acoustic cut $AcWallCut(D)$ is defined using the banded Fourier transform $\beta_{8}$, and takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.

### 2018-10-17 documentation/paper.tex
Mostly minor edits

@@
     \item Determine the most effective machine learning architecture to use in development of a discriminator.
-Furthermore, develop several neural network architectures and apply them to different input formats, in order to find which are the most effective.
+Furthermore, develop several neural network architectures and apply them to different input formats (such as audio Fourier transforms, raw waveforms, and images), in order to find which are the most effective.
 
@@
 
-In this study, all newly developed discriminators were compared to the two discrimination techniques used previously for the PICO-60 experiment: Acoustic Parameter and machine learning.
+In this study, all newly developed machine learning-based discriminators were compared to the two discrimination techniques used previously for the PICO-60 experiment: Acoustic Parameter and machine learning.
 
@@
 
-\subsubsection{Machine Learning}
+\subsubsection{Original Machine Learning Analysis}
 
@@
 
-Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original neural analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands.
+Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original machine learning analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands.
 This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.
@@
 
-A certain number of cuts are necessarily applied to all data to ensure meaningful results.
+A number of cuts are necessarily applied to all data to ensure meaningful results.
 Otherwise, significant overfitting on biases in the data is likely.
@@
 Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel.
-It can be desirable for a discriminator to handle these events correctly; however, AP does not, and neither does the neural network used in the previous PICO-60 paper.
+It can be desirable for a discriminator to handle these events correctly; however, AP alone does not, and neither does the neural network used in the previous PICO-60 paper.
 Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.
@@
 
-(Note that step \ref{exponent} only works if $p$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$.
+(Note that step \ref{exponent} only works if $\phi$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$.
 To resolve this, a piecewise exponential function is used instead, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards.
@@
             \hline
-            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% 90\% & 0.78 \\
+            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% & 90\% & 0.78 \\
             \hline
@@
 
-Iterative cluster nucleation (ICN) is a second original semi-supervised learning algorithm developed as part of this study.
+Iterative cluster nucleation (ICN) is a second semi-supervised learning algorithm developed as part of this study.
 It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.

### 2018-10-17 models/banded_frequency_network.py
Fixed input size for position-corrected data

@@ -13,7 +13,7 @@ def create_model() -> Model:
     dropout = 0.25
     activation = 'tanh'
     model = Sequential([
-        InputLayer(input_shape=(19,)),
+        InputLayer(input_shape=(16,)),
         BatchNormalization(),
         Dense(12, activation=activation, kernel_regularizer=regularizer),
         Dropout(dropout),

### 2018-10-17 training/logistic_regression_banded_train.py
Removed broken training data saving

@@ -33,19 +33,11 @@ logistic_regression.fit(training_input, training_ground_truths)
 # Print accuracy on the training and validation sets
 print('Training accuracy:', logistic_regression.score(training_input, training_ground_truths))
 print('Validation accuracy:', logistic_regression.score(validation_input, validation_ground_truths))
-# Run floating-point predictions on the training and validation sets
-training_predictions = logistic_regression.decision_function(training_input)
+# Run floating-point predictions on the validation set
 validation_predictions = logistic_regression.decision_function(validation_input)
-# Expand the arrays, adding a second axis because it's expected by the saving function
-training_predictions = np.expand_dims(training_predictions, axis=1)
+# Expand the array, adding a second axis because it's expected by the saving function
 validation_predictions = np.expand_dims(validation_predictions, axis=1)
-# Save the resulting training and validation sets to disk
-save_test(
-    event_data_set,
-    training_ground_truths,
-    training_predictions,
-    prefix='logistic_training'
-)
+# Save the resulting validation set to disk
 save_test(
     event_data_set,
     validation_ground_truths,

### 2018-10-17 training/logistic_regression_banded_train.py
Load AP predictions in logistic regression script

@@ -22,6 +22,9 @@ event_data_set = EventDataSet(
 )
 # Get the banded frequency domain data and corresponding binary ground truths
 training_input, training_ground_truths, validation_input, validation_ground_truths = event_data_set.banded_frequency_alpha_classification()
+# Get the Acoustic Parameter event predictions so they can be compared to the ground truths
+training_ap_predictions, validation_ap_predictions = [[event.logarithmic_acoustic_parameter > 0.25 for event in event_list]
+                                                      for event_list in [event_data_set.training_events, event_data_set.validation_events]]
 # Add polynomial features up to degree 3 to the input matrices
 polynomial_features = PolynomialFeatures(degree=3)
 training_input = polynomial_features.fit_transform(training_input)

### 2018-10-17 training/banded_frequency_train.py
Some stuff where I try to train with AP as ground truth

@@ -4,6 +4,8 @@
 
 import os
 
+import numpy as np
+
 from data_processing.event_data_set import EventDataSet, RunType
 from data_processing.experiment_serialization import save_test
 from models.banded_frequency_network import create_model
@@ -25,13 +27,16 @@ event_data_set = EventDataSet(
 )
 # Get the banded frequency domain data and corresponding binary ground truths
 training_input, training_ground_truths, validation_input, validation_ground_truths = event_data_set.banded_frequency_alpha_classification()
+# Get the Acoustic Parameter event predictions so they can be compared to the ground truths
+training_ap_predictions, validation_ap_predictions = [[event.logarithmic_acoustic_parameter > 0.5 for event in event_list]
+                                                      for event_list in [event_data_set.training_events, event_data_set.validation_events]]
 
 # Iterate over the defined number of epochs
 for epoch in range(EPOCHS):
     # Train the model on the loaded data set
     model.fit(
         x=training_input,
-        y=training_ground_truths,
+        y=training_ap_predictions,
         validation_data=(validation_input, validation_ground_truths),
         epochs=1
     )

### 2018-10-17 training/banded_frequency_train.py
Switched back to ordinary ground truths

@@ -36,7 +36,7 @@ for epoch in range(EPOCHS):
     # Train the model on the loaded data set
     model.fit(
         x=training_input,
-        y=training_ap_predictions,
+        y=training_ground_truths,
         validation_data=(validation_input, validation_ground_truths),
         epochs=1
     )

### 2018-10-19 documentation/paper.tex
Minor wording changes

@@
 The current practice of manually developing a discriminator function to eliminate background events is difficult when available calibration data is frequently impure and present only in small quantities.
-In this study, several different discriminator input formats and neural network architectures are applied to the task.
+In this study, several different discriminator input/preprocessing formats and neural network architectures are applied to the task.
 First, they are optimized in a supervised learning context.
@@
     \item Determine the most effective machine learning architecture to use in development of a discriminator.
-Furthermore, develop several neural network architectures and apply them to different input formats (such as audio Fourier transforms, raw waveforms, and images), in order to find which are the most effective.
+Furthermore, develop several neural network architectures and apply them to different input/preprocessing formats (such as audio Fourier transforms, raw waveforms, and images), in order to find which are the most effective.
 

### 2018-10-19 documentation/paper.tex
Removed wall cuts and stuff

@@
 
-For some experimental runs, $WallCut(D)$ was not applied. This was done for 2 reasons:
-
-\begin{enumerate}
-    \item As an exploratory effort, to determine whether or not it was possible to produce an effective discriminator without wall cuts. No such attempt was made during the original PICO-60 analysis.
-    \item To test the network's ability to handle complex acoustic characteristics including resonance.
-There is a relatively clear-cut discrimination boundary in PICO-60, but this may not be the case in other applications.
-\end{enumerate}
-
 All supervised learning configurations were trained and evaluated based on their ability to determine the origin of events: whether they are from neutron calibration source runs (predominantly neutrons) or background radiation runs (predominantly alpha particles).
@@
 A convolutional neural network was used for analyzing the raw waveform $\omega$ directly, without any preprocessing whatsoever.
-This avoids any destruction of information; theoretically, a sufficiently complex neural network should be capable of deriving any characteristics needed.
+It is furthermore referred to as $DeepConv(\omega)$.
+Direct processing of the waveform avoids any destruction of information; theoretically, a sufficiently complex neural network should be capable of deriving any characteristics needed.
 
@@
 
-Two different configurations were tested:
-\begin{enumerate}
-    \item A configuration that operates strictly on $\omega$ with no position corrections of any kind (furthermore $DeepConv(\omega)$).
-    \item A configuration with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).
-This configuration was tested both with and without $WallCut(D)$.
-\end{enumerate}
-
 \subsubsection{Results}
 
-During early tests with variations on $DeepConv(\omega)$ without $WallCut(D)$, it was revealed to have a surprising and undesirable capacity to separate run types based on meaningless information.
+During early experiments with shallower network architectures, it was revealed to have a surprising and undesirable capacity to separate run types based on meaningless information.
 It was able to discriminate between events from the AmBe neutron calibration runs and events from alpha background runs with 96\% accuracy, and to separate all combined neutron calibration runs from alpha background runs with 91\% accuracy.
@@
 
-Once the audio was cropped properly, $DeepConv(\omega)$ achieved a maximum of 77\% accuracy on validation data.
-Errors were not balanced at all; precision was 76\% and recall was 94\%, indicating that the network was overzealous to classify particles as neutrons.
-Using the same network architecture (except for the input layer), and still without $WallCut(D)$, $DeepConvWithPos(\omega, \chi)$ achieved 85\%.
-The maximum obtained on a grid search with this configuration was 91\%, although with a relatively high class-wise standard deviation of 0.61.
-Precision improved greatly, reaching 95\%, while recall was reduced, at a maximum of 84\%.
-While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume.
-
-However, there is a lack of decisiveness observed even when accuracy was relatively high, in Figure \ref{waveform_hist}. This implies that, without $WallCut(D)$, this is a suboptimal technique.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{waveform_hist}
-    \caption{\label{waveform_hist} Prediction distribution of the best $DeepConvWithPos(\omega, \chi)$ discriminator without $WallCut(D)$.
-(Validation Event Count is the number of events in the validation set that fall within a certain network prediction band.)}
-\end{figure}
-
-When $WallCut(D)$ was included, both accuracy and class-wise standard deviation improved drastically.
-The best configuration during a grid search was 96\% accurate and had a standard deviation of 0.41 (seen in Figure \ref{waveform_wall_cut_hist}).
+When this extraneous information was cropped out and the architecture was modified, both accuracy and class-wise standard deviation remained quite high.
+The best configuration during a grid search was 96\% accurate and had a standard deviation of 0.41 (seen in Figure \ref{waveform_hist}).
 This is better by both metrics than the multi-layer perceptron used in the original PICO-60 analysis.
@@
     \centering
-    \includegraphics[width=\textwidth]{waveform_wall_cut_hist}
-    \caption{\label{waveform_wall_cut_hist} Prediction distribution of the best $DeepConvWithPos(\omega, \chi)$ discriminator with $WallCut(D)$.}
+    % \includegraphics[width=\textwidth]{waveform_hist}
+    \caption{\label{waveform_hist} Prediction distribution of the best $DeepConv(\omega)$ discriminator.
+(Validation Event Count is the number of events in the validation set that fall within a certain network prediction band.)}
 \end{figure}
 
-The following tables show the results of several $DeepConvWithPos(\omega, \chi)$ discriminators produced throughout testing, with corresponding network hyperparameters.
+The following tables show the results of several $DeepConv(\omega)$ discriminators produced throughout testing, with corresponding network hyperparameters.
 
@@
     \begin{center}
-        \captionof{table}{$DeepConvWithPos(\omega, \chi)$ with $WallCut(D)$.}
+        \captionof{table}{$DeepConv(\omega)$ configurations.}
         \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
@@
             L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Max Accuracy & Precision & Recall & CWSD \\
-            \hline
-            0.003 & 0 & 24 & 3 & 17 & 95\% & 99\% & 94\% & 0.44 \\
-            % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1533824431_epoch30.json
-            \hline
-            0.003 & 0 & 48 & 5 & 17 & 96\% & 99\% & 96\% & 0.41 \\
-            % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1533825716_epoch43.json
-            \hline
-            0.003 & 0.5 & 24 & 5 & 17 & 93\% & 98\% & 92\% & 0.50 \\
-            % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group3time1533828588_epoch27.json
-            \hline
-            0.001 & 0 & 48 & 5 & 17 & 95\% & 98\% & 95\% & 0.48 \\
-            % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.001_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1533832229_epoch16.json
-            \hline
-            0.001 & 0.25 & 48 & 3 & 32 & 93\% & 94\% & 97\% & 0.62 \\
-            % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.001_dense_dropout0.25_first_layer_filters48_kernel_size3_convolutional_layers_per_group6time1533834812_epoch30.json
-            \hline
-            0.0003 & 0.5 & 24 & 5 & 32 & 95\% & 97\% & 97\% & 0.49 \\
-            % ../experimental_data/validation_results/2018_08_09_waveform_grid_search_with_wall_cuts/waveform_grid_search_l2_lambda0.0003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group6time1533849077_epoch45.json
             \hline
@@
 
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{$DeepConvWithPos(\omega, \chi)$ without $WallCut(D)$.}
-        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
-            \hline
-            L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Max Accuracy & Precision & Recall & CWSD \\
-            \hline
-            0.003 & 0 & 24 & 3 & 17 & 91\% & 95\% & 78\% & 0.61 \\
-            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.003_dense_dropout0_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1532800009_epoch78.json
-            \hline
-            0.003 & 0.5 & 24 & 5 & 17 & 91\% & 95\% & 78\% & 0.65 \\
-            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.003_dense_dropout0.5_first_layer_filters24_kernel_size5_convolutional_layers_per_group3time1532813246_epoch49.json
-            \hline
-            0.001 & 0.25 & 24 & 3 & 17 & 91\% & 88\% & 84\% & 0.66 \\
-            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.001_dense_dropout0.25_first_layer_filters24_kernel_size3_convolutional_layers_per_group3time1532825383_epoch88.json
-            \hline
-            0.001 & 0.5 & 48 & 5 & 17 & 88\% & 92\% & 73\% & 0.73 \\
-            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.001_dense_dropout0.5_first_layer_filters48_kernel_size5_convolutional_layers_per_group3time1532837124_epoch0.json
-            \hline
-            0.0003 & 0 & 48 & 5 & 32 & 78\% & 76\% & 56\% & 0.95 \\
-            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.0003_dense_dropout0_first_layer_filters48_kernel_size5_convolutional_layers_per_group6time1532846835_epoch63.json
-            \hline
-            0.0003 & 0.5 & 48 & 5 & 32 & 74\% & 67\% & 53\% & 0.92 \\
-            % ../experimental_data/validation_results/2018_07_28_waveform_grid_search/waveform_grid_search_l2_lambda0.0003_dense_dropout0.5_first_layer_filters48_kernel_size5_convolutional_layers_per_group6time1532863869_epoch10.json
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-
-These tables show that, both with and without $WallCut(D)$, simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
+This table shows that simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
 Larger kernel sizes, more filters, and more layers make performance generally worse, while higher L2 regularization does the opposite.
 Dropout seems to have a weak correlation, possibly reducing accuracy slightly.
-
-It is notable that recall is much, much poorer when wall cuts are applied.
-It seems there are more neutron events that show alpha-like audio signatures in the events near the walls, possibly due to resonance.
 
@@
 
-Once again, configurations with and without position input were tested.
-This time, position corrections were optionally included when position input was not used.
+This time, position corrections were optionally applied to the input data, as they were the input to Acoustic Parameter.
 The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$, $FourierMLP(PosCor(\beta_{N}, \chi))$ and $FourierMLPWithPos(\beta_{N}, \chi)$.
@@
 
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{$FourierMLPWithPos(\beta{N})$ trained without $WallCut(D)$.}
-        \begin{tabular}{|l|l|l|l|l|l|l|l|}
-            \hline
-            Configuration & L2 $\lambda$ & Dropout & Layers & Max Acc & Prec & Recall & CWSD \\
-            \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0 & 0 & 3 & 92\% & 93\% & 77\% & 0.63 \\
-            % ../experimental_data/validation_results/2018_07_20_no_wall_cuts/high_resolution_frequency_time1532116206_epoch111.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0 & 0 & 3 & 95\% & 89\% & 95\% & 0.49 \\
-            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search/high_res_grid_search_dropout0_l2_lambda0_hidden_layers2_time1532446602_epoch210.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0 & 2 & 96\% & 95\% & 93\% & 0.47 \\
-            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search/high_res_grid_search_dropout0_l2_lambda0.001_hidden_layers1_time1532448014_epoch149.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0 & 2 & 97\% & 93\% & 98\% & 0.47 \\
-            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search/high_res_grid_search_dropout0_l2_lambda0.01_hidden_layers1_time1532449765_epoch108.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.25 & 2 & 95\% & 97\% & 86\% & 0.54 \\
-            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search/high_res_grid_search_dropout0.25_l2_lambda0.01_hidden_layers1_time1532454347_epoch122.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.5 & 2 & 94\% & 95\% & 86\% & 0.53 \\
-            % ../experimental_data/validation_results/2018_07_24_high_resolution_frequency_grid_search//high_res_grid_search_dropout0.5_l2_lambda0.01_hidden_layers1_time1532459365_epoch243.json
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-
 It is very clear in Table \ref{fft_wall_cuts} that higher resolutions do not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
@@
 
-In an effort to resolve this, a two-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$.
+In an effort to resolve this, a two-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$, referred to as $ImageConv(\iota)$.
 The network architecture consists of a moderate number of convolutional layers (on the scale of nine) and three dense layers at the end.
 L2 regularization was used on all layers, and dropout was additionally used on the dense layers.
-No position input is used (since there are no microphones that would require position correction), so the basic configuration is $ImageConv(\iota)$.
 
@@
     \begin{center}
-        \captionof{table}{Performance of $ImageConv(\iota)$ configurations, trained without $WallCut(D)$.}
+        \captionof{table}{Performance of $ImageConv(\iota)$ configurations.}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
@@
             L2 $\lambda$ & Dropout & Convolutional Layers & Dense Layers & Max Accuracy & Precision & Recall & CWSD \\
-            \hline
-            0 & 0.5 & 4 & 3 & 77\% & 52\% & 31\% & 1.01 \\
-            % ../experimental_data/validation_results/2018_07_26_image_classification_fixed_loading/image_time1532640532_epoch8.json
-            \hline
-            0 & 0.25 & 4 & 3 & 85\% & 40\% & 38\% & 0.95 \\
-            % ../experimental_data/validation_results/2018_07_26_image_classification_less_dropout_more_epochs/image_time1532615758_epoch164.json
-            \hline
-            0.01 & 0.25 & 6 & 3 & 68\% & 68\% & 100\% & 0.98 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group2time1532895825_epoch2.json
-            \hline
-            0.003 & 0.25 & 9 & 3 & 68\% & 68\% & 100\% & 0.98 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.003_dropout0.25_convolutional_layers_per_group3time1532895118_epoch3.json
-            \hline
-            0.01 & 0.25 & 12 & 3 & 68\% & 68\% & 100\% & 0.97 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group4time1532896126_epoch0.json
-            \hline
-            0.01 & 0.5 & 12 & 3 & 68\% & 68\% & 100\% & 0.99 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.5_convolutional_layers_per_group4time1532898944_epoch1.json
             \hline

### 2018-10-19 training/waveform_grid_search.py
Disabled position input for waveform grid search

@@ -5,9 +5,6 @@
 import os
 import sys
 
-# Use only the first GPU (GTX 1080)
-os.environ['CUDA_VISIBLE_DEVICES'] = '0'
-
 from keras.layers import Conv1D, MaxPooling1D, Flatten, Dropout, Input, BatchNormalization, Dense, concatenate
 from keras.models import Model, Sequential
 from keras.regularizers import l2
@@ -29,7 +26,7 @@ event_data_set = EventDataSet(
 training_inputs, training_ground_truths, validation_inputs, validation_ground_truths = \
     event_data_set.audio_alpha_classification(
         loading_function=load_bubble_audio,
-        include_positions=True
+        include_positions=False
     )
 
 # Iterate over possible configurations for L2 and dropout regularization, number of filters for the first layer (which is scaled up for following layers), kernel size, and number of convolutional layers in most groups
@@ -49,15 +46,15 @@ for l2_lambda in [0.003, 0.001, 0.0003]:
                     print('Kernel Size', kernel_size)
                     print('Convolutional Layers in Most Groups:', convolutional_layers_per_group)
                     # Create a description string which is used for saving validation sets
-                    description = f'waveform_grid_search_l2_lambda{l2_lambda}_dense_dropout{dense_dropout}_first_layer_filters{first_layer_filters}_kernel_size{kernel_size}_convolutional_layers_per_group{convolutional_layers_per_group}'
+                    description = f'waveform_grid_search/l2_lambda{l2_lambda}/dense_dropout{dense_dropout}/first_layer_filters{first_layer_filters}/kernel_size{kernel_size}/convolutional_layers_per_group{convolutional_layers_per_group}'
 
                     # Create a one-dimensional convolutional neural network model with rectified linear activations, using the Keras functional API
                     # It should take both microphone channels and an entire clip of audio, and take the position of the bubble on all 3 axes as a secondary input
                     activation = 'tanh'
                     padding = 'same'
                     regularizer = l2(l2_lambda)
-                    audio_inputs = Input((100_000, 2))
-                    x = BatchNormalization()(audio_inputs)
+                    inputs = Input((100_000, 2))
+                    x = BatchNormalization()(inputs)
                     x = Conv1D(
                         filters=first_layer_filters,
                         kernel_size=80,
@@ -111,16 +108,12 @@ for l2_lambda in [0.003, 0.001, 0.0003]:
                         )(x)
                         x = BatchNormalization()(x)
                     x = Flatten()(x)
[…]

### 2018-10-19 documentation/paper.tex
Scrapped a lot of stuff in the report

@@
 
-$WallCut(D)$ was applied to the low-resolution Fourier transforms during early experiments (since they were used for AP) to minimize the number of variables between initial neural networks and a proven discriminator.
-
 This time, position corrections were optionally applied to the input data, as they were the input to Acoustic Parameter.
-The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$, $FourierMLP(PosCor(\beta_{N}, \chi))$ and $FourierMLPWithPos(\beta_{N}, \chi)$.
+The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and \\ $FourierMLP(PosCor(\beta_{N}, \chi))$.
 Several different resolutions $N$ of $\beta_{N}$ were used.
@@
 
-It became evident very quickly that, without $WallCut(D)$, a high resolution was not required to get a high accuracy relative to the ground truth data.
+It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
 $FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
@@
 
+Very interestingly, $FourierMLP(\beta_{8})$ performed better than $FourierMLP(PosCor(\beta_{8}), \chi))$, which produced an accuracy value of 96\%.
+Both of these results were better than any of the network architectures trained on $\omega$.
+
 \begin{figure}[H]
@@
     \includegraphics[width=\textwidth]{banded_no_pos_input_hist}
-    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best $FourierMLP(\beta_{8})$ discriminator, with $WallCut(D)$.}
-\end{figure}
-
-Very interestingly, $FourierMLP(\beta_{8})$ performed better than either $FourierMLP(PosCor(\beta_{8}), \chi)$ or \\ $FourierMLPWithPos(\beta{8}, \chi)$, which produced accuracy values of 96\% and 97\%, respectively.
-
-Unlike for the convolutional neural network trained on $\omega$, position input did not improve performance at all.
-In fact, $FourierMLP(\beta_{8})$, performed better than $FourierMLPWithPos(\beta_{8}, \chi)$, which reached only 95\% validation accuracy.
-Nevertheless, they both performed better than any of the network architectures trained on $\omega$.
-
-The higher resolution demonstrated significant power when $WallCut(D)$ was not applied, and only \\ $QualCut(MultiCut(D))$ was used.
-$FourierMLPWithPos(\beta_{20}, \chi)$ achieved only 92\% accuracy (2 percentage points lower than the same configuration without $WallCut(D)$).
-However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%.
-However, that epoch had a relatively high class-wise standard deviation, at 0.47, as seen in Figure \ref{high_freq_no_wall_hist}.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{high_freq_no_wall_hist}
-    \caption{\label{high_freq_no_wall_hist} Prediction distribution of the best $FourierMLPWithPos(\beta_{50,001}, \chi)$ discriminator, without $WallCut(D)$.}
+    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best $FourierMLP(\beta_{8})$ discriminator.}
 \end{figure}
@@
     \begin{center}
-        \captionof{table}{Neural networks trained on Fourier transforms with $WallCut(D)$.} \label{fft_wall_cuts}
+        \captionof{table}{Neural networks trained on Fourier transforms.} \label{fft_wall_cuts}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
@@
 
-Without wall cuts, the full-resolution Fourier transform produces much more effective discriminators.
-Here, L2 regularization is beneficial, possibly because there is more noise that can be overfit on.
-
-Unlike with the raw waveform, recall varies relatively little when wall cuts are applied.
-
 \subsection{Convolutional Neural Network for Image Window Analysis}
@@
 Relative to the extremely short period of time in which the bubble forms (on the scale of nanoseconds), the framerate of the camera (340Hz) is extremely slow.
-While the very early stages of bubble formation (when the sound is produced) are known to differ depending on whether the bubble was created by a nuclear recoil or an alpha particle, it is unknown whether any visually apparent differences persist when the bubble is visible.
+While the very early stages of bubble formation (when the sound is produced) are known to differ depending on whether the bubble was created by a nuclear recoil or an alpha particle, it is not experimentally known whether any visually apparent differences persist when the bubble is visible.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{image_hist}
+    % includegraphics[width=\textwidth]{image_hist}
     \caption{\label{image_hist} Prediction distribution of the best $ImageConv(\iota)$ discriminator.}
@@
     \begin{center}
-        \captionof{table}{Summary of supervised learning configurations, trained with $WallCut(D)$.}
+        \captionof{table}{Summary of supervised learning configurations.}
         \begin{tabular}{|l|l|l|l|l|}
@@
             $FourierMLP(\beta_{8})$ & 98\% & 100\% & 98\% & 0.29 \\
-            \hline
-            $FourierMLPWithPos(\beta_{8}, \chi)$ & 97\% & 98\% & 98\% & 0.43 \\
             \hline
@@
 
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Summary of supervised learning configurations, trained without $WallCut(D)$.}
-        \begin{tabular}{|l|l|l|l|l|}
-            \hline
-            Configuration & Max Accuracy & Precision & Recall & CWSD \\
-            \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 92\% & 93\% & 77\% & 0.63 \\
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 97\% & 93\% & 98\% & 0.47 \\
-            \hline
-            $DeepConv(\omega)$ & 77\% & 76\% & 94\% & 0.86 \\
-            % ../experimental_data/validation_results/2018_07_16_waveform_with_new_cuts/waveform_time1531762243_epoch14.json
-            \hline
-            $DeepConvWithPos(\omega, \chi)$ & 91\% & 95\% & 78\% & 0.61 \\
-            \hline
-            $ImageConv(\iota)$ & 85\% & 40\% & 38\% & 0.95 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
+The key conclusion to be drawn from the above tables is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
+Higher resolutions seem to make performance worse, likely due to overfitting.
+It is significantly better than the original PICO-60 neural network, and is more decisive (as measured by the class-wise standard deviation).
 
-There are 2 key conclusions to be drawn from the above tables:
+Based on this result, experimentation with semi-supervised learning proceeded using \\ $FourierMLP(PosCor(\beta_{8}, \chi))$, applying a three-layer perceptron architecture.
+There were two main reasons for this:
 
 \begin{enumerate}
-    \item With $WallCut(D)$, the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
-Higher resolutions seem to make performance worse, likely due to overfitting.
-It is significantly better than the original PICO-60 neural network, and is more decisive (as measured by the class-wise standard deviation)
-    \item Without $WallCut(D)$, full-resolution $\beta_{50,001}$ is the most effective, nearing the best performance with wall cuts at separating run types.
-All models are notably less decisive, given that the lowest standard deviation without wall cuts is 0.47, compared to 0.29 with wall cuts.
-\end{enumerate}
-
-Based on these results, experimentation with semi-supervised learning proceeded using \\ $FourierMLP(PosCor(\beta_{8}))$, applying a 3-layer perceptron architecture, and using $WallCut(D)$.
-There were three main reasons for this:
-
-\begin{enumerate}
-    \item This was a highly effective configuration with wall cuts, producing 97\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
+    \item This was a highly effective configuration, producing 97\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
     \item The use of the exact same input format as AP makes the algorithms more directly comparable.
@@
 Since the intent of these techniques is to separate particle types (the same as AP), performance evaluations (accuracy and class-wise standard deviation) use AP as a baseline.
-Note that $WallCut(D)$ was once again applied for semi-supervised learning, because $FourierMLP(PosCor(\beta_{8}))$ was used for a neural network configuration, and because AP requires application of $WallCut(D)$ to be meaningfully comparable to any neural network.
-
 To ensure that accuracy statistics are as consistent and reliable as possible, every configuration was trained three times, with a different randomly selected validation set each time.
-Accuracy, precision, recall, and class-wise standard deviation statistics are the mean over those 3 tests.
+Accuracy, precision, recall, and class-wise standard deviation statistics are the mean over those three tests.
 

### 2018-10-20 statistics/mean_ground_truth_accuracy_across_configurations.py
Print out hyperparameters for epoch with best val accuracy

@@ -13,6 +13,9 @@ RUNS_PER_CONFIGURATION = 3
 # The indices corresponding to the lines containing (mean and maximum) (training and validation) validation accuracy values within a run
 LINE_INDICES = range(7, 11)
 
+# The index containing the path to the saved validation set (including the hyperparameters)
+HYPERPARAMETER_LINE_INDEX = 11
+
 # Read all lines from standard input
 input_lines = sys.stdin.readlines()
 # Calculate the total number of runs, based on the known number of lines per run
@@ -27,6 +30,10 @@ for configuration_index in range(num_runs // RUNS_PER_CONFIGURATION):
     max_training_accumulator = 0
     mean_validation_accumulator = 0
     max_validation_accumulator = 0
+    # Create a variable to store the hyperparameter path in
+    hyperparameter_path = None
+    # Store a corresponding value for the maximum validation accuracy
+    max_validation_accuracy_for_hyperparameter_path = 0
     # Iterate over the runs within that configuration
     for run_index in range(RUNS_PER_CONFIGURATION):
         # Calculate the starting index of this run
@@ -38,11 +45,17 @@ for configuration_index in range(num_runs // RUNS_PER_CONFIGURATION):
         max_training_accumulator += max_training
         mean_validation_accumulator += mean_validation
         max_validation_accumulator += max_validation
-    # Calculate the average accuracy statistics for this configuration, and output them to the user
-    print(f'Average mean training accuracy for configuration {configuration_index}: {mean_training_accumulator / RUNS_PER_CONFIGURATION}')
-    print(f'Average maximum training accuracy for configuration {configuration_index}: {max_training_accumulator / RUNS_PER_CONFIGURATION}')
-    print(f'Average mean validation accuracy for configuration {configuration_index}: {mean_validation_accumulator / RUNS_PER_CONFIGURATION}')
-    print(f'Average maximum validation accuracy for configuration {configuration_index}: {max_validation_accumulator / RUNS_PER_CONFIGURATION}')
+        # If the maximum validation accuracy on this run is better than the previous best
+        if max_validation > max_validation_accuracy_for_hyperparameter_path:
+            # Get the path containing the hyperparameters from the defined line and store it in the variable
+            hyperparameter_path = input_lines[run_starting_index + HYPERPARAMETER_LINE_INDEX].split()[-1]
+            # Also update the maximum validation accuracy, so if future runs are less accurate they will not overwrite the path
[…]

### 2018-10-20 training/high_resolution_grid_search.py
Removed position input from high resolution grid search

@@ -23,7 +23,7 @@ event_data_set = EventDataSet(
 training_inputs, training_ground_truths, validation_inputs, validation_ground_truths = \
     event_data_set.audio_alpha_classification(
         loading_function=lambda bubble: load_bubble_frequency_domain(bubble, banded=False),
-        include_positions=True
+        include_positions=False
     )
 
 # Iterate over the valid configurations for dropout, L2 lambda, and number of hidden layers
@@ -39,17 +39,13 @@ for dropout in [0, 0.25, 0.5]:
             print('L2 Lambda:', l2_lambda)
             print('Number of Hidden Layers:', hidden_layers)
             # Create a description string containing the hyperparameters, which is used when saving test data
-            description = f'high_res_grid_search_dropout{dropout}_l2_lambda{l2_lambda}_hidden_layers{hidden_layers}_'
+            description = f'high_res_grid_search/dropout{dropout}/l2_lambda{l2_lambda}/hidden_layers{hidden_layers}'
 
             # Create a neural network model that includes several dense layers with hyperbolic tangent activations, L2 regularization, and batch normalization
             regularizer = l2(l2_lambda)
             activation = 'tanh'
-            # Create two inputs, one for the audio data and one for the position, and concatenate them together
-            audio_input = Input((100_002,))
-            # Take a separate input for the position, and concatenate it with the audio input
-            position_input = Input((3,))
-            x = concatenate([audio_input, position_input])
-            x = BatchNormalization()(x)
+            inputs = Input((100_002,))
+            x = BatchNormalization()(inputs)
             # If there are 3 hidden layers, include an earlier one with 16 neurons
             if hidden_layers == 3:
                 x = Dense(16, activation=activation, kernel_regularizer=regularizer)(x)
@@ -62,8 +58,7 @@ for dropout in [0, 0.25, 0.5]:
                 x = Dense(8, activation=activation, kernel_regularizer=regularizer)(x)
                 x = Dropout(dropout)(x)
             output = Dense(1, activation='sigmoid', kernel_regularizer=regularizer)(x)
-            # Create a model with both inputs
-            model = Model(inputs=[audio_input, position_input], outputs=output)
+            model = Model(inputs=inputs, outputs=output)
             # Output a summary of the model's architecture
             print(model.summary())
             # Use a mean squared error loss function and an Adam optimizer, and print the accuracy while training

### 2018-10-20 training/image_classification_grid_search.py
Deleted GPU selection from image classification grid search script

@@ -4,9 +4,6 @@
 
 import os
 
-# Use only the second GPU for training (GTX 1060)
-os.environ['CUDA_VISIBLE_DEVICES'] = '1'
-
 from keras.layers import Conv2D, Flatten, Dropout, InputLayer, BatchNormalization, Dense
 from keras.models import Model, Sequential
 from keras.regularizers import l2

### 2018-10-20 documentation/paper.tex
Added data table for updated waveform grid search
Included a ton and a half of citations

@@
 
+For stochastic optimization of the loss function while training, the Adam \cite{adam} optimizer was used where not otherwise specified.
+
 \section{Supervised Learning}
@@
 When this extraneous information was cropped out and the architecture was modified, both accuracy and class-wise standard deviation remained quite high.
-The best configuration during a grid search was 96\% accurate and had a standard deviation of 0.41 (seen in Figure \ref{waveform_hist}).
+The most accurate configuration during a grid search was 95\% accurate and had a standard deviation of 0.52 (seen in Figure \ref{waveform_hist}).
 This is better by both metrics than the multi-layer perceptron used in the original PICO-60 analysis.
@@
     \centering
-    % \includegraphics[width=\textwidth]{waveform_hist}
+    \includegraphics[width=\textwidth]{waveform_hist}
     \caption{\label{waveform_hist} Prediction distribution of the best $DeepConv(\omega)$ discriminator.
@@
             L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Max Accuracy & Precision & Recall & CWSD \\
+            \hline
+            0.003 & 0 & 24 & 3 & 17 & 95\% & 97\% & 97\% & 0.52 \\
+            % ../experimental_data/validation_results/waveform_grid_search/l2_lambda0.003/dense_dropout0/first_layer_filters24/kernel_size3/convolutional_layers_per_group3/time1539990193_epoch57.json
+            \hline
+            0.003 & 0.25 & 48 & 3 & 17 & 94\% & 100\% & 91\% & 0.44 \\
+            \hline
+            0.001 & 0.25 & 24 & 5 & 17 & 95\% & 99\% & 94\% & 0.46 \\
+            \hline
+            0.001 & 0.25 & 24 & 5 & 32 & 91\% & 94\% & 94\% & 0.66 \\
+            \hline
+            0.0003 & 0 & 48 & 5 & 32 & 88\% & 94\% & 89\% & 0.61 \\
+            \hline
+            0.0003 & 0 & 24 & 3 & 17 & 93\% & 96\% & 95\% & 0.59 \\
             \hline
@@
 The gradients produced by this technique are used for training in the same way as gradients calculated with a loss function.
-Backpropagation is used to calculate gradients for previous layers, and a stochastic gradient descent optimizer is used for weight updates.
+Backpropagation is used to calculate gradients for previous layers, and a stochastic gradient descent optimizer \cite{sgd} is used for weight updates.
 
@@
 During a grid search of parameters of the gravitational differentiation algorithm (in addition to the stochastic gradient descent learning rate), the highest mean accuracy over three runs of an individual configuration was 99\%.
-The highest accuracy was obtained with 256 labeled examples in the set $\varsigma$ (out of 496 total training examples).
+The highest accuracy was obtained with 256 labeled examples in the set $\varsigma$ (out of 496 total training examples), which equates to 52\% of the data being labeled.
 High accuracy statistics in this range were also obtained using 128 examples (99\% accuracy, once again) and 64 examples (97\% accuracy).
@@
         \captionof{table}{Gravitational differentiation accuracy statistics by size of $\varsigma$.} \label{grav_accuracy_by_varsigma}
-        \begin{tabular}{|l|l|l|}
+        \begin{tabular}{|l|l|l|l|}
             \hline
-            Initial $\varsigma$ Size & Max Accuracy & Mean Accuracy \\
+            Initial $\varsigma$ Size & Proportion of Data Labeled & Max Accuracy & Mean Accuracy \\
             \hline
-            32 & 83\% & 69\% \\
+            32 & 6\% & 83\% & 69\% \\
             \hline
-            64 & 97\% & 73\% \\
+            64 & 13\% & 97\% & 73\% \\
             \hline
-            128 & 99\% & 84\% \\
+            128 & 26\% & 99\% & 84\% \\
             \hline
-            256 & 99\% & 94\% \\
+            256 & 52\% & 99\% & 94\% \\
             \hline
@@
         \captionof{table}{Iterative cluster nucleation accuracy statistics by initial size of $\varsigma$.} \label{icn_accuracy_by_varsigma}
-        \begin{tabular}{|l|l|l|}
+        \begin{tabular}{|l|l|l|l|}
             \hline
-            Initial $\varsigma$ Size & Max Accuracy & Mean Accuracy \\
+            Initial $\varsigma$ Size & Proportion of Data Labeled & Max Accuracy & Mean Accuracy \\
             \hline
-            32 & 95\% & 75\% \\
+            32 & 6\% & 95\% & 75\% \\
             \hline
-            64 & 97\% & 84\% \\
+            64 & 13\% & 97\% & 84\% \\
             \hline
-            128 & 98\% & 95\% \\
+            128 & 26\% & 98\% & 95\% \\
             \hline
-            256 & 99\% & 97\% \\
+            256 & 52\% & 99\% & 97\% \\
             \hline
@@
 Keras \cite{keras}, running on a TensorFlow backend, was used for all machine learning tasks.
-NumPy and SciPy were used for linear algebra and signal processing.
-ROOT, scikit-image and scikit-learn were used for data loading and storage.
+NumPy \cite{numpy} and SciPy \cite{scipy} were used for linear algebra and signal processing.
+ROOT \cite{root}, scikit-image \cite{scikit-image}, and scikit-learn \cite{scikit-learn} were used for data loading and storage.
 Matplotlib was used for data visualization.

### 2018-10-20 training/image_classification_grid_search.py
Used proper path format

@@ -42,7 +42,7 @@ for dropout in [0, 0.25, 0.5]:
             print('Dense Dropout:', dropout)
             print('Convolutional Layers Per Group:', convolutional_layers_per_group)
             # Create a description string which is used for saving validation sets
-            description = f'image_grid_search_l2_lambda{l2_lambda}_dropout{dropout}_convolutional_layers_per_group{convolutional_layers_per_group}'
+            description = f'image_grid_search/l2_lambda{l2_lambda}/dropout{dropout}/convolutional_layers_per_group{convolutional_layers_per_group}'
 
             # Calculate the number of images there are stacked along the channels axis
             channels = END_IMAGE_INDEX - START_IMAGE_INDEX

### 2018-10-20 documentation/paper.tex
Added back image grid search that was actually done with wall cuts

@@
 Its very decisive prediction distribution can be observed in Figure \ref{banded_no_pos_input_hist}.
-Increasing the number of bands to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.
+Increasing the number of bands to 20 did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.
 Using 40 bands reduced the accuracy once again, to 91\%.
@@
 
-Throughout a variety of different network architectures, the best validation accuracy obtained was a deceptively high 85\%.
+Throughout a variety of different network architectures, the best validation accuracy obtained was 69\%.
 The corresponding training accuracy of 100\% indicates that severe overfitting is taking place (not unexpected, given the relatively small image data set).
-Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random (as seen in Figure \ref{image_hist}).
-It is likely fitting on some form of noise within $\iota$.
-The fact that throughout many trials (including a grid search), no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.
+Also, the class-wise standard deviation was a very high 0.99, which implies its decisions are nearly random (as seen in Figure \ref{image_hist}).
+It is likely fitting poorly on some form of noise within $\iota$.
+The fact that throughout many trials during a grid search, no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.
 
@@
     \centering
-    % includegraphics[width=\textwidth]{image_hist}
-    \caption{\label{image_hist} Prediction distribution of the best $ImageConv(\iota)$ discriminator.}
+    \includegraphics[width=\textwidth]{image_hist}
+    \caption{\label{image_hist} Prediction distribution of one of the best $ImageConv(\iota)$ discriminators.}
 \end{figure}
@@
     \begin{center}
-        \captionof{table}{Performance of $ImageConv(\iota)$ configurations.}
+        \captionof{table}{Performance of $ImageConv(\iota)$ configurations, trained without $WallCut(D)$.}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
@@
             L2 $\lambda$ & Dropout & Convolutional Layers & Dense Layers & Max Accuracy & Precision & Recall & CWSD \\
+            \hline
+            0 & 0 & 6 & 3 & 69\% & 69\% & 100\% & 0.99 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0_dropout0_convolutional_layers_per_group2time1532891695_epoch15.json 
+            \hline
+            0.01 & 0.25 & 6 & 3 & 68\% & 68\% & 100\% & 0.98 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group2time1532895825_epoch2.json
+            \hline
+            0.003 & 0.25 & 9 & 3 & 68\% & 68\% & 100\% & 0.98 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.003_dropout0.25_convolutional_layers_per_group3time1532895118_epoch3.json
+            \hline
+            0.01 & 0.25 & 12 & 3 & 68\% & 68\% & 100\% & 0.97 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group4time1532896126_epoch0.json
+            \hline
+            0.01 & 0.5 & 12 & 3 & 68\% & 68\% & 100\% & 0.99 \\
+            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.5_convolutional_layers_per_group4time1532898944_epoch1.json
             \hline
@@
 The high standard deviations across the board are further evidence for the hypothesis that there is insufficient information in the image data.
-
 The 100\% recall statistics are not due to high classification accuracy, but rather due to the network making predictions less than 0.5 for all examples of both classes.

### 2018-10-20 models/high_resolution_frequency_network.py
Removed position input and reverted to 20 bands

@@ -10,21 +10,17 @@ def create_model() -> Model:
     """Create and return a new instance of the high-resolution frequency network"""
     # Create a neural network composed of dense layers with dropout and L2 regularization, using hyperbolic tangent activations
     activation = 'tanh'
-    regularizer = l2(0.0003)
+    regularizer = l2(0)
     dropout = 0
     # Create two inputs, one for the audio data and one for the position, and concatenate them together
-    audio_input = Input((100_002,))
-    # Take a separate input for the position, and concatenate it with the audio input
-    position_input = Input((3,))
-    x = concatenate([audio_input, position_input])
-    x = BatchNormalization()(x)
+    inputs = Input((40,))
+    x = BatchNormalization()(inputs)
     x = Dense(12, activation=activation, kernel_regularizer=regularizer)(x)
     x = Dropout(dropout)(x)
     x = Dense(8, activation=activation, kernel_regularizer=regularizer)(x)
     x = Dropout(dropout)(x)
     output = Dense(1, kernel_regularizer=regularizer)(x)
-    # Create a model with both inputs
-    model = Model(inputs=[audio_input, position_input], outputs=output)
+    model = Model(inputs=inputs, outputs=output)
     # Output a summary of the model's architecture
     print(model.summary())
     # Use a mean squared error loss function and an Adam optimizer, and print the accuracy while training

### 2018-10-20 training/high_resolution_frequency_train.py

@@ -23,8 +23,8 @@ event_data_set = EventDataSet(
 training_inputs, training_ground_truths, validation_inputs, validation_ground_truths = \
     event_data_set.audio_alpha_classification(
         loading_function=lambda bubble:
-        load_bubble_frequency_domain(bubble, banded=False),
-        include_positions=True
+        load_bubble_frequency_domain(bubble, banded=True),
+        include_positions=False
     )
 # Create an instance of the high resolution frequency network
 model = create_model()
@@ -54,6 +54,3 @@ for epoch in range(250):
         epoch,
         prefix='high_resolution_frequency_'
     )
-    # Save the current model, named with the epoch number
-    model_path = os.path.expanduser(f'~/frequency_epoch{epoch}.h5')
-    model.save(model_path)

### 2018-10-20 data_processing/bubble_data_point.py
Switched to 40 bands; failure

@@ -33,26 +33,46 @@ CHANNELS = 8
 # The separations between the bands (in Hz) to split the frequency domain representation of each time band into
 FREQUENCY_BANDS = [
     0,
+    0.75e3,
     1.5e3,
+    2.5e3,
     3.5e3,
+    4.25e3,
     5e3,
+    6.25e3,
     7.5e3,
+    8.25e3,
     9e3,
+    1e4,
     1.25e4,
+    1.45e4,
     1.6e4,
+    1.8e4,
     2e4,
+    2.3e4,
     2.75e4,
+    3.2e4,
     3.5e4,
+    4e4,
     4.5e4,
+    5.25e4,
     6e4,
+    7e4,
     8e4,
+    9e4,
     1e5,
+    1.1e5,
     1.25e5,
+    1.35e5,
     1.5e5,
+    1.55e5,
     1.6e5,
+    1.675e5,
     1.75e5,
+    1.8e5,
     1.85e5,
-    2.0e5
+    1.925e5,
+    2e5
 ]
 
 # The number of audio samples per second in raw recordings

### 2018-10-20 models/high_resolution_frequency_network.py

@@ -11,9 +11,9 @@ def create_model() -> Model:
     # Create a neural network composed of dense layers with dropout and L2 regularization, using hyperbolic tangent activations
     activation = 'tanh'
     regularizer = l2(0)
-    dropout = 0
+    dropout = 0.001
     # Create two inputs, one for the audio data and one for the position, and concatenate them together
-    inputs = Input((40,))
+    inputs = Input((80,))
     x = BatchNormalization()(inputs)
     x = Dense(12, activation=activation, kernel_regularizer=regularizer)(x)
     x = Dropout(dropout)(x)

### 2018-10-20 documentation/paper.tex
Added high res grid search to paper
Fixed all the broken data!

@@
 This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.
-
-This process was generalized to several other banding resolutions. Higher-resolution counterparts of $\beta_{8}$ included $\beta_{20}$ and $\beta_{40}$.
 
@@
 
-In addition to the Fourier transform used in the original PICO-60 analysis, higher resolutions were computed by similar means.
-These included $\beta _{20}$ and $\beta _{40}$.
-Also, the full-resolution $\beta _{50,001}$ was input into a neural network directly, without any downscaling.
+In addition to the banded Fourier transform $\beta_{8}$ used in the original PICO-60 analysis, the full-resolution $\beta _{50,001}$ was input into a neural network directly, without any downscaling.
 
-This time, position corrections were optionally applied to the input data, as they were the input to Acoustic Parameter.
+When $\beta_{8}$ was used as input, position corrections were optionally applied to the input data, as they were the input to Acoustic Parameter.
 The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and \\ $FourierMLP(PosCor(\beta_{N}, \chi))$.
-Several different resolutions $N$ of $\beta_{N}$ were used.
 While a variety of network architectures were tested, most had a small number of layers (on the order of three), applied dropout and L2 regularization, and used batch normalization once again.
@@
 Its very decisive prediction distribution can be observed in Figure \ref{banded_no_pos_input_hist}.
-Increasing the number of bands to 20 did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.
-Using 40 bands reduced the accuracy once again, to 91\%.
 
@@
             \hline
-            $FourierMLP(PosCor(\beta_{8}))$ & 0 & 0.5 & 2 & 96\% & 98\% & 97\% & 0.42 \\
+            $FourierMLP(PosCor(\beta_{8}, \chi))$ & 0 & 0.5 & 2 & 96\% & 98\% & 97\% & 0.42 \\
             % ../experimental_data/validation_results/2018_07_16_banded_frequency_validation_passing_weighted_3_times.json
@@
             \hline
-            $FourierMLPWithPos(\beta_{8}, \chi)$ & 0 & 0.5 & 3 & 95\% & 100\% & 92\% & 0.44 \\
-            % ../experimental_data/validation_results/2018_07_17_banded_frequency_not_position_corrected_with_position_input/banded_low_resolution_time1531837503_epoch18.json
+            $FourierMLP(\beta_{50,001})$ & 0 & 0 & 2 & 95\% & 94\% & 98\% & 0.47 \\
+            % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input//dropout0/l2_lambda0/hidden_layers1/time1540053737_epoch170.json
             \hline
-            $FourierMLPWithPos(\beta_{8}, \chi)$ & 0 & 0.25 & 3 & 97\% & 98\% & 98\% & 0.43 \\
-            % ../experimental_data/validation_results/2018_07_18_banded_frequency_25\%_dropout/banded_low_resolution_time1531922044_epoch154.json
+            $FourierMLP(\beta_{50,001})$ & 0.001 & 0 & 4 & 95\% & 95\% & 98\% & 0.46 \\
+            % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0/l2_lambda0.001/hidden_layers3/time1540054370_epoch174.json
             \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0.003 & 0.25 & 3 & 93\% & 99\% & 92\% & 0.54 \\
-            % ../experimental_data/validation_results/2018_07_20_20_bands/high_resolution_frequency_time1532105605_epoch90.json
+            $FourierMLP(\beta_{50,001})$ & 0.001 & 0.5 & 4 & 94\% & 93\% & 98\% & 0.54 \\
+            % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0.5/l2_lambda0.001/hidden_layers3/time1540056735_epoch50.json
             \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 0 & 0 & 3 & 94\% & 99\% & 93\% & 0.52 \\
-            % ../experimental_data/validation_results/2018_07_20_20_bands_no_regularization/high_resolution_frequency_time1532113464_epoch172.json
-            \hline
-            $FourierMLPWithPos(\beta_{40}, \chi)$ & 0 & 0 & 3 & 91\% & 98\% & 90\% & 0.57 \\
-            % ../experimental_data/validation_results/2018_07_20_40_bands/high_resolution_frequency_time1532107386_epoch101.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.0003 & 0 & 4 & 91\% & 94\% & 93\% & 0.53 \\
-            % ../experimental_data/validation_results/2018_08_10_high_resolution_grid_search_with_wall_cuts/high_res_grid_search_dropout0_l2_lambda0.0003_hidden_layers3_time1533922958_epoch133.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0.25 & 4 & 91\% & 96\% & 89\% & 0.56 \\
-            % ../experimental_data/validation_results/2018_08_10_high_resolution_grid_search_with_wall_cuts/high_res_grid_search_dropout0.25_l2_lambda0.001_hidden_layers3_time1533924257_epoch115.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.001 & 0.5 & 3 & 89\% & 91\% & 91\% & 0.64 \\
-            % ../experimental_data/validation_results/2018_08_10_high_resolution_grid_search_with_wall_cuts/high_res_grid_search_dropout0.5_l2_lambda0.001_hidden_layers2_time1533925299_epoch169.json
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 0.01 & 0.5 & 4 & 89\% & 90\% & 93\% & 0.65 \\
-            % ../experimental_data/validation_results/2018_08_10_high_resolution_grid_search_with_wall_cuts/high_res_grid_search_dropout0.5_l2_lambda0.01_hidden_layers3_time1533925799_epoch91.json
+            $FourierMLP(\beta_{50,001})$ & 0.003 & 0.25 & 3 & 93\% & 92\% & 98\% & 0.57 \\
+            % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0.25/l2_lambda0.003/hidden_layers2/time1540055639_epoch113.json
             \hline
@@
 
-It is very clear in Table \ref{fft_wall_cuts} that higher resolutions do not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
+It is very clear in Table \ref{fft_wall_cuts} that higher resolution data does not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
 

### 2018-10-20 visualization/graph_wall_event_positions.py
Added axis labels to wall event position graph

@@ -52,7 +52,7 @@ fails_audio_line = Line2D([], [], marker='x', label='Fails audio and/or pressure
 # Display them in a legend
 plt.legend(handles=[passes_fiducial_patch, fails_fiducial_patch, passes_audio_line, fails_audio_line])
 # Label the X and Y axes
-plt.xlabel('X Position')
-plt.ylabel('Y Position')
+plt.xlabel('X Position (mm)')
+plt.ylabel('Y Position (mm)')
 # Display the graph on screen
 plt.show()

### 2018-10-20 documentation/paper.tex
Added date range of run 2 events
Labeled units and used log scale for audio graph

@@
     \includegraphics[width=\textwidth]{audio}
-    \caption{\label{} An example of the cropped audio waveform $\omega$ and full resolution Fourier transform $\beta_{50,001}$.}
+    \caption{\label{} An example of the cropped audio waveform $\omega$ and full resolution Fourier transform $\beta_{50,001}$ (recorded on 2016-12-02, run 1, event 21).}
 \end{figure}
@@
 
-The initial data set $D$ to which these cuts are applied consists of all events recorded during PICO-60 run 2.
+The initial data set $D$ to which these cuts are applied consists of all events recorded during PICO-60 run 2, between 2016-09-02 and 2017-06-20.
 

### 2018-10-20 visualization/visualize_audio.py

@@ -27,14 +27,23 @@ def visualize_and_save_audio(time_domain: np.ndarray, frequency_domain: np.ndarr
         time_domain_channel = time_domain[:, channel_index]
         plt.subplot(channels, 2, time_domain_plot_index)
         plt.plot(time_domain_channel)
-        plt.title(f'Time Domain, Channel {channel_index}')
+        plt.xlabel('Sample Time ($\mu$s)')
+        plt.ylabel(f'Piezo Sample Value (Channel {channel_index})')
         # Graph the audio in frequency domain in the plot to the right
         frequency_domain_plot_index = time_domain_plot_index + 1
         frequency_domain_channel = frequency_domain[:, channel_index]
+        # Get the frequencies corresponding to the FFT values, so they can be plotted on the X axis
+        frequencies = np.fft.rfftfreq(len(time_domain_channel))
+        # Multiply them by one million to get the values in Hz
+        frequencies_hz = frequencies * 1_000_000
         plt.subplot(channels, 2, frequency_domain_plot_index)
-        plt.plot(frequency_domain_channel)
-        plt.title(f'Frequency Domain, Channel {channel_index}')
-
+        plt.plot(frequencies_hz, frequency_domain_channel)
+        plt.xlabel('Frequency (Hz)')
+        plt.ylabel(f'Fourier Transform (Magnitude) [Channel {channel_index}]')
+        # Use a logarithmic scale for the X axis (because it corresponds to frequency)
+        plt.xscale('log')
+        # Set the Y scale manually (for some reason, the automatic one behaves strangely)
+        plt.ylim(-50_000, 250_000)
         # Take the time domain for this channel, save it as a WAV file, and notify the user
         file_path = os.path.expanduser(f'~/channel_{channel_index}.wav')
         wavfile.write(file_path, SAMPLE_RATE, time_domain_channel)

### 2018-10-21 visualization/graph_binary_predictions.py
Switched back to AP for ground truth

@@ -17,7 +17,7 @@ from utilities.verify_arguments import verify_arguments
 verify_arguments('JSON data file')
 
 # Load the data set from the file, ignoring the run type ground truths
-events, ground_truths, network_outputs = load_test(sys.argv[1])
+events, _, network_outputs = load_test(sys.argv[1])
 # Get the acoustic parameter and neural network score data from the events, if they are present (if one is, both will be)
 if hasattr(events[0], 'logarithmic_acoustic_parameter'):
     acoustic_parameters, original_neural_network_scores = zip(
@@ -30,7 +30,7 @@ else:
     acoustic_parameters = random_values
     original_neural_network_scores = random_values
 # Calculate actual neutron/alpha ground truths based on AP
-_ = np.array(acoustic_parameters) > 0.25
+ground_truths = np.array(acoustic_parameters) > 0.25
 
 # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
 for criterion_data, criterion_name, classification_threshold in zip(

### 2018-10-21 documentation/paper.tex
Made the last of the edits for now!

@@
 This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
-It is also used for $FidWallCut(D)$.
+It is also used for {\it FidWallCut}$(D)$.
 
@@
 Otherwise, significant overfitting on biases in the data is likely.
-This basic quality cut $QualCut(D)$ consists of the following restrictions:
+This basic quality cut {\it QualCut}$(D)$ consists of the following restrictions:
 
@@
 
-The complete wall cut $WallCut(D)$ is a composition of the fiducial cut $FidWallCut(D)$ (which makes use of the 3D position $\chi$), the pressure cut $PresWallCut(D)$ (which uses data from the fast pressure transducer), and the acoustic cut $AcWallCut(D)$ (which uses the banded Fourier transform $\beta _{8}$).
+The complete wall cut {\it WallCut}$(D)$ is a composition of the fiducial cut {\it FidWallCut}$(D)$ (which makes use of the 3D position $\chi$), the pressure cut {\it PresWallCut}$(D)$ (which uses data from the fast pressure transducer), and the acoustic cut {\it AcWallCut}$(D)$ (which uses the banded Fourier transform $\beta _{8}$).
 Those cuts restrict data as follows:
@@
 \begin{itemize}
-    \item The fiducial cut $FidWallCut(D)$ defines a spatial area along the walls of the vessel within which no events are accepted.
+    \item The fiducial cut {\it FidWallCut}$(D)$ defines a spatial area along the walls of the vessel within which no events are accepted.
 It determines whether events fall within this area based on the visually calculated position $\chi$.
-    \item The pressure cut $PresWallCut(D)$ restricts the pressure detected by the fast pressure transducer, without any position corrections, to be within a range of 0.3 of 1
-    \item The acoustic cut $AcWallCut(D)$ is defined using the banded Fourier transform $\beta_{8}$, and takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
+    \item The pressure cut {\it PresWallCut}$(D)$ restricts the pressure detected by the fast pressure transducer, without any position corrections, to be within a range of 0.3 of 1
+    \item The acoustic cut {\it AcWallCut}$(D)$ is defined using the banded Fourier transform $\beta_{8}$, and takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
 \end{itemize}
@@
 Given $n$ different hyperparameters to optimize and $m$ different possible values for each, $m^n$ different networks are trained (using every possible combination of hyperparameters) and tested on a validation set.
+For space reasons, only a small subset of the tested configurations are displayed in tables throughout this paper.
+
+In general, the parameters for each grid search were chosen based on some initial empirical experimentation to determine the general range within the entire parameter space that work reasonably well.
+There was a limit to how many parameters could be tested, especially in the case of semi-supervised learning, as some of the grid searches took several days to run.
 
@@
     \item In general, supervised learning systems learn to replicate biases present in the training data set.
-Thus, to evaluate the network's effectiveness at processing the input data, the run type (which is used for training) must be used as a baseline instead of AP.
+Thus, to evaluate the network's effectiveness at processing the input data, the run type (which is used for training) must be used as a baseline instead of AP (which is never provided to the network during the training process).
 \end{enumerate}
@@
 L2 regularization was used to alleviate overfitting, and batch normalization was applied to the input to stabilize training.
-The network's hyperparameters were modified significantly through manual empirical testing as well as grid searches.
+
+The following hyperparameters were tested during a grid search:
+\begin{itemize}
+    \item L2 Regularization $\lambda \in \{0.0003, 0.001, 0.003\}$
+    \item Dense Layer Dropout $\in \{0, 0.25, 0.5\}$
+    \item Convolutional Filters $\in \{24, 48\}$
+    \item Convolutional Kernel Size $\in \{3, 5\}$
+    \item Number of Convolutional Layers $\in \{17, 32\}$
+\end{itemize}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
     \centering
     \includegraphics[width=\textwidth]{waveform_hist}
-    \caption{\label{waveform_hist} Prediction distribution of the best $DeepConv(\omega)$ discriminator.
+    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ discriminator.
 (Validation Event Count is the number of events in the validation set that fall within a certain network prediction band.)}
@@
 
-The following tables show the results of several $DeepConv(\omega)$ discriminators produced throughout testing, with corresponding network hyperparameters.
+The following tables show the results of several {\it DeepConv}$(\omega)$ discriminators produced throughout testing, with corresponding network hyperparameters.
 
@@
     \begin{center}
-        \captionof{table}{$DeepConv(\omega)$ configurations.}
+        \captionof{table}{Performance of {\it DeepConv}$(\omega)$ configurations.}
         \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
@@
 
+The following hyperparameters were tested during a grid search:
+\begin{itemize}
+    \item Dropout $\in \{0, 0.25, 0.5\}$
+    \item L2 Regularization $\lambda \in \{0, 0.0003, 0.001, 0.003, 0.01\}$
+    \item Number of Layers $\in \{2, 3, 4\}$
+\end{itemize}
+
 \subsubsection{Results}
@@
 It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
-$FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
+{\it FourierMLP}$(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
 Its very decisive prediction distribution can be observed in Figure \ref{banded_no_pos_input_hist}.
 
-Very interestingly, $FourierMLP(\beta_{8})$ performed better than $FourierMLP(PosCor(\beta_{8}), \chi))$, which produced an accuracy value of 96\%.
+Very interestingly, {\it FourierMLP}$(\beta_{8})$ performed better than {\it FourierMLP}$(${\it PosCor}$(\beta_{8}), \chi))$, which produced an accuracy value of 96\%.
 Both of these results were better than any of the network architectures trained on $\omega$.
 
-\begin{figure}[H]
+\begin{figure}[h]
     \centering
     \includegraphics[width=\textwidth]{banded_no_pos_input_hist}
-    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best $FourierMLP(\beta_{8})$ discriminator.}
+    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best {\it FourierMLP}$(\beta_{8})$ discriminator.}
 \end{figure}
@@
     \begin{center}
-        \captionof{table}{Neural networks trained on Fourier transforms.} \label{fft_wall_cuts}
+        \captionof{table}{Performance of {\it FourierMLP}$(\beta_{N})$ configurations.} \label{fft_wall_cuts}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
@@
             \hline
-            $FourierMLP(PosCor(\beta_{8}, \chi))$ & 0 & 0.5 & 2 & 96\% & 98\% & 97\% & 0.42 \\
+            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ & 0 & 0.5 & 2 & 96\% & 98\% & 97\% & 0.42 \\
             % ../experimental_data/validation_results/2018_07_16_banded_frequency_validation_passing_weighted_3_times.json
             \hline
-            $FourierMLP(\beta_{8})$ & 0 & 0.5 & 2 & 98\% & 100\% & 98\% & 0.29 \\
+            {\it FourierMLP}$(\beta_{8})$ & 0 & 0.5 & 2 & 98\% & 100\% & 98\% & 0.29 \\
             % ../experimental_data/validation_results/2018_07_17_banded_frequency_not_position_corrected/banded_low_resolution_time1531831056_epoch249.json
             \hline
-            $FourierMLP(\beta_{50,001})$ & 0 & 0 & 2 & 95\% & 94\% & 98\% & 0.47 \\
+            {\it FourierMLP}$(\beta_{50,001})$ & 0 & 0 & 2 & 95\% & 94\% & 98\% & 0.47 \\
             % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input//dropout0/l2_lambda0/hidden_layers1/time1540053737_epoch170.json
             \hline
-            $FourierMLP(\beta_{50,001})$ & 0.001 & 0 & 4 & 95\% & 95\% & 98\% & 0.46 \\
+            {\it FourierMLP}$(\beta_{50,001})$ & 0.001 & 0 & 4 & 95\% & 95\% & 98\% & 0.46 \\
             % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0/l2_lambda0.001/hidden_layers3/time1540054370_epoch174.json
             \hline
-            $FourierMLP(\beta_{50,001})$ & 0.001 & 0.5 & 4 & 94\% & 93\% & 98\% & 0.54 \\
+            {\it FourierMLP}$(\beta_{50,001})$ & 0.001 & 0.5 & 4 & 94\% & 93\% & 98\% & 0.54 \\
             % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0.5/l2_lambda0.001/hidden_layers3/time1540056735_epoch50.json
             \hline
-            $FourierMLP(\beta_{50,001})$ & 0.003 & 0.25 & 3 & 93\% & 92\% & 98\% & 0.57 \\
+            {\it FourierMLP}$(\beta_{50,001})$ & 0.003 & 0.25 & 3 & 93\% & 92\% & 98\% & 0.57 \\
             % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0.25/l2_lambda0.003/hidden_layers2/time1540055639_epoch113.json
@@
 
-In an effort to resolve this, a two-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$, referred to as $ImageConv(\iota)$.
+In an effort to resolve this, a two-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$, referred to as {\it ImageConv}$(\iota)$.
 The network architecture consists of a moderate number of convolutional layers (on the scale of nine) and three dense layers at the end.
 L2 regularization was used on all layers, and dropout was additionally used on the dense layers.
+
+The following hyperparameters were tested during a grid search:
+\begin{itemize}
+    \item Dropout $\in \{0, 0.25, 0.5\}$
+    \item L2 Regularization $\lambda \in \{0, 0.0003, 0.001, 0.003, 0.006, 0.01\}$
+    \item Number of Convolutional Layers $\in \{6, 9, 12\}$
+\end{itemize}
 
@@
     \includegraphics[width=\textwidth]{image_hist}
-    \caption{\label{image_hist} Prediction distribution of one of the best $ImageConv(\iota)$ discriminators.}
+    \caption{\label{image_hist} Prediction distribution of one of the best {\it ImageConv}$(\iota)$ discriminators.}
 \end{figure}
@@
     \begin{center}
-        \captionof{table}{Performance of $ImageConv(\iota)$ configurations, trained without $WallCut(D)$.}
+        \captionof{table}{Performance of {\it ImageConv}$(\iota)$ configurations.}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
@@
 
-The following tables compare the accuracy and mean class-wise standard deviation values of certain highlight supervised learning configurations with multiple input formats, with the original PICO-60 neural network provided as a baseline.
+The following tables compare the accuracy and mean class-wise standard deviation values of certain successful supervised learning configurations with multiple input formats.
 
@@
             \hline
-            $DeepConvWithPos(\omega, \chi)$ & 96\% & 99\% & 96\% & 0.41 \\
+            {\it DeepConv}$(\omega)$ & 95\% & 97\% & 97\% & 0.52 \\
             \hline
-            $FourierMLP(PosCor(\beta_{8}))$ & 96\% & 98\% & 97\% & 0.42 \\
+            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}))$ & 96\% & 98\% & 97\% & 0.42 \\
             \hline
-            $FourierMLP(\beta_{8})$ & 98\% & 100\% & 98\% & 0.29 \\
+            {\it FourierMLP}$(\beta_{8})$ & 98\% & 100\% & 98\% & 0.29 \\
             \hline
-            $FourierMLPWithPos(\beta_{20}, \chi)$ & 94\% & 99\% & 93\% & 0.52 \\
+            {\it FourierMLP}$(\beta_{50,001})$ & 95\% & 95\% & 98\% & 0.46 \\
             \hline
-            $FourierMLPWithPos(\beta_{40}, \chi)$ & 91\% & 98\% & 90\% & 0.57 \\
-            \hline
-            $FourierMLPWithPos(\beta_{50,001}, \chi)$ & 91\% & 94\% & 93\% & 0.53 \\
-            \hline
-            Original PICO-60 Neural Network & 83\% & 83\% & 95\% & 0.83 \\
+            {\it ImageConv}$(\iota)$ & 68\% & 68\% & 100\% & 0.97 \\
             \hline
@@
 Higher resolutions seem to make performance worse, likely due to overfitting.
-It is significantly better than the original PICO-60 neural network, and is more decisive (as measured by the class-wise standard deviation).
 
-Based on this result, experimentation with semi-supervised learning proceeded using \\ $FourierMLP(PosCor(\beta_{8}, \chi))$, applying a three-layer perceptron architecture.
+Based on this result, experimentation with semi-supervised learning proceeded using \\ {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$, applying a three-layer perceptron architecture.
 There were two main reasons for this:
@@
 
-$GravDiff(p, \psi, g) = g \cdot sgn(p) \cdot abs(tanh(2(p - 0.5))) ^ \psi$
+{\it GravDiff}$(p, \psi, g) = g \cdot sgn(p) \cdot abs(tanh(2(p - 0.5))) ^ \psi$
 
@@
 
-(Note that step \ref{exponent} only works if $\phi$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$.
+(Note that step \ref{exponent} only works if $\psi$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$.
 To resolve this, a piecewise exponential function is used instead, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards.
@@
     \includegraphics[width=0.8\textwidth]{grav_diff}
-    \caption{\label{grav_diff} A visualization of $GravDiff(p, \psi, g)$ with respect to $p$ and $\psi$.}
+    \caption{\label{grav_diff} A visualization of {\it GravDiff} $(p, \psi, g)$ with respect to $p$ and $\psi$.}
 \end{figure}
@@
     \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
-    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $GravDiff(p, \psi, g)$ and record the resulting gradients for the next iteration.
+    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate {\it GravDiff}$(p, \psi, g)$ and record the resulting gradients for the next iteration.
     \item Increment $g$ by $\delta_{g}$.
@@
 
-During a grid search of parameters of the gravitational differentiation algorithm (in addition to the stochastic gradient descent learning rate), the highest mean accuracy over three runs of an individual configuration was 99\%.
+During a grid search of parameters of the gravitational differentiation algorithm, in addition to the stochastic gradient descent learning rate (a multiplier that defines how quickly and precisely optimization occurs), the highest mean accuracy over three runs of an individual configuration was 99\%.
 The highest accuracy was obtained with 256 labeled examples in the set $\varsigma$ (out of 496 total training examples), which equates to 52\% of the data being labeled.
@@
             \hline
+            278 & PICO-60 ML & 80\% & 76\% & 100\% & 0.64 \\
+            \hline
             $\infty$ & AP & 100\% & 100\% & 100\% & 0.41 \\
@@
 
-Note that in addition to achieving similar accuracy, configurations from both semi-supervised learning algorithms produce significantly more decisive predictions compared to AP, as measured by the class-wise standard deviation.
+As observed in this table, semi-supervised learning algorithms are significantly more accurate at replicating AP than the original PICO-60 machine learning analysis.
+
+Compared to AP, in addition to achieving similar accuracy, configurations from both semi-supervised learning algorithms produce significantly more decisive predictions as measured by the class-wise standard deviation.
 
@@
 ROOT \cite{root}, scikit-image \cite{scikit-image}, and scikit-learn \cite{scikit-learn} were used for data loading and storage.
-Matplotlib was used for data visualization.
+Matplotlib \cite{matplotlib} was used for data visualization.
 

### 2018-10-21 documentation/paper.tex
Mention specific events for audio and image samples

@@
     \includegraphics[width=\textwidth]{audio}
-    \caption{\label{} An example of the cropped audio waveform $\omega$ and full resolution Fourier transform $\beta_{50,001}$ (recorded on 2016-12-02, run 1, event 21).}
+    \caption{\label{} An example of the cropped audio waveform $\omega$ (on the left) and full resolution Fourier transform $\beta_{50,001}$ (on the right) for both audio channels.
+Recorded on 2016-12-02, run 1, event 21.}
 \end{figure}
@@
     \includegraphics[width=0.7\textwidth]{image_grid}
-    \caption{\label{} An example of the image sequence $\iota$.}
+    \caption{\label{} An example of the image sequence $\iota$. Recorded on 2016-09-23, run 4, event 33.}
 \end{figure}

### 2018-10-23 documentation/authors.tex
Lots of minor edits

@@
-\author{B.~Matusch}
-\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}
+\author[1]{B.~Matusch}
+\affil[1]{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}
 
-\author{C.~Amole}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[2]{C.~Amole}
+\affil[2]{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
 
-\author{M.~Ardid}
+\author[3]{M.~Ardid}
 \affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}
 
-\author{I.~J.~Arnquist}
-\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
+\author[4]{I.~J.~Arnquist}
+\affil[5]{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
 
-\author{D.~M.~Asner}
-%\affil{Brookhaven National Laboratory, Upton, New York 11973, USA}
-\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
+\author[4]{D.~M.~Asner}
 
-\author{D.~Baxter}
-\affil{
-Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
-\affil{Enrico Fermi Institute, KICP and Department of Physics, 
-University of Chicago, Chicago, Illinois 60637, USA}
-%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
+\author[5, 6]{D.~Baxter}
+\affil[5]{Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
+\affil[6]{Enrico Fermi Institute, KICP and Department of Physics, University of Chicago, Chicago, Illinois 60637, USA}
 
-\author{E.~Behnke}
-\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
+\author[7]{E.~Behnke}
+\affil[7]{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
 
-%\author{P.~Bhattacharjee}
-%\affil{Astroparticle Physics and Cosmology Division, Saha Institute of Nuclear Physics, Kolkata, India}
+\author[8]{M.~Bressler}
+\affil[8]{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}
 
-%\author{H.~Borsodi}
-%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
+\author[2]{B.~Broerman}
 
-%\author{M.~Bou-Cabo}
-%\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}
+\author[2]{G.~Cao}
 
-\author{M.~Bressler}
-\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}
+\author[5]{C.~J.~Chen}
 
-\author{B.~Broerman}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[2]{U.~Chowdhury}
 
-%\author{P.~Campion}
-%\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}
+\author[2]{K.~Clark}
 
-\author{G.~Cao}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[6]{J.~I.~Collar}
 
-\author{C.~J.~Chen}
-\affil{
-Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
+\author[9]{P.~S.~Cooper}
+\affil[9]{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
 
-\author{U.~Chowdhury}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[4, 9]{M.~Crisler}
 
-\author{K.~Clark}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
-%\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
-%\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}
+\author[2]{G.~Crowder}
 
-\author{J.~I.~Collar}
-\affil{Enrico Fermi Institute, KICP and Department of Physics,
-University of Chicago, Chicago, Illinois 60637, USA}
+\author[10]{N.A.~Cruz-Venegas}
+\affil[10]{Instituto de F\'isica, Universidad Nacional Aut\'onoma de M\'exico, M\'exico D.\:F. 01000, M\'exico}
 
-\author{P.~S.~Cooper}
-\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
+\author[5, 9]{C.~E.~Dahl}
 
-\author{M.~Crisler}
-\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
-\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
+\author[11]{M.~Das}
+\affil[11]{Astroparticle Physics and Cosmology Division, Saha Institute of Nuclear Physics, Kolkata, India}
 
-\author{G.~Crowder}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[12]{S.~Fallows}
+\affil[12]{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}
 
-\author{N.A.~Cruz-Venegas}
-\affil{Instituto de F\'isica, Universidad Nacional Aut\'onoma de M\'exico, M\'exico D.\:F. 01000, M\'exico}
+\author[13]{J.~Farine}
+\affil[13]{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
 
-\author{C.~E.~Dahl}
-\affil{
-Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
-\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
+\author[3]{I.~Felis}
 
-\author{M.~Das}
-\affil{Astroparticle Physics and Cosmology Division, Saha Institute
-of Nuclear Physics, Kolkata, India}
+\author[14]{R.~Filgas}
+\affil[14]{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}
 
-\author{S.~Fallows}
-\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}
+\author[13, 15]{F.~Girard}
+\affil[15]{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
 
-\author{J.~Farine}
-\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
+\author[2]{G.~Giroux}
 
-\author{I.~Felis}
-\affil{Departament de F\'isica Aplicada, IGIC - Universitat Polit\`ecnica de Val\`encia, Gandia 46730 Spain}
+\author[1]{J.~Hall}
 
-\author{R.~Filgas}
-\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}
+\author[2]{C.~Hardy}
 
-\author{F.~Girard}
-\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
-\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
+\author[16]{O.~Harris}
+\affil[16]{Northeastern Illinois University, Chicago, Illinois 60625, USA}
 
-\author{G.~Giroux}
-% \email{ggiroux@owl.phy.queensu.ca}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[4]{E.~W.~Hoppe}
 
-\author{J.~Hall}
-%\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
-\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}
+\author[5]{M.~Jin}
 
-\author{C.~Hardy}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[7]{L.~Klopfenstein}
 
-\author{O.~Harris}
-\affil{Northeastern Illinois University, Chicago, Illinois 60625, USA}
+\author[12]{C.~B.~Krauss}
 
-\author{E.~W.~Hoppe}
-\affil{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
+\author[15]{M.~Laurin}
 
-\author{M.~Jin}
-\affil{
-Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
+\author[1, 13]{I.~Lawson}
 
-\author{L.~Klopfenstein}
-\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
+\author[13]{A.~Leblanc}
 
-\author{C.~B.~Krauss}
-\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}
+\author[7]{I.~Levine}
 
-\author{M.~Laurin}
-\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
+\author[9]{W.~H.~Lippincott}
 
-\author{I.~Lawson}
-\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
-\affil{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}
+\author[14]{F.~Mamedov}
 
-\author{A.~Leblanc}
-\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
+\author[17]{D.~Maurya}
+\affil[17]{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}
 
-\author{I.~Levine}
-\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
+\author[12]{P.~Mitra}
 
-\author{W.~H.~Lippincott}
-\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
+\author[2]{C.~Moore}
 
-\author{F.~Mamedov}
-\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}
+\author[7]{T.~Nania}
 
-\author{D.~Maurya}
-\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}
+\author[8]{R.~Neilson}
 
-\author{P.~Mitra}
-\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}
+\author[2]{A.~J.~Noble}
 
-\author{C.~Moore}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[7]{P.~Oedekerk}
 
-\author{T.~Nania}
-\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
+\author[6]{A.~Ortega}
 
-\author{R.~Neilson}
-\affil{Department of Physics, Drexel University, Philadelphia, Pennsylvania 19104, USA}
+\author[12]{M.-C.~Piro}
 
-\author{A.~J.~Noble}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[15]{A.~Plante}
 
-\author{P.~Oedekerk}
-\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
+\author[13]{R.~Podviyanuk}
 
-%\author{S.~Olson}
-%\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
+\author[17]{S.~Priya}
 
-\author{A.~Ortega}
-\affil{Enrico Fermi Institute, KICP and Department of Physics,
-University of Chicago, Chicago, Illinois 60637, USA}
+\author[15]{A.~E.~Robinson}
 
-\author{M.-C.~Piro}
-\affil{Department of Physics, University of Alberta, Edmonton, T6G 2E1, Canada}
+\author[11]{S.~Sahoo}
 
-\author{A.~Plante}
-\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
+\author[13]{O.~Scallon}
 
-\author{R.~Podviyanuk}
-\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
+\author[11]{S.~Seth}
 
-\author{S.~Priya}
-\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}
+\author[9]{A.~Sonnenschein}
 
-\author{A.~E.~Robinson}
-\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
-%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
+\author[15]{N.~Starinski}
 
-%\author{A.~Roeder}
-%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
+\author[14]{I.~\v{S}tekl}
 
-%\author{R.~Rucinski}
-%\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
+\author[2]{T.~Sullivan}
 
-\author{S.~Sahoo}
-\affil{Astroparticle Physics and Cosmology Division, Saha Institute
-of Nuclear Physics, Kolkata, India}
+\author[15]{F.~Tardif}
 
-\author{O.~Scallon}
-\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
+\author[10, 13]{E.~V\'azquez-J\'auregui}
 
-\author{S.~Seth}
-\affil{Astroparticle Physics and Cosmology Division, Saha Institute
-of Nuclear Physics, Kolkata, India}
+\author[7]{N.~Walkowski}
 
-\author{A.~Sonnenschein}
-\affil{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
+\author[13]{U.~Wichoski}
 
-\author{N.~Starinski}
-\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
+\author[17]{Y.~Yan}
 
-\author{I.~\v{S}tekl}
-\affil{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Prague, Cz-12800, Czech Republic}
+\author[15]{V.~Zacek}
 
-\author{T.~Sullivan}
-\affil{Department of Physics, Queen's University, Kingston, K7L 3N6, Canada}
-
-\author{F.~Tardif}
-\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
-
-\author{E.~V\'azquez-J\'auregui}
-\affil{Instituto de F\'isica, Universidad Nacional Aut\'onoma de M\'exico, M\'exico D.\:F. 01000, M\'exico}
-\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
-
-\author{N.~Walkowski}
-\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
-
-%\author{J.~Wells}
-%\affil{Department of Physics, Indiana University South Bend, South Bend, Indiana 46634, USA}
-
-\author{U.~Wichoski}
-\affil{Department of Physics, Laurentian University, Sudbury, P3E 2C6, Canada}
-
-\author{Y.~Yan}
-\affil{Bio-Inspired Materials and Devices Laboratory (BMDL), Center for Energy Harvesting Material and Systems (CEHMS), Virginia Tech, Blacksburg, Virginia 24061, USA}
-
-\author{V.~Zacek}
-\affil{D\'epartement de Physique, Universit\'e de Montr\'eal, Montr\'eal, H3C 3J7, Canada}
-
-\author{J.~Zhang}
-\affil{
-Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
+\author[5]{J.~Zhang}

### 2018-10-23 documentation/paper.tex
Minor formatting resolutions
Lots of minor edits

@@
 \usepackage{caption}
+\usepackage{lineno}
+\linenumbers
 
@@
 \end{abstract}
+
+\pagebreak
 
@@
 The lowest-level data derived from the piezos is the raw waveform $\omega$.
-This consists of a series of samples collected every $2.5 \times 10^{-6}$ seconds.
+This consists of a series of samples, collected at the rate of one per microsecond.
 They are each represented as a 16-bit integer.
@@
     \includegraphics[width=\textwidth]{audio}
-    \caption{\label{} An example of the cropped audio waveform $\omega$ (on the left) and full resolution Fourier transform $\beta_{50,001}$ (on the right) for both audio channels.
+    \caption{\label{} An example of the cropped audio waveform $\omega$ (left) and full resolution Fourier transform $\beta_{50,001}$ (right) for both audio channels.
 Recorded on 2016-12-02, run 1, event 21.}
@@
 
-For each event that is captured, four cameras within the PICO-60 apparatus each capture a sequence of 71 images before and after formation of a bubble is detected using the cameras and an image-based trigger.
+For each event that is captured, four cameras within the PICO-60 apparatus each capture a sequence of 41 images before and after formation of a bubble is detected using the cameras and an image-based trigger.
 These raw images contain a large amount of extraneous information; they encompass the entire vessel.
 To reduce the input information, the image window sequence $\iota$ includes 50\texttimes50 cropped windows around the position of the bubble.
-The 71 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
+The 41 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
 Five frames from before the recording trigger are included in $\iota$, because the bubble does not cause a trigger until it is already of a significant size.
@@
 
-The initial data set $D$ to which these cuts are applied consists of all events recorded during PICO-60 run 2, between 2016-09-02 and 2017-06-20.
+The initial data set $D$ to which these cuts are applied consists of all events recorded in the PICO-60 detector between 2016-09-02 and 2017-06-20.
 
@@
     \item The run was not collected during engineering or testing
-    \item Recording was triggered by the camera
+    \item The recording process was triggered by the camera (as opposed to a manual trigger, et cetera)
     \item AP is not erroneously large and negative ($log_{10}(AP)>-100$)
-    \item Recorded more than 25 seconds after reaching target pressure
+    \item The event was recorded more than 25 seconds after reaching target pressure
     \item The bubble position $\chi$ was successfully calculated ($[\chi_{X}, \chi_{Y}, \chi_{Z}]\neq[-100, -100, -100]$)
@@
 
-PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles and WIMP candidates never scatter.
+PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles and WIMP candidates never multiply scatter.
 Thus, no discriminator is required to handle these events, so they are removed.
-The bubble multiplicity cut $MultiCut(D)$ consists of the following restrictions:
+The bubble multiplicity cut {\it MultiCut}$(D)$ consists of the following restrictions:
 

### 2018-10-23 visualization/visualize_audio.py
Changed scale to better fit data

@@ -39,11 +39,11 @@ def visualize_and_save_audio(time_domain: np.ndarray, frequency_domain: np.ndarr
         plt.subplot(channels, 2, frequency_domain_plot_index)
         plt.plot(frequencies_hz, frequency_domain_channel)
         plt.xlabel('Frequency (Hz)')
-        plt.ylabel(f'Fourier Transform (Magnitude) [Channel {channel_index}]')
+        plt.ylabel(f'Fourier Transform Magnitude (Channel {channel_index})')
         # Use a logarithmic scale for the X axis (because it corresponds to frequency)
         plt.xscale('log')
         # Set the Y scale manually (for some reason, the automatic one behaves strangely)
-        plt.ylim(-50_000, 250_000)
+        plt.ylim(-10_000, 220_000)
         # Take the time domain for this channel, save it as a WAV file, and notify the user
         file_path = os.path.expanduser(f'~/channel_{channel_index}.wav')
         wavfile.write(file_path, SAMPLE_RATE, time_domain_channel)

### 2018-10-23 documentation/paper.tex
Replaced graph; updated caption

@@
 Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because there is only background noise prior to this window, and a clipped signal produced by hydraulics repressurizing the vessel after.
-(These sections contain no information and can be undesirably fit on by neural networks.)
+These sections contain no information and can be undesirably fit on by neural networks.
+(For context, there are 250,000 samples in the entire waveform, and approximately half of them are recorded prior to the camera trigger.)
 
@@
     \includegraphics[width=\textwidth]{audio}
-    \caption{\label{} An example of the cropped audio waveform $\omega$ (left) and full resolution Fourier transform $\beta_{50,001}$ (right) for both audio channels.
+    \caption{\label{} An example of the audio waveform $\omega$ with initial background and hydraulic noise cropped (left) and the resulting full resolution Fourier transform $\beta_{50,001}$ (right) for both audio channels.
 Recorded on 2016-12-02, run 1, event 21.}

### 2018-10-29 documentation/paper.tex
Included exact dates and times for example data
Included CAD of the PICO-60 detector
Removed line numbers
Added description of detector

@@
 \usepackage{caption}
-\usepackage{lineno}
-\linenumbers
 
@@
 
+\subsubsection{Introduction}
+
 The PICO-60 bubble chamber experiment, created for the detection of weakly interacting massive particles (WIMPs) \cite{pico}, was completed in 2017.
 It provides an excellent basis on which to develop and test the efficacy of new machine learning discriminators, relative to the techniques used during initial analysis of the experiment.
+
+\subsubsection{About the Detector}
+
+The detector was filled with 52 kg of liquid C$_3$F$_8$, which was held at a constant temperature by a large chilled water bath.
+The pressure of the target liquid, held inside a synthetic quartz inner vessel, was very accurately controlled with propylene glycol as hydraulic fluid inside a stainless steel pressure vessel (Figure \ref{pico_cad}).
+
+The target was superheated by reducing its pressure until it entered a metastable state in which the vapor phase became energetically accessible.
+In the presence of a sufficiently dense energy deposition, a bubble would be nucleated and would grow to visible size.
+Four cameras stereoscopically monitored the chamber for these bubbles, issuing a recompression trigger upon their appearance.
+At this time, acoustic traces are recorded that capture the sound of bubble formation.
+The rising pressure caused the vapor bubble to condense as the fluid reverted to a non-superheated state.
+Following this, the hydraulic system prepared to return the detector to the sensitive superheated state by once again expanding to a lower pressure.
+
+Alpha decays along the U/Th chains present the dominant remaining background which must be distinguished from nuclear recoils acoustically.
+
+\subsubsection{Properties of Events}
 
@@
 Consequently, it is imperative to isolate only those events associated with nuclear recoils.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{pico_cad}
+    \caption{\label{pico_cad} CAD rendering of the PICO-60 detector as configured for its operation with C$_3$F$_8$.}
+\end{figure}
 
@@
     \caption{\label{} An example of the audio waveform $\omega$ with initial background and hydraulic noise cropped (left) and the resulting full resolution Fourier transform $\beta_{50,001}$ (right) for both audio channels.
-Recorded on 2016-12-02, run 1, event 21.}
+Recorded on 2016-12-03 at 04:22:45 UTC.}
 \end{figure}
@@
     \includegraphics[width=0.7\textwidth]{image_grid}
-    \caption{\label{} An example of the image sequence $\iota$. Recorded on 2016-09-23, run 4, event 33.}
+    \caption{\label{} An example of the image sequence $\iota$. Recorded on 2016-09-23 at 22:15:44 UTC.}
 \end{figure}

### 2018-11-09 documentation/paper.tex
Revert "Removed line numbers"
This reverts commit f92047a24ee0d2d07078fc81964fba54a51d80b0.
Edits to paper

@@
 
-\section{Objective}
+\section{Objective} \label{objective}
 
@@
 
-For stochastic optimization of the loss function while training, the Adam \cite{adam} optimizer was used where not otherwise specified.
+Where not otherwise specified, the following hyperparameters were used universally:
+\begin{itemize}
+    \item Adam \cite{adam} stochastic optimizer
+    \item Mean squared error loss function
+    \item Batch size of 32
+    \item $tanh$ activation function
+\end{itemize}
 
@@
 Starting with a relatively small amount of labeled data from two classes (nuclear recoils and alpha particles, in this case), predictions are run on a set of unlabeled data, and those predictions are used to produce further training data.
-However, there are two key differences:
+However, there are three key differences:
 
@@
     \begin{center}
-        \captionof{table}{Summary of the most successful semi-supervised learning configuration.}
+        \captionof{table}{Summary of the most successful semi-supervised learning configurations.}
         \begin{tabular}{|l|l|l|l|l|l|}
@@
 As observed in this table, semi-supervised learning algorithms are significantly more accurate at replicating AP than the original PICO-60 machine learning analysis.
+Compared to AP, in addition to achieving similar accuracy, configurations from both semi-supervised learning algorithms produce significantly more decisive predictions as measured by the class-wise standard deviation.
 
-Compared to AP, in addition to achieving similar accuracy, configurations from both semi-supervised learning algorithms produce significantly more decisive predictions as measured by the class-wise standard deviation.
+The following table compares the overall average performance of each of the semi-supervised learning configurations for different labeled set sizes.
+
+\begin{minipage}{\textwidth}
+    \begin{center}
+        \captionof{table}{Summary of the average accuracy values for each semi-supervised learning technique.} \label{average_semi}
+        \begin{tabular}{|l|l|l|}
+            \hline
+            Initial $\varsigma$ Size & Technique & Mean Accuracy \\
+            \hline
+            32 & GD & 69\% \\
+            \hline
+            64 & GD & 73\% \\
+            \hline
+            128 & GD & 84\% \\
+            \hline
+            256 & GD & 94\% \\
+            \hline
+            32 & ICN & 75\% \\
+            \hline
+            64 & ICN & 84\% \\
+            \hline
+            128 & ICN & 95\% \\
+            \hline
+            256 & ICN & 97\% \\
+            \hline
+        \end{tabular}
+    \end{center}
+\end{minipage}
 
@@
 This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like the Acoustic Parameter technique.
-This conclusion has been achieved through two rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems with the goal of replicating AP as accurately as possible.
+This conclusion has been achieved through two rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems with the goal of producing a high-accuracy discriminator to complement AP.
 
@@
 While both techniques produced some very effective discriminators, iterative cluster nucleation was found to be the more effective of the two.
-It replicated AP with a mean of 97\% accuracy, using 256 labeled training examples.
+As seen in Table \ref{average_semi}, it replicated AP with a mean of 97\% accuracy, using 256 labeled training examples.
 This indicates that the technique shows significant promise.
 
-While there is still of course a place for conventionally developed discriminators such as AP, they are difficult to develop until a large quantity of calibration data has been collected and the physical principles of the experiment are understood well.
-The innovations developed in this study suggest that semi-supervised learning can facilitate very quick development of discriminators immediately after the first calibration runs are conducted on a new experimental apparatus.
+As discussed in Section \ref{objective}, discriminators such as AP are difficult to develop until a large quantity of calibration data has been collected and the physical principles of the experiment are understood well.
+The innovations developed in this study suggest that semi-supervised learning can complement AP, by facilitating very quick development of discriminators immediately after the first calibration runs are conducted on a new experimental apparatus.
 This has the potential to allow the teams working on these experiments to iterate more quickly and better understand intermediate results, before all data has been analyzed in depth.

### 2018-11-13 documentation/paper.tex
Lots of general edits to the paper

@@
 \usepackage{caption}
+\usepackage{hyperref}
 \usepackage{lineno}
@@
 
+\setcounter{section}{-1}
+
+\section{Structure}
+
+This paper is organized as follows:
+\begin{itemize}
+    \item In Section 1, we outline generally the PICO-60 experiment and the context in which machine learning is applicable.
+    \item In Section 2, we discuss the specific properties of the experiment and its data, as well as the existing discriminator.
+    \item In Section 3, we discuss various machine learning techniques that are applied to develop a new discriminator.
+    \item In Section 4, we document experimentation with supervised learning, applied to various data formats.
+    \item In Section 5, we document experimentation with semi-supervised learning with the goal of improving data efficiency.
+    \item In Section 6, we conclude this study and discuss its implications.
+    \item In Section 7, we overview the technologies and software used for this study.
+\end{itemize}
+
 \section{Objective} \label{objective}
 
-In the PICO-60 experiment, and in most other experiments striving to directly detect dark matter, one of the most important and challenging hurdles to overcome is that of background events.
+In the PICO-60 \cite{pico} experiment, and in most other experiments striving to directly detect WIMP dark matter \cite{wimp}, one of the most important and challenging hurdles to overcome is that of background events.
 Unwanted particles from a variety of sources can produce event signatures that appear very similar to those which are expected to be created by dark matter candidates.
@@
 
-    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only run types as training data.
+    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only calibration/background data collection run types (henceforth "run types") as training data.
 While run types are known to be mostly correlated with particle types, the relatively large percentage of impurities has the potential to hinder conventional learning techniques.
@@
 
-The architecture described may have wider applications to several similar cases, such as those that require a discriminator which is difficult to construct using conventional techniques.
+The architectures described above may have wider applications to several similar cases, such as those that require a discriminator which is difficult to construct using conventional techniques.
 
@@
 
-The PICO-60 bubble chamber experiment, created for the detection of weakly interacting massive particles (WIMPs) \cite{pico}, was completed in 2017.
+The PICO-60 \cite{pico} bubble chamber experiment, created for the detection of weakly interacting massive particles (WIMPs), was completed in 2017.
 It provides an excellent basis on which to develop and test the efficacy of new machine learning discriminators, relative to the techniques used during initial analysis of the experiment.
@@
 
-Alpha decays along the U/Th chains present the dominant remaining background which must be distinguished from nuclear recoils acoustically.
-
-\subsubsection{Properties of Events}
-
-WIMP candidates in the cross-section and mass ranges to which the PICO-60 detector is sensitive are predicted to produce nuclear recoils with the same properties as those induced by neutrons.
-The key difference is that neutrons frequently (but not always) scatter several times and produce multiple bubbles, where the extremely small predicted cross-sections of WIMP candidates mean that they should almost invariably create just one bubble.
-This means single-bubble neutron events can be used to optimize a discriminator to detect WIMP events.
-
-The approximate expected ratio between single-bubble and multiple-bubble events generated by neutrons is known.
-A measured quantity of single-bubble events generated by nuclear recoils, in excess of this prediction, would be indicative of WIMP interactions.
-Consequently, it is imperative to isolate only those events associated with nuclear recoils.
+Alpha decays along the U/Th chains present the dominant remaining background which must be distinguished from nuclear recoils acoustically \cite{isotopes}.
 
@@
 \end{figure}
+
+\subsubsection{Properties of Events}
+
+The nuclear recoils produced by WIMP candidates (in the cross-section and mass ranges to which the PICO-60 detector is sensitive) are predicted to be indistinguishable from those induced by neutrons.
+The key difference is that neutrons frequently (but not always) scatter several times and produce multiple bubbles, where the extremely small predicted cross-sections of WIMP candidates mean that they should almost invariably create just one bubble.
+This means single-bubble neutron events can be used to optimize a discriminator to detect WIMP events.
+
+The approximate expected ratio between single-bubble and multiple-bubble events generated by neutrons is known based on Monte Carlo simulation and extrapolations from experience with previous detectors.
+A measured quantity of single-bubble events generated by nuclear recoils, in excess of this prediction, would be indicative of WIMP interactions.
+Consequently, it is imperative to isolate only those events associated with nuclear recoils.
 
@@
 Before AP can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$).
-It must also be preprocessed with the position correction function $PosCor(\beta_{8}, \chi)$, which adjusts the overall volume of an audio recording according to its position in the vessel.
-This is done because the acoustic characteristics differ significantly depending on the position of the event in the vessel.
+It must also be preprocessed with the position correction function (discussed in Section \ref{camera_derived}), which adjusts the overall volume of an audio recording according to its position in the vessel.
+This is done because the acoustic characteristics differ significantly depending on the position of the event in the vessel \cite{pico}.
 
-Finally, a set of data cuts must be applied to remove classes of events which are not accurately classified using AP alone. These classes include wall event and multiple-bubble events.
+Finally, a set of data cuts must be applied to remove classes of events which are not accurately classified using AP alone.
+These classes include events originating close to the walls of the detector, and multiple-bubble events.
 
@@
 The lowest-level data derived from the piezos is the raw waveform $\omega$.
-This consists of a series of samples, collected at the rate of one per microsecond.
-They are each represented as a 16-bit integer.
+This consists of a series of 250,000 samples, collected at the rate of one per microsecond.
+Each sample is represented as a 16-bit integer.
 Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because there is only background noise prior to this window, and a clipped signal produced by hydraulics repressurizing the vessel after.
 These sections contain no information and can be undesirably fit on by neural networks.
-(For context, there are 250,000 samples in the entire waveform, and approximately half of them are recorded prior to the camera trigger.)
+(For context, approximately half of the samples are recorded prior to the camera trigger.)
 
@@
 
-\subsubsection{Camera-Derived}
+\subsubsection{Camera-Derived} \label{camera_derived}
 
@@
 
-For each event that is captured, four cameras within the PICO-60 apparatus each capture a sequence of 41 images before and after formation of a bubble is detected using the cameras and an image-based trigger.
+Bubble events are detected using the four cameras within the PICO-60 apparatus, with an image-based trigger.
+When this happens, each camera records a sequence of 41 images before and after formation.
 These raw images contain a large amount of extraneous information; they encompass the entire vessel.
@@
 This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
-It is also used for {\it FidWallCut}$(D)$.
 
@@
 
-PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles and WIMP candidates never multiply scatter.
+PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles are stopped electromagnetically, and WIMP candidates have a negligible probability of multiply scattering because of their extremely low predicted cross section.
 Thus, no discriminator is required to handle these events, so they are removed.
@@
 
-Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel.
+Events that occur near the walls of the vessel have acoustic properties which are very different from events nearer the center of the vessel.
 It can be desirable for a discriminator to handle these events correctly; however, AP alone does not, and neither does the neural network used in the previous PICO-60 paper.
@@
     \includegraphics[width=\textwidth]{wall_event_positions}
-    \caption{\label{} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
+    \caption{\label{wall_cuts_visualization} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
 \end{figure}
@@
 
+In Figure \ref{wall_cuts_visualization}, the relative and combined effectiveness of each of these cuts can be seen.
+It is clear that most of the time, both of these cuts agree, but there are certain cases in which only one or the other marks an event.
+
 \section{Machine Learning Techniques}
@@
 
-While machine learning has existed for many years, recent developments in artificial neural networks have opened up a wide variety of applications and fields in which it can now be used.
+While neural networks have existed for many years, recent developments in computing power, most importantly GPU acceleration \cite{gpu}, have opened up a wide variety of applications and fields in which they can now be used.
 
@@
 
-This characteristic has already shown wide-ranging implications in many fields.
+This characteristic has already shown wide-ranging implications in many fields, such as autonomous driving \cite{end-to-end} and detection of heart arrhythmia \cite{arrhythmia}.
 For one, it has the potential to drastically increase the speed of iteration for the people working on a project, since retraining a machine learning system when a hardware or software variable changes is much faster than calibrating a human-designed model.
@@
 
-Two original semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation.
+Two original semi-supervised algorithms were developed and implemented for this study: gravitational differentiation (GD) and iterative cluster nucleation (ICN).
 Both are discussed in detail in Section \ref{semi_supervised}.
@@
 
-$S=std(N \cup A)$
-
-$CWSD=(std(N \div S) + std(A \div S)) \div 2$
+\begin{equation}
+    S=std(N \cup A)
+\end{equation}
+\begin{equation}
+    \mbox{\it CWSD}=(std(N \div S) + std(A \div S)) \div 2
+\end{equation}
 
@@
 This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are for each class.
-Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off those specific values, will produce a low class-wise standard deviation.
+Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any given $x$ and $y$), with minimal variance off those specific values, will produce a low class-wise standard deviation.
 
@@
 
-The following tables show the results of several {\it DeepConv}$(\omega)$ discriminators produced throughout testing, with corresponding network hyperparameters.
+Table \ref{deepconv_table} shows the results of several {\it DeepConv}$(\omega)$ discriminators produced throughout testing, with corresponding network hyperparameters.
 
@@
     \begin{center}
-        \captionof{table}{Performance of {\it DeepConv}$(\omega)$ configurations.}
+        \captionof{table}{Performance of {\it DeepConv}$(\omega)$ configurations.} \label{deepconv_table} 
         \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
@@
 
-The following tables compare various multi-layer perceptron discriminators, with and without wall cuts.
+Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators, with and without wall cuts.
 
@@
 
-It is very clear in Table \ref{fft_wall_cuts} that higher resolution data does not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
+It is very clear that higher resolution data does not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
 
@@
     \begin{center}
-        \captionof{table}{Performance of {\it ImageConv}$(\iota)$ configurations.}
+        \captionof{table}{Performance of {\it ImageConv}$(\iota)$ configurations.} \label{imageconv_table}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
@@
 
-The high standard deviations across the board are further evidence for the hypothesis that there is insufficient information in the image data.
+In Table \ref{imageconv_table}, high standard deviations across the board are further evidence for the hypothesis that there is insufficient information in the image data.
 The 100\% recall statistics are not due to high classification accuracy, but rather due to the network making predictions less than 0.5 for all examples of both classes.
@@
 
-The following tables compare the accuracy and mean class-wise standard deviation values of certain successful supervised learning configurations with multiple input formats.
+Table \ref{super_overview} compares the accuracy and mean class-wise standard deviation values of certain successful supervised learning configurations with multiple input formats.
 
@@
     \begin{center}
-        \captionof{table}{Summary of supervised learning configurations.}
+        \captionof{table}{Summary of supervised learning configurations.} \label{super_overview}
         \begin{tabular}{|l|l|l|l|l|}
@@
 
-The key conclusion to be drawn from the above tables is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
+The key conclusion to be drawn from the above table is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
 Higher resolutions seem to make performance worse, likely due to overfitting.
@@
 
-Gradients are calculated using the gravitational differentiation function $GravDiff(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$.
+Gradients are calculated using the gravitational differentiation function $\mathrm{GravDiff}(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$.
 The function is defined as follows:
 
-{\it GravDiff}$(p, \psi, g) = g \cdot sgn(p) \cdot abs(tanh(2(p - 0.5))) ^ \psi$
+\begin{equation}
+    \mathrm{GravDiff}(p, \psi, g) = g \cdot \mathrm{sgn}(p) \cdot \mathrm{abs}(\mathrm{tanh}(2(p - 0.5))) ^ \psi
+\end{equation}
 
@@
     \includegraphics[width=0.8\textwidth]{grav_diff}
-    \caption{\label{grav_diff} A visualization of {\it GravDiff} $(p, \psi, g)$ with respect to $p$ and $\psi$.}
+    \caption{\label{grav_diff} A visualization of $\mathrm{GravDiff}(p, \psi, g)$ with respect to $p$ and $\psi$.}
 \end{figure}
@@
     \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
-    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate {\it GravDiff}$(p, \psi, g)$ and record the resulting gradients for the next iteration.
+    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $\mathrm{GravDiff}(p, \psi, g)$ and record the resulting gradients for the next iteration.
     \item Increment $g$ by $\delta_{g}$.
@@
 
-The following table represents the performance of several high-accuracy gravitational differentiation configurations tested during the grid search.
+Table \ref{grav_overview} represents the performance of several high-accuracy gravitational differentiation configurations tested during the grid search.
 The accuracy statistic is the mean over three tests of the same configuration.
@@
     \begin{center}
-        \captionof{table}{Performance of gravitational differentiation configurations.}
+        \captionof{table}{Performance of gravitational differentiation configurations.} \label{grav_overview}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
@@
 
-The following table represents the performance of a variety of high-accuracy iterative cluster nucleation configurations tested during the grid search.
+Table \ref{icn_overview} represents the performance of a variety of high-accuracy iterative cluster nucleation configurations tested during the grid search.
 As with gravitational differentiation, the accuracy statistic is the mean over three tests of the same configuration.
@@
     \begin{center}
-        \captionof{table}{Performance of iterative cluster nucleation configurations.}
+        \captionof{table}{Performance of iterative cluster nucleation configurations.} \label{icn_overview}
         \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
@@
 
-The following table directly compares performance of the most successful semi-supervised learning configurations, with AP provided as a baseline for comparison.
+Table \ref{semi_success_overview} directly compares performance of the most successful semi-supervised learning configurations, with AP provided as a baseline for comparison.
 Remember that ``Mean AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.
@@
     \begin{center}
-        \captionof{table}{Summary of the most successful semi-supervised learning configurations.}
+        \captionof{table}{Summary of the most successful semi-supervised learning configurations.} \label{semi_success_overview}
         \begin{tabular}{|l|l|l|l|l|l|}
@@
 
-The following table compares the overall average performance of each of the semi-supervised learning configurations for different labeled set sizes.
+Table \ref{average_semi} compares the overall average performance of each of the semi-supervised learning configurations for different labeled set sizes.
 

### 2018-11-20 documentation/paper.tex
Lots of changes from review

@@
 \input{authors.tex}
-\date{August 2018}
+\date{November 2018}
 \maketitle
@@
     \item In Section 1, we outline generally the PICO-60 experiment and the context in which machine learning is applicable.
-    \item In Section 2, we discuss the specific properties of the experiment and its data, as well as the existing discriminator.
+    \item In Section 2, we discuss the specific properties of the experiment and its data, as well as the existing signal versus background discriminator.
     \item In Section 3, we discuss various machine learning techniques that are applied to develop a new discriminator.
@@
 
-In the PICO-60 \cite{pico} experiment, and in most other experiments striving to directly detect WIMP dark matter \cite{wimp}, one of the most important and challenging hurdles to overcome is that of background events.
+In the PICO-60 \cite{pico} experiment, and in other experiments striving to directly detect WIMP dark matter \cite{wimp}, one of the most important and challenging hurdles to overcome is that of background events.
 Unwanted particles from a variety of sources can produce event signatures that appear very similar to those which are expected to be created by dark matter candidates.
@@
 
-    \item A major intent of many dark matter experiments, which is to reduce the number of background events to the lowest level possible, means that there will be a very small number detected.
+    \item A major intent of many dark matter experiments is to reduce the number of background events to the lowest level possible.
+This means that there will be a very small number detected.
 This places a significant constraint on the amount of data that is available to optimize such a discriminator.
@@
 
-    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only calibration/background data collection run types (henceforth "run types") as training data.
+    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only calibration/background data collection run types (henceforth ``run types'') as training data.
 While run types are known to be mostly correlated with particle types, the relatively large percentage of impurities has the potential to hinder conventional learning techniques.
@@
 The architectures described above may have wider applications to several similar cases, such as those that require a discriminator which is difficult to construct using conventional techniques.
+The following are possible reasons for this limitation:
 
@@
 The detector was filled with 52 kg of liquid C$_3$F$_8$, which was held at a constant temperature by a large chilled water bath.
-The pressure of the target liquid, held inside a synthetic quartz inner vessel, was very accurately controlled with propylene glycol as hydraulic fluid inside a stainless steel pressure vessel (Figure \ref{pico_cad}).
+The pressure of the target liquid, held inside a synthetic quartz inner vessel, was precisely accurately controlled with propylene glycol as hydraulic fluid inside a stainless steel pressure vessel (Figure \ref{pico_cad}).
 
@@
 Four cameras stereoscopically monitored the chamber for these bubbles, issuing a recompression trigger upon their appearance.
-At this time, acoustic traces are recorded that capture the sound of bubble formation.
+At this time, acoustic traces were recorded that captured the sound of bubble formation, and a Dytran 2006V1 fast pressure transducer \cite{amole_thesis} monitored the rise.
 The rising pressure caused the vapor bubble to condense as the fluid reverted to a non-superheated state.
@@
 Before AP can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$).
-It must also be preprocessed with the position correction function (discussed in Section \ref{camera_derived}), which adjusts the overall volume of an audio recording according to its position in the vessel.
+It must also be preprocessed with the position correction function (discussed in Section \ref{camera_derived}), which adjusts the overall amplitude of an audio recording according to its position in the vessel.
 This is done because the acoustic characteristics differ significantly depending on the position of the event in the vessel \cite{pico}.
@@
 PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles are stopped electromagnetically, and WIMP candidates have a negligible probability of multiply scattering because of their extremely low predicted cross section.
-Thus, no discriminator is required to handle these events, so they are removed.
+Thus, multi-bubble events can be safely removed.
 The bubble multiplicity cut {\it MultiCut}$(D)$ consists of the following restrictions:
@@
 In Figure \ref{wall_cuts_visualization}, the relative and combined effectiveness of each of these cuts can be seen.
-It is clear that most of the time, both of these cuts agree, but there are certain cases in which only one or the other marks an event.
+It is clear that most of the time, both of these cuts agree, but there are certain cases in which only one or the other eliminates an event.
+
+Please note that events appearing to be in the middle of the vessel in Figure \ref{wall_cuts_visualization} are in actually at the top and bottom of the detector, as only the X and Y axes are displayed.
 
@@
 \begin{enumerate}
-    \item In future experiments, when no AP equivalent is available, impure data is the only information available for training. The network structure should be selected to accommodate this.
+    \item During early operations of future experiments, when no AP equivalent is yet available, impure data is the only information available for training.
+The network structure should be selected to accommodate this.
     \item In general, supervised learning systems learn to replicate biases present in the training data set.
@@
 
-This table shows that simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
+Table \ref{deepconv_table} shows that simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
 Larger kernel sizes, more filters, and more layers make performance generally worse, while higher L2 regularization does the opposite.
@@
 Preprocessing the audio by applying a Fourier transform may be advantageous.
-If the frequency distribution and overall volume are indeed the most important factors for discrimination, the network can gather them straight from the Fourier transform $\beta_{N}$ rather than having to analyze $\omega$ to extract this information.
+If the frequency distribution and overall amplitude are indeed the most important factors for discrimination, the network can gather them straight from the Fourier transform $\beta_{N}$ rather than having to analyze $\omega$ to extract this information.
 
@@
 When $\beta_{8}$ was used as input, position corrections were optionally applied to the input data, as they were the input to Acoustic Parameter.
-The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and \\ $FourierMLP(PosCor(\beta_{N}, \chi))$.
+The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and $FourierMLP(PosCor(\beta_{N}, \chi))$.
 While a variety of network architectures were tested, most had a small number of layers (on the order of three), applied dropout and L2 regularization, and used batch normalization once again.
@@
 Based on this result, experimentation with semi-supervised learning proceeded using \\ {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$, applying a three-layer perceptron architecture.
-There were two main reasons for this:
+There were three main reasons for this:
 
@@
     \item The use of the exact same input format as AP makes the algorithms more directly comparable.
+    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because of the different position distributions of the two classes.
 \end{enumerate}
@@
 To ensure that accuracy statistics are as consistent and reliable as possible, every configuration was trained three times, with a different randomly selected validation set each time.
-Accuracy, precision, recall, and class-wise standard deviation statistics are the mean over those three tests.
+(Using more than three training runs would have taken prohibitive amounts of time.) Accuracy, precision, recall, and class-wise standard deviation statistics are the mean over those three tests.
 
@@
     \item Apply the sigmoidal hyperbolic tangent, producing a value of 0 in the center and asymptotic slopes to -1 and 1 at the edges.
-    \item \label{exponent} What is really desired is a shallow slope in the center (so low-confidence network outputs near 0.5 will produce very shallow output gradients).
-Apply a relatively large exponent which is $\psi$ (in the range of 3 to 11) to squash values close to 0 closer to 0.
+    \item \label{exponent} Apply a relatively large exponent which is $\psi$ (in the range of 3 to 11) to squash values close to 0 closer to 0.
+This ensures a shallow slope in the center (so low-confidence network outputs near 0.5 will produce very small output gradients).
 \end{enumerate}
@@
     \includegraphics[width=\textwidth]{grav_grid_search}
-    \caption{\label{grav_grid_search} Accuracy of best gravitational differentiation training epoch compared to AP.}
+    \caption{\label{grav_grid_search} Predictions of best gravitational differentiation model compared to AP (on randomly selected validation set).}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{icn_grid_search}
-    \caption{\label{icn_grid_search} Accuracy of best iterative cluster nucleation training epoch compared to AP.}
+    \caption{\label{icn_grid_search} Predictions of best iterative cluster nucleation model compared to AP (on randomly selected validation set).}
 \end{figure}
@@
 As discussed in Section \ref{objective}, discriminators such as AP are difficult to develop until a large quantity of calibration data has been collected and the physical principles of the experiment are understood well.
-The innovations developed in this study suggest that semi-supervised learning can complement AP, by facilitating very quick development of discriminators immediately after the first calibration runs are conducted on a new experimental apparatus.
-This has the potential to allow the teams working on these experiments to iterate more quickly and better understand intermediate results, before all data has been analyzed in depth.
+The innovations developed in this study suggest that semi-supervised learning can efficiently replicate AP, thus facilitating very quick development of discriminators immediately after the first calibration runs are conducted on a new experimental apparatus.
+This has the potential to allow the teams working on these experiments to iterate more quickly, and to better understand intermediate results.
 
@@
 
+\section{Acknowledgements}
+
+The PICO collaboration wishes to thank SNOLAB and its staff for support through underground space, logistical and technical services.
+SNOLAB operations are supported by the Canada Foundation for Innovation and the Province of Ontario Ministry of Research and Innovation, with underground access provided by Vale at the Creighton mine site.
+We wish to acknowledge the support of the Natural Sciences and Engineering Research Council of Canada (NSERC) and the Canada Foundation for Innovation (CFI) for funding.
+We acknowledge the support from National Science Foundation (NSF) (Grant 0919526, 1506337, 1242637 and 1205987).
+We acknowledge that this work is supported by the U.S.
+Department of Energy (DOE) Office of Science, Office of High Energy Physics (under award DE-SC-0012161), by the DOE Office of Science Graduate Student Research (SCGSR) award,  by DGAPA-UNAM (PAPIIT No.\:IA100316 and IA100118) and Consejo Nacional de Ciencia y Tecnolog\'ia (CONACyT, M\'exico, Grant No.\:252167), by the Department of Atomic Energy (DAE), the Government of India, under the Center of AstroParticle Physics II project (,CAPP-II) at SAHA Institute of Nuclear Physics (SINP), European Regional Development Fund-Project ``Engineering applications of microworld physics'' (No.\:CZ.02.1.01/0.0/0.0/16\_019/0000766),
+and the Spanish Ministerio de Econom\'ia y Competitividad, Consolider MultiDark (Grant CSD2009-00064).
+This work is partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through NSF grant 1125897, and an endowment from the Kavli Foundation and its founder Fred Kavli.
+We also wish to acknowledge the support from Fermi National Accelerator Laboratory under Contract No.\:DE-AC02-07CH11359, and Pacific Northwest National Laboratory, which is operated by Battelle for the U.S.
+Department of Energy under Contract No.\:DE-AC05-76RL01830.
+We also thank Compute Canada (\url{www.computecanada.ca}) and the Centre for Advanced Computing, ACENET, Calcul Qu\'ebec, Compute Ontario and WestGrid for the computational support.
+
 \printbibliography

### 2018-11-20 documentation/paper.tex
Clarified WIMP search runs

@@
         \item Neutron calibration sets, which are known to contain approximately 90\% neutrons, and
-        \item Background radiation runs, which contain approximately 99\% alpha particles
+        \item WIMP search runs, which contain approximately 99\% alpha particles
     \end{enumerate}
@@
 
-Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original machine learning analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands.
+Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original machine learning analysis) was computed by integrating the resonant energy over all frequencies within each of a set of frequency bands ranging from 1 kHz to 300 kHz \cite{pico}.
 This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.

### 2018-11-21 documentation/paper.tex
Added footnotes for authors
Addressed more comments

@@
 \usepackage{lineno}
+
 \linenumbers
@@
 \renewcommand*{\Affilfont}{\footnotesize}
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
 
@@
         \item Neutron calibration sets, which are known to contain approximately 90\% neutrons, and
-        \item WIMP search runs, which contain approximately 99\% alpha particles
+        \item Previously blinded WIMP search runs and diagnostic background runs, which contain approximately 99\% alpha particles.
     \end{enumerate}
@@
 The detector was filled with 52 kg of liquid C$_3$F$_8$, which was held at a constant temperature by a large chilled water bath.
-The pressure of the target liquid, held inside a synthetic quartz inner vessel, was precisely accurately controlled with propylene glycol as hydraulic fluid inside a stainless steel pressure vessel (Figure \ref{pico_cad}).
+The pressure of the target liquid, held inside a synthetic quartz inner vessel, was precisely controlled with propylene glycol as hydraulic fluid inside a stainless steel pressure vessel (Figure \ref{pico_cad}).
 
@@
 \end{figure}
+
+(Note that in Figure \ref{grav_grid_search}, AP is normalized differently than in the original PICO-60 study.
+In this normalized spectrum, the decision boundary between neutrons and alpha particles is at 0.25.)
 

### 2018-11-21 documentation/paper.tex
More comments

@@
     \item A detailed model of the physical environment is often used to produce an accurate discriminator.
-This takes a long time, and will have to be updated whenever the physical variables of the experimental apparatus change.
+This takes a long time to develop and compute, and will have to be updated whenever the physical variables of the experimental apparatus change.
 
@@
 Furthermore, develop several neural network architectures and apply them to different input/preprocessing formats (such as audio Fourier transforms, raw waveforms, and images), in order to find which are the most effective.
-
     \item Determine whether these techniques can be extended, applying two original algorithms based on semi-supervised learning, to develop an effective discriminator function based on incomplete and inaccurate initial information.
+\end{enumerate}
 
@@
 \end{enumerate}
-\end{enumerate}
-
-The architectures described above may have wider applications to several similar cases, such as those that require a discriminator which is difficult to construct using conventional techniques.
+
+The architectures described in this paper may have wider applications to several similar cases, such as those that require a discriminator which is difficult to construct using conventional techniques.
 The following are possible reasons for this limitation:
@@
 Four cameras stereoscopically monitored the chamber for these bubbles, issuing a recompression trigger upon their appearance.
-At this time, acoustic traces were recorded that captured the sound of bubble formation, and a Dytran 2006V1 fast pressure transducer \cite{amole_thesis} monitored the rise.
+At this time, acoustic traces were recorded that captured the sound of bubble formation, and a Dytran 2006V1 fast pressure transducer \cite{amole_thesis} monitored the pressure rise.
 The rising pressure caused the vapor bubble to condense as the fluid reverted to a non-superheated state.
@@
 
+Superheated fluid detectors have an intrinsic insensitivity to the relatively diffuse energy depositions from electron recoils \cite{pico}, and neutron backgrounds are mitigated by shielding the detector volume in a 20 t water tank.
 Alpha decays along the U/Th chains present the dominant remaining background which must be distinguished from nuclear recoils acoustically \cite{isotopes}.
@@
 
-The Acoustic Parameter method (furthermore AP) is the technique that was used for event discrimination in the original PICO-60 analysis.
+An Acoustic Parameter method (furthermore AP) is the technique that was used for event discrimination in the original PICO-60 analysis.
 A function was defined that discriminates between alpha particles and nuclear recoils based on audio data.
@@
 Discriminators can use a variety of different data formats to make predictions.
-These are collected from the two essential sensors present in the PICO-60 apparatus: piezoelectric microphones (piezos) and cameras.
+These are collected from the two essential sensors present in the PICO-60 apparatus: two piezoelectric microphones (piezos) and cameras.
 
@@
 
-The lowest-level data derived from the piezos is the raw waveform $\omega$.
+The lowest-level data derived from the two piezos in the PICO-60 apparatus is the raw waveform $\omega$.
 This consists of a series of 250,000 samples, collected at the rate of one per microsecond.
@@
 Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because there is only background noise prior to this window, and a clipped signal produced by hydraulics repressurizing the vessel after.
-These sections contain no information and can be undesirably fit on by neural networks.
+These sections contain no information and can be undesirably fitted by neural networks.
 (For context, approximately half of the samples are recorded prior to the camera trigger.)
@@
 Since DFTs produce sequences of complex numbers, the magnitude of each element was computed.
-The direct output from such a DFT is the full-resolution Fourier transform $\beta _{50,001}$, which consists of a sequence of numbers half the length of the raw waveform $\omega$.
-
-Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original machine learning analysis) was computed by integrating the resonant energy over all frequencies within each of a set of frequency bands ranging from 1 kHz to 300 kHz \cite{pico}.
+The direct output from such a DFT is the full-resolution Fourier transform $\beta _{50,001}$, which consists of a sequence of 50,001 numbers (half the length of the raw waveform $\omega$).
+
+Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original machine learning analysis) was computed by integrating the resonant energy over all frequencies within each of a set of eight frequency bands ranging from 1 kHz to 300 kHz \cite{pico}.
 This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.
+
+Both $\beta _{50,001}$ and $\beta_{8}$ were used as inputs to machine learning models applied in this study.
 
@@
 To reduce the input information, the image window sequence $\iota$ includes 50\texttimes50 cropped windows around the position of the bubble.
-The 41 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble.
-Five frames from before the recording trigger are included in $\iota$, because the bubble does not cause a trigger until it is already of a significant size.
+The 41 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 frames immediately around the formation of the bubble.
+Two to three frames from before the recording trigger are included in $\iota$, because the bubble does not cause a trigger until it is already of a significant size.
 
@@
 The 3-dimensional position $\chi$ of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera.
+
 This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
+$PosCor(\beta _{8}, \chi)$ is used for training several machine learning models throughout this study.
+Details of its use are in Section \ref{mlp}.
 
@@
 
-The initial data set $D$ to which these cuts are applied consists of all events recorded in the PICO-60 detector between 2016-09-02 and 2017-06-20.
+The initial data set $D$ to which these cuts are applied consists of all events recorded in the PICO-60 detector, at the 3.3keV threshold, between 2016-09-02 and 2017-06-20.
 
@@
 
-PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles are stopped electromagnetically, and WIMP candidates have a negligible probability of multiply scattering because of their extremely low predicted cross section.
+PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles are stopped electromagnetically, and WIMP candidates have a negligible probability of multiple scattering because of their extremely low predicted cross section.
 Thus, multi-bubble events can be safely removed.
@@
 
-Please note that events appearing to be in the middle of the vessel in Figure \ref{wall_cuts_visualization} are in actually at the top and bottom of the detector, as only the X and Y axes are displayed.
+Some events eliminated by fiducial cuts appear to be in the middle of the vessel in Figure \ref{wall_cuts_visualization}.
+These events are at the top or bottom of the detector (along the Z axis, which is not displayed in the 2D plot).
+
+The data set, after all of these cuts are applied, consists of 624 events. During each training run, 128 randomly selected events are set aside for validation, leaving a training set of 496 events.
 
@@
 
-Performance of machine learning systems was evaluated using two metrics: classification accuracy and class-wise standard deviation.
+Performance of the machine learning systems was evaluated using two metrics: classification accuracy and class-wise standard deviation.
 
@@
 When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
-While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, AP has a significantly wider range.
+While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation on the last layer, AP has a significantly wider range.
 Normalization of the union prevents this from creating a bias where AP would produce a higher standard deviation with a similarly proportioned error.
@@
     \item Number of Convolutional Layers $\in \{17, 32\}$
+    \item Number of Dense Layers $= 3$
 \end{itemize}
@@
 
-\subsection{Multi-Layer Perceptron for Fourier Transform Analysis}
+\subsection{Multi-Layer Perceptron for Fourier Transform Analysis} \label{mlp}
 
@@
 The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and $FourierMLP(PosCor(\beta_{N}, \chi))$.
-While a variety of network architectures were tested, most had a small number of layers (on the order of three), applied dropout and L2 regularization, and used batch normalization once again.
+While a variety of network architectures were tested, most had a small number of dense layers (on the order of three), applied dropout and L2 regularization, and used batch normalization once again.
 
@@
 
-It is very clear that higher resolution data does not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.
+It is very clear that higher resolution data does not improve accuracy in a relatively shallow network. Regularization, both dropout and L2, appears to have relatively little effect.
 

### 2018-11-22 documentation/paper.tex
Addressed lots and lots of comments

@@
 As part of the PICO-60 experiment, some testing was conducted to determine whether machine learning (in the form of a multi-layer perceptron) could be a viable technique to consider for a discrimination function.
-The results, while not definitive, indicated that the general technique warranted further investigation.
+The results, published in the 2017 PICO-60 paper \cite{pico}, indicated that the general technique warranted further investigation.
+
+The neural network in this study used the same $\beta_{8}$ input format as was applied in this study (detailed in Section \ref{piezo_derived}), with pre-trigger noise subtracted.
+Position corrections (also in Section \ref{piezo_derived}) were not applied; rather, the position of the event in the vessel is additionally used as input.
+AP was used to restrict the data used as input to the neural network.
+In addition to neutrons and alpha particles, gamma events were used as training and testing data for this neural network.
+
+Software for this experiment was implemented using the MATLAB Neural Network Toolbox \cite{matlab}.
 
@@
 
-\subsubsection{Piezo-Derived}
+\subsubsection{Piezo-Derived} \label{piezo_derived}
 
@@
     \item Number of Convolutional Layers $\in \{17, 32\}$
-    \item Number of Dense Layers $= 3$
+    \item Number of Dense Layers = 3 (with neuron counts 64, 16, and 1)
 \end{itemize}
@@
 
-During early experiments with shallower network architectures, it was revealed to have a surprising and undesirable capacity to separate run types based on meaningless information.
-It was able to discriminate between events from the AmBe neutron calibration runs and events from alpha background runs with 96\% accuracy, and to separate all combined neutron calibration runs from alpha background runs with 91\% accuracy.
-This was determined to be caused by the network detecting and fitting on biases in the white noise at the beginning and hydraulic recompression sounds at the end of the raw waveform.
-This extraneous information was subsequently cropped out.
-
-When this extraneous information was cropped out and the architecture was modified, both accuracy and class-wise standard deviation remained quite high.
-The most accurate configuration during a grid search was 95\% accurate and had a standard deviation of 0.52 (seen in Figure \ref{waveform_hist}).
-This is better by both metrics than the multi-layer perceptron used in the original PICO-60 analysis.
+The most accurate configuration during a waveform CNN grid search was 95\% accurate and had a standard deviation of 0.52 (seen in Figure \ref{waveform_hist}).
 
@@
     \item L2 Regularization $\lambda \in \{0, 0.0003, 0.001, 0.003, 0.01\}$
-    \item Number of Layers $\in \{2, 3, 4\}$
+    \item Number of Dense Layers $\in \{2, 3, 4\}$
 \end{itemize}
@@
     \item Number of Convolutional Layers $\in \{6, 9, 12\}$
+    \item Number of Dense Layers = 3 (with neuron counts 64, 16, and 1)
 \end{itemize}
@@
     \includegraphics[width=\textwidth]{image_hist}
-    \caption{\label{image_hist} Prediction distribution of one of the best {\it ImageConv}$(\iota)$ discriminators.}
+    \caption{\label{image_hist} Prediction distribution of one of the best {\it ImageConv}$(\iota)$ discriminators. (Note the truncated X axis.)}
 \end{figure}
@@
 \begin{enumerate}
-    \item This was a highly effective configuration, producing 97\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
+    \item This was a highly effective configuration, producing 96\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
     \item The use of the exact same input format as AP makes the algorithms more directly comparable.
@@
 
+Because all data in the PICO-60 experiment is labeled, unlabeled training data sets are randomly sampled from the 496 training examples remaining (after the validation set is selected).
+
 Since the intent of these techniques is to separate particle types (the same as AP), performance evaluations (accuracy and class-wise standard deviation) use AP as a baseline.
+All graphs in this section use AP's predictions to define the sets of neutrons and alpha particles.
+
 To ensure that accuracy statistics are as consistent and reliable as possible, every configuration was trained three times, with a different randomly selected validation set each time.
@@
 
-\begin{figure}[h]
+\subsubsection{Application}
+
+This function is applied in a training algorithm that is parameterized by the set of imperfectly labeled data $\varsigma$, the set of unlabeled training data $\upsilon$, a binary classification neural network $NN(x)$ (where $x$ is the input data format), and the gravitational multiplier increment $\delta_g$.
+The distortion power $\psi$ is assumed to be constant throughout a training run, and the gravitational multiplier $g$ is initialized at 0, gradually increasing during training.
+It iterates as follows:
+
+\begin{enumerate}
+    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
+    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $\mathrm{GravDiff}(p, \psi, g)$ and record the resulting gradients for the next iteration.
+    \item Increment $g$ by $\delta_{g}$.
+At the beginning, when $g = 0$, training will be entirely based on the labeled set $\varsigma$, and progressively, over the course of a run, the gravitational effect increases with $g$.
+\end{enumerate}
+
+This system was trained using a multi-layer perceptron with an 8-band Fourier transform input for $NN(x)$, because this was one of the most successful configurations in supervised learning.
+It produced the highest accuracy as well as the lowest class-wise standard deviation.
+
+\begin{figure}[H]
     \centering
@@
 \end{figure}
-
-\subsubsection{Application}
-
-This function is applied in a training algorithm that is parameterized by the set of imperfectly labeled data $\varsigma$, the set of unlabeled training data $\upsilon$, a binary classification neural network $NN(x)$ (where $x$ is the input data format), and the gravitational multiplier increment $\delta_g$.
-The distortion power $\psi$ is assumed to be constant throughout a training run, and the gravitational multiplier $g$ is initialized at 0, gradually increasing during training.
-It iterates as follows:
-
-\begin{enumerate}
-    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
-    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $\mathrm{GravDiff}(p, \psi, g)$ and record the resulting gradients for the next iteration.
-    \item Increment $g$ by $\delta_{g}$.
-At the beginning, when $g = 0$, training will be entirely based on the labeled set $\varsigma$, and progressively, over the course of a run, the gravitational effect increases with $g$.
-\end{enumerate}
-
-This system was trained using a multi-layer perceptron with an 8-band Fourier transform input for $NN(x)$, because this was one of the most successful configurations in supervised learning.
-It produced the highest accuracy as well as the lowest class-wise standard deviation.
 
@@
 Table \ref{semi_success_overview} directly compares performance of the most successful semi-supervised learning configurations, with AP provided as a baseline for comparison.
-Remember that ``Mean AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.
+Remember that ``Mean AP Accuracy'' is calculated relative to AP's predictions, rather than replication of the ground truths.
 
@@
     \begin{center}
-        \captionof{table}{Summary of the most successful semi-supervised learning configurations.} \label{semi_success_overview}
+        \captionof{table}{Summary of the most successful semi-supervised learning models.} \label{semi_success_overview}
         \begin{tabular}{|l|l|l|l|l|l|}
@@
             \hline
-            $\infty$ & AP & 100\% & 100\% & 100\% & 0.41 \\
+            N/A & AP & 100\% & 100\% & 100\% & 0.41 \\
             \hline
@@
 All programming for this study was done in Python 3.
-Keras \cite{keras}, running on a TensorFlow backend, was used for all machine learning tasks.
+Keras \cite{keras}, running on a TensorFlow \cite{tensorflow} backend, was used for all machine learning tasks.
 NumPy \cite{numpy} and SciPy \cite{scipy} were used for linear algebra and signal processing.

### 2018-11-22 documentation/authors.tex
Lots of changes from review
Added footnotes for authors
Addressed lots and lots of comments
More edits

@@
-\author[1]{B.~Matusch}
+\author[1]{B.~Matusch \thanks{Corresponding: \url{brendon.matusch@snolab.ca}}}
 \affil[1]{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}
@@
 \author[4]{I.~J.~Arnquist}
-\affil[5]{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
+\affil[4]{Pacific Northwest National Laboratory, Richland, Washington 99354, USA}
 
-\author[4]{D.~M.~Asner}
+\author[4]{D.~M.~Asner \thanks{now at Brookhaven National Laboratory}}
 
@@
 
-\author[2]{U.~Chowdhury}
+\author[2]{U.~Chowdhury \thanks{now at Canadian Nuclear Laboratories}}
 
@@
 \affil[9]{Fermi National Accelerator Laboratory, Batavia, Illinois 60510, USA}
+
+\author[4]{C.~Cowles}
 
@@
 
+\author[4]{C.~M.~Jackson}
+
 \author[5]{M.~Jin}
@@
 \author[9]{W.~H.~Lippincott}
+
+\author[4]{B.~Loer}
 
@@
 
+\author[4]{K.~Wierman}
+
 \author[17]{Y.~Yan}
@@
 
-\author[5]{J.~Zhang}
+\author[5]{J.~Zhang \thanks{Argonne National Laboratory}}

### 2018-11-22 documentation/paper.tex
More edits

@@
 
+\DeclareUnicodeCharacter{0301}{'}
+
 \linenumbers
@@
 When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
-While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation on the last layer, AP has a significantly wider range.
+While the neural network's outputs are bounded in the range of 0 to 1 with a sigmoid activation on the last layer, AP has a significantly wider range.
 Normalization of the union prevents this from creating a bias where AP would produce a higher standard deviation with a similarly proportioned error.
@@
 The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually.
-This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are for each class.
+This is an indication of how tightly clustered or widely dispersed the discriminator's predictions are for each class.
 Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any given $x$ and $y$), with minimal variance off those specific values, will produce a low class-wise standard deviation.
@@
 
-(Note that in Figure \ref{grav_grid_search}, AP is normalized differently than in the original PICO-60 study.
-In this normalized spectrum, the decision boundary between neutrons and alpha particles is at 0.25.)
+Note that in Figure \ref{grav_grid_search}, AP is normalized using a similar method to the original PICO-60 study \cite{pico}.
+In this normalized spectrum, the decision boundary between neutrons and alpha particles is at 0.25.
 

### 2018-11-22 documentation/authors.tex
Added more authors and email

@@
-\author[1]{B.~Matusch \thanks{Corresponding: \url{brendon.matusch@snolab.ca}}}
+\author[1]{B.~Matusch\thanks{Corresponding: \url{brendon.matusch@snolab.ca} or \url{analysis@picoexperiment.com}}}
 \affil[1]{SNOLAB, Lively, Ontario, P3Y 1N2, Canada}
@@
 
-\author[4]{D.~M.~Asner \thanks{now at Brookhaven National Laboratory}}
+\author[4]{D.~M.~Asner\thanks{now at Brookhaven National Laboratory}}
 
@@
 
-\author[2]{U.~Chowdhury \thanks{now at Canadian Nuclear Laboratories}}
+\author[2]{U.~Chowdhury\thanks{now at Canadian Nuclear Laboratories}}
 
@@
 
+\author[13]{T.~Hillier}
+
 \author[4]{E.~W.~Hoppe}
@@
 \author[7]{I.~Levine}
+
+\author[13]{C.~Licciardi}
 
@@
 
+\author[13]{E.~Weima}
+
 \author[13]{U.~Wichoski}
@@
 
-\author[5]{J.~Zhang \thanks{Argonne National Laboratory}}
+\author[5]{J.~Zhang\thanks{now at Argonne National Laboratory}}

### 2018-11-27 documentation/paper.tex
Added more authors and email
Final changes

@@
 \usepackage{hyperref}
-\usepackage{lineno}
-
-\DeclareUnicodeCharacter{0301}{'}
-
-\linenumbers
 
@@
 
-Superheated fluid detectors have an intrinsic insensitivity to the relatively diffuse energy depositions from electron recoils \cite{pico}, and neutron backgrounds are mitigated by shielding the detector volume in a 20 t water tank.
+Superheated fluid detectors have an intrinsic insensitivity to the relatively diffuse energy depositions from electron recoils \cite{pico}, and neutron backgrounds are mitigated by shielding the detector volume in a 20~t water tank.
 Alpha decays along the U/Th chains present the dominant remaining background which must be distinguished from nuclear recoils acoustically \cite{isotopes}.
@@
     \item The use of the exact same input format as AP makes the algorithms more directly comparable.
-    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because of the different position distributions of the two classes.
+    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because there are different position distributions of the two classes; neutron events during calibration runs occur more frequently near the wall of the detector where the calibration source is placed.
+
+    Note that it is possible some this bias leaks through to some degree if position corrections are not perfect; acoustics may be noticeably different near the walls of the vessel in particular.
 \end{enumerate}
@@
 Matplotlib \cite{matplotlib} was used for data visualization.
+
+Implementation code for all algorithms discussed in this paper can be found at \url{https://github.com/brendon-ai/dark-matter}.
 

### 2018-11-27 documentation/paper.tex
Clarified biases

@@
 
-\subsubsection{Original Machine Learning Analysis}
+\subsubsection{Original Machine Learning Analysis} \label{og_analysis}
 
@@
 In addition to neutrons and alpha particles, gamma events were used as training and testing data for this neural network.
+
+Note that it is possible there are position-related biases present in this training data, because neutron events during calibration runs occur more frequently near the wall of the detector where the calibration source is placed.
 
@@
     \item The use of the exact same input format as AP makes the algorithms more directly comparable.
-    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because there are different position distributions of the two classes; neutron events during calibration runs occur more frequently near the wall of the detector where the calibration source is placed.
-
-    Note that it is possible some this bias leaks through to some degree if position corrections are not perfect; acoustics may be noticeably different near the walls of the vessel in particular.
+    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because there are different position distributions of the two classes (as described in Section \ref{og_analysis}).
 \end{enumerate}

### 2018-11-27 documentation/paper.tex
Switched to BibTeX

@@
 \usepackage{authblk}
-\usepackage[backend=biber, style=alphabetic, sorting=ynt]{biblatex}
+\usepackage[backend=bibtex, style=alphabetic, sorting=ynt]{biblatex}
 \usepackage{graphicx}
@@
 This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
-$PosCor(\beta _{8}, \chi)$ is used for training several machine learning models throughout this study.
+$PosCor(\beta _{8}, \chi)$ is used for training several machine learning models throughout this study, because it is expected to alleviate the biases mentioned in Section \ref{og_analysis}.
 Details of its use are in Section \ref{mlp}.

### 2018-11-27 documentation/paper.tex
Language clarifications

--- 2018-11-27 documentation/paper.tex
@@
 Note that it is possible there are position-related biases present in this training data, because neutron events during calibration runs occur more frequently near the wall of the detector where the calibration source is placed.
+These biases are expected to be alleviated by application of position corrections which normalize the audio amplitude; these corrections are detailed in Section \ref{piezo_derived}.
 

### 2018-11-30 documentation/paper.tex
Final updates for arXiv
Removed number from Overview
Updated text to get rid of GT comparison; updated first graph

@@
 
-\setcounter{section}{-1}
-
-\section{Structure}
+\section*{Overview}
 
@@
     \item Determine whether these techniques can be extended, applying two original algorithms based on semi-supervised learning, to develop an effective discriminator function based on incomplete and inaccurate initial information.
+Furthermore, determine whether semi-supervised learning is of any benefit.
 \end{enumerate}
@@
 Numerically, this corresponds to the number of validation examples the network classifies correctly.
-The baseline for correctness is the run type during supervised learning experiments, and the AP prediction for semi-supervised experiments.
+The baseline for correctness is the AP prediction.
 
@@
 
-Numerically, this is a variable defined below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively (or calibration and background sets respectively, depending on which is used as ground truth data).
+Numerically, this is a variable defined below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
 It calculates the spread of the network's predictions for each ground truth class.
@@
 
+Accuracy statistics cited throughout the rest of this paper are averaged over late epochs (after accuracy has reached an asymptote) of several configurations in the highest-accuracy region of the hyperparameter search space, rather than over a single configuration.
+This alleviates the accuracy variance caused by randomly chosen validation sets.
+
 \section{Supervised Learning}
@@
     \item A convolutional neural network trained on the image window data $\iota$.
-\end{enumerate}
-
-All supervised learning configurations were trained and evaluated based on their ability to determine the origin of events: whether they are from neutron calibration source runs (predominantly neutrons) or background radiation runs (predominantly alpha particles).
-This metric was used for training and testing instead of AP for two reasons:
-
-\begin{enumerate}
-    \item During early operations of future experiments, when no AP equivalent is yet available, impure data is the only information available for training.
-The network structure should be selected to accommodate this.
-    \item In general, supervised learning systems learn to replicate biases present in the training data set.
-Thus, to evaluate the network's effectiveness at processing the input data, the run type (which is used for training) must be used as a baseline instead of AP (which is never provided to the network during the training process).
 \end{enumerate}
@@
 
-The most accurate configuration during a waveform CNN grid search was 95\% accurate and had a standard deviation of 0.52 (seen in Figure \ref{waveform_hist}).
+The most accurate configuration during a waveform CNN grid search was 93.8\% accurate and had a standard deviation of 0.43 (seen in Figure \ref{waveform_hist}).
+The average accuracy over 3 configurations in the best-performing region of the hyperparameter search space (with 24 filters, L2 $\lambda$ of 0.003, 32 convolutional layers, and a kernel size of 3) was 87.5\%.
+Performance statistics of these configurations are shown in Table \ref{deepconv_table}.
 
@@
     \includegraphics[width=\textwidth]{waveform_hist}
-    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ discriminator.
-(Validation Event Count is the number of events in the validation set that fall within a certain network prediction band.)}
+    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ configuration.
+Validation Event Count is the number of events in the validation set that fall within a certain network prediction band.
+The black vertical line represents the AP decision boundary.}
 \end{figure}
-
-Table \ref{deepconv_table} shows the results of several {\it DeepConv}$(\omega)$ discriminators produced throughout testing, with corresponding network hyperparameters.
 
@@
     \begin{center}
-        \captionof{table}{Performance of {\it DeepConv}$(\omega)$ configurations.} \label{deepconv_table} 
+        \captionof{table}{Performance of {\it DeepConv}$(\omega)$ configurations in the best-performing region of the hyperparameter search space.} \label{deepconv_table} 
         \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
             \hline
-            L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Max Accuracy & Precision & Recall & CWSD \\
-            \hline
-            0.003 & 0 & 24 & 3 & 17 & 95\% & 97\% & 97\% & 0.52 \\
-            % ../experimental_data/validation_results/waveform_grid_search/l2_lambda0.003/dense_dropout0/first_layer_filters24/kernel_size3/convolutional_layers_per_group3/time1539990193_epoch57.json
-            \hline
-            0.003 & 0.25 & 48 & 3 & 17 & 94\% & 100\% & 91\% & 0.44 \\
-            \hline
-            0.001 & 0.25 & 24 & 5 & 17 & 95\% & 99\% & 94\% & 0.46 \\
-            \hline
-            0.001 & 0.25 & 24 & 5 & 32 & 91\% & 94\% & 94\% & 0.66 \\
-            \hline
-            0.0003 & 0 & 48 & 5 & 32 & 88\% & 94\% & 89\% & 0.61 \\
-            \hline
-            0.0003 & 0 & 24 & 3 & 17 & 93\% & 96\% & 95\% & 0.59 \\
+            L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Accuracy & Precision & Recall & CWSD \\
+            \hline
+            0.003 & 0 & 24 & 3 & 32 & 93.8\% & 94.0\% & 96.3\% & 0.50 \\
+            \hline
+            0.003 & 0.25 & 24 & 3 & 32 & 93.8\% & 94.0\% & 96.3\% & 0.43 \\
+            \hline
+            0.003 & 0.5 & 24 & 3 & 32 & 75.0\% & 71.9\% & 100.0\% & 0.71 \\
             \hline
@@
 \end{minipage}
-
-Table \ref{deepconv_table} shows that simple network architectures seem to produce better accuracy, likely due to reduced overfitting.
-Larger kernel sizes, more filters, and more layers make performance generally worse, while higher L2 regularization does the opposite.
-Dropout seems to have a weak correlation, possibly reducing accuracy slightly.
+% cat ../experimental_data/training_logs/2018_11_24_waveform_ap_similarity_last_epoch.out | grep filters24 | grep lambda0.003 | grep group6 | grep kernel_size3 | ./mean_ap_accuracy_across_configurations.py
 

### 2018-11-30 visualization/graph_binary_predictions.py

@@ -99,5 +99,7 @@ plt.legend(handles=[background_patch, calibration_patch])
 # Label the X and Y axes
 plt.xlabel('Logarithmic Acoustic Parameter')
 plt.ylabel('Neural Network Prediction')
+# Draw a vertical line to represent the AP decision boundary
+plt.axvline(0.25, color='black')
 # Display the graph on screen
 plt.show()

### 2018-11-30 documentation/paper.tex
Added updated data for FFT network

@@
 
-In addition to the banded Fourier transform $\beta_{8}$ used in the original PICO-60 analysis, the full-resolution $\beta _{50,001}$ was input into a neural network directly, without any downscaling.
-
-When $\beta_{8}$ was used as input, position corrections were optionally applied to the input data, as they were the input to Acoustic Parameter.
+In addition to the banded Fourier transform $\beta_{8}$ used in the original PICO-60 analysis, the full-resolution $\beta_{50,001}$ was input into a neural network directly, without any downscaling.
+
+When $\beta_{8}$ was used as input, position corrections were applied to the input data, as they were used for the input to Acoustic Parameter.
 The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and $FourierMLP(PosCor(\beta_{N}, \chi))$.
+
 While a variety of network architectures were tested, most had a small number of dense layers (on the order of three), applied dropout and L2 regularization, and used batch normalization once again.
@@
     \includegraphics[width=\textwidth]{banded_no_pos_input_hist}
-    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best {\it FourierMLP}$(\beta_{8})$ discriminator.}
+    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best {\it FourierMLP}$(PosCor(\beta_{8}))$ discriminator.}
 \end{figure}
 
-Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators, with and without wall cuts.
+Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators from the most effective region of the hyperparameter search space (L2 $\lambda$ of 0.003 for $\beta_8$ input, and L2 $\lambda$ of 0.001 and dropout of 0 for $\beta_{50,001}$ input).
+It is very clear that higher resolution data does not improve accuracy in a relatively shallow network.
 
@@
     \begin{center}
-        \captionof{table}{Performance of {\it FourierMLP}$(\beta_{N})$ configurations.} \label{fft_wall_cuts}
+        \captionof{table}{Performance of {\it FourierMLP}$(\beta_{N})$ configurations in the best-performing region of the hyperparameter search space.} \label{fft_wall_cuts}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
             \hline
-            Configuration & L2 $\lambda$ & Dropout & Layers & Max Acc & Prec & Recall & CWSD \\
-            \hline
-            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ & 0 & 0.5 & 2 & 96\% & 98\% & 97\% & 0.42 \\
-            % ../experimental_data/validation_results/2018_07_16_banded_frequency_validation_passing_weighted_3_times.json
-            \hline
-            {\it FourierMLP}$(\beta_{8})$ & 0 & 0.5 & 2 & 98\% & 100\% & 98\% & 0.29 \\
-            % ../experimental_data/validation_results/2018_07_17_banded_frequency_not_position_corrected/banded_low_resolution_time1531831056_epoch249.json
-            \hline
-            {\it FourierMLP}$(\beta_{50,001})$ & 0 & 0 & 2 & 95\% & 94\% & 98\% & 0.47 \\
-            % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input//dropout0/l2_lambda0/hidden_layers1/time1540053737_epoch170.json
-            \hline
-            {\it FourierMLP}$(\beta_{50,001})$ & 0.001 & 0 & 4 & 95\% & 95\% & 98\% & 0.46 \\
-            % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0/l2_lambda0.001/hidden_layers3/time1540054370_epoch174.json
-            \hline
-            {\it FourierMLP}$(\beta_{50,001})$ & 0.001 & 0.5 & 4 & 94\% & 93\% & 98\% & 0.54 \\
-            % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0.5/l2_lambda0.001/hidden_layers3/time1540056735_epoch50.json
-            \hline
-            {\it FourierMLP}$(\beta_{50,001})$ & 0.003 & 0.25 & 3 & 93\% & 92\% & 98\% & 0.57 \\
-            % ../experimental_data/validation_results/2018_10_20_high_resolution_grid_search_without_position_input/dropout0.25/l2_lambda0.003/hidden_layers2/time1540055639_epoch113.json
+            Input Format & L2 $\lambda$ & Dropout & Layers & Accuracy & Precision & Recall & CWSD \\
+            \hline
+            $\beta_8$ & 0.003 & 0 & 2 & 96.2\% & 94.1\% & 99.8\% & 0.28 \\
+            \hline
+            $\beta_8$ & 0.003 & 0.25 & 2 & 96.7\% & 94.9\% & 99.5\% & 0.27 \\
+            \hline
+            $\beta_8$ & 0.003 & 0.5 & 2 & 96.7\% & 95.0\% & 99.9\% & 0.30 \\
+            \hline
+            $\beta_{50,001}$ & 0.001 & 0 & 2 & 89.1\% & 84.0\% & 100.0\% & 0.47 \\
+            \hline
+            $\beta_{50,001}$ & 0.001 & 0 & 3 & 92.2\% & 88.8\% & 100.0\% & 0.44 \\
+            \hline
+            $\beta_{50,001}$ & 0.001 & 0 & 4 & 93.0\% & 89.8\% & 100.0\% & 0.40 \\
             \hline
@@
 \end{minipage}
-
-It is very clear that higher resolution data does not improve accuracy in a relatively shallow network. Regularization, both dropout and L2, appears to have relatively little effect.
+% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples256 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_11_24_high_resolution_ap_similarity_last_epoch.out | grep dropout0/ | grep l2_lambda0.001 | ./mean_ap_accuracy_across_configurations.py
 

### 2018-11-30 documentation/paper.tex
Updated graphs with enforced range of 0 to 1

@@
     \includegraphics[width=\textwidth]{banded_no_pos_input_hist}
-    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best {\it FourierMLP}$(PosCor(\beta_{8}))$ discriminator.}
+    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best {\it FourierMLP}$(${\it PosCor}$(\beta_{8}))$ configuration.}
 \end{figure}
 
-Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators from the most effective region of the hyperparameter search space (L2 $\lambda$ of 0.003 for $\beta_8$ input, and L2 $\lambda$ of 0.001 and dropout of 0 for $\beta_{50,001}$ input).
+Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators from the most effective region of the hyperparameter search space (L2 $\lambda$ of 0.003 and 2 layers for $\beta_8$ input, and L2 $\lambda$ of 0.001 and dropout of 0 for $\beta_{50,001}$ input).
 It is very clear that higher resolution data does not improve accuracy in a relatively shallow network.

### 2018-11-30 visualization/graph_binary_predictions.py

@@ -101,5 +101,7 @@ plt.xlabel('Logarithmic Acoustic Parameter')
 plt.ylabel('Neural Network Prediction')
 # Draw a vertical line to represent the AP decision boundary
 plt.axvline(0.25, color='black')
+# Restrict the network prediction axis to the range of 0 to 1
+plt.ylim(0, 1)
 # Display the graph on screen
 plt.show()

### 2018-11-30 documentation/paper.tex
Removed irrelevant sentence about histogram
Updated information for image network

@@
     \includegraphics[width=\textwidth]{waveform_hist}
-    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ configuration.
-Validation Event Count is the number of events in the validation set that fall within a certain network prediction band.
-The black vertical line represents the AP decision boundary.}
+    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ configuration. The black vertical line represents the AP decision boundary.}
 \end{figure}
@@
 
-Throughout a variety of different network architectures, the best validation accuracy obtained was 69\%.
-The corresponding training accuracy of 100\% indicates that severe overfitting is taking place (not unexpected, given the relatively small image data set).
-Also, the class-wise standard deviation was a very high 0.99, which implies its decisions are nearly random (as seen in Figure \ref{image_hist}).
-It is likely fitting poorly on some form of noise within $\iota$.
-The fact that throughout many trials during a grid search, no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.
+Every network architecture tested produced validation accuracy between 62\% and 63\%.
+As seen in Figure \ref{image_hist}, all examples were given the exact same prediction.
+This provides strong evidence that images provide insufficient information for any kind of meaningful discrimination.
 
@@
     \includegraphics[width=\textwidth]{image_hist}
-    \caption{\label{image_hist} Prediction distribution of one of the best {\it ImageConv}$(\iota)$ discriminators. (Note the truncated X axis.)}
+    \caption{\label{image_hist} Prediction distribution of an {\it ImageConv}$(\iota)$ configuration. (53 out of 54 configurations had exactly the same accuracy.)}
 \end{figure}
@@
         \captionof{table}{Performance of {\it ImageConv}$(\iota)$ configurations.} \label{imageconv_table}
-        \begin{tabular}{|l|l|l|l|l|l|l|l|}
-            \hline
-            L2 $\lambda$ & Dropout & Convolutional Layers & Dense Layers & Max Accuracy & Precision & Recall & CWSD \\
-            \hline
-            0 & 0 & 6 & 3 & 69\% & 69\% & 100\% & 0.99 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0_dropout0_convolutional_layers_per_group2time1532891695_epoch15.json 
-            \hline
-            0.01 & 0.25 & 6 & 3 & 68\% & 68\% & 100\% & 0.98 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group2time1532895825_epoch2.json
-            \hline
-            0.003 & 0.25 & 9 & 3 & 68\% & 68\% & 100\% & 0.98 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.003_dropout0.25_convolutional_layers_per_group3time1532895118_epoch3.json
-            \hline
-            0.01 & 0.25 & 12 & 3 & 68\% & 68\% & 100\% & 0.97 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group4time1532896126_epoch0.json
-            \hline
-            0.01 & 0.5 & 12 & 3 & 68\% & 68\% & 100\% & 0.99 \\
-            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.5_convolutional_layers_per_group4time1532898944_epoch1.json
+        \begin{tabular}{|l|l|l|l|l|l|l|}
+            \hline
+            L2 $\lambda$ & Dropout & Convolutional Layers & Accuracy & Precision & Recall & CWSD \\
+            \hline
+            0 & 0 & 6 & 62.7\% & 62.9\% & 99.5\% & 1.00 \\
+            \hline
+            0.0003 & 0 & 12 & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
+            \hline
+            0.001 & 0 & 9 & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
             \hline

### 2018-11-30 documentation/paper.tex
Updated supervised learning summary table

@@
 When $\beta_{8}$ was used as input, position corrections were applied to the input data, as they were used for the input to Acoustic Parameter.
-The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and $FourierMLP(PosCor(\beta_{N}, \chi))$.
+(With the position correction method currently used for PICO-60, it is currently not practical to apply it to other input formats.)
+
+The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and \\ $FourierMLP(PosCor(\beta_{N}, \chi))$.
 
@@
 It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
-{\it FourierMLP}$(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
+{\it FourierMLP}$(PosCor(\beta_{8}))$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
 Its very decisive prediction distribution can be observed in Figure \ref{banded_no_pos_input_hist}.
-
-Very interestingly, {\it FourierMLP}$(\beta_{8})$ performed better than {\it FourierMLP}$(${\it PosCor}$(\beta_{8}), \chi))$, which produced an accuracy value of 96\%.
-Both of these results were better than any of the network architectures trained on $\omega$.
 
@@
 
-Table \ref{super_overview} compares the accuracy and mean class-wise standard deviation values of certain successful supervised learning configurations with multiple input formats.
+Table \ref{super_overview} compares the mean accuracy, precision, recall, and class-wise standard deviation values of each of the input formats.
 
@@
     \begin{center}
-        \captionof{table}{Summary of supervised learning configurations.} \label{super_overview}
+        \captionof{table}{Comparison of supervised learning input formats.} \label{super_overview}
         \begin{tabular}{|l|l|l|l|l|}
             \hline
-            Configuration & Max Accuracy & Precision & Recall & CWSD \\
-            \hline
-            {\it DeepConv}$(\omega)$ & 95\% & 97\% & 97\% & 0.52 \\
-            \hline
-            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}))$ & 96\% & 98\% & 97\% & 0.42 \\
-            \hline
-            {\it FourierMLP}$(\beta_{8})$ & 98\% & 100\% & 98\% & 0.29 \\
-            \hline
-            {\it FourierMLP}$(\beta_{50,001})$ & 95\% & 95\% & 98\% & 0.46 \\
-            \hline
-            {\it ImageConv}$(\iota)$ & 68\% & 68\% & 100\% & 0.97 \\
+            Configuration & Accuracy & Precision & Recall & CWSD \\
+            \hline
+            {\it DeepConv}$(\omega)$ & 87.5\% & 71.9\% & 100.0\% & 0.55 \\
+            \hline
+            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ & 96.5\% & 94.6\% & 99.7\% & 0.28 \\
+            \hline
+            {\it FourierMLP}$(\beta_{50,001})$ & 91.4\% & 87.8\% & 100.0\% & 0.44 \\
+            \hline
+            {\it ImageConv}$(\iota)$ & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
             \hline

### 2018-11-30 statistics/mean_ap_accuracy_across_configurations.py

@@ -32,8 +32,11 @@ for line in statistic_lines:
     # Add the statistics tuple to the corresponding list in the dictionary
     statistics[configuration].append(statistics_tuple)
 
-# Create a list to add mean accuracy values to
+# Create lists to add mean statistics to
 mean_accuracy_values = []
+mean_precision_values = []
+mean_recall_values = []
+mean_cwsd_values = []
 # Iterate over the configuration keys of the dictionary
 for configuration in statistics:
     # Zip the list of statistics and calculate the mean of each individual statistic
@@ -41,7 +44,14 @@ for configuration in statistics:
     # Convert the disagreement value to accuracy, and add it to the list for mean calculation
     mean_accuracy = 1 - (mean_disagreements / VALIDATION_EXAMPLES)
     mean_accuracy_values.append(mean_accuracy)
+    # Add other statistics to their respective lists
+    mean_precision_values.append(mean_precision)
+    mean_recall_values.append(mean_recall)
+    mean_cwsd_values.append(mean_standard_deviation)
     # Output the relevant statistics for this configuration to the user
     print('Configuration:', configuration, 'Accuracy:', mean_accuracy, 'CWSD:', mean_standard_deviation, 'Precision:', mean_precision, 'Recall:', mean_recall)
-# Print the mean accuracy over all configurations
+# Print the mean statistics over all configurations
 print('Overall mean accuracy:', sum(mean_accuracy_values) / len(mean_accuracy_values))
+print('Overall mean precision:', sum(mean_precision_values) / len(mean_precision_values))
+print('Overall mean recall:', sum(mean_recall_values) / len(mean_recall_values))
+print('Overall mean CWSD:', sum(mean_cwsd_values) / len(mean_cwsd_values))

### 2018-12-01 statistics/grid_search_ap_similarity_statistics.py
Added AP similarity data for last 300 epochs (all training examples)

@@ -50,9 +50,7 @@ file_paths = [path for path in glob.glob(os.path.expanduser(sys.argv[1]), recurs
 # Filter the paths, taking only the ones that are in the last few epochs
 epochs = [int(path.split('epoch')[1].split('.json')[0]) for path in file_paths]
 max_epoch = max(epochs)
-file_paths = [path for path, epoch in zip(file_paths, epochs) if epoch == max_epoch]
-# Temporary hack: only take those file paths that represent validation sets and not gravitational training sets
-file_paths = [path for path in file_paths if 'validation' in path]
+file_paths = [path for path, epoch in zip(file_paths, epochs) if epoch > max_epoch - 300]
 # Get the run identifier corresponding to each of the file paths (not including the specific epoch)
 run_identifiers = [file_path.split('time')[0] for file_path in file_paths]
 # Convert the identifiers to a set so that they are all unique (there are many duplicated)

### 2018-12-01 documentation/paper.tex
Lots of clarifications of various things

@@
 
+(Side note: intuitively, it would make sense to provide the position $\chi$ directly to the neural network when corrections cannot be applied.
+However, this cannot presently be done because $\chi$ is a bias for the calibration run type, as described in Section \ref{og_analysis}; neutron sources are placed on one side of the vessel, so neutron events occur more often near that side.)
+
 The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and \\ $FourierMLP(PosCor(\beta_{N}, \chi))$.
@@
 
-Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators from the most effective region of the hyperparameter search space (L2 $\lambda$ of 0.003 and 2 layers for $\beta_8$ input, and L2 $\lambda$ of 0.001 and dropout of 0 for $\beta_{50,001}$ input).
-It is very clear that higher resolution data does not improve accuracy in a relatively shallow network.
+Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators from the best-performing region of the hyperparameter search space.
 
@@
             \hline
-            $\beta_8$ & 0.003 & 0 & 2 & 96.2\% & 94.1\% & 99.8\% & 0.28 \\
-            \hline
-            $\beta_8$ & 0.003 & 0.25 & 2 & 96.7\% & 94.9\% & 99.5\% & 0.27 \\
-            \hline
-            $\beta_8$ & 0.003 & 0.5 & 2 & 96.7\% & 95.0\% & 99.9\% & 0.30 \\
+            $\beta_8$ & 0 & 0.5 & 2 & 98.1\% & 97.1\% & 100.0\% & 0.24 \\
+            \hline
+            $\beta_8$ & 0.001 & 0.5 & 2 & 96.6\% & 94.7\% & 99.6\% & 0.27 \\
+            \hline
+            $\beta_8$ & 0.003 & 0.5 & 2 & 96.3\% & 94.3\% & 100.0\% & 0.28 \\
             \hline
@@
 \end{minipage}
-% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples256 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_12_01_banded_all_examples_ap_similarity_last_300_epochs.out | grep dropout0.5 | ./mean_ap_accuracy_across_configurations.py
 % cat ../experimental_data/training_logs/2018_11_24_high_resolution_ap_similarity_last_epoch.out | grep dropout0/ | grep l2_lambda0.001 | ./mean_ap_accuracy_across_configurations.py
@@
 
-\subsection{Performance Summary}
-
-Table \ref{super_overview} compares the mean accuracy, precision, recall, and class-wise standard deviation values of each of the input formats.
+\subsection{Performance Comparison}
+
+Table \ref{super_overview} compares the mean accuracy, precision, recall, and class-wise standard deviation values (of the highest-performance region of the hyperparameter search space) for each of the input formats.
 
@@
             \hline
-            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ & 96.5\% & 94.6\% & 99.7\% & 0.28 \\
+            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ & 97.0\% & 95.4\% & 99.9\% & 0.26 \\
             \hline
@@
 The key conclusion to be drawn from the above table is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
-Higher resolutions seem to make performance worse, likely due to overfitting.
+The raw waveform and high-resolution Fourier transform make performance worse.
+This may be caused by one of the following three reasons:
+
+\begin{itemize}
+    \item Insufficient network depth to process higher resolutions
+    \item Overfitting on extraneous data
+    \item Position corrections are essential to good performance
+\end{itemize}
+
+The last reason is very likely.
+At a fundamental level, $\beta _8$ does not contain any more information than $\beta _{50,001}$, so there is no obvious reason why (with the right network architecture) it should not perform at least as well.
+While repurposing the position corrections to work on arbitrary-resolution data is outside the scope of this study, it seems a promising subject of future investigation.
 
 Based on this result, experimentation with semi-supervised learning proceeded using \\ {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$, applying a three-layer perceptron architecture.
-There were three main reasons for this:
+There were two main reasons for this:
 
 \begin{enumerate}
-    \item This was a highly effective configuration, producing 96\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
-    \item The use of the exact same input format as AP makes the algorithms more directly comparable.
-    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because there are different position distributions of the two classes (as described in Section \ref{og_analysis}).
+    \item This was a highly effective configuration, producing 97.0\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
+    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because alpha and neutron events have different position distributions (as described in Section \ref{og_analysis}).
 \end{enumerate}

### 2018-12-01 documentation/paper.tex
Updated information for GD

@@
     \includegraphics[width=\textwidth]{waveform_hist}
-    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ configuration. The black vertical line represents the AP decision boundary.}
+    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ configuration, relative to AP, on a randomly selected validation set.
+The black vertical line represents the AP decision boundary.}
 \end{figure}
@@
 Because all data in the PICO-60 experiment is labeled, unlabeled training data sets are randomly sampled from the 496 training examples remaining (after the validation set is selected).
+Labeled set sizes of 32, 64, 128, and 256 were tested; the remainder of the 496 examples would be used as unlabeled data.
 
@@
 
-During a grid search of parameters of the gravitational differentiation algorithm, in addition to the stochastic gradient descent learning rate (a multiplier that defines how quickly and precisely optimization occurs), the highest mean accuracy over three runs of an individual configuration was 99\%.
+During a grid search of parameters of the gravitational differentiation algorithm, in addition to .
 The highest accuracy was obtained with 256 labeled examples in the set $\varsigma$ (out of 496 total training examples), which equates to 52\% of the data being labeled.
@@
 
+As with supervised learning techniques, a grid search was run for various parameters of the gravitational differentiation algorithm, in addition to the stochastic gradient descent learning rate (a multiplier that defines how quickly and precisely optimization occurs).
+Table \ref{grav_accuracy_by_varsigma} shows the mean asymptotic accuracy over the best-performing region of the hyperparameter search space, for each size of the initial labeled set $\varsigma$.
+Note that optimal hyperparameters vary by set size.
+
+It is evident that larger labeled sets produce better performance.
+However, note that a $\varsigma$ size of 128 produces accuracy of 97.0\%, the same as the banded supervised system with all 496 labeled examples.
+That is, GD achieves the same accuracy with 74\% of the labels removed.
+
 \begin{minipage}{\textwidth}
@@
         \captionof{table}{Gravitational differentiation accuracy statistics by size of $\varsigma$.} \label{grav_accuracy_by_varsigma}
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            Initial $\varsigma$ Size & Proportion of Data Labeled & Max Accuracy & Mean Accuracy \\
-            \hline
-            32 & 6\% & 83\% & 69\% \\
-            \hline
-            64 & 13\% & 97\% & 73\% \\
-            \hline
-            128 & 26\% & 99\% & 84\% \\
-            \hline
-            256 & 52\% & 99\% & 94\% \\
+        \begin{tabular}{|l|l|l|l|l|l|}
+            \hline
+            Initial $\varsigma$ Size & Proportion of Data Labeled & Accuracy & Precision & Recall & CWSD \\
+            \hline
+            32 & 6\% & 71.3\% & 70.1\% & 98.0\% & 0.73 \\
+            \hline
+            64 & 13\% & 79.7\% & 75.5\% & 98.6\% & 0.63 \\
+            \hline
+            128 & 26\% & 97.0\% & 96.3\% & 99.6\% & 0.23 \\
+            \hline
+            256 & 52\% & 99.2\% & 99.0\% & 99.7\% & 0.11 \\
             \hline
@@
 \end{minipage}
-
-Table \ref{grav_overview} represents the performance of several high-accuracy gravitational differentiation configurations tested during the grid search.
-The accuracy statistic is the mean over three tests of the same configuration.
+% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples256 | grep rate0.03 | grep power3 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples128 | grep rate0.03 | grep increment0.0005 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples64 | grep power11 | grep rate0.001 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples32 | grep power11 | grep increment0.001 | ./mean_ap_accuracy_across_configurations.py
+
+Table \ref{grav_overview} represents the performance of several of the individual configurations incorporated into the mean accuracy values in Table \ref{grav_accuracy_by_varsigma}.
+All accuracy statistics are the mean over three tests of the same configuration.
 
@@
     \begin{center}
-        \captionof{table}{Performance of gravitational differentiation configurations.} \label{grav_overview}
+        \captionof{table}{Performance of gravitational differentiation configurations in the best-performing region of the hyperparameter search space.} \label{grav_overview}
         \begin{tabular}{|l|l|l|l|l|l|l|l|}
             \hline
-            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Mean AP Accuracy & Precision & Recall & CWSD \\
-            \hline
-            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% & 90\% & 0.78 \\
-            \hline
-            32 & 0.0005 & 0.003 & 9 & 83\% & 80\% & 99\% & 0.63 \\
-            \hline
-            64 & 0.0005 & 0.003 & 9 & 77\% & 74\% & 97\% & 0.71 \\
-            \hline
-            64 & 0.003 & 0.03 & 7 & 97\% & 96\% & 100\% & 0.18 \\
-            \hline
-            128 & 0.008 & 0.001 & 9 & 74\% & 71\% & 95\% & 0.70 \\
-            \hline
-            128 & 0.0005 & 0.03 & 11 & 99\% & 99\% & 99\% & 0.17 \\
-            \hline
-            256 & 0.005 & 0.01 & 11 & 97\% & 96\% & 99\% & 0.25 \\
-            \hline
-            256 & 0.001 & 0.03 & 3 & 99\% & 99\% & 100\% & 0.10 \\
+            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Accuracy & Precision & Recall & CWSD \\
+            \hline
+            32 & 0.001 & 0.001 & 11 & 79.0\% & 78.2\% & 94.8\% & 0.74 \\
+            \hline
+            32 & 0.001 & 0.003 & 11 & 59.9\% & 59.9\% & 100.0\% & 0.78 \\
+            \hline
+            64 & 0.005 & 0.001 & 11 & 82.8\% & 79.0\% & 99.1\% & 0.58 \\
+            \hline
+            64 & 0.008 & 0.001 & 11 & 73.9\% & 71.0\% & 99.6\% & 0.71 \\
+            \hline
+            128 & 0.0005 & 0.03 & 3 & 98.3\% & 97.7\% & 99.6\% & 0.21 \\
+            \hline
+            128 & 0.0005 & 0.03 & 9 & 96.6\% & 95.1\% & 99.6\% & 0.29 \\
+            \hline
+            256 & 0.008 & 0.03 & 3 & 99.1\% & 99.0\% & 99.5\% & 0.16 \\
+            \hline
+            256 & 0.001 & 0.03 & 3 & 99.7\% & 99.6\% & 99.6\% & 0.05 \\
             \hline
@@
     \includegraphics[width=\textwidth]{grav_grid_search}
-    \caption{\label{grav_grid_search} Predictions of best gravitational differentiation model compared to AP (on randomly selected validation set).}
+    \caption{\label{grav_grid_search} Prediction distribution of the best gravitational differentiation configuration.}
 \end{figure}

### 2018-12-01 visualization/graph_binary_predictions.py

@@ -101,7 +101,7 @@ plt.xlabel('Logarithmic Acoustic Parameter')
 plt.ylabel('Neural Network Prediction')
 # Draw a vertical line to represent the AP decision boundary
 plt.axvline(0.25, color='black')
-# Restrict the network prediction axis to the range of 0 to 1
-plt.ylim(0, 1)
+# # Restrict the network prediction axis to the range of 0 to 1
+# plt.ylim(0, 1)
 # Display the graph on screen
 plt.show()

### 2018-12-01 documentation/paper.tex
Added data for ICN

@@
             \hline
-            Initial $\varsigma$ Size & Proportion of Data Labeled & Accuracy & Precision & Recall & CWSD \\
+            $\varsigma$ Size & Proportion of Data Labeled & Accuracy & Precision & Recall & CWSD \\
             \hline
@@
         \captionof{table}{Iterative cluster nucleation accuracy statistics by initial size of $\varsigma$.} \label{icn_accuracy_by_varsigma}
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            Initial $\varsigma$ Size & Proportion of Data Labeled & Max Accuracy & Mean Accuracy \\
-            \hline
-            32 & 6\% & 95\% & 75\% \\
-            \hline
-            64 & 13\% & 97\% & 84\% \\
-            \hline
-            128 & 26\% & 98\% & 95\% \\
-            \hline
-            256 & 52\% & 99\% & 97\% \\
+        \begin{tabular}{|l|l|l|l|l|l|}
+            \hline
+            Initial $\varsigma$ Size & Proportion of Data Labeled & Accuracy & Precision & Recall & CWSD \\
+            \hline
+            32 & 6\% & 85.2\% & 83.4\% & 98.4\% & 0.54 \\
+            \hline
+            64 & 13\% & 94.3\% & 92.2\% & 99.8\% & 0.34 \\
+            \hline
+            128 & 26\% & 98.2\% & 97.7\% & 99.4\% & 0.21 \\
+            \hline
+            256 & 52\% & 98.2\% & 97.6\% & 99.5\% & 0.22 \\
             \hline
@@
 \end{minipage}
+% cat ../experimental_data/training_logs/2018_10_26_icn_ap_similarity_last_300_epochs.out | grep examples256 | grep dropout0.5 | grep lambda0.001/ | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_icn_ap_similarity_last_300_epochs.out | grep examples128 | grep threshold0.02 | grep multiplier1.05 | grep dropout0.5 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_icn_ap_similarity_last_300_epochs.out | grep examples64 | grep threshold0.02 | grep lambda0/ | grep multiplier1.05 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_icn_ap_similarity_last_300_epochs.out | grep examples32 | grep lambda0/ | grep dropout0/ | ./mean_ap_accuracy_across_configurations.py
 
@@
             \hline
-            Initial $\varsigma$ Size & L2 $\lambda$ & Dropout & Initial $j$ & $k$ & Mean AP Accuracy & Precision & Recall & CWSD \\
-            \hline
-            32 & 0 & 0.25 & 0.01 & 1.025 & 80\% & 79\% & 99\% & 0.60 \\
-            \hline
-            32 & 0 & 0 & 0.01 & 1.025 & 95\% & 94\% & 99\% & 0.38 \\
-            \hline
-            64 & 0.001 & 0 & 0.01 & 1.05 & 97\% & 96\% & 99\% & 0.28 \\
-            \hline
-            64 & 0 & 0.25 & 0.02 & 1.025 & 97\% & 95\% & 100\% & 0.27 \\
-            \hline
-            128 & 0 & 0.5 & 0.01 & 1.05 & 92\% & 90\% & 100\% & 0.35 \\
-            \hline
-            128 & 0 & 0.5 & 0.02 & 1.05 & 98\% & 98\% & 100\% & 0.21 \\
-            \hline
-            256 & 0 & 0.5 & 0.01 & 1.05 & 99\% & 99\% & 100\% & 0.15 \\
-            \hline
-            256 & 0.003 & 0.5 & 0.01 & 1.05 & 98\% & 98\% & 98\% & 0.26 \\
+            Initial $\varsigma$ Size & L2 $\lambda$ & Dropout & Initial $j$ & $k$ & Accuracy & Precision & Recall & CWSD \\
+            \hline
+            32 & 0 & 0 & 0.01 & 1.025 & 94.9\% & 94.2\% & 98.3\% & 0.40 \\
+            \hline
+            32 & 0 & 0 & 0.02 & 1.05 & 79.7\% & 79.0\% & 96.1\% & 0.66 \\
+            \hline
+            64 & 0 & 0 & 0.02 & 1.05 & 95.4\% & 93.0\% & 99.9\% & 0.34 \\
+            \hline
+            64 & 0 & 0.5 & 0.02 & 1.05 & 93.5\% & 91.2\% & 100.0\% & 0.31 \\
+            \hline
+            128 & 0.001 & 0.5 & 0.02 & 1.05 & 97.7\% & 97.4\% & 99.2\% & 0.21 \\
+            \hline
+            128 & 0 & 0.5 & 0.02 & 1.05 & 98.5\% & 98.0\% & 99.6\% & 0.21 \\
+            \hline
+            256 & 0.001 & 0.5 & 0.01 & 1.025 & 97.5\% & 96.7\% & 99.3\% & 0.28 \\
+            \hline
+            256 & 0.001 & 0.5 & 0.02 & 1.05 & 98.4\% & 97.6\% & 100.0\% & 0.20 \\
             \hline
@@
 
-\subsection{Performance Summary}
+\subsection{Performance Comparison}
 
 Table \ref{semi_success_overview} directly compares performance of the most successful semi-supervised learning configurations, with AP provided as a baseline for comparison.
-Remember that ``Mean AP Accuracy'' is calculated relative to AP's predictions, rather than replication of the ground truths.
 
@@
             \hline
-            Initial $\varsigma$ Size & Technique & Mean AP Accuracy & Mean Precision & Mean Recall & Mean CWSD \\
+            Initial $\varsigma$ Size & Technique & Accuracy & Precision & Recall & CWSD \\
             \hline

### 2018-12-01 documentation/paper.tex
Completed edits for semi-supervised

@@
 
-\subsubsection{Results}
+\subsubsection{Results} \label{super_results}
 
@@
     \begin{center}
-        \captionof{table}{Comparison of supervised learning input formats.} \label{super_overview}
+        \captionof{table}{Comparison of supervised learning input formats, alongside the original PICO-60 ML analysis and AP.} \label{super_overview}
         \begin{tabular}{|l|l|l|l|l|}
@@
             {\it ImageConv}$(\iota)$ & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
+            \hline
+            PICO-60 ML & 80.7\% & 76.9\% & 100.0\% & 0.65 \\
+            \hline
+            AP & 100\% & 100\% & 100\% & 0.43 \\
             \hline
@@
     \includegraphics[width=\textwidth]{icn_grid_search}
-    \caption{\label{icn_grid_search} Predictions of best iterative cluster nucleation model compared to AP (on randomly selected validation set).}
+    \caption{\label{icn_grid_search} Prediction distribution of the best iterative cluster nucleation configuration.}
 \end{figure}
 
-Once again, the parameters inherent to the ICN algorithm, as well as dropout and L2 regularization, do not have a meaningful polynomial correlation with accuracy.
-Note that there are several hyperparameters not specifically optimized in the context of iterative cluster nucleation, including the number of layers and numbers of neurons within those layers.
-The correlation between those hyperparameters and accuracy is subject to future research.
-
-The mean accuracy for a given initial $\varsigma$ is again important.
-In Table \ref{icn_accuracy_by_varsigma}, it is clear that with small initial $\varsigma$ sets, the training process is very unreliable, much like gravitational differentiation.
-However, the mean accuracy is much closer to the maximum for initial $\varsigma$ sets of 128 and 256, indicating that the training process of ICN is significantly more consistent than gravitational differentiation.
+In Table \ref{icn_accuracy_by_varsigma}, it is clear that for a labeled set of 256, ICN performs worse than GD (which reached 99.2\%).
+However, the accuracy for 128 examples in $\varsigma$ is almost exactly the same as that for 256, and is still higher than the 97.0\% reached with supervised learning on 496 training examples (see Section \ref{super_results}).
+Moreover, for even smaller training sets of 64 and 32, ICN continues to perform better than GD.
 
@@
 
-Table \ref{semi_success_overview} directly compares performance of the most successful semi-supervised learning configurations, with AP provided as a baseline for comparison.
+Table \ref{semi_success_overview} directly compares performance of all semi-supervised learning configurations, with supervised learning, the original PICO-60 ML analysis, and AP provided as baselines for comparison.
+For every $\varsigma$ size except 32, ICN performs better than supervised learning; for 128 and 256, GD also performs better.
+For 32, 64, and 128 labeled examples, ICN proves to be more effective than GD; however, when 256 examples are labeled, GD provides the best accuracy, at 99.2\%.
+
+Perhaps most notably, iterative cluster nucleation with only 128 labeled examples achieves accuracy 1.2 percentage points higher than supervised learning with all 496 training examples labeled.
+It also makes more confident predictions (with a class-wise standard deviation of only 0.21, as opposed to 0.26).
+
+Note that the supervised multi-layer perceptron trained on $\beta _8$ originally used labels for all 496 training examples.
+Here, it is re-trained with smaller training sets so that it can be more directly compared to semi-supervised learning (which does not have access to labels for the entire set).
 
@@
             \hline
-            32 & GD & 83\% & 80\% & 99\% & 0.63 \\
-            \hline
-            32 & ICN & 95\% & 94\% & 99\% & 0.38 \\
-            \hline
-            64 & GD & 97\% & 96\% & 100\% & 0.18 \\
-            \hline
-            64 & ICN & 97\% & 96\% & 99\% & 0.28 \\
-            \hline
-            128 & GD & 99\% & 99\% & 99\% & 0.17 \\
-            \hline
-            128 & ICN & 98\% & 98\% & 100\% & 0.21 \\
-            \hline
-            256 & GD & 99\% & 99\% & 100\% & 0.10 \\
-            \hline
-            256 & ICN & 99\% & 99\% & 100\% & 0.15 \\
-            \hline
-            278 & PICO-60 ML & 80\% & 76\% & 100\% & 0.64 \\
-            \hline
-            N/A & AP & 100\% & 100\% & 100\% & 0.41 \\
+            32 & GD & 71.3\% & 70.1\% & 98.0\% & 0.73 \\
+            \hline
+            32 & ICN & 85.2\% & 83.4\% & 98.4\% & 0.54 \\
+            \hline
+            32 & Supervised & 90.4\% & 89.0\% & 97.1\% & 0.53 \\
+            \hline
+            64 & GD & 79.7\% & 75.5\% & 98.6\% & 0.63 \\
+            \hline
+            64 & ICN & 94.3\% & 92.2\% & 99.8\% & 0.34 \\
+            \hline
+            64 & Supervised & 92.7\% & 90.4\% & 98.9\% & 0.41 \\
+            \hline
+            128 & GD & 97.0\% & 96.3\% & 99.6\% & 0.23 \\
+            \hline
+            128 & ICN & 98.2\% & 97.7\% & 99.4\% & 0.21 \\
+            \hline
+            128 & Supervised & 94.7\% & 92.6\% & 99.5\% & 0.37 \\
+            \hline
+            256 & GD & 99.2\% & 99.0\% & 99.7\% & 0.11 \\
+            \hline
+            256 & ICN & 98.2\% & 97.6\% & 99.5\% & 0.22 \\
+            \hline
+            256 & Supervised & 96.5\% & 94.6\% & 99.7\% & 0.28 \\
+            \hline
+            496 & Supervised & 97.0\% & 95.4\% & 99.9\% & 0.26 \\
+            \hline
+            278 & PICO-60 ML & 80.7\% & 76.9\% & 100.0\% & 0.65 \\
+            \hline
+            N/A & AP & 100\% & 100\% & 100\% & 0.43 \\
             \hline
@@
 \end{minipage}
-
-As observed in this table, semi-supervised learning algorithms are significantly more accurate at replicating AP than the original PICO-60 machine learning analysis.
-Compared to AP, in addition to achieving similar accuracy, configurations from both semi-supervised learning algorithms produce significantly more decisive predictions as measured by the class-wise standard deviation.
-
-Table \ref{average_semi} compares the overall average performance of each of the semi-supervised learning configurations for different labeled set sizes.
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Summary of the average accuracy values for each semi-supervised learning technique.} \label{average_semi}
-        \begin{tabular}{|l|l|l|}
-            \hline
-            Initial $\varsigma$ Size & Technique & Mean Accuracy \\
-            \hline
-            32 & GD & 69\% \\
-            \hline
-            64 & GD & 73\% \\
-            \hline
-            128 & GD & 84\% \\
-            \hline
-            256 & GD & 94\% \\
-            \hline
-            32 & ICN & 75\% \\
-            \hline
-            64 & ICN & 84\% \\
-            \hline
-            128 & ICN & 95\% \\
-            \hline
-            256 & ICN & 97\% \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
+% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples256 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples128 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples64 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
+% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples32 | grep dropout0.25 | ./mean_ap_accuracy_across_configurations.py
 
@@
 Two semi-supervised learning algorithms were developed, applying the best configuration found for supervised learning.
-While both techniques produced some very effective discriminators, iterative cluster nucleation was found to be the more effective of the two.
-As seen in Table \ref{average_semi}, it replicated AP with a mean of 97\% accuracy, using 256 labeled training examples.
-This indicates that the technique shows significant promise.
+Both techniques produced very effective discriminators, performing better than supervised learning.
+Iterative cluster nucleation proved more data-efficient with smaller labeled training sets, while gravitational differentiation consistently produced the highest accuracy overall when 256 labeled examples were used.
 

### 2018-12-02 documentation/paper.tex
Converted PNG graphs to PDF
Lots of minor fixes and deletions

@@
 It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
-{\it FourierMLP}$(PosCor(\beta_{8}))$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29.
-Its very decisive prediction distribution can be observed in Figure \ref{banded_no_pos_input_hist}.
+{\it FourierMLP}$(PosCor(\beta_{8}))$ managed a mean of 97.0\% validation accuracy over the best-performing region of the hyperparameter search space, and a mean class-wise standard deviation of 0.26.
+Its fairly decisive prediction distribution can be observed in Figure \ref{banded_no_pos_input_hist}.
 
@@
     \includegraphics[width=\textwidth]{banded_no_pos_input_hist}
-    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best {\it FourierMLP}$(${\it PosCor}$(\beta_{8}))$ configuration.}
+    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ configuration.}
 \end{figure}
@@
 
-The key conclusion to be drawn from the above table is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
+The key conclusion to be drawn from Table \ref{super_overview} is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
 The raw waveform and high-resolution Fourier transform make performance worse.
@@
 \begin{enumerate}
-    \item This was a highly effective configuration, producing 97.0\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
+    \item This was a highly effective general configuration, producing a mean of 97.0\% accuracy and a class-wise standard deviation of 0.26, much better than the original PICO-60 neural network on that data set.
     \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because alpha and neutron events have different position distributions (as described in Section \ref{og_analysis}).
@@
 \begin{itemize}
-    \item For examples with confident predictions close to 0 or 1: Create a large gradient that pulls the prediction closer to 0 or 1 through gradient descent (like gravity).
+    \item For examples with confident predictions close to 0 or 1: Create a large gradient that pulls the prediction closer to 0 or 1 through gradient descent (analogous to gravity).
     \item For examples with low-confidence predictions close to 0.5: Create a gradient near 0 that does not affect training significantly.
@@
 
-It is a distortion of the hyperbolic tangent that flattens the central range and comparatively exaggerates the asymptote on either side. The equation above can be described in the following steps:
+It is a transformation of the hyperbolic tangent that flattens the central range and comparatively exaggerates the asymptote on either side. The equation above can be described in the following steps:
 
@@
 
-\begin{figure}[H]
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=0.8\textwidth]{grav_diff}
+    \includegraphics[width=\textwidth]{grav_diff}
     \caption{\label{grav_diff} A visualization of $\mathrm{GravDiff}(p, \psi, g)$ with respect to $p$ and $\psi$.}
@@
 \subsubsection{Results}
-
-During a grid search of parameters of the gravitational differentiation algorithm, in addition to .
-The highest accuracy was obtained with 256 labeled examples in the set $\varsigma$ (out of 496 total training examples), which equates to 52\% of the data being labeled.
-High accuracy statistics in this range were also obtained using 128 examples (99\% accuracy, once again) and 64 examples (97\% accuracy).
-32 examples produced a much lower maximum accuracy, at only 83\%.
-From this, it is evident that it is possible to achieve very high accuracy using this technique.
-It is also clear that, in general, larger initial sets produce higher performance.
-
-However, the four hyperparameters optimized during the grid search were observed to have a weak polynomial correlation with the network's accuracy.
-This implies that the training process is generally insensitive to variations in those particular network parameters.
-
-This lack of correlation also means that the mean accuracy over all configurations with a given $\varsigma$ size is a meaningful statistic, since the validation accuracy for an individual configuration is difficult to confidently predict.
-These statistics, shown in Table \ref{grav_accuracy_by_varsigma}, show that for all four set sizes, there is a fairly wide spread between maximum and mean accuracy.
 
@@
 It is evident that larger labeled sets produce better performance.
-However, note that a $\varsigma$ size of 128 produces accuracy of 97.0\%, the same as the banded supervised system with all 496 labeled examples.
+Gravitational differentiation with 256 examples is considerably better than any supervised learning system.
+Additionally, note that a $\varsigma$ size of 128 produces accuracy of 97.0\%, the same as the banded supervised system with all 496 labeled examples.
 That is, GD achieves the same accuracy with 74\% of the labels removed.
@@
 
-During a grid search of these parameters (in addition to regularization parameters of the network), the best mean accuracy over three runs of an individual configuration was 99\%, using 256 initial labeled examples in $\varsigma$.
-(The network's predictions on the best performing epoch can be observed in Figure \ref{icn_grid_search}.) With 128 examples, the maximum was 98\%; with 64 it was 97\%; and with 32, 95\%.
-As with gravitational differentiation, larger set sizes improve accuracy significantly.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{icn_grid_search}
-    \caption{\label{icn_grid_search} Prediction distribution of the best iterative cluster nucleation configuration.}
-\end{figure}
-
 In Table \ref{icn_accuracy_by_varsigma}, it is clear that for a labeled set of 256, ICN performs worse than GD (which reached 99.2\%).
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{icn_grid_search}
+    \caption{\label{icn_grid_search} Prediction distribution of the best iterative cluster nucleation configuration.}
+\end{figure}
+
 \subsection{Performance Comparison}
@@
 This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like the Acoustic Parameter technique.
-This conclusion has been achieved through two rounds of optimization, the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems with the goal of producing a high-accuracy discriminator to complement AP.
-
-In the realm of supervised learning, it was made clear that the banded Fourier transform is the most effective input format.
-It produced significantly higher accuracy and tighter prediction distributions than neural networks trained on either the raw audio waveform or image window data.
+This conclusion has been achieved through two rounds of optimization: the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems with the goal of producing a high-accuracy discriminator to complement AP.
+
+In the realm of supervised learning, it was made clear that the banded Fourier transform with position corrections applied is the most effective input format.
+It produced significantly higher accuracy and tighter prediction distributions than neural networks trained on the full-resolution Fourier transform, the raw audio waveform, or the image window data.
 There is strong evidence that image data in particular does not contain sufficient information to be used for discrimination.

### 2019-01-24 documentation/report.tex
Started work on ISEF report

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{textcomp}
\usepackage{parskip}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage[backend=bibtex, style=alphabetic, sorting=ynt]{biblatex}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{caption}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{titlesec}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyhf{}
\lfoot{\scriptsize January 2019}
\cfoot{\scriptsize Brendon Matusch---Improving Particle Classification in WIMP Experiments}
\rfoot{\scriptsize Page~\thepage~of~\pageref{LastPage}}

\graphicspath{{./images/}}

\addbibresource{paper.bib}

\titlespacing{\section}{0pt}{0pt}{0pt}
\titlespacing{\subsection}{0pt}{0pt}{0pt}
\titlespacing{\subsubsection}{0pt}{0pt}{0pt}

\begin{document}

\begin{LARGE}
    \begin{center}
        Improving Particle Classification in WIMP Dark Matter Detection Experiments Using Neural Networks
    \end{center}
\end{LARGE}

\doublespacing

\section{Introduction}

In all experiments for detection of Weakly Interacting Massive Particle (WIMP) dark matter, it is essential to develop a function that can distinguish events caused by WIMP candidates from those caused by background radiation. Manually developing such a classifier is challenging, time-consuming, and necessitates detailed physical modeling.

Machine learning (ML) has the potential to automate this and accelerate experimentation, in addition to detecting patterns that humans cannot. However, impure calibration data hinders training of models, and unusual detector topologies make data challenging to process.

Concretely, I worked with two dark matter experiments: the PICO-60 bubble chamber \cite{pico}, and the DEAP-3600 liquid argon scintillator \cite{deap}. In both experiments, alpha particles comprise a problematic class of background radiation. In PICO-60, WIMP-like neutron calibration events are used as training data; however, alpha particles mixed into this calibration data pose a challenge to effective training. In DEAP-3600, the detector format, containing 255 spherically arranged photomultipliers, makes data processing challenging.

In both experiments, conventional classifiers are typically used. In PICO-60, it is known to be nearly optimal; however, ML was additionally used (for its automation benefits), reaching a mean of 80.2\% accuracy. In DEAP-3600, in a simulated environment, the conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91\% of simulated WIMP events.

The objective of my study was to develop novel machine learning algorithms that perform better at this task than previous methods, in the PICO-60 and DEAP-3600 dark matter experiments, and I succeeded at this task. I have lead-authored an academic paper \cite{me} on my PICO-60 research, which the PICO collaboration has reviewed and approved.

\singlespacing
\printbibliography

\end{document}

### 2019-01-24 documentation/report.tex
Too long but not bad draft

@@
 \usepackage{titlesec}
+\usepackage{enumitem}
+\usepackage{subfig}
 
@@
 
+\setlist[enumerate]{noitemsep, nolistsep}
+\setlist[itemize]{noitemsep, nolistsep}
+
 \graphicspath{{./images/}}
@@
 
+\doublespacing
+
 \begin{document}
 
+\begin{center}
     \begin{LARGE}
-    \begin{center}
         Improving Particle Classification in WIMP Dark Matter Detection Experiments Using Neural Networks
-    \end{center}
     \end{LARGE}
 
-\doublespacing
+    Brendon Matusch
+\end{center}
 
@@
 Manually developing such a classifier is challenging, time-consuming, and necessitates detailed physical modeling.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+    \centering
+    \fbox{\includegraphics[width=0.38\textwidth]{pico_cad_cropped}}
+    \caption{CAD rendering of the PICO-60 detector as configured for its operation with C$_3$F$_8$.}
+\end{wrapfigure}
 
@@
 In PICO-60, WIMP-like neutron calibration events are used as training data; however, alpha particles mixed into this calibration data pose a challenge to effective training.
-In DEAP-3600, the detector format, containing 255 spherically arranged photomultipliers, makes data processing challenging.
+In DEAP-3600, the detector format, containing 255 spherically arranged photomultiplier tubes (PMTs), making data processing challenging.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+    \centering
+    \fbox{\includegraphics[width=0.38\textwidth]{deap_cad}}
+    \caption{CAD rendering of the DEAP-3600 liquid argon scintillator.}
+\end{wrapfigure}
 
@@
 
+\section{Procedure}
+
+\subsection{PICO-60}
+
+For PICO-60, I developed and compared two sets of classification algorithms:
+
+\begin{enumerate}
+    \item I experimented with conventional supervised learning (where a model is trained on a set of inputs with expected outputs).
+I explored the following data formats to learn which was the best-performing solution:
+    \begin{itemize}
+        \item An 8-band Fourier transform. I applied a dense neural network with dropout and L2 regularization (which allowed for less overfitting to impure data).
+        \item A higher-resolution Fourier transform of the audio (all 50,001 data points). I once again applied a dense neural network.
+        \item A raw audio waveform with a very deep 1D convolutional neural network (CNN) inspired by Dai et al \cite{verydeepconvnets}.
+I hoped to learn whether data preprocessing was required for good performance.
+        \item Images captured by cameras in the detector. I trained a 2D CNN on these, to learn whether they contain any relevant information.
+    \end{itemize}
+    \begin{figure}[ht]
+        \centering
+        \subfloat{\fbox{\includegraphics[width=0.9\textwidth]{audio_cropped}}}
+        \qquad
+        \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{image_grid}}}
+        \caption{Examples of the audio waveform and Fourier transform (above) and the images captured by cameras (below).}
+    \end{figure}
+    \item The impurity in the training data was alleviated with regularization but not entirely solved.
+For a more optimal solution, I developed two entirely new semi-supervised learning algorithms.
+The particle labels (WIMP-like nuclear recoils, or alpha events) were removed from a large portion of the data, and my algorithms learned more accurate labels for this data.
+    \begin{itemize}
+        \item Iterative Cluster Nucleation is inspired by unsupervised "clustering" algorithms.
+First, a neural network is trained on a labeled set.
+After some time, it runs predictions on the unlabeled set.
+Those unlabeled examples with the most confident predictions (close to 0 or 1) are added to their corresponding cluster in the labeled set; these synthesized labels are more accurate than the original labels.
+        \item Gravitational Differentiation is a more analog redesign.
+Using an original piecewise exponential function for calculating final-layer derivatives, called GravDiff, unlabeled examples are caused to "gravitate" toward more accurate predictions based on the confidence of the neural network's predictions.
+    \end{itemize}
+\end{enumerate}
+
+\subsection{DEAP-3600}
+
+In DEAP-3600, the key challenge is the aforementioned unusual detector format: a sphere tiled with a hexagonal lattice of PMTs.
+While it is fundamentally an image, conventional CNNs are intended only for flat rectangular images.
+I attempted to solve the problem in three different ways:
+
+\begin{enumerate}
+    \item I tried simply inputting photon counts from each PMT into a multi-layer perceptron, which has no awareness of spatial arrangement.
+    \item The spherical image challenge is much like that cartographers face when making a map of the spherical Earth.
+I developed and applied a Mercator-inspired cylindrical projection to the spherical data, and used a 2D CNN.
+    \item I developed an entirely new type of CNN, called a topological CNN.
+Rather than a 2D image of square pixels, kernels convolve over an arbitrary topology in any number of dimensions (in this case, a hexagonal lattice on the surface of a sphere).
+\end{enumerate}
+
+\begin{figure}[ht]
+    \centering
+    \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{map_projection}}}
+    \qquad
+    \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{topological}}}
+    \caption{Diagrams of the map projection (left) and topological CNN (right) systems.}
+\end{figure}
+
+\subsection{Implementation}
+
+ML algorithms developed in this study incorporated the Adam \cite{adam} and stochastic gradient descent \cite{sgd} optimizers, and applied dropout \cite{dropout} and L2 regularization.
+
+All programming for this study was done in Python 3.
+Keras \cite{keras}, running on a TensorFlow \cite{tensorflow} backend, was used for all machine learning tasks.
+NumPy \cite{numpy} and SciPy \cite{scipy} were used for linear algebra and signal processing.
+ROOT \cite{root}, scikit-image \cite{scikit-image}, and scikit-learn \cite{scikit-learn} were used for data loading and storage.
+Matplotlib \cite{matplotlib} was used for data visualization.
+
+\section{Results}
+
+\subsection{Statistical Practices}
+
+All performance statistics refer to a test set where specified, and a validation set elsewhere.
+Models were trained on a training set, and a separate randomly selected validation set was used to select the best-performing technique and corresponding hyperparameters (with a grid search).
+A test set was used to verify the performance of the model ultimately chosen.
+
+\subsection{PICO-60}
+
+\begin{itemize}
+    \item The dense neural network, with banded Fourier transform input data, increased accuracy to 97.3\% on validation data, up from a previous best of 80.3\% with ML.
+    \item The full resolution Fourier transform unexpectedly worsened accuracy to 92.2\%.
+    \item The raw waveform with a 1D very deep CNN reached 94.5\% accuracy.
+    \item Iterative cluster nucleation, using the dense network and banded input data, improved validation accuracy to 99.1\%.
+    \item Gravitational differentiation was even more accurate, at 99.7\%.
+    \item The best gravitational differentiation configuration was not as accurate on final test data as on validation data, but was much better than supervised learning, at 98.3\%.
+\end{itemize}
+
+\subsection{DEAP-3600}
+
+The three architectures I tested for neck alpha identification in DEAP-3600 were evaluated based on their ability to reduce the rate of false positives (potential WIMP candidates misidentified as neck alphas) compared to the previous 91\% result from conventional methods.
+Only models with the same (or lower) 0.4\% false negative rate were considered.
+
+\begin{itemize}
+    \item The dense neural network always produced 100\% false positives.
+    \item The topological CNN produced a false positive rate of 92.6\% (also worse than 91.0\%).
+    \item The cylindrical projection method with a 2D CNN improved the false positive rate from 91.0\% to 75.7\% (a 63.0\% reduction)!
+\end{itemize}
+
+\section{Conclusion}
+
+These results confirm that I achieved my goal.
+Indeed, semi-supervised learning for PICO-60 performed better than both the previous best result with machine learning, and my own supervised learning work.
+This significantly improves the accuracy that can be obtained quickly and without any manual optimization, allowing the team working on the experiment to iterate more quickly without the overhead of developing a conventional classifier.
+
+The reduction in the false positive rate for DEAP-3600 provides evidence that machine learning is capable of improving the efficiency of this experiment.
+This may reduce the operation time required to collect sufficient data pointing towards the existence or non-existence of WIMP dark matter.
+
+The classifiers developed during this study demonstrate great promise for machine learning in dark matter detection.
+Fundamentally, the problems I have solved with respect to PICO-60 and DEAP-3600 are not specific to these two experiments; they are common across many others.
+Thus, my algorithms are applicable in the broader field of dark matter detection.
+
 \singlespacing

### 2019-01-24 documentation/report.tex
Almost there

--- 2019-01-24 documentation/report.tex
@@
 
-In all experiments for detection of Weakly Interacting Massive Particle (WIMP) dark matter, it is essential to develop a function that can distinguish events caused by WIMP candidates from those caused by background radiation.
-Manually developing such a classifier is challenging, time-consuming, and necessitates detailed physical modeling.
+In all experiments for detection of WIMP dark matter, it is essential to develop a classifier that can distinguish WIMP events from background radiation, in particular alpha particles.
+Manually developing such a classifier necessitates time-consuming physical modeling.
 
@@
 
-Concretely, I worked with two dark matter experiments: the PICO-60 bubble chamber \cite{pico}, and the DEAP-3600 liquid argon scintillator \cite{deap}.
-In both experiments, alpha particles comprise a problematic class of background radiation.
-In PICO-60, WIMP-like neutron calibration events are used as training data; however, alpha particles mixed into this calibration data pose a challenge to effective training.
-In DEAP-3600, the detector format, containing 255 spherically arranged photomultipliers, makes data processing challenging.
+I worked with two dark matter experiments: the PICO-60 bubble chamber \cite{pico}, and the DEAP-3600 liquid argon scintillator \cite{deap}.
+In PICO-60, impurities in calibration data pose an obstacle to effective training.
+In DEAP-3600, the detector format, containing 255 spherically arranged photomultiplier tubes (PMTs), makes data processing challenging.
 
-In both experiments, conventional classifiers are typically used.
-In PICO-60, it is known to be nearly optimal; however, ML was additionally used (for its automation benefits), reaching a mean of 80.2\% accuracy.
-In DEAP-3600, in a simulated environment, the conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91\% of simulated WIMP events.
+In PICO-60, supervised machine learning was used, reaching a mean of 80.2\% accuracy.
+In DEAP-3600, in a simulated environment, a conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
+My objective was to develop novel machine learning algorithms that are more accurate.
 
-The objective of my study was to develop novel machine learning algorithms that perform better at this task than previous methods, in the PICO-60 and DEAP-3600 dark matter experiments, and I succeeded at this task.
-I have lead-authored an academic paper \cite{me} on my PICO-60 research, which the PICO collaboration has reviewed and approved.
+\section{Procedure}
+
+\subsection{PICO-60}
+
+For PICO-60, I developed and compared two sets of classification algorithms:
+
+\begin{enumerate}
+    \item I experimented further with supervised learning, exploring the following data formats to learn which was the best-performing solution:
+    \begin{itemize}
+        \item An 8-band Fourier transform. I applied a dense neural network with dropout \cite{dropout} and L2 regularization (which allowed for less overfitting to impure data).
+        \item A full-resolution Fourier transform of the audio (all 50,001 data points). I once again applied a dense neural network with the Adam \cite{adam} optimizer.
+        \item A raw audio waveform with a very deep 1D convolutional neural network (CNN) inspired by Dai et al \cite{verydeepconvnets}.
+        \item Images captured by cameras in the detector. I trained a 2D CNN on these, to learn whether they contain any relevant information.
+    \end{itemize}
+    \begin{figure}[ht]
+        \centering
+        \subfloat{\fbox{\includegraphics[width=0.9\textwidth]{audio_cropped}}}
+        \qquad
+        \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{image_grid}}}
+        \caption{Examples of the audio waveform and Fourier transform (above) and the images captured by cameras (below).}
+    \end{figure}
+    \item To even better handle impurities, I developed two novel semi-supervised learning algorithms.
+The impure alpha/WIMP-like labels were removed from a large portion of the training data, and my algorithms learned more accurate labels for this data.
+    \begin{itemize}
+        \item Iterative Cluster Nucleation is inspired by unsupervised clustering algorithms.
+A neural network is trained on a labeled set, and runs inference on the unlabeled set.
+Unlabeled examples with predictions closest to 0 or 1 are added to the labeled set; the newly synthesized labels are more accurate than the original labels.
+        \item Gravitational Differentiation uses a novel piecewise exponential function to calculate final-layer gradients. Unlabeled data ``gravitates'' to more accurate labels.
+    \end{itemize}
+\end{enumerate}
+
+\subsection{DEAP-3600}
+
+In DEAP-3600, the key challenge is the aforementioned unusual detector format: a sphere tiled with a hexagonal lattice of PMTs. I approached this problem in three different ways:
+
+\begin{enumerate}
+    \item I tried simply inputting photon counts from each PMT into a multi-layer perceptron.
+    \item I developed and applied a Mercator-like cylindrical projection, and used a 2D CNN.
+    \item I developed an entirely new type of CNN, called a topological CNN. Rather than a 2D image of square pixels, kernels convolve over an arbitrary topology.
+\end{enumerate}
+
+\begin{figure}[ht]
+    \centering
+    \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{map_projection}}}
+    \qquad
+    \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{topological}}}
+    \caption{Diagrams of the map projection (left) and topological CNN (right) systems.}
+\end{figure}
+
+\section{Results}
+
+\subsection{Statistical Practices}
+
+All models were trained on a training set, and a separate randomly selected validation set was used to select the best-performing technique and corresponding hyperparameters (with a grid search).
+A test set was used to verify the performance of the model ultimately chosen.
+
+\subsection{PICO-60}
+
+\begin{itemize}
+    \item The dense neural network, with banded Fourier transform data, increased validation accuracy to 97.3\% on validation data, up from the previous ML accuracy of 80.2\%.
+    \item The full-resolution Fourier transform unexpectedly worsened accuracy to 92.2\%.
+    \item The raw waveform with a 1D very deep CNN reached 94.5\% accuracy.
+    \item There was no useful information in the image data; only 63.0\% accuracy was achieved.
+    \item Iterative cluster nucleation, using banded FT input data, improved accuracy to 99.1\%.
+    \item Gravitational differentiation was even more accurate, at 99.7\%.
+    \item The best gravitational differentiation configuration was not as accurate on final test data as on validation data, but was much better than supervised learning, at 98.3\%.
+\end{itemize}
+
+\begin{figure}[ht]
+    \centering
+    \fbox{\includegraphics[width=0.8\textwidth]{pico_final_results}}
+    \caption{Validation accuracy, precision, recall, and loss for each technique in PICO-60. Error bars represent the 8th and 92nd percentiles.}
+\end{figure}
+
+\subsection{DEAP-3600}
+
+The three algorithms I developed were evaluated based on their reduction in the rate of false positives (simulated WIMP events misidentified as neck alphas) compared to the conventional 91.0\% result.
+Only models with at least 99.6\% neck alpha removal were considered.
+
+\begin{itemize}
+    \item The dense neural network always produced 100\% false positives.
+    \item The topological CNN produced a false positive rate of 92.6\% (also worse than 91.0\%).
+    \item The cylindrical projection method with a 2D CNN improved the false positive rate from 91.0\% to 75.7\% (a 63.0\% reduction)!
+\end{itemize}
+
+\begin{figure}[ht]
+    \centering
+    \fbox{\includegraphics[width=0.8\textwidth]{deap_final_results}}
+    \caption{Alpha background and WIMP removal rates for the highest-accuracy hyperparameter configuration of each technique in DEAP-3600.}
+\end{figure}
+
+\section{Conclusion}
+
+These results confirm that I achieved my goal.
+Indeed, semi-supervised learning for PICO-60 performed better than both the previous best result with machine learning, and my own supervised learning work.
+This significantly improves the accuracy that can be obtained quickly and without any manual optimization, allowing the team working on the experiment to iterate more quickly without the overhead of developing a conventional classifier.
+
+The reduction in the false positive rate for DEAP-3600 provides evidence that machine learning is capable of improving the efficiency of this experiment.
+This may reduce the operation time required to collect sufficient data pointing towards the existence or non-existence of WIMP dark matter.
+
+The classifiers developed during this study demonstrate great promise for machine learning in dark matter detection.
+Fundamentally, the problems I have solved with respect to PICO-60 and DEAP-3600 are not specific to these two experiments; they are common across many others.
+Thus, my algorithms are applicable in the broader field of dark matter detection.
+
+\section{Acknowledgements}
+
+Many thanks to Dr.
+Nigel Smith, Dr.
+Ken Clark, Dr.
+Carsten Krauss, Dr.
+Scott Fallows, Dr.
+Eric V\'azquez-J\'auregui, and Dr.
+Pierre Gorel of SNOLAB for introducing me to PICO-60 and DEAP-3600, and generously providing access to data for training.
+
+\section*{Bibliography}
+
+\begin{itemize}
+    \item All programming for this study was done in Python 3 \cite{python}.
+    \item Keras \cite{keras}, running on a TensorFlow \cite{tensorflow} backend, was used for all machine learning tasks.
+    \item NumPy \cite{numpy} and SciPy \cite{scipy} were used for linear algebra and signal processing.
+    \item ROOT \cite{root}, scikit-image \cite{scikit-image}, and scikit-learn \cite{scikit-learn} were used for data loading and storage.
+    \item Matplotlib \cite{matplotlib} was used for data visualization.
+\end{itemize}
 

### 2019-01-24 visualization/visualize_overall_performance_deap.py

@@ -27,14 +27,14 @@ for array, mean in [(alpha_high, alpha_mean), (alpha_low, alpha_mean), (wimp_hig
             array[index] = np.abs(array[index] - mean[index])
 
 # The names of the configurations corresponding to all these lists
-CONFIGURATIONS = ['Conventional Analysis', 'Dense Neural Network', 'Map Projection CNN', 'Topological CNN', 'Map Projection CNN Final Test']
+CONFIGURATIONS = ['Conventional Analysis', 'Dense Neural Network', 'Map Projection CNN', 'Topological CNN', 'Map Projection CNN Test']
 
 # Get the locations for the bars
 BAR_WIDTH = 0.4
 alpha_locations = np.arange(len(CONFIGURATIONS))
 wimp_locations = alpha_locations + BAR_WIDTH
 # Create a figure with a predefined size
-_, ax = plt.subplots(figsize=(8, 8))
+_, ax = plt.subplots(figsize=(8, 6))
 # Use percentage labels for the Y axis
 ax.yaxis.set_major_formatter(mtick.PercentFormatter())
 # Draw a bar graph using the performance statistics

### 2019-01-24 visualization/visualize_overall_performance_pico.py

@@ -39,8 +39,8 @@ for array, mean in [(accuracy_high, accuracy_mean), (accuracy_low, accuracy_mean
         else:
             array[index] = np.abs(array[index] - mean[index])
 # The names of the configurations corresponding to all these lists
-CONFIGURATIONS = ['Original NN Analysis', 'Banded DFT', 'Full-Resolution DFT', 'Raw Waveform CNN', 'Image CNN',
-                  'Iterative Cluster Nucleation', 'Gravitational Differentiation', 'Gravitational Differentiation Final Test']
+CONFIGURATIONS = ['Prev. ML Study', 'Banded FT', 'Full-Resolution FT', 'Waveform CNN', 'Image CNN',
+                  'I. Cluster Nucleation', 'Gravitational Diff.', 'Gravitational Diff. Test']
 
 # Get the locations for the bars
 BAR_WIDTH = 0.2
@@ -49,7 +49,7 @@ precision_locations = accuracy_locations + BAR_WIDTH
 recall_locations = precision_locations + BAR_WIDTH
 mse_locations = recall_locations + BAR_WIDTH
 # Create a figure with a predefined size
-_, ax = plt.subplots(figsize=(12, 8))
+_, ax = plt.subplots(figsize=(8, 6))
 # Use percentage labels for the Y axis
 ax.yaxis.set_major_formatter(mtick.PercentFormatter())
 # Draw a bar graph using the performance statistics
@@ -57,7 +57,7 @@ plt.bar(accuracy_locations, accuracy_mean * 100, width=BAR_WIDTH, yerr=(accuracy
 plt.bar(precision_locations, precision_mean * 100, width=BAR_WIDTH, yerr=(precision_low * 100, precision_low * 100), label='Precision')
 plt.bar(recall_locations, recall_mean * 100, width=BAR_WIDTH, yerr=(recall_low * 100, recall_high * 100), label='Recall')
 # Now set the configuration names, with words angled so they fit
-plt.xticks(accuracy_locations + (BAR_WIDTH * 1.5), CONFIGURATIONS, rotation=10)
+plt.xticks(accuracy_locations + (BAR_WIDTH * 1.5), CONFIGURATIONS, rotation=15)
 # Start at 60% so the difference is more obvious
 plt.ylim(60, 100)
 # Label the Y axis to specify what the numbers mean

### 2019-01-26 documentation/report.tex
Pretty happy with it

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{textcomp}
 \usepackage{parskip}
 \usepackage{amsmath}
-\usepackage{authblk}
-\usepackage[backend=bibtex, style=alphabetic, sorting=ynt]{biblatex}
+\usepackage{biblatex}
 \usepackage{graphicx}
@@
 \usepackage{float}
-\usepackage{caption}
 \usepackage{setspace}
@@
 \usepackage{subfig}
+\usepackage{hyperref}
+\usepackage{soul}
 
@@
 \titlespacing{\subsubsection}{0pt}{0pt}{0pt}
+
+\setlength{\belowcaptionskip}{-12pt}
+
+\titleformat*{\section}{\large\bfseries}
+\titleformat*{\subsection}{\normalsize\bfseries}
+\titleformat*{\subsubsection}{\normalsize\bfseries}
 
@@
 In DEAP-3600, in a simulated environment, a conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
-My objective was to develop novel machine learning algorithms that are more accurate.
+
+My objective was to develop novel machine learning algorithms that are more accurate, and I succeeded at this goal!
+I composed a 26-page research whitepaper \cite{me} about my work on PICO-60, which has been approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}.
+
+\ul{Please note that the entire PICO collaboration are listed on the research whitepaper because of their contributions to the original PICO-60 dark matter experiment.
+They did not contribute to this study; I completed and documented it entirely independently.}
 
@@
     \centering
-    \fbox{\includegraphics[width=0.8\textwidth]{deap_final_results}}
+    \fbox{\includegraphics[width=0.6\textwidth]{deap_final_results}}
     \caption{Alpha background and WIMP removal rates for the highest-accuracy hyperparameter configuration of each technique in DEAP-3600.}
@@
 These results confirm that I achieved my goal.
-Indeed, semi-supervised learning for PICO-60 performed better than both the previous best result with machine learning, and my own supervised learning work.
-This significantly improves the accuracy that can be obtained quickly and without any manual optimization, allowing the team working on the experiment to iterate more quickly without the overhead of developing a conventional classifier.
+Indeed, compared to previous machine learning studies, semi-supervised learning significantly improved the accuracy that can be obtained quickly and without any manual optimization.
+This should allow physicists to iterate more quickly, without the overhead of developing a conventional classifier.
 
-The reduction in the false positive rate for DEAP-3600 provides evidence that machine learning is capable of improving the efficiency of this experiment.
-This may reduce the operation time required to collect sufficient data pointing towards the existence or non-existence of WIMP dark matter.
+The reduction in the false positive rate for DEAP-3600 provides evidence that machine learning can improve the efficiency of this experiment, possibly reducing the operation time required to collect sufficient evidence for or against the existence of WIMP dark matter.
 
-The classifiers developed during this study demonstrate great promise for machine learning in dark matter detection.
-Fundamentally, the problems I have solved with respect to PICO-60 and DEAP-3600 are not specific to these two experiments; they are common across many others.
-Thus, my algorithms are applicable in the broader field of dark matter detection.
+The problems I have solved are not specific to PICO-60 and DEAP-3600; they are common across many dark matter detection experiments. Thus, my algorithms are broadly applicable.
 
@@
 
+\singlespacing
+\printbibliography
+\doublespacing
+
 \section*{Bibliography}
@@
 
-\singlespacing
-\printbibliography
-
 \end{document}

### 2019-01-26 visualization/visualize_overall_performance_deap.py

@@ -34,19 +34,21 @@ BAR_WIDTH = 0.4
 alpha_locations = np.arange(len(CONFIGURATIONS))
 wimp_locations = alpha_locations + BAR_WIDTH
 # Create a figure with a predefined size
-_, ax = plt.subplots(figsize=(8, 6))
+fig, ax = plt.subplots(figsize=(6, 3))
 # Use percentage labels for the Y axis
 ax.yaxis.set_major_formatter(mtick.PercentFormatter())
 # Draw a bar graph using the performance statistics
 plt.bar(alpha_locations, alpha_mean * 100, width=BAR_WIDTH, yerr=(alpha_low * 100, alpha_high * 100), label='Alpha Removal')
 plt.bar(wimp_locations, wimp_mean * 100, width=BAR_WIDTH, yerr=(wimp_low * 100, wimp_high * 100), label='WIMP Removal')
 # Now set the configuration names, with words angled so they fit
-plt.xticks(wimp_locations, CONFIGURATIONS, rotation=10)
+plt.xticks(wimp_locations, CONFIGURATIONS, rotation=15)
 # Start at 60% so the difference is more obvious
 plt.ylim(60, 100)
 # Label the Y axis to specify what the numbers mean
 plt.ylabel('Percentage Removal')
 # Create a legend using the predefined labels
 ax.legend(loc='lower left')
+# Rearrange the plot to fit the labels
+fig.tight_layout()
 # Display the graph on screen
 plt.show()

### 2019-01-26 visualization/visualize_overall_performance_pico.py

@@ -49,7 +49,7 @@ precision_locations = accuracy_locations + BAR_WIDTH
 recall_locations = precision_locations + BAR_WIDTH
 mse_locations = recall_locations + BAR_WIDTH
 # Create a figure with a predefined size
-_, ax = plt.subplots(figsize=(8, 6))
+fig, ax = plt.subplots(figsize=(8, 4))
 # Use percentage labels for the Y axis
 ax.yaxis.set_major_formatter(mtick.PercentFormatter())
 # Draw a bar graph using the performance statistics
@@ -64,12 +64,14 @@ plt.ylim(60, 100)
 plt.ylabel('Accuracy/Precision/Recall')
 # Twin the Y axis and plot the mean squared error
 ax_mse = ax.twinx()
-plt.bar(mse_locations, mse_mean, width=BAR_WIDTH, yerr=(mse_low, mse_high), label='Mean Squared Error Loss', color='C3')
+plt.bar(mse_locations, mse_mean, width=BAR_WIDTH, yerr=(mse_low, mse_high), label='MSE', color='C3')
 # Label the secondary Y axis as well
-plt.ylabel('Mean Squared Error Loss')
+plt.ylabel('Mean Squared Error (MSE) Loss')
 # Combine all of the lines from both Y axes together, and create a legend
 main_lines, main_labels = ax.get_legend_handles_labels()
 mse_lines, mse_labels = ax_mse.get_legend_handles_labels()
-ax_mse.legend(main_lines + mse_lines, main_labels + mse_labels, loc='lower left')
+ax_mse.legend(main_lines + mse_lines, main_labels + mse_labels, loc='center right')
+# Rearrange the plot to fit the labels
+fig.tight_layout()
 # Display the graph on screen
 plt.show()

### 2019-01-26 documentation/report.tex
Improved various things

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{parskip}
 \usepackage{amsmath}
@@
 \usepackage{graphicx}
-\usepackage{wrapfig}
 \usepackage{float}
@@
 
-Machine learning (ML) has the potential to automate this and accelerate experimentation, in addition to detecting patterns that humans cannot.
+Machine learning (ML) has the potential to automate this and accelerate experimentation, and also to detect patterns that humans cannot.
 However, impure calibration data hinders training of models, and unusual detector topologies make data challenging to process.
@@
 I worked with two dark matter experiments: the PICO-60 bubble chamber \cite{pico}, and the DEAP-3600 liquid argon scintillator \cite{deap}.
-In PICO-60, impurities in calibration data pose an obstacle to effective training.
+In PICO-60, background alpha and WIMP-like neutron calibration datasets are used for training; however, there is an impurity of 10\% alphas in the neutron set, hindering training.
 In DEAP-3600, the detector format, containing 255 spherically arranged photomultiplier tubes (PMTs), makes data processing challenging.
 
-In PICO-60, supervised machine learning was used, reaching a mean of 80.2\% accuracy.
+Previously in PICO-60, a conventional classifier was developed and is thought to be 100\% accurate.
+However, machine learning is also applicable because it is quicker to develop and re-calibrate.
+A supervised neural network was used, reaching a mean of 80.2\% accuracy.
+
 In DEAP-3600, in a simulated environment, a conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
@@
     \begin{itemize}
-        \item An 8-band Fourier transform. I applied a dense neural network with dropout \cite{dropout} and L2 regularization (which allowed for less overfitting to impure data).
+        \item An 8-band Fourier transform of audio recorded in the bubble chamber.
+I applied a dense neural network with dropout \cite{dropout} and L2 regularization (which allowed for less overfitting to impure data).
         \item A full-resolution Fourier transform of the audio (all 50,001 data points). I once again applied a dense neural network with the Adam \cite{adam} optimizer.
@@
 Unlabeled examples with predictions closest to 0 or 1 are added to the labeled set; the newly synthesized labels are more accurate than the original labels.
-        \item Gravitational Differentiation uses a novel piecewise exponential function to calculate final-layer gradients. Unlabeled data ``gravitates'' to more accurate labels.
+        \item Gravitational Differentiation uses a novel piecewise exponential function to calculate final-layer gradients.
+Unlabeled examples ``gravitate'' toward more accurate predictions based on the confidence of the neural network's predictions.
     \end{itemize}
@@
 
-\section{Results}
-
 \subsection{Statistical Practices}
@@
 A test set was used to verify the performance of the model ultimately chosen.
+
+\section{Results}
 
@@
     \centering
-    \fbox{\includegraphics[width=0.8\textwidth]{pico_final_results}}
+    \fbox{\includegraphics[width=0.95\textwidth]{pico_final_results}}
     \caption{Validation accuracy, precision, recall, and loss for each technique in PICO-60. Error bars represent the 8th and 92nd percentiles.}
@@
 
-The three algorithms I developed were evaluated based on their reduction in the rate of false positives (simulated WIMP events misidentified as neck alphas) compared to the conventional 91.0\% result.
-Only models with at least 99.6\% neck alpha removal were considered.
+The three algorithms I developed were evaluated based on their reduction in the rate of false positives (simulated WIMP events misidentified as neck alphas) compared to the 91.0\% result achieved with a conventional classifier.
+Only models with at least 99.6\% neck alpha removal (in line with the conventional classifier) were considered.
 
 \begin{itemize}
-    \item The dense neural network always produced 100\% false positives.
+    \item The dense neural network always produced 100\% false positives (worse than 91.0\%).
     \item The topological CNN produced a false positive rate of 92.6\% (also worse than 91.0\%).
@@
 These results confirm that I achieved my goal.
-Indeed, compared to previous machine learning studies, semi-supervised learning significantly improved the accuracy that can be obtained quickly and without any manual optimization.
-This should allow physicists to iterate more quickly, without the overhead of developing a conventional classifier.
+While the conventional classifier in PICO-60 is believed to be perfectly accurate, my semi-supervised learning methods have significantly improved the accuracy that can be obtained quickly and without manual optimization.
+This should allow physicists working on future iterations of PICO to iterate more quickly, without modifying a conventional classifier whenever some aspect of the experiment changes.
 
@@
 
-The problems I have solved are not specific to PICO-60 and DEAP-3600; they are common across many dark matter detection experiments. Thus, my algorithms are broadly applicable.
+Fundamentally, the problems I have solved are not specific to PICO-60 and DEAP-3600; they are common across many dark matter detection experiments.
+Thus, my algorithms are broadly applicable, and demonstrate promise in the field as a whole.
 
@@
 
-Many thanks to Dr.
-Nigel Smith, Dr.
-Ken Clark, Dr.
-Carsten Krauss, Dr.
-Scott Fallows, Dr.
-Eric V\'azquez-J\'auregui, and Dr.
-Pierre Gorel of SNOLAB for introducing me to PICO-60 and DEAP-3600, and generously providing access to data for training.
+Many thanks to Dr.~Nigel~Smith, Dr.~Ken~Clark, Dr.~Carsten~Krauss, Dr.~Scott~Fallows, Dr.~Eric~V\'azquez-J\'auregui, and Dr.~Pierre~Gorel for introducing me to PICO-60 and DEAP-3600, and generously providing access to data for training.
 
@@
 \doublespacing
+\pagebreak
 

### 2019-01-26 visualization/graph_binary_predictions.py

@@ -34,9 +34,9 @@ ground_truths = np.array(acoustic_parameters) > 0.25
 
 # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
 for criterion_data, criterion_name, classification_threshold in zip(
-    copy.deepcopy([original_neural_network_scores]),
-    ['original neural network scores'],
-    [0.5]
+    copy.deepcopy([acoustic_parameters]),
+    ['acoustic parameters'],
+    [0.25]
 ):
     # Convert the raw outputs to binary classifications, based on whether they are greater than or equal to the threshold
     classifications = np.array([data_point >= classification_threshold for data_point in criterion_data])
@@ -103,5 +103,5 @@ plt.ylabel('Neural Network Prediction')
 plt.axvline(0.25, color='black')
 # # Restrict the network prediction axis to the range of 0 to 1
 # plt.ylim(0, 1)
-# Display the graph on screen
-plt.show()
+# # Display the graph on screen
+# plt.show()

### 2019-01-26 visualization/visualize_overall_performance_deap.py

@@ -8,9 +8,9 @@ import matplotlib.pyplot as plt
 import matplotlib.ticker as mtick
 
 # Create lists containing the performance statistics (in the form (mean, 8th percentile, 92nd percentile))
-ALPHA = [(0.996, None, None), (0.992453871216953, 0.9843091334894615, 1.0), (0.9986762886158047, 0.996774193548387, 1.0),
+ALPHA = [(0.996, None, None), (1.0, 1.0, 1.0), (0.9986762886158047, 0.996774193548387, 1.0),
          (0.9985087719298246, 0.9921052631578948, 1.0), (0.9959030977982591, 0.9914285714285715, 1.0)]
-WIMP = [(0.91, None, None), (0.7835294571076522, 0.6601297764960347, 0.8723531004603303), (0.7672341098533079, 0.74375, 0.79915611814346),
+WIMP = [(0.91, None, None), (1.0, 1.0, 1.0), (0.7672341098533079, 0.74375, 0.79915611814346),
         (0.9301859015688803, 0.6060606060606061, 1.0), (0.75745328295003, 0.7313432835820896, 0.7927631578947368)]
 # Zip all of these lists into separate high and low arrays
 alpha_mean, alpha_low, alpha_high = np.array(list(zip(*ALPHA)))
@@ -27,7 +27,7 @@ for array, mean in [(alpha_high, alpha_mean), (alpha_low, alpha_mean), (wimp_hig
             array[index] = np.abs(array[index] - mean[index])
 
 # The names of the configurations corresponding to all these lists
-CONFIGURATIONS = ['Conventional Analysis', 'Dense Neural Network', 'Map Projection CNN', 'Topological CNN', 'Map Projection CNN Test']
+CONFIGURATIONS = ['Conventional Clas.', 'Dense Neural Network', 'Map Projection CNN', 'Topological CNN', 'Map Projection CNN Test']
 
 # Get the locations for the bars
 BAR_WIDTH = 0.4

### 2019-01-26 visualization/visualize_overall_performance_pico.py

@@ -13,14 +13,18 @@ import matplotlib.ticker as mtick
 # Recall: [1.0, 1.0, 1.0, 1.0, 1.0, 1.0]
 # MSE (CWSD): [0.60794566699542, 0.5299834700383025, 0.6476162519709378, 0.5931055604078271, 0.608065520294811, 0.5913339464426839]
 
+# RAW DATA FOR ACOUSTIC PARAMETER
+# Everything except MSE is 1.0
+# MSE (CWSD): [0.5820416895527105, 0.5300608757628296, 0.569641996488072, 0.5903858973218383, 0.5402959832385111, 0.5353142091844783]
+
 # Create lists containing the performance statistics (in the form (mean, 8th percentile, 92nd percentile))
-ACCURACY = [(0.8020833333333334, 0.7578125, 0.8359375), (0.9713107638888889, 0.9295572916666667, 0.9921875), (0.921875, None, None),
+ACCURACY = [(1.0, 1.0, 1.0), (0.8020833333333334, 0.7578125, 0.8359375), (0.9713107638888889, 0.9295572916666667, 0.9921875), (0.921875, None, None),
             (0.9453125, None, None), (0.629746835443038, None, None), (0.9914062499999999, 0.98203125, 1.0), (0.9973958333333334, 0.9921875, 1.0), (0.9828645833333334, 0.9690625, 1.0)]
-PRECISION = [(0.7524213015423129, 0.7, 0.7962962962962963), (0.9603736240440659, 0.8927487791089421, 1.0),
+PRECISION = [(1.0, 1.0, 1.0), (0.7524213015423129, 0.7, 0.7962962962962963), (0.9603736240440659, 0.8927487791089421, 1.0),
              (0.8876404494382022, None, None), (0.9213483146067416, None, None), (0.629746835443038, None, None), (0.990066119642678, 0.9826983589280337, 1.0), (0.996031746031746, 0.9880952380952381, 1.0), (0.9874765392592432, 0.9726728395061726, 1.0)]
-RECALL = [(1.0, 1.0, 1.0), (0.9948717948717949, 0.9846153846153846, 1.0), (1.0, None, None), (1.0, None, None), (1.0, None, None),
+RECALL = [(1.0, 1.0, 1.0), (1.0, 1.0, 1.0), (0.9948717948717949, 0.9846153846153846, 1.0), (1.0, None, None), (1.0, None, None), (1.0, None, None),
[…]

### 2019-02-14 documentation/report.tex
Minor fix
Edits for peer review and stuff; updated research plan

@@
 My objective was to develop novel machine learning algorithms that are more accurate, and I succeeded at this goal!
-I composed a 26-page research whitepaper \cite{me} about my work on PICO-60, which has been approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}.
+I composed a 28-page research whitepaper \cite{me} about my work on PICO-60, which has been reviewed and approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}.
+It is currently undergoing peer review for eventual publication in the journal \textit{Computer Physics Communications}.
 
@@
         \item A full-resolution Fourier transform of the audio (all 50,001 data points). I once again applied a dense neural network with the Adam \cite{adam} optimizer.
-        \item A raw audio waveform with a very deep 1D convolutional neural network (CNN) inspired by Dai et al \cite{verydeepconvnets}.
+        \item A raw audio waveform with a very deep 1D convolutional neural network (CNN) inspired by Dai et al.\ \cite{verydeepconvnets}.
         \item Images captured by cameras in the detector. I trained a 2D CNN on these, to learn whether they contain any relevant information.
@@
     \item The dense neural network, with banded Fourier transform data, increased validation accuracy to 97.3\% on validation data, up from the previous ML accuracy of 80.2\%.
-    \item The full-resolution Fourier transform unexpectedly worsened accuracy to 92.2\%.
+    \item The full-resolution Fourier transform produced validation accuracy of 92.2\%.
     \item The raw waveform with a 1D very deep CNN reached 94.5\% accuracy.
@@
 
-\singlespacing
-\printbibliography
-\doublespacing
-\pagebreak
+\section{References}
 
-\section*{Bibliography}
+\printbibliography[heading=none]
+
+\section{Bibliography}
 

### 2019-02-24 documentation/report.tex
Ok draft

@@
 \usepackage{biblatex}
+\usepackage{parskip}
 \usepackage{graphicx}
 \usepackage{float}
-\usepackage{setspace}
 \usepackage{fancyhdr}
@@
 
-\setlist[enumerate]{noitemsep, nolistsep}
-\setlist[itemize]{noitemsep, nolistsep}
-
 \graphicspath{{./images/}}
@@
 \titlespacing{\subsubsection}{0pt}{0pt}{0pt}
-
-\setlength{\belowcaptionskip}{-12pt}
-
-\titleformat*{\section}{\large\bfseries}
-\titleformat*{\subsection}{\normalsize\bfseries}
-\titleformat*{\subsubsection}{\normalsize\bfseries}
-
-\doublespacing
 
@@
 \ul{Please note that the entire PICO collaboration are listed on the research whitepaper because of their contributions to the original PICO-60 dark matter experiment.
-They did not contribute to this study; I completed and documented it entirely independently.}
+They did not contribute to this study; I completed and documented it independently.}
 
@@
 
-For PICO-60, I developed and compared two sets of classification algorithms:
+For PICO-60, I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.
+
+\subsubsection{Supervised Learning}
+
+I experimented further with supervised learning, exploring the following data formats to learn which was the best-performing solution:
 
 \begin{enumerate}
-    \item I experimented further with supervised learning, exploring the following data formats to learn which was the best-performing solution:
-    \begin{itemize}
     \item An 8-band Fourier transform of audio recorded in the bubble chamber.
@@
     \item Images captured by cameras in the detector. I trained a 2D CNN on these, to learn whether they contain any relevant information.
-    \end{itemize}
+\end{enumerate}
+
 \begin{figure}[ht]
     \centering
-        \subfloat{\fbox{\includegraphics[width=0.9\textwidth]{audio_cropped}}}
+    \subfloat{\fbox{\includegraphics[width=0.95\textwidth]{audio_cropped}}}
     \qquad
@@
 \end{figure}
-    \item To even better handle impurities, I developed two novel semi-supervised learning algorithms.
+
+\subsubsection{Semi-Supervised Learning}
+
+To even better handle impurities, I developed two novel semi-supervised learning algorithms.
 The impure alpha/WIMP-like labels were removed from a large portion of the training data, and my algorithms learned more accurate labels for this data.
-    \begin{itemize}
-        \item Iterative Cluster Nucleation is inspired by unsupervised clustering algorithms.
-A neural network is trained on a labeled set, and runs inference on the unlabeled set.
-Unlabeled examples with predictions closest to 0 or 1 are added to the labeled set; the newly synthesized labels are more accurate than the original labels.
-        \item Gravitational Differentiation uses a novel piecewise exponential function to calculate final-layer gradients.
-Unlabeled examples ``gravitate'' toward more accurate predictions based on the confidence of the neural network's predictions.
-    \end{itemize}
+
+\begin{enumerate}
+    \item Iterative Cluster Nucleation is inspired by unsupervised "clustering" algorithms, and takes advantage of the ``confidence'' of the neural network's predictions (that is, how close the predictions are to 0 or 1).
+
+    First, a neural network is trained on a labeled set.
+After 30 epochs, it runs predictions on the unlabeled set.
+As the network trains, the confidence of these predictions increases.
+When $p < t$ or $p > 1 - t$ for a given prediction $p$ and a (small) confidence threshold $t$, the example is labeled correspondingly and added to the labeled set.
+
+    It is hypothesized that accuracy will gradually \textit{improve} as the neural network's most confident predictions are fed back into the network as newly labeled training data.
+
+    \item Gravitational Differentiation is an analog alternative to Iterative Cluster Nucleation.
+Using a novel piecewise exponential function for calculating final-layer derivatives, called $\mathrm{GravDiff}(p, \psi, g)$, unlabeled examples are caused to "gravitate" toward more accurate predictions based on the confidence of the neural network's predictions.
+
+    The function is a transformation of the hyperbolic tangent that flattens the central range and comparatively exaggerates the asymptote on either side, defined as follows:
+    \begin{center}
+        $\mathrm{GravDiff}(p, \psi, g) = g \cdot \mathrm{sgn}(p) \cdot \lvert \mathrm{tanh}(2p - 1) \rvert ^ \psi$
+    \end{center}
+    where $p$ is the network's prediction, $\psi$ determines the rate at which the output approaches 0 as the confidence of the prediction decreases, and $g$ is the learning rate multiplier.
+The output is the final-layer gradient to be used in backpropagation.
 \end{enumerate}
@@
 
-All models were trained on a training set, and a separate randomly selected validation set was used to select the best-performing technique and corresponding hyperparameters (with a grid search).
-A test set was used to verify the performance of the model ultimately chosen.
+Models were optimized using grid searches, in which every configuration within a hyperparameter space is trained and evaluated.
+All models were trained on a training set, and a separate randomly selected validation set was used to select the best-performing technique and corresponding hyperparameters.
+
+Finally, the best-performing model was retrained on a newly selected training set, and a test set was used to verify its performance.
+In DEAP-3600, two test sets were used: one consisting of simulated data, and another consisting of 30 examples collected from the real-world detector (neutron calibration data, and background radiation).
 
@@
 
-\begin{itemize}
-    \item The dense neural network, with banded Fourier transform data, increased validation accuracy to 97.3\% on validation data, up from the previous ML accuracy of 80.2\%.
-    \item The full-resolution Fourier transform produced validation accuracy of 92.2\%.
-    \item The raw waveform with a 1D very deep CNN reached 94.5\% accuracy.
-    \item There was no useful information in the image data; only 63.0\% accuracy was achieved.
-    \item Iterative cluster nucleation, using banded FT input data, improved accuracy to 99.1\%.
-    \item Gravitational differentiation was even more accurate, at 99.7\%.
-    \item The best gravitational differentiation configuration was not as accurate on final test data as on validation data, but was much better than supervised learning, at 98.3\%.
-\end{itemize}
+As seen in Figure \ref{pico_final_results}, the banded FFT input format was the most effective of all of the supervised learning models (including the previous study), reaching 97.3\% accuracy on validation data.
+The image data was shown to contain no meaningful information; it permitted only 63.0\% accuracy.
+
+Semi-supervised learning produced better accuracy than supervised learning.
+Gravitational differentation was the best overall model, reaching 99.7\% accuracy on validation data and 98.3\% on test data.
+This confirms my hypothesis; it is indeed beneficial to train on the neural network's most confident predictions.
 
@@
     \centering
-    \fbox{\includegraphics[width=0.95\textwidth]{pico_final_results}}
-    \caption{Validation accuracy, precision, recall, and loss for each technique in PICO-60. Error bars represent the 8th and 92nd percentiles.}
+    \fbox{\includegraphics[width=0.9\textwidth]{pico_final_results}}
+    \caption{\label{pico_final_results} Validation accuracy, precision, recall, and loss for each technique in PICO-60. Error bars represent the 8th and 92nd percentiles.}
 \end{figure}
@@
 
-\begin{itemize}
-    \item The dense neural network always produced 100\% false positives (worse than 91.0\%).
-    \item The topological CNN produced a false positive rate of 92.6\% (also worse than 91.0\%).
-    \item The cylindrical projection method with a 2D CNN improved the false positive rate from 91.0\% to 75.7\% (a 63.0\% reduction)!
-\end{itemize}
+Both the dense neural network and topological CNN produced a higher rate of false positives than the conventional classifier.
+However, as seen in Figure \ref{deap_final_results}, the cylindrical projection method improved the false positive rate from 91.0\% to 75.7\% (a 63.0\% reduction)!
+
+When tested on real-world data, the cylindrical projection method removed 73.9\% of nuclear recoil (neutron calibration) events (in line with the simulation).
+
+However, only 91.2\% of background radiation events were removed.
+This indicates either that the simulation is not an accurate representation of the real-world detector, or (less probably) that WIMPs are present in the background data.
 
@@
     \centering
-    \fbox{\includegraphics[width=0.6\textwidth]{deap_final_results}}
-    \caption{Alpha background and WIMP removal rates for the highest-accuracy hyperparameter configuration of each technique in DEAP-3600.}
+    \fbox{\includegraphics[width=0.55\textwidth]{deap_final_results}}
+    \caption{\label{deap_final_results} Background and WIMP removal rates for the highest-accuracy hyperparameter configuration of each technique in DEAP-3600.}
 \end{figure}
@@
 The reduction in the false positive rate for DEAP-3600 provides evidence that machine learning can improve the efficiency of this experiment, possibly reducing the operation time required to collect sufficient evidence for or against the existence of WIMP dark matter.
+
+Additionally, the unusual predictions made on real-world data provide important diagnostic information on the DEAP-3600 simulation (assuming that WIMPs are not present).
 
@@
 
+\pagebreak
+
 \section{Bibliography}

### 2019-02-24 visualization/visualize_overall_performance_deap.py

@@ -27,14 +27,14 @@ for array, mean in [(alpha_high, alpha_mean), (alpha_low, alpha_mean), (wimp_hig
             array[index] = np.abs(array[index] - mean[index])
 
 # The names of the configurations corresponding to all these lists
-CONFIGURATIONS = ['Conventional Clas.', 'Dense Neural Network', 'Map Projection CNN', 'Topological CNN', 'Map CNN Test (Sim)', 'Map CNN Test (Real)']
+CONFIGURATIONS = ['Conventional Clas.', 'Dense Neural Network', 'Map Projection CNN', 'Topological CNN', 'Map CNN Test (Sim)', 'Map CNN Test (Real,\nBackground May\nInclude WIMPs)']
 
 # Get the locations for the bars
 BAR_WIDTH = 0.4
 alpha_locations = np.arange(len(CONFIGURATIONS))
 wimp_locations = alpha_locations + BAR_WIDTH
 # Create a figure with a predefined size
-fig, ax = plt.subplots(figsize=(6, 3))
+fig, ax = plt.subplots(figsize=(6, 4))
 # Use percentage labels for the Y axis
 ax.yaxis.set_major_formatter(mtick.PercentFormatter())
 # Draw a bar graph using the performance statistics

### 2019-02-24 visualization/visualize_overall_performance_pico.py

@@ -53,7 +53,7 @@ precision_locations = accuracy_locations + BAR_WIDTH
 recall_locations = precision_locations + BAR_WIDTH
 mse_locations = recall_locations + BAR_WIDTH
 # Create a figure with a predefined size
-fig, ax = plt.subplots(figsize=(10, 4))
+fig, ax = plt.subplots(figsize=(9, 4))
 # Use percentage labels for the Y axis
 ax.yaxis.set_major_formatter(mtick.PercentFormatter())
 # Draw a bar graph using the performance statistics

### 2019-02-24 documentation/report.tex
Properly explained DEAP algorithms

@@
 
-In DEAP-3600, the key challenge is the aforementioned unusual detector format: a sphere tiled with a hexagonal lattice of PMTs. I approached this problem in three different ways:
+In DEAP-3600, the key challenge is the aforementioned unusual detector format: a sphere (with an opening on the top) tiled with a hexagonal lattice of PMTs.
+I approached this problem in three different ways:
 
@@
     \item I tried simply inputting photon counts from each PMT into a multi-layer perceptron.
-    \item I developed and applied a Mercator-like cylindrical projection, and used a 2D CNN.
-    \item I developed an entirely new type of CNN, called a topological CNN. Rather than a 2D image of square pixels, kernels convolve over an arbitrary topology.
+
+    \item I developed and applied a Mercator-like cylindrical projection, by mapping each row of nodes in the sphere to a row in a rectangular image.
+Cubic spline interpolation was used to map latitude lines on the sphere (which vary in length) to equal-length rows on the rectangular image.
+
+    \item I developed a new type of CNN, called a topological CNN. Rather than a 2D image of square pixels, kernels convolve over an arbitrary topology.
+
+    While a conventional convolutional layer downscales a rectangular image to a smaller rectangular image by multiplying square regions by a set of corresponding weights, a topological CNN (as applied to this problem) downscales the approximately spherical hexagonal mesh to a smaller but topologically equivalent hexagonal mesh.
 \end{enumerate}

### 2019-02-26 documentation/report.tex
Resized charts to fit text
Made everything much clearer

@@
 
+\titleformat*{\section}{\large\bfseries}
+\titleformat*{\subsection}{\normalsize\bfseries}
+\titleformat*{\subsubsection}{\normalsize\bfseries}
+
 \begin{document}
@@
 \begin{center}
-    \begin{LARGE}
+    \begin{Large}
         Improving Particle Classification in WIMP Dark Matter Detection Experiments Using Neural Networks
-    \end{LARGE}
+    \end{Large}
 
@@
 
-In DEAP-3600, in a simulated environment, a conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
+In DEAP-3600, a simulation was developed prior to my work, because it is impractical to collect large amounts of calibration data.
+A conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of simulated WIMP events.
 
@@
     \centering
-    \subfloat{\fbox{\includegraphics[width=0.95\textwidth]{audio_cropped}}}
+    \subfloat{\fbox{\includegraphics[width=0.8\textwidth]{audio_cropped}}}
     \qquad
@@
 \begin{enumerate}
-    \item Iterative Cluster Nucleation is inspired by unsupervised "clustering" algorithms, and takes advantage of the ``confidence'' of the neural network's predictions (that is, how close the predictions are to 0 or 1).
+    \item Iterative Cluster Nucleation is inspired by unsupervised clustering algorithms, and takes advantage of the ``confidence'' of the neural network's predictions (that is, how close the predictions are to 0 or 1).
 
@@
 
-    \item I developed and applied a Mercator-like cylindrical projection, by mapping each row of nodes in the sphere to a row in a rectangular image.
+    \item I developed and applied a Mercator-like cylindrical projection, which maps a sphere onto a rectangle.
 Cubic spline interpolation was used to map latitude lines on the sphere (which vary in length) to equal-length rows on the rectangular image.
@@
     \centering
-    \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{map_projection}}}
+    \fbox{\includegraphics[width=0.45\textwidth]{map_projection}}
     \qquad
-    \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{topological}}}
-    \caption{Diagrams of the map projection (left) and topological CNN (right) systems.}
+    \fbox{\includegraphics[width=0.45\textwidth]{topological}}
+    \caption{Diagrams of the cylindrical projection (left) and topological CNN (right) systems.}
 \end{figure}
@@
 
-As seen in Figure \ref{pico_final_results}, the banded FFT input format was the most effective of all of the supervised learning models (including the previous study), reaching 97.3\% accuracy on validation data.
-The image data was shown to contain no meaningful information; it permitted only 63.0\% accuracy.
+As seen in Figure \ref{pico_final_results}, the banded FFT format was the most effective of the supervised learning models (including the previous study), reaching 97.3\% accuracy on validation data.
+The image data was shown to contain no meaningful information, permitting only 63.0\% accuracy.
 
 Semi-supervised learning produced better accuracy than supervised learning.
-Gravitational differentation was the best overall model, reaching 99.7\% accuracy on validation data and 98.3\% on test data.
+Gravitational differentiation was the best overall model, reaching 99.7\% accuracy on validation data and 98.3\% on test data.
 This confirms my hypothesis; it is indeed beneficial to train on the neural network's most confident predictions.
@@
     \centering
-    \fbox{\includegraphics[width=0.9\textwidth]{pico_final_results}}
+    \fbox{\includegraphics[width=0.85\textwidth]{pico_final_results}}
     \caption{\label{pico_final_results} Validation accuracy, precision, recall, and loss for each technique in PICO-60. Error bars represent the 8th and 92nd percentiles.}
@@
 
-Both the dense neural network and topological CNN produced a higher rate of false positives than the conventional classifier.
-However, as seen in Figure \ref{deap_final_results}, the cylindrical projection method improved the false positive rate from 91.0\% to 75.7\% (a 63.0\% reduction)!
+Both the dense neural network and topological CNN produced a higher rate of false positives than the conventional classifier, meaning that they misclassified more WIMPs as alpha particles.
+However, as seen in Figure \ref{deap_final_results}, the cylindrical projection method \textit{reduced} the false positive rate from 91.0\% to 75.7\%.
 
-When tested on real-world data, the cylindrical projection method removed 73.9\% of nuclear recoil (neutron calibration) events (in line with the simulation).
+This successful network model was tested on the limited amount of available real-world data.
+It removed 73.9\% of WIMP-like (neutron calibration) events, which is in line with the 75.7\% obtained on simulated data.
 
-However, only 91.2\% of background radiation events were removed.
-This indicates either that the simulation is not an accurate representation of the real-world detector, or (less probably) that WIMPs are present in the background data.
+However, only 91.2\% of background events (expected to be neck alphas) were removed, quite far off the 99.6\% reached on simulated data. This indicates one of three possibilities:
+
+\begin{enumerate}
+    \item The simulation is not an accurate representation of the real-world detector (some physical dynamic has not been simulated), and thus the network did not generalize.
+    \item There is WIMP dark matter present in the set of background events.
+    \item There is an unanticipated form of background radiation present.
+\end{enumerate}
 
@@
     \centering
-    \fbox{\includegraphics[width=0.55\textwidth]{deap_final_results}}
+    \fbox{\includegraphics[width=0.5\textwidth]{deap_final_results}}
+    \qquad
+    \begin{tabular}[b]{|l|l|l|}
+        \hline
+        Method & Backgnd. & WIMP \\
+        \hline
+        Conventional & 99.6\% & 91.0\% \\
+        \hline
+        Dense NN & 100\% & 100\% \\
+        \hline
+        Cylindrical & 99.9\% & 76.7\% \\
+        \hline
+        Topological & 99.9\% & 93.0\% \\
+        \hline
+        Cyl Sim Test & 99.6\% & 75.7\% \\
+        \hline
+        Cyl Real Test & 91.2\% & 73.9\% \\
+        \hline
+    \end{tabular}
     \caption{\label{deap_final_results} Background and WIMP removal rates for the highest-accuracy hyperparameter configuration of each technique in DEAP-3600.}

### 2019-02-26 data_processing/store_deap_data.py

@@ -1,123 +1,0 @@
-#!/usr/bin/env python2
-"""Load the DEAP simulation and real-word data from ROOT files and save in a portable scikit-learn Joblib format"""
-# Created by Brendon Matusch, August 2018
-# Written in Python 2 because the DEAP-specific RAT distribution required for loading the relevant files binds to Python 2
-
-from __future__ import print_function
-
-import glob
-import itertools
-import os
-
-import numpy as np
-# Import RAT even though it is not used directly, because it modifies ROOT
-import rat
-import ROOT
-from sklearn.externals import joblib
-
-# The paths the simulated and real-world DEAP data is saved in
-SIMULATED_PATH = os.path.expanduser('~/deap_data.pkl')
-REAL_WORLD_PATH = os.path.expanduser('~/real_deap_data.pkl')
-
-# The total number of PMTs, which is used as the length of various arrays
-PMT_COUNT = 255
-
-
-def load_data_from_file(file_path, cut_position):
-    """Given a path to a ROOT file, load all of the relevant data from it and return it as a list of tuples"""
-    # Load the data file from the full path, and get the main tree
-    data_file = ROOT.TFile(file_path)
-    tree = data_file.Get('T')
-    # Create a list to add all PMT photon count and timing arrays to
-    pmt_data_arrays = []
-    # Iterate over all events within the tree
-    for event in tree:
-        # Skip events for which the calibrated branch is not present (it contains the number of photons)
-        if not event.ds.ExistCAL():
-            continue
-        # If the number of photons is outside the desired range, skip the event
-        photon_count = event.ds.cal[0].GetQPE()
-        if photon_count < 80 or photon_count > 240:
-            continue
-        # If there are supposed to be cuts done on the position to make sure the event is either near the center or in the neck
-        if cut_position:
-            # If the event's approximated position is outside the vessel
-            if event.ds.ev[0].mblikelihood.pos.Mag() >= 630:
-                # Skip this event; it is probably not a neck alpha
-                continue
-        # Try to get the calibration sub-tree, which is the only data available for real-world events
-        # It is a vector with either 0 or 1 elements
-        try:
-            calibration = event.ds.cal[0]
-        # If it is not present, this event was not detected, so it can be skipped
-        except IndexError:
-            continue
[…]

### 2019-02-26 visualization/visualize_overall_performance_deap.py

@@ -27,7 +27,7 @@ for array, mean in [(alpha_high, alpha_mean), (alpha_low, alpha_mean), (wimp_hig
             array[index] = np.abs(array[index] - mean[index])
 
 # The names of the configurations corresponding to all these lists
-CONFIGURATIONS = ['Conventional Clas.', 'Dense Neural Network', 'Map Projection CNN', 'Topological CNN', 'Map CNN Test (Sim)', 'Map CNN Test (Real,\nBackground May\nInclude WIMPs)']
+CONFIGURATIONS = ['Conventional', 'Dense NN', 'Cylindrical Proj.', 'Topological CNN', 'Cylindrical Test (Sim)', 'Cylindrical CNN Test (Real,\nBackground May\nInclude WIMPs)']
 
 # Get the locations for the bars
 BAR_WIDTH = 0.4

### 2019-02-26 documentation/report.tex
Fixed paper

@@
 \usepackage{soul}
+\usepackage{color}
+\usepackage{colortbl}
 
@@
 
-\titleformat*{\section}{\large\bfseries}
+\titleformat*{\section}{\normalsize\bfseries}
 \titleformat*{\subsection}{\normalsize\bfseries}
@@
 \begin{center}
-    \begin{Large}
-        Improving Particle Classification in WIMP Dark Matter Detection Experiments Using Neural Networks
-    \end{Large}
-
-    Brendon Matusch
+    \begin{large}
+        Improving Particle Classification in WIMP Dark Matter Detection Experiments Using Neural Networks -- Brendon Matusch
+    \end{large}
 \end{center}
@@
 
-In all experiments for detection of WIMP dark matter, it is essential to develop a classifier that can distinguish WIMP events from background radiation, in particular alpha particles.
-Manually developing such a classifier necessitates time-consuming physical modeling.
+In all experiments for detection of WIMP dark matter, it is essential to develop a classifier that can distinguish potential WIMP events from background radiation, in particular alpha particles.
+Manually developing a classifier necessitates time-consuming physical modeling.
 
@@
 Previously in PICO-60, a conventional classifier was developed and is thought to be 100\% accurate.
-However, machine learning is also applicable because it is quicker to develop and re-calibrate.
-A supervised neural network was used, reaching a mean of 80.2\% accuracy.
+A supervised neural network (NN) was also previously developed, achieving a mean accuracy of 80.2\% -- not usable as a practical replacement for conventional methods.
 
-In DEAP-3600, a simulation was developed prior to my work, because it is impractical to collect large amounts of calibration data.
-A conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of simulated WIMP events.
+In DEAP-3600, a physical simulation was developed (prior to my work) to provide calibration data; physical limitations make it impractical to collect large amounts of calibration data.
+In the simulation, the conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
 
-My objective was to develop novel machine learning algorithms that are more accurate, and I succeeded at this goal!
+My objective was to develop accurate machine learning algorithms that are more accurate, and I succeeded at this goal!
 I composed a 28-page research whitepaper \cite{me} about my work on PICO-60, which has been reviewed and approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}.
@@
 
-\ul{Please note that the entire PICO collaboration are listed on the research whitepaper because of their contributions to the original PICO-60 dark matter experiment.
-They did not contribute to this study; I completed and documented it independently.}
+\ul{All PICO researchers are listed on my paper for their work on the original PICO-60 experiment. They did not contribute to this study; I completed and documented it independently.}
 
@@
 
-For PICO-60, I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.
+For PICO-60, my intent was to develop a highly accurate classification algorithm that is intolerant to substantial impurities in calibration data.
+I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.
 
@@
 
-I experimented further with supervised learning, exploring the following data formats to learn which was the best-performing solution:
+I experimented with supervised learning, exploring the following data formats to determine which facilitates the best-performing solution:
 
 \begin{enumerate}
-    \item An 8-band Fourier transform of audio recorded in the bubble chamber.
-I applied a dense neural network with dropout \cite{dropout} and L2 regularization (which allowed for less overfitting to impure data).
+    \item An 8-band Fourier transform of audio recorded in the detector. I applied a dense NN with dropout \cite{dropout} and L2 regularization (which reduce overfitting to impure data).
     \item A full-resolution Fourier transform of the audio (all 50,001 data points). I once again applied a dense neural network with the Adam \cite{adam} optimizer.
-    \item A raw audio waveform with a very deep 1D convolutional neural network (CNN) inspired by Dai et al.\ \cite{verydeepconvnets}.
-    \item Images captured by cameras in the detector. I trained a 2D CNN on these, to learn whether they contain any relevant information.
+    \item A raw audio waveform, with a very deep 1D convolutional neural network (CNN) inspired by Dai et al.\ \cite{verydeepconvnets}.
+    \item Event images, with a 2D CNN (to determine whether they contain useful information).
 \end{enumerate}
@@
     \centering
-    \subfloat{\fbox{\includegraphics[width=0.8\textwidth]{audio_cropped}}}
+    \subfloat{\fbox{\includegraphics[width=0.95\textwidth]{audio_cropped}}}
     \qquad
-    \subfloat{\fbox{\includegraphics[width=0.45\textwidth]{image_grid}}}
-    \caption{Examples of the audio waveform and Fourier transform (above) and the images captured by cameras (below).}
+    \subfloat{\fbox{\includegraphics[width=0.35\textwidth]{image_grid}}}
+    \caption{Examples of the audio and image data.}
 \end{figure}
@@
 
-To even better handle impurities, I developed two novel semi-supervised learning algorithms.
-The impure alpha/WIMP-like labels were removed from a large portion of the training data, and my algorithms learned more accurate labels for this data.
+To better handle impurities, I developed two novel semi-supervised learning techniques.
+The concept of both algorithms is to iteratively generate more accurate labels for a set of inaccurately labeled training data, permitting less overfitting and greater validation accuracy than simple regularization would allow.
 
@@
 In DEAP-3600, the key challenge is the aforementioned unusual detector format: a sphere (with an opening on the top) tiled with a hexagonal lattice of PMTs.
-I approached this problem in three different ways:
+The major source of background radiation is ``neck alphas'', which strike atoms in the gas filling the neck above the spherical detector body.
+These must be eliminated from the experimental data set.
+
+I approached this problem in three different ways, listed below.
+Techniques \ref{map_entry} and \ref{topological_entry} (Figure \ref{deap_methods}) take advantage of geometric and spatial patterns in PMT activations associated with neck alpha events, by extending the CNN concept for use in spherical and arbitrary topologies.
 
 \begin{enumerate}
-    \item I tried simply inputting photon counts from each PMT into a multi-layer perceptron.
+    \item I inputted photon counts from each PMT into a multi-layer perceptron.
 
@@
 Cubic spline interpolation was used to map latitude lines on the sphere (which vary in length) to equal-length rows on the rectangular image.
+Once again, photon counts were used as input data.
+\label{map_entry}
 
-    \item I developed a new type of CNN, called a topological CNN. Rather than a 2D image of square pixels, kernels convolve over an arbitrary topology.
+    \item I developed a new type of CNN, called a topological CNN. Rather than a 2D image of square pixels, kernels convolve over an arbitrary topology. \label{topological_entry}
 
@@
     \centering
-    \fbox{\includegraphics[width=0.45\textwidth]{map_projection}}
+    \fbox{\includegraphics[width=0.425\textwidth]{map_projection}}
     \qquad
-    \fbox{\includegraphics[width=0.45\textwidth]{topological}}
-    \caption{Diagrams of the cylindrical projection (left) and topological CNN (right) systems.}
+    \fbox{\includegraphics[width=0.475\textwidth]{topological}}
+    \caption{\label{deap_methods} Diagrams of the cylindrical projection (left) and topological CNN (right) systems.}
 \end{figure}
@@
 Models were optimized using grid searches, in which every configuration within a hyperparameter space is trained and evaluated.
-All models were trained on a training set, and a separate randomly selected validation set was used to select the best-performing technique and corresponding hyperparameters.
+The best-performing technique and hyperparameters were selected based on repeated random sub-sampling of validation sets.
 
@@
 
-As seen in Figure \ref{pico_final_results}, the banded FFT format was the most effective of the supervised learning models (including the previous study), reaching 97.3\% accuracy on validation data.
-The image data was shown to contain no meaningful information, permitting only 63.0\% accuracy.
+As seen in Figure \ref{pico_final_results}, the banded Fourier transform format was the most effective of the supervised learning models (including the previous study), achieving 97.3\% validation accuracy.
+The image data was shown to contain no useful information, reaching only 63.0\% accuracy.
 
 Semi-supervised learning produced better accuracy than supervised learning.
-Gravitational differentiation was the best overall model, reaching 99.7\% accuracy on validation data and 98.3\% on test data.
+Gravitational differentiation was the best overall model, achieving 99.7\% accuracy on validation data and 98.3\% on test data.
 This confirms my hypothesis; it is indeed beneficial to train on the neural network's most confident predictions.
@@
     \centering
-    \fbox{\includegraphics[width=0.85\textwidth]{pico_final_results}}
+    \fbox{\includegraphics[width=0.95\textwidth]{pico_final_results}}
     \caption{\label{pico_final_results} Validation accuracy, precision, recall, and loss for each technique in PICO-60. Error bars represent the 8th and 92nd percentiles.}
@@
 
-The three algorithms I developed were evaluated based on their reduction in the rate of false positives (simulated WIMP events misidentified as neck alphas) compared to the 91.0\% result achieved with a conventional classifier.
-Only models with at least 99.6\% neck alpha removal (in line with the conventional classifier) were considered.
+The three algorithms I developed were evaluated based on:
 
-Both the dense neural network and topological CNN produced a higher rate of false positives than the conventional classifier, meaning that they misclassified more WIMPs as alpha particles.
-However, as seen in Figure \ref{deap_final_results}, the cylindrical projection method \textit{reduced} the false positive rate from 91.0\% to 75.7\%.
+\begin{itemize}
+    \item Identification of neck alphas with 99.6\% precision (same as the conventional classifier).
+    \item Reduction in the rate of false positives (simulated WIMP events misidentified as neck alphas) compared to the 91.0\% result achieved with a conventional classifier.
+\end{itemize}
 
-This successful network model was tested on the limited amount of available real-world data.
+All techniques were validated on simulated data.
+Both the dense neural network and topological CNN techniques were unsuccessful; they produced higher false positive rates than the conventional classifier, meaning that they misclassified more WIMPs as alpha particles.
+
+However, as seen in Figure \ref{deap_final_results}, the cylindrical projection method retained a 99.6\% neck alpha identification rate, and also successfully \textit{reduced} the false positive rate from 91.0\% to 75.7\%.
+
+This successful network model was further tested on the limited amount of available real-world experimental data.
 It removed 73.9\% of WIMP-like (neutron calibration) events, which is in line with the 75.7\% obtained on simulated data.
 
-However, only 91.2\% of background events (expected to be neck alphas) were removed, quite far off the 99.6\% reached on simulated data. This indicates one of three possibilities:
+However, only 91.2\% of background events (expected to be neck alphas) were removed, below the 99.6\% reached on simulated data. This indicates one of three possibilities:
 
@@
     \centering
-    \fbox{\includegraphics[width=0.5\textwidth]{deap_final_results}}
-    \qquad
-    \begin{tabular}[b]{|l|l|l|}
+    \definecolor{lightcyan}{rgb}{0.8, 1, 1}
+    \begin{tabular}[b]{|l|l|l|l|}
         \hline
-        Method & Backgnd. & WIMP \\
+        \rowcolor{lightcyan}
+        Technique & Data Source & Background & False Positive Rate \\
+        \rowcolor{lightcyan}
+        & & Removal Rate & (WIMP Misclassifications) \\
         \hline
-        Conventional & 99.6\% & 91.0\% \\
+        Conventional Classifier & Simulation & 99.6\% & 91.0\% \\
         \hline
-        Dense NN & 100\% & 100\% \\
+        Dense Neural Network & Simulation & 100\% & 100\% \\
         \hline
-        Cylindrical & 99.9\% & 76.7\% \\
+        Cylindrical Projection & Simulation & 99.9\% & 76.7\% \\
         \hline
-        Topological & 99.9\% & 93.0\% \\
+        Topological CNN & Simulation & 99.9\% & 93.0\% \\
         \hline
-        Cyl Sim Test & 99.6\% & 75.7\% \\
+        Cylindrical Projection & Simulation (Test) & 99.6\% & 75.7\% \\
         \hline
-        Cyl Real Test & 91.2\% & 73.9\% \\
+        Cylindrical Projection & Real-World & 91.2\% & 73.9\% \\
         \hline
     \end{tabular}
-    \caption{\label{deap_final_results} Background and WIMP removal rates for the highest-accuracy hyperparameter configuration of each technique in DEAP-3600.}
+    \caption{\label{deap_final_results} Average background and WIMP/neutron removal rates for the highest-accuracy hyperparameter configuration of each technique in DEAP-3600.}
 \end{figure}

### 2019-02-28 documentation/report.tex
Updated report

@@
 
-These results confirm that I achieved my goal.
-While the conventional classifier in PICO-60 is believed to be perfectly accurate, my semi-supervised learning methods have significantly improved the accuracy that can be obtained quickly and without manual optimization.
-This should allow physicists working on future iterations of PICO to iterate more quickly, without modifying a conventional classifier whenever some aspect of the experiment changes.
+These results confirm that I met my goals on both the PICO-60 and DEAP-3600 experiments.
 
-The reduction in the false positive rate for DEAP-3600 provides evidence that machine learning can improve the efficiency of this experiment, possibly reducing the operation time required to collect sufficient evidence for or against the existence of WIMP dark matter.
+On PICO-60, the new semi-supervised learning algorithms performed extremely well, exceeding 98\% accuracy while permitting impure calibration data and requiring little manual analysis or optimization.
 
-Additionally, the unusual predictions made on real-world data provide important diagnostic information on the DEAP-3600 simulation (assuming that WIMPs are not present).
+On DEAP-3600, my cylindrical projection, combined with a 2D CNN, performed better than the conventional classifier on simulated data, reducing the WIMP misclassification rate from 91.0\% to 75.7\% while matching the conventional classifier's 99.6\% background removal rate.
+
+The predictions made on real-world data demonstrate clear potential.
+Differences between the simulated and real-world data provide potentially useful diagnostic information on the DEAP-3600 simulation (assuming that WIMPs are not present).
 

### 2019-03-01 documentation/report.tex
Minor grammar edits

@@
 
-My objective was to develop accurate machine learning algorithms that are more accurate, and I succeeded at this goal!
+My objective was to develop machine learning algorithms that are more accurate, and I succeeded at this goal!
 I composed a 28-page research whitepaper \cite{me} about my work on PICO-60, which has been reviewed and approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}.
@@
 
-For PICO-60, my intent was to develop a highly accurate classification algorithm that is intolerant to substantial impurities in calibration data.
+For PICO-60, my intent was to develop a highly accurate classification algorithm that is tolerant to substantial impurities in calibration data.
 I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.
@@
     \centering
-    \fbox{\includegraphics[width=0.95\textwidth]{pico_final_results}}
+    \fbox{\includegraphics[width=0.85\textwidth]{pico_final_results}}
     \caption{\label{pico_final_results} Validation accuracy, precision, recall, and loss for each technique in PICO-60. Error bars represent the 8th and 92nd percentiles.}
@@
 \begin{itemize}
-    \item Identification of neck alphas with 99.6\% precision (same as the conventional classifier).
+    \item Identification of neck alphas with a minimum of 99.6\% precision (the same as the previously developed conventional classifier).
     \item Reduction in the rate of false positives (simulated WIMP events misidentified as neck alphas) compared to the 91.0\% result achieved with a conventional classifier.
@@
         \hline
-        Conventional Classifier & Simulation & 99.6\% & 91.0\% \\
+        Conventional Classifier & Simulation (Val.) & 99.6\% & 91.0\% \\
         \hline
-        Dense Neural Network & Simulation & 100\% & 100\% \\
+        Dense Neural Network & Simulation (Val.) & 100\% & 100\% \\
         \hline
-        Cylindrical Projection & Simulation & 99.9\% & 76.7\% \\
+        Cylindrical Projection & Simulation (Val.) & 99.9\% & 76.7\% \\
         \hline
-        Topological CNN & Simulation & 99.9\% & 93.0\% \\
+        Topological CNN & Simulation (Val.) & 99.9\% & 93.0\% \\
         \hline

### 2019-03-01 documentation/report.tex
Renamed ICN

@@
 \begin{enumerate}
-    \item Iterative Cluster Nucleation is inspired by unsupervised clustering algorithms, and takes advantage of the ``confidence'' of the neural network's predictions (that is, how close the predictions are to 0 or 1).
+    \item Iterative Clustering is inspired by unsupervised clustering algorithms, and takes advantage of the ``confidence'' of the neural network's predictions (that is, how close the predictions are to 0 or 1).
 
@@
 
-    \item Gravitational Differentiation is an analog alternative to Iterative Cluster Nucleation.
+    \item Gravitational Differentiation is an analog alternative to Iterative Clustering.
 Using a novel piecewise exponential function for calculating final-layer derivatives, called $\mathrm{GravDiff}(p, \psi, g)$, unlabeled examples are caused to "gravitate" toward more accurate predictions based on the confidence of the neural network's predictions.

### 2019-03-01 visualization/visualize_overall_performance_pico.py

@@ -44,7 +44,7 @@ for array, mean in [(accuracy_high, accuracy_mean), (accuracy_low, accuracy_mean
             array[index] = np.abs(array[index] - mean[index])
 # The names of the configurations corresponding to all these lists
 CONFIGURATIONS = ['Conventional Clas.', 'Prev. ML Study', 'Banded FT', 'Full-Resolution FT', 'Waveform CNN', 'Image CNN',
-                  'I. Cluster Nucleation', 'Gravitational Diff.', 'Gravitational Diff. Test']
+                  'Iter. Clustering', 'Gravitational Diff.', 'Grav. Diff. Test']
 
 # Get the locations for the bars
 BAR_WIDTH = 0.2

### 2019-03-01 documentation/report.tex
Rewrote half of it

@@
 
-In all experiments for detection of WIMP dark matter, it is essential to develop a classifier that can distinguish potential WIMP events from background radiation, in particular alpha particles.
-Manually developing a classifier necessitates time-consuming physical modeling.
+In all experiments for detection of WIMP dark matter, it is essential to develop a classifier that can distinguish potential WIMP events from background radiation.
+Most often, classifiers are developed manually, via physical modeling and empirical optimization.
+This is problematic for two reasons: it takes a great deal of time and effort away from developing the experiment, and the resulting classifiers often perform suboptimally (which means that a greater amount of expensive run time is required to obtain a confident experimental result).
 
-Machine learning (ML) has the potential to automate this and accelerate experimentation, and also to detect patterns that humans cannot.
-However, impure calibration data hinders training of models, and unusual detector topologies make data challenging to process.
+Machine learning has the potential to automate this and accelerate experimentation, and also to detect patterns that humans cannot.
+However, two major challenges, which are shared among several dark matter experiments, stand in the way: impure calibration data, which hinders training of models, and unpredictable physical dynamics within the detector itself.
 
-I worked with two dark matter experiments: the PICO-60 bubble chamber \cite{pico}, and the DEAP-3600 liquid argon scintillator \cite{deap}.
-In PICO-60, background alpha and WIMP-like neutron calibration datasets are used for training; however, there is an impurity of 10\% alphas in the neutron set, hindering training.
-In DEAP-3600, the detector format, containing 255 spherically arranged photomultiplier tubes (PMTs), makes data processing challenging.
+My objective was to develop a set of machine learning techniques that resolve these two problems, and thus more efficiently generate highly accurate classifiers.
 
-Previously in PICO-60, a conventional classifier was developed and is thought to be 100\% accurate.
-A supervised neural network (NN) was also previously developed, achieving a mean accuracy of 80.2\% -- not usable as a practical replacement for conventional methods.
+I was able to obtain raw data for two dark matter experiments which exhibit these challenges: the PICO-60 bubble chamber \cite{pico}, and the DEAP-3600 liquid argon scintillator \cite{deap}.
 
-In DEAP-3600, a physical simulation was developed (prior to my work) to provide calibration data; physical limitations make it impractical to collect large amounts of calibration data.
-In the simulation, the conventional classifier removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
+In PICO-60, background alpha and WIMP-like neutron calibration datasets are used for training; however, there is an impurity of 10\% alphas in the neutron set.
+While a conventional classifier was developed (and is believed to be 100\% accurate), machine learning in the form of a supervised neural network (NN) has also been previously explored, because of the benefits of automation.
+Unfortunately, it achieved a mean accuracy of only 80.2\% -- not usable as a practical replacement for conventional methods in future iterations of the experiment.
 
-My objective was to develop machine learning algorithms that are more accurate, and I succeeded at this goal!
+In DEAP-3600, photons are absorbed and re-emitted by a wavelength shifting medium in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) arranged in a hexagonal lattice on the surface of the spherical detector.
+The unpredictability severely limits the accuracy of conventional classifiers; the best classifier so far removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
+Because of physical limitations, simulated data is used for calibration, with a limited amount of real-world experimental data available for testing.
+
 I composed a 28-page research whitepaper \cite{me} about my work on PICO-60, which has been reviewed and approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}.
@@
 
-\subsection{PICO-60}
+\subsection{Resolving Impurities (PICO-60)}
 
-For PICO-60, my intent was to develop a highly accurate classification algorithm that is tolerant to substantial impurities in calibration data.
+In the context of PICO-60, my intent was to develop a highly accurate classification algorithm that is tolerant to substantial impurities in calibration data.
 I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.
@@
 
-\subsection{DEAP-3600}
+\subsection{Resolving Unpredictable Photon Dynamics (DEAP-3600)}
 
-In DEAP-3600, the key challenge is the aforementioned unusual detector format: a sphere (with an opening on the top) tiled with a hexagonal lattice of PMTs.
-The major source of background radiation is ``neck alphas'', which strike atoms in the gas filling the neck above the spherical detector body.
-These must be eliminated from the experimental data set.
+In DEAP-3600, the key challenge is the aforementioned unpredictable photon dynamics.
+The major source of background radiation is ``neck alphas'', which strike atoms in the gas filling the neck above the spherical detector body; because of the unpredictable direction of photon emission, it is non-trivial to calculate the position of these events.
 
-I approached this problem in three different ways, listed below.
+Because of the inherently spatial nature of this problem, I hypothesized it would be important to most effectively process the spherical arrangement of PMTs in the DEAP-3600 detector.
+I evaluated three approaches to this problem (two of which are original), listed below.
 Techniques \ref{map_entry} and \ref{topological_entry} (Figure \ref{deap_methods}) take advantage of geometric and spatial patterns in PMT activations associated with neck alpha events, by extending the CNN concept for use in spherical and arbitrary topologies.
@@
     \item I developed and applied a Mercator-like cylindrical projection, which maps a sphere onto a rectangle.
-Cubic spline interpolation was used to map latitude lines on the sphere (which vary in length) to equal-length rows on the rectangular image.
+Cubic spline interpolation was used to map latitude lines on the hexagonal lattice (which vary in length) to equal-length rows on the rectangular image.
 Once again, photon counts were used as input data.
@@
 
-The three algorithms I developed were evaluated based on:
+I evaluated three algorithms for resolution of the unpredictable photon dynamics present in DEAP-3600. Two of the algorithms are original. The three algorithms I developed were evaluated based on:
 

### 2019-03-02 documentation/report.tex
Fixes

@@
 I was able to obtain raw data for two dark matter experiments which exhibit these challenges: the PICO-60 bubble chamber \cite{pico}, and the DEAP-3600 liquid argon scintillator \cite{deap}.
+For each experiment, I developed and compared three general-purpose algorithms intended to resolve its inherent challenge (impurity and unpredictable dynamics, respectively).
 
@@
 
-In DEAP-3600, photons are absorbed and re-emitted by a wavelength shifting medium in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) arranged in a hexagonal lattice on the surface of the spherical detector.
-The unpredictability severely limits the accuracy of conventional classifiers; the best classifier so far removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
+In DEAP-3600, photons are absorbed by a wavelength shifting medium and re-emitted in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) arranged in a hexagonal lattice on the surface of the spherical detector.
+The randomness severely limits the accuracy of conventional classifiers; the best classifier so far removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
 Because of physical limitations, simulated data is used for calibration, with a limited amount of real-world experimental data available for testing.
@@
 \end{enumerate}
-
-\begin{figure}[ht]
-    \centering
-    \subfloat{\fbox{\includegraphics[width=0.95\textwidth]{audio_cropped}}}
-    \qquad
-    \subfloat{\fbox{\includegraphics[width=0.35\textwidth]{image_grid}}}
-    \caption{Examples of the audio and image data.}
-\end{figure}
 
@@
 In DEAP-3600, the key challenge is the aforementioned unpredictable photon dynamics.
-The major source of background radiation is ``neck alphas'', which strike atoms in the gas filling the neck above the spherical detector body; because of the unpredictable direction of photon emission, it is non-trivial to calculate the position of these events.
+The major source of background radiation is ``neck alphas'', which strike atoms in the gas filling the neck above the spherical detector body; because of the unpredictable direction of photon emission, it is non-trivial to identify this class of events.
 
@@
 
-These results confirm that I met my goals on both the PICO-60 and DEAP-3600 experiments.
+These results confirm that I met my objective: my new techniques resolved the two fundamental challenges I identified, permitting significantly improved performance, requiring little manual analysis or optimization.
 
-On PICO-60, the new semi-supervised learning algorithms performed extremely well, exceeding 98\% accuracy while permitting impure calibration data and requiring little manual analysis or optimization.
+On PICO-60, the new semi-supervised learning algorithms performed extremely well, exceeding 98\% accuracy while nonetheless permitting impure calibration data.
 
 On DEAP-3600, my cylindrical projection, combined with a 2D CNN, performed better than the conventional classifier on simulated data, reducing the WIMP misclassification rate from 91.0\% to 75.7\% while matching the conventional classifier's 99.6\% background removal rate.
-
-The predictions made on real-world data demonstrate clear potential.
+Additionally, the predictions made on real-world data demonstrate clear potential.
 Differences between the simulated and real-world data provide potentially useful diagnostic information on the DEAP-3600 simulation (assuming that WIMPs are not present).

### 2019-03-02 documentation/report.tex
Formatting resolutions

@@
 \lfoot{\scriptsize January 2019}
-\cfoot{\scriptsize Brendon Matusch---Improving Particle Classification in WIMP Experiments}
+\cfoot{\scriptsize Improving Particle Classification in WIMP Experiments -- Brendon Matusch}
 \rfoot{\scriptsize Page~\thepage~of~\pageref{LastPage}}
@@
 
-In DEAP-3600, photons are absorbed by a wavelength shifting medium and re-emitted in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) arranged in a hexagonal lattice on the surface of the spherical detector.
-The randomness severely limits the accuracy of conventional classifiers; the best classifier so far removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
-Because of physical limitations, simulated data is used for calibration, with a limited amount of real-world experimental data available for testing.
+In DEAP-3600, photons are absorbed by a wavelength shifting medium and re-emitted in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) on the surface of the spherical detector.
+The randomness severely limits the accuracy of conventional classifiers; the best so far removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
+Because of physical limitations, simulated data is used for calibration, with 30 real-world experimental events available for testing.
 
-I composed a 28-page research whitepaper \cite{me} about my work on PICO-60, which has been reviewed and approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}.
-It is currently undergoing peer review for eventual publication in the journal \textit{Computer Physics Communications}.
+I have written a research paper \cite{me} about my work on PICO-60, which has been approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}.
+It is currently undergoing peer review for publication in \textit{Computer Physics Communications}.
 
@@
 
-In the context of PICO-60, my intent was to develop a highly accurate classification algorithm that is tolerant to substantial impurities in calibration data.
-I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.
+In the context of PICO-60, my intent was to develop a highly accurate classification algorithm
+
+that is tolerant to substantial impurities in calibration data. I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.
 
@@
 
-Because of the inherently spatial nature of this problem, I hypothesized it would be important to most effectively process the spherical arrangement of PMTs in the DEAP-3600 detector.
+Because of the inherently spatial nature of this problem, I hypothesized it would be important to most effectively process the arrangement of PMTs in the DEAP-3600 detector: a hexagonal lattice on the surface of a sphere.
 I evaluated three approaches to this problem (two of which are original), listed below.
@@
 
-I evaluated three algorithms for resolution of the unpredictable photon dynamics present in DEAP-3600. Two of the algorithms are original. The three algorithms I developed were evaluated based on:
+The three I developed to resolve the problem of unpredictable dynamics in DEAP-3600 were evaluated based on:
 

### 2019-03-02 documentation/report.tex
Final edits

@@
 
-My objective was to develop a set of machine learning techniques that resolve these two problems, and thus more efficiently generate highly accurate classifiers.
+My objective was to develop a set of machine learning techniques that address these two problems, and thus more efficiently generate highly accurate classifiers.
 
@@
 
-In DEAP-3600, photons are absorbed by a wavelength shifting medium and re-emitted in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) on the surface of the spherical detector.
-The randomness severely limits the accuracy of conventional classifiers; the best so far removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events.
+In DEAP-3600, photons are absorbed by a wavelength shifting medium and re-emitted in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) around the spherical detector.
+The randomness severely limits the accuracy of conventional classifiers; in a simulation, the best so far removes 99.6\% of alpha background, while also (undesirably) removing 91.0\% of WIMP events.
 Because of physical limitations, simulated data is used for calibration, with 30 real-world experimental events available for testing.
@@
 
-\subsection{PICO-60}
+\subsection{Resolving Impurities (PICO-60)}
 
@@
 
-\subsection{DEAP-3600}
+\subsection{Resolving Unpredictable Photon Dynamics (DEAP-3600)}
 
@@
 
-This successful network model was further tested on the limited amount of available real-world experimental data.
-It removed 73.9\% of WIMP-like (neutron calibration) events, which is in line with the 75.7\% obtained on simulated data.
+This successful network model was further tested on the limited amount of available real-world experimental data. It removed 73.9\% of WIMP-like (neutron calibration) events,
+
+which is in line with the 75.7\% obtained on simulated data.
 
@@
 
-These results confirm that I met my objective: my new techniques resolved the two fundamental challenges I identified, permitting significantly improved performance, requiring little manual analysis or optimization.
+These results confirm that I met my objective: my new techniques successfully addressed the two fundamental challenges I identified, permitting significantly improved performance, requiring little manual analysis or optimization.
 

### 2019-03-02 documentation/report.tex
Grammar edits

@@
     \item Gravitational Differentiation is an analog alternative to Iterative Clustering.
-Using a novel piecewise exponential function for calculating final-layer derivatives, called $\mathrm{GravDiff}(p, \psi, g)$, unlabeled examples are caused to "gravitate" toward more accurate predictions based on the confidence of the neural network's predictions.
+Using a novel piecewise exponential function for calculating final-layer derivatives, called $\mathrm{GravDiff}(p, \psi, g)$, unlabeled examples are caused to ``gravitate'' toward more accurate predictions based on the confidence of the neural network's predictions.
 
@@
 Semi-supervised learning produced better accuracy than supervised learning.
-Gravitational differentiation was the best overall model, achieving 99.7\% accuracy on validation data and 98.3\% on test data.
-This confirms my hypothesis; it is indeed beneficial to train on the neural network's most confident predictions.
+Gravitational Differentiation was the best overall technique, achieving 99.7\% accuracy on validation data and 98.3\% on test data.
+This confirms my hypothesis; it is indeed beneficial to incorporate the neural network's most confident predictions into its training data.
 
@@
 
-These results confirm that I met my objective: my new techniques successfully addressed the two fundamental challenges I identified, permitting significantly improved performance, requiring little manual analysis or optimization.
+These results confirm that I met my objective: my new techniques successfully addressed the two fundamental challenges I identified, permitting significantly improved performance while requiring little manual analysis or optimization.
 
@@
 Additionally, the predictions made on real-world data demonstrate clear potential.
-Differences between the simulated and real-world data provide potentially useful diagnostic information on the DEAP-3600 simulation (assuming that WIMPs are not present).
+Differences between the simulated and real-world data provide potentially important diagnostic information on the DEAP-3600 simulation (assuming that WIMPs are not present).
 

### 2019-07-15 documentation/paper.tex
Updated paper for resubmission

@@
 The current practice of manually developing a discriminator function to eliminate background events is difficult when available calibration data is frequently impure and present only in small quantities.
-In this study, several different discriminator input/preprocessing formats and neural network architectures are applied to the task.
+In this study, neural networks trained on calibration and background data are applied to the task.
 First, they are optimized in a supervised learning context.
-Next, two novel semi-supervised learning algorithms are trained, and found to replicate the Acoustic Parameter (AP) discriminator previously used in PICO-60 with a mean of 97\% accuracy.
+Next, semi-supervised learning algorithms are applied; a standard algorithm based on pseudo-labels is compared to a novel algorithm called gravitational differentiation.
+The latter is found to replicate the Acoustic Parameter (AP) discriminator previously used in PICO-60 with a mean of 97\% accuracy.
 \end{abstract}
@@
     \item In Section 3, we discuss various machine learning techniques that are applied to develop a new discriminator.
-    \item In Section 4, we document experimentation with supervised learning, applied to various data formats.
+    \item In Section 4, we document experimentation with supervised learning.
     \item In Section 5, we document experimentation with semi-supervised learning with the goal of improving data efficiency.
@@
 \begin{enumerate}
-    \item Determine the most effective machine learning architecture to use in development of a discriminator.
-Furthermore, develop several neural network architectures and apply them to different input/preprocessing formats (such as audio Fourier transforms, raw waveforms, and images), in order to find which are the most effective.
-    \item Determine whether these techniques can be extended, applying two original algorithms based on semi-supervised learning, to develop an effective discriminator function based on incomplete and inaccurate initial information.
-Furthermore, determine whether semi-supervised learning is of any benefit.
+    \item Develop a supervised learning system with the objective of enabling better performance an a previous machine learning study (discussed in Section \ref{og_analysis}).
+    \item Using the supervised learning model as a baseline, explore semi-supervised learning with additional unlabeled data.
+Compare the existing method of pseudo-labels with a novel method called gravitational differentiation.
 \end{enumerate}
@@
 
-These conversion, preprocessing, and data cutting steps were reapplied during this study, and are described in Section \ref{data_formats}.
+These conversion, preprocessing, and data cutting steps were reapplied during this study, and are described in Section \ref{data_format}.
 
@@
 
-The neural network in this study used the same $\beta_{8}$ input format as was applied in this study (detailed in Section \ref{piezo_derived}), with pre-trigger noise subtracted.
-Position corrections (also in Section \ref{piezo_derived}) were not applied; rather, the position of the event in the vessel is additionally used as input.
+The neural network in this study used the same $\beta_{8}$ input format as was applied in this study (detailed in Section \ref{data_format}), with pre-trigger noise subtracted.
+Position corrections (also in Section \ref{data_format}) were not applied; rather, the position of the event in the vessel is additionally used as input.
 AP was used to restrict the data used as input to the neural network.
@@
 Note that it is possible there are position-related biases present in this training data, because neutron events during calibration runs occur more frequently near the wall of the detector where the calibration source is placed.
-These biases are expected to be alleviated by application of position corrections which normalize the audio amplitude; these corrections are detailed in Section \ref{piezo_derived}.
+These biases are expected to be alleviated by application of position corrections which normalize the audio amplitude; these corrections are detailed in Section \ref{data_format}.
 
@@
 
-\subsection{Data Sources and Formats} \label{data_formats}
+\subsection{Data Format} \label{data_format}
 
-Discriminators can use a variety of different data formats to make predictions.
-These are collected from the two essential sensors present in the PICO-60 apparatus: two piezoelectric microphones (piezos) and cameras.
+The lowest-level data derived from the two piezos in the PICO-60 apparatus is the raw waveform.
+This consists of a series of 250,000 samples, collected at the rate of one per microsecond.
+The raw waveform was converted into the frequency domain by means of a one-dimensional Discrete Fourier Transform (DFT) for real input.
+Since DFTs produce sequences of complex numbers, the magnitude of each element was computed.
 
-\subsubsection{Piezo-Derived} \label{piezo_derived}
-
-\textit{Raw Waveform}
-
-The lowest-level data derived from the two piezos in the PICO-60 apparatus is the raw waveform $\omega$.
-This consists of a series of 250,000 samples, collected at the rate of one per microsecond.
-Each sample is represented as a 16-bit integer.
-Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because there is only background noise prior to this window, and a clipped signal produced by hydraulics repressurizing the vessel after.
-These sections contain no information and can be undesirably fitted by neural networks.
-(For context, approximately half of the samples are recorded prior to the camera trigger.)
-
-\textit{Fourier Transform}
-
-The raw waveform $\omega$ was converted into the frequency domain by means of a one-dimensional Discrete Fourier Transform (DFT) for real input.
-Since DFTs produce sequences of complex numbers, the magnitude of each element was computed.
-The direct output from such a DFT is the full-resolution Fourier transform $\beta _{50,001}$, which consists of a sequence of 50,001 numbers (half the length of the raw waveform $\omega$).
-
-Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original machine learning analysis) was computed by integrating the resonant energy over all frequencies within each of a set of eight frequency bands ranging from 1 kHz to 300 kHz \cite{pico}.
+Beyond this, the banded Fourier transform (which is the input to the AP function and all machine learning models) was computed by integrating the resonant energy over all frequencies within each of a set of eight frequency bands ranging from 1 kHz to 300 kHz \cite{pico}.
 This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.
 
-Both $\beta _{50,001}$ and $\beta_{8}$ were used as inputs to machine learning models applied in this study.
+The 3-dimensional position of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera.
+This is used in the banded frequency position correction function, which corrects the banded Fourier transform for variations in amplitude which depend on the position of the bubble.
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{audio}
-    \caption{\label{} An example of the audio waveform $\omega$ with initial background and hydraulic noise cropped (left) and the resulting full resolution Fourier transform $\beta_{50,001}$ (right) for both audio channels.
-Recorded on 2016-12-03 at 04:22:45 UTC.}
-\end{figure}
-
-\subsubsection{Camera-Derived} \label{camera_derived}
-
-\textit{Image Window Sequence}
-
-Bubble events are detected using the four cameras within the PICO-60 apparatus, with an image-based trigger.
-When this happens, each camera records a sequence of 41 images before and after formation.
-These raw images contain a large amount of extraneous information; they encompass the entire vessel.
-To reduce the input information, the image window sequence $\iota$ includes 50\texttimes50 cropped windows around the position of the bubble.
-The 41 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 frames immediately around the formation of the bubble.
-Two to three frames from before the recording trigger are included in $\iota$, because the bubble does not cause a trigger until it is already of a significant size.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.7\textwidth]{image_grid}
-    \caption{\label{} An example of the image sequence $\iota$. Recorded on 2016-09-23 at 22:15:44 UTC.}
-\end{figure}
-
-\textit{3D Position}
-
-The 3-dimensional position $\chi$ of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera.
-
-This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.
-$PosCor(\beta _{8}, \chi)$ is used for training several machine learning models throughout this study.
-Details of its use are in Section \ref{mlp}.
+The banded Fourier transform, after this correction, is henceforth referred to as $\beta$.
 
@@
     \item The event was recorded more than 25 seconds after reaching target pressure
-    \item The bubble position $\chi$ was successfully calculated ($[\chi_{X}, \chi_{Y}, \chi_{Z}]\neq[-100, -100, -100]$)
+    \item The bubble position was successfully calculated ($[X, Y, Z]\neq[-100, -100, -100]$)
 \end{itemize}
@@
 
-The complete wall cut {\it WallCut}$(D)$ is a composition of the fiducial cut {\it FidWallCut}$(D)$ (which makes use of the 3D position $\chi$), the pressure cut {\it PresWallCut}$(D)$ (which uses data from the fast pressure transducer), and the acoustic cut {\it AcWallCut}$(D)$ (which uses the banded Fourier transform $\beta _{8}$).
+The complete wall cut {\it WallCut}$(D)$ is a composition of the fiducial cut {\it FidWallCut}$(D)$ (which makes use of the 3D position), the pressure cut {\it PresWallCut}$(D)$ (which uses data from the fast pressure transducer), and the acoustic cut {\it AcWallCut}$(D)$ (which uses the banded Fourier transform $\beta$).
 Those cuts restrict data as follows:
@@
     \item The fiducial cut {\it FidWallCut}$(D)$ defines a spatial area along the walls of the vessel within which no events are accepted.
-It determines whether events fall within this area based on the visually calculated position $\chi$.
+It determines whether events fall within this area based on the visually calculated 3D position.
     \item The pressure cut {\it PresWallCut}$(D)$ restricts the pressure detected by the fast pressure transducer, without any position corrections, to be within a range of 0.3 of 1
-    \item The acoustic cut {\it AcWallCut}$(D)$ is defined using the banded Fourier transform $\beta_{8}$, and takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
+    \item The acoustic cut {\it AcWallCut}$(D)$ is defined using the banded Fourier transform $\beta$, and takes advantage of differences in the frequency distribution (specifically the first and second frequency bands of the first and third piezos) of wall events and non-wall events.
 \end{itemize}
@@
 
-The data set, after all of these cuts are applied, consists of 624 events. During each training run, 128 randomly selected events are set aside for validation, leaving a training set of 496 events.
+The data set, after all of these cuts are applied, consists of 624 events.
+During each training run, 128 randomly selected events are set aside for validation or testing (more details in Section \ref{optimization}), leaving a training set of 496 events.
 
@@
 
-Two original semi-supervised algorithms were developed and implemented for this study: gravitational differentiation (GD) and iterative cluster nucleation (ICN).
+Two semi-supervised algorithms were implemented for this study: the existing method pseudo-labels \cite{pseudo}, and a novel method, gravitational differentiation.
 Both are discussed in detail in Section \ref{semi_supervised}.
@@
 
-Performance of the machine learning systems was evaluated using two metrics: classification accuracy and class-wise standard deviation.
+Performance of the machine learning systems was evaluated using four metrics: classification accuracy, precision, recall, and mean squared error.
 
@@
 
-Class-wise standard deviation, abbreviated CWSD, captures how decisive the network is: whether its prediction is confidently high or low, or is only slightly closer to one edge of the spectrum.
+Mean squared error represents the ``confidence'' of the network's predictions: how close they are to 0 or 1.
+It is simply calculated as $\frac{1}{n} \sum _{i=1} ^n (p_i - y_i)^2$, where $p$ is the prediction vector, $y$ is the ground truth vector, and both have a length of $n$.
 
-Numerically, this is a variable defined below, where $N$ and $A$ are the sets of outputs of the supervised learning discriminator in question, corresponding to the sets of neutrons and alpha particles respectively.
-It calculates the spread of the network's predictions for each ground truth class.
-This means that a decisive discriminator, which produces a wide separation between the two classes, is preferred over one that produces a nebulous cloud of outputs with a seemingly arbitrary decision boundary.
+\subsection{Optimization Process} \label{optimization}
 
-\begin{equation}
-    S=std(N \cup A)
-\end{equation}
-\begin{equation}
-    \mbox{\it CWSD}=(std(N \div S) + std(A \div S)) \div 2
-\end{equation}
-
-The first step is to calculate the standard deviation $S$ of the union of $N$ and $A$.
-This gives an indication of the scale of the overall distribution.
-When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1.
-While the neural network's outputs are bounded in the range of 0 to 1 with a sigmoid activation on the last layer, AP has a significantly wider range.
-Normalization of the union prevents this from creating a bias where AP would produce a higher standard deviation with a similarly proportioned error.
-
-The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually.
-This is an indication of how tightly clustered or widely dispersed the discriminator's predictions are for each class.
-Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any given $x$ and $y$), with minimal variance off those specific values, will produce a low class-wise standard deviation.
-
-\subsection{Optimization Process}
-
-Each of the learning algorithms used in this study applies a neural network.
-For each of the general configurations (multi-layer perceptron, one-dimensional convolutional neural network, et cetera) there are many possible specific architectures with different hyperparameters, which have to be optimized to improve performance.
-Specific hyperparameters that were optimized include:
+For any neural network, there are many possible architectures with different hyperparameters, which have to be optimized to improve performance.
+Specific hyperparameters that were optimized in this study include:
 
@@
 Several of these hyperparameters are optimized at a time, using a grid search.
-Given $n$ different hyperparameters to optimize and $m$ different possible values for each, $m^n$ different networks are trained (using every possible combination of hyperparameters) and tested on a validation set.
-For space reasons, only a small subset of the tested configurations are displayed in tables throughout this paper.
+Given $n$ different hyperparameters to optimize and $m_i$ possible values for hyperparameter $i$, $\prod _{i=1} ^n m_i$ different networks are trained (using every possible combination of hyperparameters) and tested on a validation set.
 
-In general, the parameters for each grid search were chosen based on some initial empirical experimentation to determine the general range within the entire parameter space that work reasonably well.
-There was a limit to how many parameters could be tested, especially in the case of semi-supervised learning, as some of the grid searches took several days to run.
+In general, the parameters for each grid search were chosen based on some initial empirical experimentation to determine a range of parameters within the parameter space that works reasonably well.
+This range was then used as the search space for grid searches.
 
-Where not otherwise specified, the following hyperparameters were used universally:
-\begin{itemize}
-    \item Adam \cite{adam} stochastic optimizer
-    \item Mean squared error loss function
-    \item Batch size of 32
-    \item $tanh$ activation function
-\end{itemize}
+The best model out of a grid search was selected based on validation accuracy.
+For each run, 496 events were used for training, and 128 for validation.
+Each network configuration was trained and evaluated three times, and accuracy was averaged over the last ten epochs of these three runs.
 
-Accuracy statistics cited throughout the rest of this paper are averaged over late epochs (after accuracy has reached an asymptote) of several configurations in the highest-accuracy region of the hyperparameter search space, rather than over a single configuration.
-This alleviates the accuracy variance caused by randomly chosen validation sets.
+Finally, the best model selected on these criteria was re-trained and re-evaluated on a newly randomly selected test set of 128 examples.
+Like for validation sets, this process is repeated three times and results are averaged.
+All performance statistics cited throughout this paper are calculated on test data in this way.
 
@@
 
-The primary objective, in experimentation with supervised learning, is to determine the most effective combination of input format and neural network architecture to use in replicating a discriminator function.
-Three major configurations were tested:
+The primary objective, in experimentation with supervised learning, is to evaluate different combinations of hyperparameters to use in replicating a discriminator function.
+The resulting architecture will be used as a baseline against which to compare semi-supervised learning models.
 
-\begin{enumerate}
-    \item A convolutional neural network trained on the raw waveform $\omega$.
-    \item A dense neural network trained on banded Fourier transforms $\beta_{N}$.
-    \item A convolutional neural network trained on the image window data $\iota$.
-\end{enumerate}
+\subsection{Structure}
 
-\subsection{Convolutional Neural Network for Raw Waveform Analysis}
+A variety of network architectures were tested for this task.
+Most had a small number of dense layers (on the order of three), applied dropout and L2 regularization, and used batch normalization once again.
 
-\subsubsection{Structure}
-
-A convolutional neural network was used for analyzing the raw waveform $\omega$ directly, without any preprocessing whatsoever.
-It is furthermore referred to as $DeepConv(\omega)$.
-Direct processing of the waveform avoids any destruction of information; theoretically, a sufficiently complex neural network should be capable of deriving any characteristics needed.
-
-For this task, a very deep (20 layers in its shallowest configuration) one-dimensional fully convolutional neural network was applied.
-The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.
-L2 regularization was used to alleviate overfitting, and batch normalization was applied to the input to stabilize training.
-
-The following hyperparameters were tested during a grid search:
-\begin{itemize}
-    \item L2 Regularization $\lambda \in \{0.0003, 0.001, 0.003\}$
-    \item Dense Layer Dropout $\in \{0, 0.25, 0.5\}$
-    \item Convolutional Filters $\in \{24, 48\}$
-    \item Convolutional Kernel Size $\in \{3, 5\}$
-    \item Number of Convolutional Layers $\in \{17, 32\}$
-    \item Number of Dense Layers = 3 (with neuron counts 64, 16, and 1)
-\end{itemize}
-
-\subsubsection{Results}
-
-The most accurate configuration during a waveform CNN grid search was 93.8\% accurate and had a standard deviation of 0.43 (seen in Figure \ref{waveform_hist}).
-The average accuracy over 3 configurations in the best-performing region of the hyperparameter search space (with 24 filters, L2 $\lambda$ of 0.003, 32 convolutional layers, and a kernel size of 3) was 87.5\%.
-Performance statistics of these configurations are shown in Table \ref{deepconv_table}.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{waveform_hist}
-    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ configuration, relative to AP, on a randomly selected validation set.
-The black vertical line represents the AP decision boundary.}
-\end{figure}
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Performance of {\it DeepConv}$(\omega)$ configurations in the best-performing region of the hyperparameter search space.} \label{deepconv_table} 
-        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
-            \hline
-            L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Accuracy & Precision & Recall & CWSD \\
-            \hline
-            0.003 & 0 & 24 & 3 & 32 & 93.8\% & 94.0\% & 96.3\% & 0.50 \\
-            \hline
-            0.003 & 0.25 & 24 & 3 & 32 & 93.8\% & 94.0\% & 96.3\% & 0.43 \\
-            \hline
-            0.003 & 0.5 & 24 & 3 & 32 & 75.0\% & 71.9\% & 100.0\% & 0.71 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-% cat ../experimental_data/training_logs/2018_11_24_waveform_ap_similarity_last_epoch.out | grep filters24 | grep lambda0.003 | grep group6 | grep kernel_size3 | ./mean_ap_accuracy_across_configurations.py
-
-\subsection{Multi-Layer Perceptron for Fourier Transform Analysis} \label{mlp}
-
-\subsubsection{Structure}
-
-Preprocessing the audio by applying a Fourier transform may be advantageous.
-If the frequency distribution and overall amplitude are indeed the most important factors for discrimination, the network can gather them straight from the Fourier transform $\beta_{N}$ rather than having to analyze $\omega$ to extract this information.
-
-In addition to the banded Fourier transform $\beta_{8}$ used in the original PICO-60 analysis, the full-resolution $\beta_{50,001}$ was input into a neural network directly, without any downscaling.
-
-When $\beta_{8}$ was used as input, position corrections were applied to the input data, as they were used for the input to Acoustic Parameter.
-(With the position correction method currently used for PICO-60, it is currently not practical to apply it to other input formats.)
-
-(Side note: intuitively, it would make sense to provide the position $\chi$ directly to the neural network when corrections cannot be applied.
-However, this cannot presently be done because $\chi$ is a bias for the calibration run type, as described in Section \ref{og_analysis}; neutron sources are placed on one side of the vessel, so neutron events occur more often near that side.)
-
-The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$ and \\ $FourierMLP(PosCor(\beta_{N}, \chi))$.
-
-While a variety of network architectures were tested, most had a small number of dense layers (on the order of three), applied dropout and L2 regularization, and used batch normalization once again.
-
-The following hyperparameters were tested during a grid search:
+The following hyperparameter range was selected based on manual empirical testing, and was thus exhaustively searched:
 \begin{itemize}
@@
     \item Number of Dense Layers $\in \{2, 3, 4\}$
+    \item Optimizer = Adam \cite{adam}
+    \item Batch Size = 32
 \end{itemize}
 
-\subsubsection{Results}
-
-It became evident very quickly that a high resolution was not required to get a high accuracy relative to the ground truth data.
-{\it FourierMLP}$(PosCor(\beta_{8}))$ managed a mean of 97.0\% validation accuracy over the best-performing region of the hyperparameter search space, and a mean class-wise standard deviation of 0.26.
-Its fairly decisive prediction distribution can be observed in Figure \ref{banded_no_pos_input_hist}.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{banded_no_pos_input_hist}
-    \caption{\label{banded_no_pos_input_hist} Prediction distribution of the best {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ configuration.}
-\end{figure}
-
-Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators from the best-performing region of the hyperparameter search space.
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Performance of {\it FourierMLP}$(\beta_{N})$ configurations in the best-performing region of the hyperparameter search space.} \label{fft_wall_cuts}
-        \begin{tabular}{|l|l|l|l|l|l|l|l|}
-            \hline
-            Input Format & L2 $\lambda$ & Dropout & Layers & Accuracy & Precision & Recall & CWSD \\
-            \hline
-            $\beta_8$ & 0 & 0.5 & 2 & 98.1\% & 97.1\% & 100.0\% & 0.24 \\
-            \hline
-            $\beta_8$ & 0.001 & 0.5 & 2 & 96.6\% & 94.7\% & 99.6\% & 0.27 \\
-            \hline
-            $\beta_8$ & 0.003 & 0.5 & 2 & 96.3\% & 94.3\% & 100.0\% & 0.28 \\
-            \hline
-            $\beta_{50,001}$ & 0.001 & 0 & 2 & 89.1\% & 84.0\% & 100.0\% & 0.47 \\
-            \hline
-            $\beta_{50,001}$ & 0.001 & 0 & 3 & 92.2\% & 88.8\% & 100.0\% & 0.44 \\
-            \hline
-            $\beta_{50,001}$ & 0.001 & 0 & 4 & 93.0\% & 89.8\% & 100.0\% & 0.40 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-% cat ../experimental_data/training_logs/2018_12_01_banded_all_examples_ap_similarity_last_300_epochs.out | grep dropout0.5 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_11_24_high_resolution_ap_similarity_last_epoch.out | grep dropout0/ | grep l2_lambda0.001 | ./mean_ap_accuracy_across_configurations.py
-
-\subsection{Convolutional Neural Network for Image Window Analysis}
-
-\subsubsection{Structure}
-
-It is an open question whether or not there is any information in the image data $\iota$ that could be used to distinguish between particle classes.
-Relative to the extremely short period of time in which the bubble forms (on the scale of nanoseconds), the framerate of the camera (340Hz) is extremely slow.
-While the very early stages of bubble formation (when the sound is produced) are known to differ depending on whether the bubble was created by a nuclear recoil or an alpha particle, it is not experimentally known whether any visually apparent differences persist when the bubble is visible.
-
-In an effort to resolve this, a two-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$, referred to as {\it ImageConv}$(\iota)$.
-The network architecture consists of a moderate number of convolutional layers (on the scale of nine) and three dense layers at the end.
-L2 regularization was used on all layers, and dropout was additionally used on the dense layers.
-
-The following hyperparameters were tested during a grid search:
-\begin{itemize}
-    \item Dropout $\in \{0, 0.25, 0.5\}$
-    \item L2 Regularization $\lambda \in \{0, 0.0003, 0.001, 0.003, 0.006, 0.01\}$
-    \item Number of Convolutional Layers $\in \{6, 9, 12\}$
-    \item Number of Dense Layers = 3 (with neuron counts 64, 16, and 1)
-\end{itemize}
-
-\subsubsection{Results} \label{super_results}
-
-Every network architecture tested produced validation accuracy between 62\% and 63\%.
-As seen in Figure \ref{image_hist}, all examples were given the exact same prediction.
-This provides strong evidence that images provide insufficient information for any kind of meaningful discrimination.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{image_hist}
-    \caption{\label{image_hist} Prediction distribution of an {\it ImageConv}$(\iota)$ configuration. (53 out of 54 configurations had exactly the same accuracy.)}
-\end{figure}
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Performance of {\it ImageConv}$(\iota)$ configurations.} \label{imageconv_table}
-        \begin{tabular}{|l|l|l|l|l|l|l|}
-            \hline
-            L2 $\lambda$ & Dropout & Convolutional Layers & Accuracy & Precision & Recall & CWSD \\
-            \hline
-            0 & 0 & 6 & 62.7\% & 62.9\% & 99.5\% & 1.00 \\
-            \hline
-            0.0003 & 0 & 12 & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
-            \hline
-            0.001 & 0 & 9 & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-
-In Table \ref{imageconv_table}, high standard deviations across the board are further evidence for the hypothesis that there is insufficient information in the image data.
-The 100\% recall statistics are not due to high classification accuracy, but rather due to the network making predictions less than 0.5 for all examples of both classes.
-
-\subsection{Performance Comparison}
-
-Table \ref{super_overview} compares the mean accuracy, precision, recall, and class-wise standard deviation values (of the highest-performance region of the hyperparameter search space) for each of the input formats.
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Comparison of supervised learning input formats, alongside the original PICO-60 ML analysis and AP.} \label{super_overview}
-        \begin{tabular}{|l|l|l|l|l|}
-            \hline
-            Configuration & Accuracy & Precision & Recall & CWSD \\
-            \hline
-            {\it DeepConv}$(\omega)$ & 87.5\% & 71.9\% & 100.0\% & 0.55 \\
-            \hline
-            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ & 97.0\% & 95.4\% & 99.9\% & 0.26 \\
-            \hline
-            {\it FourierMLP}$(\beta_{50,001})$ & 91.4\% & 87.8\% & 100.0\% & 0.44 \\
-            \hline
-            {\it ImageConv}$(\iota)$ & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
-            \hline
-            PICO-60 ML & 80.7\% & 76.9\% & 100.0\% & 0.65 \\
-            \hline
-            AP & 100\% & 100\% & 100\% & 0.43 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-
-The key conclusion to be drawn from Table \ref{super_overview} is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format.
-The raw waveform and high-resolution Fourier transform make performance worse.
-This may be caused by one of the following three reasons:
-
-\begin{itemize}
-    \item Insufficient network depth to process higher resolutions
-    \item Overfitting on extraneous data
-    \item Position corrections are essential to good performance
-\end{itemize}
-
-The last reason is very likely.
-At a fundamental level, $\beta _8$ does not contain any more information than $\beta _{50,001}$, so there is no obvious reason why (with the right network architecture) it should not perform at least as well.
-While repurposing the position corrections to work on arbitrary-resolution data is outside the scope of this study, it seems a promising subject of future investigation.
-
-Based on this result, experimentation with semi-supervised learning proceeded using \\ {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$, applying a three-layer perceptron architecture.
-There were two main reasons for this:
-
-\begin{enumerate}
-    \item This was a highly effective general configuration, producing a mean of 97.0\% accuracy and a class-wise standard deviation of 0.26, much better than the original PICO-60 neural network on that data set.
-    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because alpha and neutron events have different position distributions (as described in Section \ref{og_analysis}).
-\end{enumerate}
+Results from this grid search are discussed in Section \ref{results}.
 
@@
 
-Two techniques for semi-supervised learning have been developed for this study: gravitational differentiation and iterative cluster nucleation.
+Two techniques for semi-supervised learning have been used for this study: pseudo-labels (an existing method), and gravitational differentiation (a new method).
 The essence of both algorithms is a positive feedback loop.
@@
 
+As mentioned above, all data in the PICO-60 experiment is labeled.
+Thus, the key application of these algorithms is not in the PICO-60, but rather in future WIMP detection experiments where not all data is labeled, or where part of the data contains a large proportion of impurities and it is thus beneficial to train without those labels.
+
 Since the intent of these techniques is to separate particle types (the same as AP), performance evaluations (accuracy and class-wise standard deviation) use AP as a baseline.
@@
 
-To ensure that accuracy statistics are as consistent and reliable as possible, every configuration was trained three times, with a different randomly selected validation set each time.
-(Using more than three training runs would have taken prohibitive amounts of time.) Accuracy, precision, recall, and class-wise standard deviation statistics are the mean over those three tests.
+\subsection{Pseudo-Labels}
+
+\subsubsection{Concept}
+
+Pseudo-labels is an existing semi-supervised learning algorithm developed by Lee \cite{pseudo}.
+It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
+
+The basic concept is as follows: Starting with a relatively small amount of labeled data from two classes (nuclear recoils and alpha particles, in this case), predictions are run on a set of unlabeled data, and those predictions are used to produce further training data.
+
+Specifically, labels are calculated based on the confidence of the predictions on unlabeled data.
+If the confidence of a prediction is within a certain threshold $j$ of 0 or 1, a binary \textit{pseudo-label} is generated and used to label the training example.
+
+\subsubsection{Algorithm}
+
+The algorithm is initialized as follows:
+
+\begin{enumerate}
+    \item Take a subset of the training data, referred to as the seed set $\varsigma$. Use this as the beginning of the training set.
+    \item Remove classifications from any other available training examples. These create the unlabeled set $\upsilon$.
+    \item Compile and randomly initialize the weights of a binary classification neural network $NN(x)$.
+\end{enumerate}
+
+Parameterized by the initial seed threshold $j$ (on the order of 0.01) and the seed threshold multiplier $k$ (on the order of 1.05), the algorithm follows these iterative steps:
+
+\begin{enumerate}
+    \item Train $NN(x)$ for 30 epochs on $\varsigma$, using the imperfect ground truth values available.
+    \item Using the partially trained weights of $NN(x)$, run inference on the entirety of $\upsilon$, producing a set of predictions $p$.
+    \item Find predictions within $p$ that are within a distance of $j$ of either 0 or 1; as this is a binary classifier, such a prediction represents high confidence.
+Remove any such examples from $\upsilon$ and add them to $\varsigma$.
+The principle is that, when $NN(x)$ is trained on a relatively small set for short period of time, the few predictions within $p$ that are very confident are highly likely to be correct.
+    \item If no examples have been removed from $\upsilon$ and added to $\varsigma$, multiply $j$ by $k$ in place.
+This is done to increase the acceptance rate later in the training process, when most easily classifiable examples have been added to $\varsigma$.
+Otherwise, gridlock would occur, where certain examples could not be confidently classified given the training set.
+\end{enumerate}
 
@@
 Gravitational differentiation (GD) is a novel technique, developed during this study, for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.
+
+The motivation for this method is the fact that pseudo-labels are entirely binary; they are 0 or 1 regardless of the confidence of the prediction that produced them (as long as the prediction is within the defined threshold).
+In contrast, gravitational differentiation is an analog method in which more confident predictions result in progressively greater influences on the training process.
 
@@
 
-\subsubsection{Results}
-
-As with supervised learning techniques, a grid search was run for various parameters of the gravitational differentiation algorithm, in addition to the stochastic gradient descent learning rate (a multiplier that defines how quickly and precisely optimization occurs).
-Table \ref{grav_accuracy_by_varsigma} shows the mean asymptotic accuracy over the best-performing region of the hyperparameter search space, for each size of the initial labeled set $\varsigma$.
-Note that optimal hyperparameters vary by set size.
-
-It is evident that larger labeled sets produce better performance.
-Gravitational differentiation with 256 examples is considerably better than any supervised learning system.
-Additionally, note that a $\varsigma$ size of 128 produces accuracy of 97.0\%, the same as the banded supervised system with all 496 labeled examples.
-That is, GD achieves the same accuracy with 74\% of the labels removed.
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Gravitational differentiation accuracy statistics by size of $\varsigma$.} \label{grav_accuracy_by_varsigma}
-        \begin{tabular}{|l|l|l|l|l|l|}
-            \hline
-            $\varsigma$ Size & Proportion of Data Labeled & Accuracy & Precision & Recall & CWSD \\
-            \hline
-            32 & 6\% & 71.3\% & 70.1\% & 98.0\% & 0.73 \\
-            \hline
-            64 & 13\% & 79.7\% & 75.5\% & 98.6\% & 0.63 \\
-            \hline
-            128 & 26\% & 97.0\% & 96.3\% & 99.6\% & 0.23 \\
-            \hline
-            256 & 52\% & 99.2\% & 99.0\% & 99.7\% & 0.11 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples256 | grep rate0.03 | grep power3 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples128 | grep rate0.03 | grep increment0.0005 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples64 | grep power11 | grep rate0.001 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples32 | grep power11 | grep increment0.001 | ./mean_ap_accuracy_across_configurations.py
-
-Table \ref{grav_overview} represents the performance of several of the individual configurations incorporated into the mean accuracy values in Table \ref{grav_accuracy_by_varsigma}.
-All accuracy statistics are the mean over three tests of the same configuration.
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Performance of gravitational differentiation configurations in the best-performing region of the hyperparameter search space.} \label{grav_overview}
-        \begin{tabular}{|l|l|l|l|l|l|l|l|}
-            \hline
-            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Accuracy & Precision & Recall & CWSD \\
-            \hline
-            32 & 0.001 & 0.001 & 11 & 79.0\% & 78.2\% & 94.8\% & 0.74 \\
-            \hline
-            32 & 0.001 & 0.003 & 11 & 59.9\% & 59.9\% & 100.0\% & 0.78 \\
-            \hline
-            64 & 0.005 & 0.001 & 11 & 82.8\% & 79.0\% & 99.1\% & 0.58 \\
-            \hline
-            64 & 0.008 & 0.001 & 11 & 73.9\% & 71.0\% & 99.6\% & 0.71 \\
-            \hline
-            128 & 0.0005 & 0.03 & 3 & 98.3\% & 97.7\% & 99.6\% & 0.21 \\
-            \hline
-            128 & 0.0005 & 0.03 & 9 & 96.6\% & 95.1\% & 99.6\% & 0.29 \\
-            \hline
-            256 & 0.008 & 0.03 & 3 & 99.1\% & 99.0\% & 99.5\% & 0.16 \\
-            \hline
-            256 & 0.001 & 0.03 & 3 & 99.7\% & 99.6\% & 99.6\% & 0.05 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{grav_grid_search}
-    \caption{\label{grav_grid_search} Prediction distribution of the best gravitational differentiation configuration.}
-\end{figure}
-
-Note that in Figure \ref{grav_grid_search}, AP is normalized using a similar method to the original PICO-60 study \cite{pico}.
-In this normalized spectrum, the decision boundary between neutrons and alpha particles is at 0.25.
-
-\subsection{Iterative Cluster Nucleation}
-
-\subsubsection{Concept}
-
-Iterative cluster nucleation (ICN) is a second semi-supervised learning algorithm developed as part of this study.
-It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
-
-The basic concept is similar to that of gravitational differentiation.
-Starting with a relatively small amount of labeled data from two classes (nuclear recoils and alpha particles, in this case), predictions are run on a set of unlabeled data, and those predictions are used to produce further training data.
-However, there are three key differences:
-
-\begin{enumerate}
-    \item In iterative cluster nucleation, an unlabeled example is not included in the training process at all, until a very confident prediction is made on it.
-Once that occurs, it is added to the training set, expanding one of the two ``clusters'' of training data.
-    \item Once an example is added to the active training set, it is never removed.
-This ensures that the highly confident predictions made at the very beginning continue to influence the training process.
-    \item The unlabeled examples, once added to the active training set, are weighted just as strongly as examples from the original labeled data set.
-This ensures imperfect labels in the labeled data do not outweigh the confident and more likely correct predictions from later in the training process.
-\end{enumerate}
-
-\subsubsection{Algorithm}
-
-The algorithm is initialized as follows:
-
-\begin{enumerate}
-    \item Take a subset of the training data, referred to as the seed set $\varsigma$. Use this as the beginning of the training set.
-    \item Remove classifications from any other available training examples. These create the unlabeled set $\upsilon$.
-    \item Compile and randomly initialize the weights of a binary classification neural network $NN(x)$.
-\end{enumerate}
-
-Parameterized by the initial seed threshold $j$ (on the order of 0.01) and the seed threshold multiplier $k$ (on the order of 1.05), the algorithm follows these iterative steps:
-
-\begin{enumerate}
-    \item Train $NN(x)$ for 30 epochs on $\varsigma$, using the imperfect ground truth values available.
-    \item Using the partially trained weights of $NN(x)$, run inference on the entirety of $\upsilon$, producing a set of predictions $p$.
-    \item Find predictions within $p$ that are within a distance of $j$ of either 0 or 1; as this is a binary classifier, such a prediction represents high confidence.
-Remove any such examples from $\upsilon$ and add them to $\varsigma$.
-The principle is that, when $NN(x)$ is trained on a relatively small set for short period of time, the few predictions within $p$ that are very confident are highly likely to be correct.
-    \item If no examples have been removed from $\upsilon$ and added to $\varsigma$, multiply $j$ by $k$ in place.
-This is done to increase the acceptance rate later in the training process, when most easily classifiable examples have been added to $\varsigma$.
-Otherwise, gridlock would occur, where certain examples could not be confidently classified given the training set.
-\end{enumerate}
-
-\subsubsection{Results}
-
-In Table \ref{icn_accuracy_by_varsigma}, it is clear that for a labeled set of 256, ICN performs worse than GD (which reached 99.2\%).
-However, the accuracy for 128 examples in $\varsigma$ is almost exactly the same as that for 256, and is still higher than the 97.0\% reached with supervised learning on 496 training examples (see Section \ref{super_results}).
-Moreover, for even smaller training sets of 64 and 32, ICN continues to perform better than GD.
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Iterative cluster nucleation accuracy statistics by initial size of $\varsigma$.} \label{icn_accuracy_by_varsigma}
-        \begin{tabular}{|l|l|l|l|l|l|}
-            \hline
-            Initial $\varsigma$ Size & Proportion of Data Labeled & Accuracy & Precision & Recall & CWSD \\
-            \hline
-            32 & 6\% & 85.2\% & 83.4\% & 98.4\% & 0.54 \\
-            \hline
-            64 & 13\% & 94.3\% & 92.2\% & 99.8\% & 0.34 \\
-            \hline
-            128 & 26\% & 98.2\% & 97.7\% & 99.4\% & 0.21 \\
-            \hline
-            256 & 52\% & 98.2\% & 97.6\% & 99.5\% & 0.22 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-% cat ../experimental_data/training_logs/2018_10_26_icn_ap_similarity_last_300_epochs.out | grep examples256 | grep dropout0.5 | grep lambda0.001/ | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_icn_ap_similarity_last_300_epochs.out | grep examples128 | grep threshold0.02 | grep multiplier1.05 | grep dropout0.5 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_icn_ap_similarity_last_300_epochs.out | grep examples64 | grep threshold0.02 | grep lambda0/ | grep multiplier1.05 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_icn_ap_similarity_last_300_epochs.out | grep examples32 | grep lambda0/ | grep dropout0/ | ./mean_ap_accuracy_across_configurations.py
-
-Table \ref{icn_overview} represents the performance of a variety of high-accuracy iterative cluster nucleation configurations tested during the grid search.
-As with gravitational differentiation, the accuracy statistic is the mean over three tests of the same configuration.
-
-\begin{minipage}{\textwidth}
-    \begin{center}
-        \captionof{table}{Performance of iterative cluster nucleation configurations.} \label{icn_overview}
-        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
-            \hline
-            Initial $\varsigma$ Size & L2 $\lambda$ & Dropout & Initial $j$ & $k$ & Accuracy & Precision & Recall & CWSD \\
-            \hline
-            32 & 0 & 0 & 0.01 & 1.025 & 94.9\% & 94.2\% & 98.3\% & 0.40 \\
-            \hline
-            32 & 0 & 0 & 0.02 & 1.05 & 79.7\% & 79.0\% & 96.1\% & 0.66 \\
-            \hline
-            64 & 0 & 0 & 0.02 & 1.05 & 95.4\% & 93.0\% & 99.9\% & 0.34 \\
-            \hline
-            64 & 0 & 0.5 & 0.02 & 1.05 & 93.5\% & 91.2\% & 100.0\% & 0.31 \\
-            \hline
-            128 & 0.001 & 0.5 & 0.02 & 1.05 & 97.7\% & 97.4\% & 99.2\% & 0.21 \\
-            \hline
-            128 & 0 & 0.5 & 0.02 & 1.05 & 98.5\% & 98.0\% & 99.6\% & 0.21 \\
-            \hline
-            256 & 0.001 & 0.5 & 0.01 & 1.025 & 97.5\% & 96.7\% & 99.3\% & 0.28 \\
-            \hline
-            256 & 0.001 & 0.5 & 0.02 & 1.05 & 98.4\% & 97.6\% & 100.0\% & 0.20 \\
-            \hline
-        \end{tabular}
-    \end{center}
-\end{minipage}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{icn_grid_search}
-    \caption{\label{icn_grid_search} Prediction distribution of the best iterative cluster nucleation configuration.}
-\end{figure}
-
-\subsection{Performance Comparison}
+\section{Results} \label{results}
 

### 2019-07-22 documentation/paper.tex
Some minor changes

@@
 Before AP can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$).
-It must also be preprocessed with the position correction function (discussed in Section \ref{camera_derived}), which adjusts the overall amplitude of an audio recording according to its position in the vessel.
+It must also be preprocessed with the position correction function (discussed in Section \ref{data_format}), which adjusts the overall amplitude of an audio recording according to its position in the vessel.
 This is done because the acoustic characteristics differ significantly depending on the position of the event in the vessel \cite{pico}.
@@
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
     \centering
@@
     \item Optimizer = Adam \cite{adam}
+    \item Loss = Mean Squared Error (found empirically to outperform binary cross-entropy)
+    \item Activation = tanh (found empirically to outperform ReLU)
     \item Batch Size = 32
@@
 
-Results from this grid search are discussed in Section \ref{results}.
+As discussed in Section \ref{optimization}, the best model was selected from this grid search based on accuracy and tested. Results from all grid searches are discussed in Section \ref{results}.
 
@@
     \item Remove classifications from any other available training examples. These create the unlabeled set $\upsilon$.
-    \item Compile and randomly initialize the weights of a binary classification neural network $NN(x)$.
+    \item Compile and randomly initialize the weights of a binary classification neural network $N(x)$.
 \end{enumerate}
@@
 \begin{enumerate}
-    \item Train $NN(x)$ for 30 epochs on $\varsigma$, using the imperfect ground truth values available.
-    \item Using the partially trained weights of $NN(x)$, run inference on the entirety of $\upsilon$, producing a set of predictions $p$.
+    \item Train $N(x)$ for 30 epochs on $\varsigma$, using the imperfect ground truth values available.
+    \item Using the partially trained weights of $N(x)$, run inference on the entirety of $\upsilon$, producing a set of predictions $p$.
     \item Find predictions within $p$ that are within a distance of $j$ of either 0 or 1; as this is a binary classifier, such a prediction represents high confidence.
 Remove any such examples from $\upsilon$ and add them to $\varsigma$.
-The principle is that, when $NN(x)$ is trained on a relatively small set for short period of time, the few predictions within $p$ that are very confident are highly likely to be correct.
+The principle is that, when $N(x)$ is trained on a relatively small set for short period of time, the few predictions within $p$ that are very confident are highly likely to be correct.
     \item If no examples have been removed from $\upsilon$ and added to $\varsigma$, multiply $j$ by $k$ in place.
@@
 
-This function is applied in a training algorithm that is parameterized by the set of imperfectly labeled data $\varsigma$, the set of unlabeled training data $\upsilon$, a binary classification neural network $NN(x)$ (where $x$ is the input data format), and the gravitational multiplier increment $\delta_g$.
+This function is applied in a training algorithm that is parameterized by the set of imperfectly labeled data $\varsigma$, the set of unlabeled training data $\upsilon$, a binary classification neural network $N(x)$ (where $x$ is the input data format), and the gravitational multiplier increment $\delta_g$.
 The distortion power $\psi$ is assumed to be constant throughout a training run, and the gravitational multiplier $g$ is initialized at 0, gradually increasing during training.
@@
 \begin{enumerate}
-    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
-    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $\mathrm{GravDiff}(p, \psi, g)$ and record the resulting gradients for the next iteration.
+    \item Train $N(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
+    \item Calculate predictions $p$ with $N(x)$ on the entirety of $\upsilon$. Calculate $\mathrm{GravDiff}(p, \psi, g)$ and record the resulting gradients for the next iteration.
     \item Increment $g$ by $\delta_{g}$.
@@
 
-This system was trained using a multi-layer perceptron with an 8-band Fourier transform input for $NN(x)$, because this was one of the most successful configurations in supervised learning.
+This system was trained using a multi-layer perceptron with an 8-band Fourier transform input for $N(x)$, because this was one of the most successful configurations in supervised learning.
 It produced the highest accuracy as well as the lowest class-wise standard deviation.
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
     \centering
@@
 
-Implementation code for all algorithms discussed in this paper can be found at \url{https://github.com/brendon-ai/dark-matter}.
+Implementation code for all algorithms discussed in this paper can be found at \url{https://github.com/bfmat/dark-matter}.
 

### 2019-07-25 documentation/paper.tex
Work on paper

@@
 
-Two semi-supervised algorithms were implemented for this study: the existing method pseudo-labels \cite{pseudo}, and a novel method, gravitational differentiation.
+Two semi-supervised algorithms were implemented for this study: the existing method of pseudo-labels \cite{pseudo}, and a novel method, gravitational differentiation.
 Both are discussed in detail in Section \ref{semi_supervised}.
@@
     \item Dropout $\in \{0, 0.25, 0.5\}$
-    \item L2 Regularization $\lambda \in \{0, 0.0003, 0.001, 0.003, 0.01\}$
-    \item Number of Dense Layers $\in \{2, 3, 4\}$
+    \item L2 regularization $\lambda \in \{0, 0.0003, 0.001, 0.003, 0.01\}$
+    \item Number of dense layers $\in \{2, 3, 4\}$
     \item Optimizer = Adam \cite{adam}
-    \item Loss = Mean Squared Error (found empirically to outperform binary cross-entropy)
+    \item Loss = mean squared error (found empirically to outperform binary cross-entropy)
     \item Activation = tanh (found empirically to outperform ReLU)
-    \item Batch Size = 32
+    \item Batch size = 32
 \end{itemize}
@@
 Because all data in the PICO-60 experiment is labeled, unlabeled training data sets are randomly sampled from the 496 training examples remaining (after the validation set is selected).
-Labeled set sizes of 32, 64, 128, and 256 were tested; the remainder of the 496 examples would be used as unlabeled data.
+Labeled set sizes of 128 and 256 were tested; the remainder of the 496 examples would be used as unlabeled data.
 
@@
 
-It is a transformation of the hyperbolic tangent that flattens the central range and comparatively exaggerates the asymptote on either side. The equation above can be described in the following steps:
+It is a transformation of the hyperbolic tangent that flattens the central range and comparatively exaggerates the asymptote on either side, as visualized in Figure \ref{grav_diff}.
+The equation above can be described in the following steps:
 
@@
 
-This system was trained using a multi-layer perceptron with an 8-band Fourier transform input for $N(x)$, because this was one of the most successful configurations in supervised learning.
-It produced the highest accuracy as well as the lowest class-wise standard deviation.
-
-\begin{figure}[!ht]
+\begin{figure}[H]
     \centering
-    \includegraphics[width=\textwidth]{grav_diff}
+    \includegraphics[width=0.85\textwidth]{grav_diff}
     \caption{\label{grav_diff} A visualization of $\mathrm{GravDiff}(p, \psi, g)$ with respect to $p$ and $\psi$.}
@@
 
-Table \ref{semi_success_overview} directly compares performance of all semi-supervised learning configurations, with supervised learning, the original PICO-60 ML analysis, and AP provided as baselines for comparison.
-For every $\varsigma$ size except 32, ICN performs better than supervised learning; for 128 and 256, GD also performs better.
-For 32, 64, and 128 labeled examples, ICN proves to be more effective than GD; however, when 256 examples are labeled, GD provides the best accuracy, at 99.2\%.
-
-Perhaps most notably, iterative cluster nucleation with only 128 labeled examples achieves accuracy 1.2 percentage points higher than supervised learning with all 496 training examples labeled.
-It also makes more confident predictions (with a class-wise standard deviation of only 0.21, as opposed to 0.26).
-
-Note that the supervised multi-layer perceptron trained on $\beta _8$ originally used labels for all 496 training examples.
-Here, it is re-trained with smaller training sets so that it can be more directly compared to semi-supervised learning (which does not have access to labels for the entire set).
+Table \ref{results_table} compares performance on test data of the best-performing architectures for each algorithm and training set size.
+These architectures were selected based on accuracy on validation data.
 
@@
     \begin{center}
-        \captionof{table}{Summary of the most successful semi-supervised learning models.} \label{semi_success_overview}
         \begin{tabular}{|l|l|l|l|l|l|}
             \hline
-            Initial $\varsigma$ Size & Technique & Accuracy & Precision & Recall & CWSD \\
+            Algorithm & Initial $\varsigma$ Size & Accuracy & Precision & Recall \\
             \hline
-            32 & GD & 71.3\% & 70.1\% & 98.0\% & 0.73 \\
+            Original PICO-60 ML & 278 & 80.7\% & 76.9\% & 100.0\% \\
             \hline
-            32 & ICN & 85.2\% & 83.4\% & 98.4\% & 0.54 \\
+            Acoustic Parameter & N/A & 100\% & 100\% & 100\% \\
             \hline
-            32 & Supervised & 90.4\% & 89.0\% & 97.1\% & 0.53 \\
+            Supervised & 128 & 96.79\% & 95.85\% & 99.09\% \\
             \hline
-            64 & GD & 79.7\% & 75.5\% & 98.6\% & 0.63 \\
+            Supervised & 256 & 96.97\% & 95.99\% & 99.59\% \\
             \hline
-            64 & ICN & 94.3\% & 92.2\% & 99.8\% & 0.34 \\
+            Pseudo-Labels & 128 & 94.89\% & 93.31\% & 99.36\% \\
             \hline
-            64 & Supervised & 92.7\% & 90.4\% & 98.9\% & 0.41 \\
+            Pseudo-Labels & 256 & 97.01\% & 95.30\% & 100.00\% \\
             \hline
-            128 & GD & 97.0\% & 96.3\% & 99.6\% & 0.23 \\
+            Gravitational Differentiation & 128 & 94.78\% & 96.68\% & 99.28\% \\
             \hline
-            128 & ICN & 98.2\% & 97.7\% & 99.4\% & 0.21 \\
-            \hline
-            128 & Supervised & 94.7\% & 92.6\% & 99.5\% & 0.37 \\
-            \hline
-            256 & GD & 99.2\% & 99.0\% & 99.7\% & 0.11 \\
-            \hline
-            256 & ICN & 98.2\% & 97.6\% & 99.5\% & 0.22 \\
-            \hline
-            256 & Supervised & 96.5\% & 94.6\% & 99.7\% & 0.28 \\
-            \hline
-            496 & Supervised & 97.0\% & 95.4\% & 99.9\% & 0.26 \\
-            \hline
-            278 & PICO-60 ML & 80.7\% & 76.9\% & 100.0\% & 0.65 \\
-            \hline
-            N/A & AP & 100\% & 100\% & 100\% & 0.43 \\
+            Gravitational Differentiation & 256 & 98.29\% & 98.75\% & 99.44\% \\
             \hline
         \end{tabular}
+        \captionof{table}{Summary of the most successful semi-supervised learning models.} \label{results_table}
     \end{center}
 \end{minipage}
-% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples256 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples128 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples64 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
-% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples32 | grep dropout0.25 | ./mean_ap_accuracy_across_configurations.py
+
+It can be seen in Table \ref{results_table} that with 128 labeled examples and 368 unlabeled examples, supervised learning outperforms both pseudo-labels and gravitational differentiation.
+This is likely due to the semi-supervised learning algorithms reinforcing non-representative biases present in the small set of data.
+
+However, when 256 labeled examples and 240 unlabeled examples are used, it is observed that pseudo-labels performs almost identically to supervised learning, and that gravitational differentiation outperforms both existing algorithms.
+This indicates that, when the right balance between labeled and unlabeled examples is used, the confidence of predictions on unlabeled examples as used in gravitational differentiation is a meaningful metric and can be used to further optimize a neural network alongside a conventional loss function.
+
+The test set predictions of gravitational differentiation trained on 256 labeled examples can be seen in Figure \ref{grav_test_hist}, with Acoustic Parameter on the X axis for comparison.
+The decision boundaries of AP and the neural network are represented by black lines.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{grav_test_hist}
+    \caption{\label{grav_test_hist} Test set predictions of a gravitational differentiation model, compared against AP.}
+\end{figure}
 
@@
 
-This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like the Acoustic Parameter technique.
+This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like the AP technique.
 This conclusion has been achieved through two rounds of optimization: the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems with the goal of producing a high-accuracy discriminator to complement AP.

### 2019-07-25 visualization/graph_binary_predictions.py

@@ -17,7 +17,7 @@ from utilities.verify_arguments import verify_arguments
 verify_arguments('JSON data file')
 
 # Load the data set from the file, ignoring the run type ground truths
-events, ground_truths, network_outputs = load_test(sys.argv[1])
+events, _, network_outputs = load_test(sys.argv[1])
 # Get the acoustic parameter and neural network score data from the events, if they are present (if one is, both will be)
 if hasattr(events[0], 'logarithmic_acoustic_parameter'):
     acoustic_parameters, original_neural_network_scores = zip(
@@ -30,7 +30,7 @@ else:
     acoustic_parameters = random_values
     original_neural_network_scores = random_values
 # Calculate actual neutron/alpha ground truths based on AP
-#ground_truths = np.array(acoustic_parameters) > 0.25
+ground_truths = np.array(acoustic_parameters) > 0.25
 
 # Iterate over the three criteria standard deviations will be calculated for, with corresponding names and classification thresholds
 for criterion_data, criterion_name, classification_threshold in zip(
@@ -99,8 +99,9 @@ plt.legend(handles=[background_patch, calibration_patch])
 # Label the X and Y axes
 plt.xlabel('Logarithmic Acoustic Parameter')
 plt.ylabel('Neural Network Prediction')
-# Draw a vertical line to represent the AP decision boundary
+# Draw a vertical line to represent the AP decision boundary, and a horizontal line to represent the NN decision boundary
 plt.axvline(0.25, color='black')
+plt.axhline(0.5, color='black')
 # # Restrict the network prediction axis to the range of 0 to 1
 # plt.ylim(0, 1)
 # Display the graph on screen

### 2019-07-25 documentation/paper.tex
More cleanup on paper

@@
 \usepackage{hyperref}
+\usepackage{lineno}
+
+\linenumbers
 
@@
 Next, semi-supervised learning algorithms are applied; a standard algorithm based on pseudo-labels is compared to a novel algorithm called gravitational differentiation.
-The latter is found to replicate the Acoustic Parameter (AP) discriminator previously used in PICO-60 with a mean of 97\% accuracy.
+The latter is found to replicate the Acoustic Parameter (AP) discriminator previously used in PICO-60 with 98.3\% accuracy.
 \end{abstract}
@@
 \begin{itemize}
-    \item In Section 1, we outline generally the PICO-60 experiment and the context in which machine learning is applicable.
-    \item In Section 2, we discuss the specific properties of the experiment and its data, as well as the existing signal versus background discriminator.
-    \item In Section 3, we discuss various machine learning techniques that are applied to develop a new discriminator.
-    \item In Section 4, we document experimentation with supervised learning.
-    \item In Section 5, we document experimentation with semi-supervised learning with the goal of improving data efficiency.
-    \item In Section 6, we conclude this study and discuss its implications.
-    \item In Section 7, we overview the technologies and software used for this study.
+    \item In Section \ref{objective}, we outline generally the PICO-60 experiment and the context in which machine learning is applicable.
+    \item In Section \ref{previous_analysis}, we discuss the specific properties of the experiment and its data, as well as the existing signal versus background discriminator.
+    \item In Section \ref{ml_techniques}, we discuss various machine learning techniques that are applied to develop a new discriminator.
+    \item In Section \ref{supervised}, we document experimentation with supervised learning.
+    \item In Section \ref{semi_supervised}, we document experimentation with semi-supervised learning in order to take advantage of additional unlabeled data.
+    \item In Section \ref{results}, we document the results obtained with each of the techniques explored, and compare their performance.
+    \item In Section \ref{conclusions}, we conclude this study and discuss its implications.
+    \item In Section \ref{technologies}, we overview the technologies and software used for this study.
 \end{itemize}
@@
 
-\section{Previous PICO-60 Analysis}
+\section{Previous PICO-60 Analysis} \label{previous_analysis}
 
@@
 
-\subsection{Discrimination Techniques}
+\subsection{Discrimination Techniques} \label{discrimination_techniques}
 
@@
 
-\section{Machine Learning Techniques}
+\section{Machine Learning Techniques} \label{ml_techniques}
 
@@
 
-\section{Supervised Learning}
+\section{Supervised Learning} \label{supervised}
 
@@
 
+The following hyperparameter range was exhaustively searched:
+\begin{itemize}
+    \item Dropout $\in \{0, 0.25, 0.5\}$
+    \item L2 regularization $\lambda \in \{0, 0.0003, 0.001, 0.003, 0.01\}$
+    \item $j \in \{0.01, 0.02\}$
+    \item $k \in \{1.025, 1.05\}$
+\end{itemize}
+
 \subsection{Gravitational Differentiation}
@@
     \centering
-    \includegraphics[width=0.85\textwidth]{grav_diff}
+    \includegraphics[width=0.75\textwidth]{grav_diff}
     \caption{\label{grav_diff} A visualization of $\mathrm{GravDiff}(p, \psi, g)$ with respect to $p$ and $\psi$.}
 \end{figure}
+
+The following hyperparameter range was exhaustively searched:
+\begin{itemize}
+    \item Dropout = 0.25
+    \item L2 regularization $\lambda = 0.001$
+    \item $\delta_g \in \{0.0005, 0.001, 0.003, 0.005, 0.008\}$
+    \item $\psi \in \{3, 5, 7, 9, 11\}$
+    \item Learning rate $\in \{0.001, 0.003, 0.01, 0.03\}$
+\end{itemize}
 
@@
             \hline
-            Acoustic Parameter & N/A & 100\% & 100\% & 100\% \\
+            Acoustic Parameter & N/A & 100.0\% & 100.0\% & 100.0\% \\
             \hline
-            Supervised & 128 & 96.79\% & 95.85\% & 99.09\% \\
+            Supervised & 128 & 96.8\% & 95.9\% & 99.1\% \\
             \hline
-            Supervised & 256 & 96.97\% & 95.99\% & 99.59\% \\
+            Supervised & 256 & 97.0\% & 96.0\% & 99.6\% \\
             \hline
-            Pseudo-Labels & 128 & 94.89\% & 93.31\% & 99.36\% \\
+            Pseudo-Labels & 128 & 94.9\% & 93.3\% & 99.4\% \\
             \hline
-            Pseudo-Labels & 256 & 97.01\% & 95.30\% & 100.00\% \\
+            Pseudo-Labels & 256 & 97.0\% & 95.3\% & 100.0\% \\
             \hline
-            Gravitational Differentiation & 128 & 94.78\% & 96.68\% & 99.28\% \\
+            Gravitational Differentiation & 128 & 92.0\% & 96.3\% & 99.4\% \\
             \hline
-            Gravitational Differentiation & 256 & 98.29\% & 98.75\% & 99.44\% \\
+            Gravitational Differentiation & 256 & 98.3\% & 98.8\% & 99.4\% \\
             \hline
@@
     \centering
-    \includegraphics[width=\textwidth]{grav_test_hist}
+    \includegraphics[width=0.9\textwidth]{grav_test_hist}
     \caption{\label{grav_test_hist} Test set predictions of a gravitational differentiation model, compared against AP.}
@@
 
-\section{Conclusions}
+\section{Conclusions} \label{conclusions}
 
 This study has demonstrated that it is possible, using semi-supervised learning, to automatically optimize a discriminator with performance comparable to one developed and tuned by humans like the AP technique.
-This conclusion has been achieved through two rounds of optimization: the first to explore and evaluate various network structures and input formats using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems with the goal of producing a high-accuracy discriminator to complement AP.
+This conclusion has been achieved through two rounds of optimization: the first to explore and evaluate various network architectures using supervised learning, and the second to incorporate the most successful configuration into semi-supervised learning systems with the goal of producing a high-accuracy discriminator to complement AP.
 
-In the realm of supervised learning, it was made clear that the banded Fourier transform with position corrections applied is the most effective input format.
-It produced significantly higher accuracy and tighter prediction distributions than neural networks trained on the full-resolution Fourier transform, the raw audio waveform, or the image window data.
-There is strong evidence that image data in particular does not contain sufficient information to be used for discrimination.
-
-Two semi-supervised learning algorithms were developed, applying the best configuration found for supervised learning.
-Both techniques produced very effective discriminators, performing better than supervised learning.
-Iterative cluster nucleation proved more data-efficient with smaller labeled training sets, while gravitational differentiation consistently produced the highest accuracy overall when 256 labeled examples were used.
+After optimization of a supervised learning discriminator, two semi-supervised learning methods were implemented: the conventional algorithm of pseudo-counts, and a new algorithm, gravitational differentiation.
+Using 128 labeled examples, they were found to perform worse than supervised learning.
+When 256 labeled training examples were used, both supervised learning and pseudo-counts reached 97.0\% accuracy, while gravitational differentiation reached 98.3\%, a 43\% reduction in misclassifications.
+This demonstrates the potential of semi-supervised learning to take advantage of unlabeled training examples to produce a machine learning discriminator competitive with manually developed discriminator.
 
@@
 
-\section{Technologies}
+\section{Technologies} \label{technologies}
 
@@
 We acknowledge that this work is supported by the U.S.
-Department of Energy (DOE) Office of Science, Office of High Energy Physics (under award DE-SC-0012161), by the DOE Office of Science Graduate Student Research (SCGSR) award,  by DGAPA-UNAM (PAPIIT No.\:IA100316 and IA100118) and Consejo Nacional de Ciencia y Tecnolog\'ia (CONACyT, M\'exico, Grant No.\:252167), by the Department of Atomic Energy (DAE), the Government of India, under the Center of AstroParticle Physics II project (,CAPP-II) at SAHA Institute of Nuclear Physics (SINP), European Regional Development Fund-Project ``Engineering applications of microworld physics'' (No.\:CZ.02.1.01/0.0/0.0/16\_019/0000766),
-and the Spanish Ministerio de Econom\'ia y Competitividad, Consolider MultiDark (Grant CSD2009-00064).
+Department of Energy (DOE) Office of Science, Office of High Energy Physics (under award DE-SC-0012161), by the DOE Office of Science Graduate Student Research (SCGSR) award,  by DGAPA-UNAM (PAPIIT No.\:IA100316 and IA100118) and Consejo Nacional de Ciencia y Tecnolog\'ia (CONACyT, M\'exico, Grant No.\:252167), by the Department of Atomic Energy (DAE), the Government of India, under the Center of AstroParticle Physics II project (CAPP-II) at SAHA Institute of Nuclear Physics (SINP), European Regional Development Fund-Project ``Engineering applications of microworld physics'' (No.\:CZ.02.1.01/0.0/0.0/16\_019/0000766), and the Spanish Ministerio de Econom\'ia y Competitividad, Consolider MultiDark (Grant CSD2009-00064).
 This work is partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through NSF grant 1125897, and an endowment from the Kavli Foundation and its founder Fred Kavli.
